### 2020-08-25 notasparaarticulo.tex
notas para el primer artículo basado en los archivos de trabajo (detallado) y (resumen). Contiene la obtención de la función de distribución en el equilibrio de un gas ideal cuántico usando el método variacional a la funcional H que hemos definido, la demostración del teorema H usando la aproximación de primer orden (parecida a la de Huang pero no igual).También se incluye la obtención de la función de distribución en el equilibrio para un gas ideal clásico usando la funcion H de Boltzmann y la demostración del teorema H clásico con este método.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage{amsmath}

\title{Quantum H theorem, Variational Entropy, and Relaxation Processes}
\author{labfqot}
\date{March 2020}

\begin{document}

\maketitle

\section{Abstract}
We review in detail the quantum formulation of the Boltzamnns' H theorem. Starting from a H functional defined in the energy space, by means a variational procedure, we demonstrate the analogous of the quantum H theorem. At difference of the Tolmans' procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous system, and show that when $\frac{dH}{dt}=0$ the system reaches the equilibrium condition and then the proposed H functional becomes the well known expression of the entropy of an ideal quantum gas. The time evolution of the system occurs in such a way that $\frac{dH}{dt}<0$. Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium.

%Variational entropy is defined as new $H$ functional that, performing varitational procedures, the Bose-Einstein and Fermi-Dirac distributions are obtained in the quantum scheme and the Maxwell-Boltzman distribution in the classical scheme. The $H$ theorem is proved in the quantum and classical schemes making use of the system is in a relaxation process. In t ssical scheme, a new $H$ functional is proposed according to the $H$ functional proposed by Boltzman. Besides, a ent expression for quantum systems is found in the varational procedure.%

\section{Introduction}
The evolution of the systems is a topic analyzed in many schemes in physics. The time reversal invariance is shown in classical mechanics as a symmetry of systems, that is, for any classical system, if one inverts the flux of time, the system keeps the same evolution\cite{goldstein}. That symmetry is presented in quantum mechanics \cite{sakurai}. That symmetry was broken in statistical mechanics when Boltzmann introduced the $H$-theorem. Using the Boltzmann homogeneous transport equation, the molecular chaos hypothesis and a homogeneous distribution function, Boltzmann showed that the evolution of the system is always in one preferential direction \cite{huang}.\\
In other hand, the Poincaré's theorem establishes that any system placed in any initial conditions, in a certain finite time, the system will be in the same initial conditions, making a paradox with the $H$-theorem \cite{paradox1, paradox2}. That paradox was solved by Boltzmann calculating the time that the system takes to arrive to the initial conditions. Than number is extremely big, even more that the time of life of the universe.
The behavior of the $H$ functional in times bigger than collisions, will be soft and decreasing. When the change of $H$ is zero, the system will be in equilibrium and the corresponding distribution function will be the Maxwell-Boltzmann distribution\cite{huang, patrick}. \\
In Quantum Mechanics, Tolman established the quantum $H$ theorem \cite{tolman} using the probability transition relation, the random phases hypothesis and a homogeneous distribution function. The lack of the quantum Boltzmann transport equation in analogy to classical statistical mechanics has stopped the development of the quantum transport formulation and the quantum non-equilibrium formulation. In order to find answers to conceptual questions for the quantum $H$ theorem, the scientific community has been studying it. The quantum $H$ theorem was proved using not only binary conditions. That work uses n-colissions to prove the quantum $H$ theorem, this idea completes in a classical scheme the $H$ theorem \cite{binary}. There are some special conditions in quantum mechanics where the second law of thermodynamics is violated \cite{contradictions}, making that the quantum $H$ theorem is not thoroughly understood.\\
In transport phenomena, the Boltzmann transport equation is solved using the zero and first order approximation. When one uses the zero order approximation, one finds that the total heat flux in the system is zero, that is a contradiction. However, when one applies the first order approximation, one finds the evolution operator expressed by the left expression from the Boltzmann transport equation. That evolution operator tends to zero when time tends to infinity, this is the equilibrium condition\cite{reif}. This result, that describes inhomogeneties, is not expressed in the $H$ theorem.\\
In order to contribute to a consistent $H$ theorem, in this paper, the quantum and classical $H$ theorem will be proved proposing a new $H$ functional for both cases considering a non-homogeneous distribution function and a system divided into imaginary cells. With those assumptions, particles transport is allowed inside the system. Performing variational procedures, one finds that, in the quantum case, Bose-Einstein and Fermi-Dirac distributions minimize the $H$ functional and, in the classical case, Maxwell-Boltzmann distribution minimize the other $H$ functional. A expression of entropy to quantum systems will be obtained in the variational procedure. In other hand, making use of the hypothesis that the system is in a relaxation process and the non-homogeneous distribution function, the $H$ theorem will be proved for both cases.

\section{Equilibrium}
We begin by defining the form of the functional $H$
\begin{eqnarray}
    H(t)&=&\sum_{M} \int_V \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]\mathrm{d}V \label{entropy}.
\end{eqnarray}
This functional, in equilibrium should be a density of entropy, which we will call variational entropy.
This functional describes a system with global energy E, global number of free
quantum particles (which may be fermions for upper sing or bosons for lower sing) N, global volume V, and temperature T. The
global system in turn consists of a set of $K$ small cells, whose local distribution functions, $f_{Mn}(\epsilon_{n},t)$, are
identially zero outside the cell’s volume. The distribution function will depend on the position $\bar{r}$, the energy per cell $\epsilon_{n}$ and time $t$. We will use the uppercase latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 . . . , K$. Without loss of generality, we choose these cells to have
local identical volumes (\textit{i.e.} $V_M = V/K, \forall M$). Each cell has its own local chemical potential, $\mu_M$, local
$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution.
function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
\begin{eqnarray}
    H_M&=& \int_V \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]\mathrm{d}V \label{entropycell},\\
    {\mathcal{N}}_M&=&\int_V \sum_{n}f_{Mn}(\epsilon_{n} ,t)\mathrm{d}V, \nonumber \\
{\mathcal{E}}_M&=&\int_V \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\mathrm{d}V.
\end{eqnarray}
When the system is in equilibrium, the quantities $\bar{\mathcal{N}}_M$ and $\bar{\mathcal{E}}_M$ will not depend on the cell number. This is
\begin{equation}
   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
\end{equation}
The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
Consequently, we can assume that the available quantum states do not depend on the cell features,
therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$. In addition, the system is subjected to the microcanonical restrictions
\begin{eqnarray}
    &&\sum_{M} \left[\int_V \sum_{n}f_{Mn}(\epsilon_{n} ,t)\mathrm{d}V\right]=\sum_{M} {\mathcal{N}}_{M}=N; \nonumber \\
    &&\sum_{M}\left[\int_V \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{Mn}\mathrm{d}V\right]=\sum_M {\mathcal{E}}_M=E, \label{restriccions1}
\end{eqnarray}
where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
Inspired by the Hamilton principle, H will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining
\begin{eqnarray}
&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
\end{eqnarray}
where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
\begin{equation}
    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
\end{equation}{}
If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, one obtains
\begin{equation}
    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
\end{equation}{}
and finally
\begin{eqnarray}
    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
\end{eqnarray}{}
It is necessary to remark that the microcanonical ensemble is used, but the Bose-Einstein and Fermi-Dirac distribution functions are obtained from the grant canonical ensemble. This result is a consequence of describing the system by its cells. The appearing of the chemical potential is a consequence of this description. Equiprobability assumption from the microcanonical ensemble is recovered because of the chemical potential and the temperature are equal and constants in equilibrium, then the total distribution function is a constant.\\    
From here, one can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
\begin{equation}
      H(t)=\sum_M \sum_n \int_V \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\mathrm{d}V \pm \int_V \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right]\mathrm{d}V\label{H-entropy}.
  \end{equation}
  This result lets us calculate the entropy of a quantum system with inhomogenities, this functional could be used as a thermodynamic variable to describe quantum systems.
\section{The change of H for relaxation processes}

The functional H has the following form
\begin{equation}
    H(t)=\sum_{M} \int_V\sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right]\mathrm{d}V \label{entropy2},
\end{equation}
where, for simplicity, we have defined $f(\bar{r},\epsilon_{n},t)\equiv f_{Mn}$.
These functional has the following microcanonical restrictions
\begin{eqnarray}
        \int_V\sum_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int_V\sum_{n}\epsilon_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
  \end{eqnarray}{}
  where $\bar \mathcal{ N}$ and $\bar \mathcal{E}$ are the mean particle number and the mean energy by cell in equilibrium
  \begin{equation}
      \bar{\mathcal{N}}=\int_V \sum_n \bar{f}_n; \ \ \ \ \bar{\mathcal{E}}=\int_V \sum_n \epsilon_n\bar{f}_n
  \end{equation}
  and $\Delta_M,\delta_M$ could be seen as a deviation from the mean number of particles and mean energy by cell from the equilibrium respectively, with $\Delta_M(t)\ll N$ and $\delta_M(t) \ll E$. Those condition described a system out of equilibrium but no so far from it.  \\
If one performs the derivative of the functional H with respect to time
\begin{equation}
   \frac{dH(t)}{dt}= \sum_n \sum_M \int_{V}\dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right]\mathrm{d}V.\label{deltaH}
\end{equation}{}
Using the first-order approximation, that is
\begin{eqnarray}
   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}g_{Mn}, \label{firstorder}
\end{eqnarray}{}
one obtains

\begin{eqnarray}
    \frac{dH(t)}{dt}&=&\sum_n \sum_M \int_V\bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM}\mathrm{d}V \nonumber \\
    &=&\sum_n \sum_M \int_V\bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\mathrm{d}V.\nonumber \\
    \label{cambioH1}
\end{eqnarray}{}
Here the deviation $g_{Mn}=g_{Mn}(\bar{r},\epsilon_{n},t)$.\\
It's necessary to remark that, in the case of fermions
\begin{eqnarray}
   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}.
\end{eqnarray}{}
The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$. Furthermore, whether $\bar f_{n}=0$, this is, all energy levels are occupied, then the system will not have inhomogenities because of the exclusion principle and as consequence of this,  $g_{nM}=0$.\\
In the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$. In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$. With those relations, one can approximate the logarithm functions to their first-order Taylor series. Besides, this approximation is for both signs for $g_{nM}$.
That approximation applied to logarithm functions in (\ref{cambioH1}) around $\bar f_n g_{nM}=0$ is
\begin{equation}
    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}. \label{lnapproximation}
\end{equation}{}
with the previous approximation, (\ref{cambioH1}) will be
\begin{eqnarray}
    \frac{dH}{dt}&=&\sum_n \sum_M \int_V\bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\}\mathrm{d}V \nonumber \\
    &&-\sum_{n}\sum_{M}\int_V\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\mathrm{d}V.\label{cambioH2}
\end{eqnarray}{}
Making use of the expression (\ref{relation}),
(\ref{cambioH2}) casts into
\begin{eqnarray}
    \frac{dH}{dt}&=&\sum_n \sum_M \int_V\bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\mathrm{d}V. \nonumber \\
    \label{cambioH3}
\end{eqnarray}{}
On the other hand, one can obtain from the restrictions the following expressions
\begin{eqnarray}
    &&\int_V\sum_n \bar{f}_n g_{nM}\mathrm{d}V=\Delta_M(t) \ \  \Rightarrow \ \ \int_V \sum_n \bar{f}_n \dot{g}_{nM}\mathrm{d}V=\dot{\Delta}_M(t), \nonumber \\
    &&\sum_n \int_V \bar{f}_n g_{nM}\epsilon_n\mathrm{d}V=\delta_M(t) \ \  \Rightarrow \ \ \int_V \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n\mathrm{d}V=\dot{\delta}_M(t)
\end{eqnarray}{}
and as a consequence of $\sum_M \Delta_M(t)=\sum_M \delta_M(t)=0$, one finds
\begin{equation}
    \sum_M \dot{\Delta}_M(t)=\sum_M \dot{\delta}_{M}(t)=0.
\end{equation}{}
Substituting the previous expression to (\ref{cambioH3}) one obtains
\begin{eqnarray}
   \frac{dH}{dt}&=& \bar{\alpha}\sum_M \dot{\Delta}_M(t)+\bar{\beta}\sum_M \dot{\delta}_M(t)+\sum_n \sum_M \int_V\left \{ \bar{f}_{n}g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\mathrm{d}V \nonumber \\
   &=&\sum_n  \bar{f}_{n}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V= \sum_n \left(\frac{1\pm e^{\bar{\alpha}+\bar{\beta}\epsilon_n}}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_n}\pm 1} \right)\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V. \nonumber \\ \label{cambioH4}
\end{eqnarray}{}
Now, the summation over M will be expressed in two summations
\begin{equation}
    \frac{dH}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L}\int_V g^{+}_{nJ}\dot{g}^{+}_{nJ}\mathrm{d}V+\sum^{N}_J \int_V g^{-}_{nJ}\dot{g}^{-}_{nJ}\mathrm{d}V \right), \label{cambioH5}
\end{equation}{}
where $L+N=M$. Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or missing of particles or energy respectively.\\
On the one hand, $\dot{g}^{+}_{nJ}<0$ as a consequence of lost of particles or energy and so , $g^{+}_{nJ}>0$. On the other hand, $\dot{g}^{-}_{nJ}>0$ as a consequence of gain of particles or energy and therefore $g^{-}_{nJ}<0$. As a result of the previous idea
\begin{eqnarray}
   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
\end{eqnarray}{}
\\
and consequently, (\ref{cambioH5}) obtains the following form
\begin{equation}
    \frac{dH}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L}\int_V |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\mathrm{d}V+\sum^{N}_J \int_V |g^{-}_{nJ}||\dot{g}^{-}_{nJ}|\mathrm{d}V \right). \label{cambioH6}
\end{equation}{}
In conclusion, $\frac{dH}{dt}<0$.\\
When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) that $\frac{dH}{dt}=0$.
with the previous prove, in analogous to the classical scheme, the homogeneous distribution function hypothesis must be eliminated to be consistent with this prove. Now, in the following section, variational procedure and the relaxation processes will be used in the classical scheme to prove the $H$ theorem eliminating the homogeneous distribution function hypothesis.

\section{Classical scheme}
\subsection{Variational procedure}
We proposed the following $H$ functional

\begin{equation}
   H(t)=\int_{}^{}\int_{}^{} f(r,v,t) \ln f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{CH2}.
\end{equation}{}
The functional (\ref{CH2}) contains a no-homogeneous distribution function but keeps the same form of the Boltzman's $H$ functional. Excluding the homogeneous distribution function hypothesis, one uses the same procedure, that it was shown in the previous section, to prove the $H$-theorem for relaxation processes.\\ 
If one considers the microcanonical ensemble conditions 
\begin{equation}
    \int_{}^{}\int_{}^{}f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N, \ \ \ \int_{}^{}\int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v \mathrm{d}^3r=E \label{micro},
\end{equation}{}
one obtains the distribution function that minimizes $H$
\begin{eqnarray}
\frac{\delta H}{\delta f(r',v')}&=&\int_{}^{}\frac{\delta}{\delta f(r',v')}\left[f(r,v)\ln f(r,v)  \right]\mathrm{d}^3v \mathrm{d}^3r-\int_{}^{}\alpha(r)\int_{}^{}\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
&&-\int_{}^{}\beta(r) \int_{}^{}\epsilon(v)\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
&=&\ln f(r',v')+1-\alpha(r')-\beta(r') \epsilon(v')=0.
\end{eqnarray}{}
isolating the distribution function 
\begin{eqnarray}
\ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f(r',v')=e^{\alpha(r') +\beta(r') \epsilon(v')-1} \nonumber \\
&=&Ce^{\alpha(r')+\beta(r') \epsilon(v') } \label{relacion1},
\end{eqnarray}{}
where $C$ is a constant. It is trivial that using the variational procedures, one obtains the form of the Maxwell-Boltzmann distribution. \\
As $H$ in equilibrium is a density of entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
\begin{equation}
    f(v)=Ce^{\alpha+\beta \epsilon(v)}.
\end{equation}{}

\subsubsection{Relaxation processes}
Now one uses the restrictions (\ref{restrictionoutside}) but in the continuous case
\begin{equation}
    \int_{}^{}f(r,v,t)\mathrm{d}^3v=\mathcal{N}+\Delta(r,t); \ \ \ \int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v=\mathcal{E}+\Omega(r,t)\label{restriccionescambio},
\end{equation}{}
where $\Delta(r,t)$ y $\Omega(r,t)$ could be seen as a deviation from the mean number of particles and mean energy by differential element $d^3r$.
Performing the derivative of the functional $H$ (\ref{CH2})with respect to time
\begin{equation}
    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\left[ 1+\ln f(r,v,t) \right]\dot f(r,v,t) \mathrm{d}^3v \mathrm{d}^3r \label{dH1}.
\end{equation}{}
Applying the first order approximation (\ref{firstorder}) to (\ref{dH1}), one obtains
\begin{equation}
    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g(r,v,t) \right\} \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,1}.
\end{equation}{}
Using the same approximation (\ref{lnapproximation})
to (\ref{dH1,1}) 
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g(r,v,t) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r,
\end{eqnarray}{}
using (\ref{relacion1}) with homogeneous multipliers
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r+\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,2}. \nonumber \\
\end{eqnarray}{}
If one uses the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), one obtains the same result. Then, applying that result to (\ref{dH1,2}), one finds
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r.
\end{eqnarray}{}
Using the same consideration in the previous section to $g$ and $\dot g$, one proves that $\frac{dH}{dt}<0$. If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH}{dt}=0$.

\section{Discussion}
For quantum ideal gases, the H-theorem has been proved performing the variational method. In the classical scheme, the $H$-theorem is proved using the Boltzmann transport equation for a homogeneous distribution function. Classical $H$-theorem don't show the idea of particles transport in a system. That assumption in quantum mechanics presented by Tolman \cite{tolman} remains. This work eliminates that assumption defining the $H$ functional (\ref{entropy}) including a non-homogeneous distribution function. The prove of the quantum $H$-theorem in this work includes the idea of particles transport in relaxation processes using the first order approximation. This idea has not found in the literature. As consequence of this, all ideal gases placed in non-equilibrium state with inhomogeneities, when time tends to infinity, the system tends to the equilibrium state. This result complements the classical $H$-theorem due to the lack of the non-homogeneous distribution function. In addition, the classical $H$ theorem was analysed to include the particles transport and complete the classical $H$-theorem modifying the classical $H$ functional proposed by Boltzmann. Those proves only are applied to relaxation processes but the fact of including a non-homogeneous distribution function is a new way to complete the $H$ theorem for non-equilibrium systems. 
In other hand, the equation (\ref{H-entropy}) lets us calculate the entropy for a quantum system and describe the evolution of that system using the $H$ functional as a thermodynamic variable for a non-equilibrium system. Furthermore, thermodynamic variables to describe non-equilibrium systems is a problem in discussion. The previous idea adds new knowledge to this problem.

\begin{thebibliography}{}
\bibitem{goldstein}
H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).

\bibitem{sakurai}
J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.

\bibitem{huang}
K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).

\bibitem{paradox1}
Harvey R. Brown and Wayne Myrvold, {\it Boltzmann’s H-theorem, its limitations, and thebirth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.

\bibitem{paradox2}
Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.

\bibitem{patrick}
R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).

\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.

\bibitem{binary}
Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.

\bibitem{contradictions}
Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.

\bibitem{reif}
F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.

\end{thebibliography}{}
\end{document}

### 2020-10-09 notasparaarticulo.tex
classic limit was intruduced, conclusions and introduction were modified and the H functional was modified and references were added.

@@
 \usepackage[english]{babel}
-\usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage{xcolor}
 
-\title{Quantum H theorem, Variational Entropy, and Relaxation Processes}
+\title{Quantum H theorem, variational entropy, and relaxation processes}
 \author{labfqot}
@@
 \section{Abstract}
-We review in detail the quantum formulation of the Boltzamnns' H theorem.
-Starting from a H functional defined in the energy space, by means a variational procedure, we demonstrate the analogous of the quantum H theorem.
-At difference of the Tolmans' procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous system, and show that when $\frac{dH}{dt}=0$ the system reaches the equilibrium condition and then the proposed H functional becomes the well known expression of the entropy of an ideal quantum gas.
-The time evolution of the system occurs in such a way that $\frac{dH}{dt}<0$.
-Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium.
-
-%Variational entropy is defined as new $H$ functional that, performing varitational procedures, the Bose-Einstein and Fermi-Dirac distributions are obtained in the quantum scheme and the Maxwell-Boltzman distribution in the classical scheme.
-The $H$ theorem is proved in the quantum and classical schemes making use of the system is in a relaxation process.
-In t ssical scheme, a new $H$ functional is proposed according to the $H$ functional proposed by Boltzman.
-Besides, a ent expression for quantum systems is found in the varational procedure.%
+We review in detail the quantum formulation of the Boltzamnn's H theorem.
+Starting from a H functional defined in the energy space.
+By means a variational procedure, we demonstrate the analogous of the quantum H theorem.
+At difference of the Tolman's procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous system, and show that when $\frac{dH}{dt}=0$, the system reaches the equilibrium condition and then proposed H functional becomes the well known expression of the entropy of an ideal quantum gas.
+The time evolution of the system happens in such a way that $\frac{dH}{dt}<0$.
+Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium, Finally, in the limit of high energy levels, the Boltzmann's $H$ functional is recovered.
 
 \section{Introduction}
-The evolution of the systems is a topic analyzed in many schemes in physics.
-The time reversal invariance is shown in classical mechanics as a symmetry of systems, that is, for any classical system, if one inverts the flux of time, the system keeps the same evolution\cite{goldstein}.
-That symmetry is presented in quantum mechanics \cite{sakurai}.
-That symmetry was broken in statistical mechanics when Boltzmann introduced the $H$-theorem.
-Using the Boltzmann homogeneous transport equation, the molecular chaos hypothesis and a homogeneous distribution function, Boltzmann showed that the evolution of the system is always in one preferential direction \cite{huang}.\\
-In other hand, the Poincaré's theorem establishes that any system placed in any initial conditions, in a certain finite time, the system will be in the same initial conditions, making a paradox with the $H$-theorem \cite{paradox1, paradox2}.
-That paradox was solved by Boltzmann calculating the time that the system takes to arrive to the initial conditions.
-Than number is extremely big, even more that the time of life of the universe.
-The behavior of the $H$ functional in times bigger than collisions, will be soft and decreasing.
+
+\textcolor{red}{In the literature, the basis for equilibrium systems in the classical scheme is well-established.
+From thermodynamics to statistical mechanics, the procedures for equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature \cite{huang, reif, patrick}.
+This is not the case of out of equilibrium systems.
+There are some well-established theoretical schemes to deal with out of equilibrium systems in the classical scheme, such as the Onsager formulation (called linear thermodynamics), some few other restricted developments proposed outstanding theoretical authors \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.
+The Kinetic Theory of dilute gases describes the behavior of a dilute classical gas by means the Boltzmann transport equation.
+Furthermore, the evolution to a equilibrium state for any system can be proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
+However, there are some some aspects in the Boltzmann's $H$ theorem that deserve some review in order to accomplish better consistency and is still in discussion [6-10].
+%\cite{paradox1, paradox2, hamiltonian, H-theorem-violation, H-theorem-and-entropy}.
+In order to obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the $H$ theorem and the Boltzmann transport equation are necessary such that in the appropriate limit, the  Boltzmann's $H$ theorem and the Boltzmann transport equation should be recovered.\\
+One of the first developments in this field was developed by Tolman \cite{tolman}.
+He establishes the quantum version of the $H$ theorem using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
+The equation of motion for the occupation number was obtained by performing the time perturbation theory.
+However, in order to obtain the flux relations in analogy to the Boltzmann's method, it is required a quantum version of the transport equation that includes a non-homogeneous distribution function.\\
+In another hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (the analogous to the Boltzmann transport equation) [12-15].
+%\cite{entropic-framework, thermal-harmonic, quantum-transport, quantum-transport-reservoirs}.
+Besides some papers construct the $H$ functional and prove the $H$ theorem \cite{htheorem2, quantum1, quantum2}.
+However, the homogeneous-hypothesis and the correspondence principle are not clear or discussed.
+Moreover the quantum $H$ theorem, the quantum entropy and the second law of thermodynamics in quantum mechanics is still in discussion [19-26].\\ %\cite{quantum-entropy, quantum-colapse, quantum3, quantum4, binary, contradictions,ergodic-theorem, hydrodynamic-simulation}.\\
+In order to solve those problems, in this work we define a new $H$ functional, which name is variational entropy.
+This variational entropy includes as main variable the distribution function of an ideal quantum gas (Bosons or Fermions particles) inside a volume $V$ divided into cells.
+Before developing the non-equilibrium description, the statistical mechanics results for equilibrium systems are recovered from the variational entropy.
+Furthermore we prove that the behavior of the variational entropy in time is less or equal to zero for systems that are close enough to the equilibrium state.
+This prove corresponds with the $H$ theorem.
+In addition, the Boltzmann transport equation is obtained from the Boltzmann's $H$ functional and, in analogy with this, a quantum version of the Boltzmann transport equation is obtained from the variational entropy.
+Finally, in order to be consistence with the correspondence principle, the Boltzmann's $H$ functional is recovered from the variational entropy in the limit of the high level energies.}
+
+%It is well known for the Physics community that the set of schemes, methods, techniques, and approximations of Statistical Physics to deal with equilibrium systems, are relatively well established.
+Liouvilles' theorem, Ergodic hypothesis and the main corollary of the H theorem, are some of the conceptual basis on which the theoretical structures of Statistical Thermodynamics have been grown and are sustained \cite{huang, reif, patrick}.
+This is not the case of the Statistical Physics of out of equilibrium systems, in which Poincares' theorem, Ergodic hypothesis, and H theorem establish the general conceptual basis of the theoretical developments, but these later are not as general and satisfactory as those for equilibrium [ref].
+There are of course some well stablished theoretical schemes to deal with out of equilibrium systems like those developed by Onsager known as linear thermodynamics, and some few other restricted developments proposed by some outstanding theoretical authors \cite{kei, onsager}.
+Nevertheless, even the very fundamentals of out of equilibrium Statistical physics like Boltzmanns' H Theorem, still presents some aspects that deserve some review in order to accomplish better consistency [ref].
+In the framework of classical mechanics, the prove of the Boltzmanns' H theorem in the scheme of Kinetic Theory of dilute gases, is founded on the Boltzmann Transport  Equation, which in turn is supported on the Molecular Chaos Hypothesis.
+%It is commonly accepted that BHT solves in a more complete way the Poicares' paradox and concilies the the theoretical conflict between a time reversible microscopic dynamics and an inherent macroscopic one \cite{paradox1, paradox2}.
+The quantum version of H Theorem, still currently is under analysis and discussion \cite{quantum1, quantum2, quantum3, quantum4}.
+
+%The behavior of the $H$ functional in times bigger than collisions, will be soft and decreasing.
 When the change of $H$ is zero, the system will be in equilibrium and the corresponding distribution function will be the Maxwell-Boltzmann distribution\cite{huang, patrick}.
-\\
-In Quantum Mechanics, Tolman established the quantum $H$ theorem \cite{tolman} using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
+\\%
+
+%revisar la redaccion y coherencia de estas ideas% 
+%One of those discussions is the propose of Tolman.
+He established the quantum $H$ theorem \cite{tolman} using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
 The lack of the quantum Boltzmann transport equation in analogy to classical statistical mechanics has stopped the development of the quantum transport formulation and the quantum non-equilibrium formulation.
-In order to find answers to conceptual questions for the quantum $H$ theorem, the scientific community has been studying it.
-The quantum $H$ theorem was proved using not only binary conditions.
-That work uses n-colissions to prove the quantum $H$ theorem, this idea completes in a classical scheme the $H$ theorem \cite{binary}.
+Another discission is the prove of the quantum $H$ theorem using n-colissions.
+This idea completes the $H$ theorem in a classical and quantum scheme \cite{binary}.
 There are some special conditions in quantum mechanics where the second law of thermodynamics is violated \cite{contradictions}, making that the quantum $H$ theorem is not thoroughly understood.\\
-In transport phenomena, the Boltzmann transport equation is solved using the zero and first order approximation.
+%In transport phenomena, the Boltzmann transport equation is solved using the zero and first order approximation.
 When one uses the zero order approximation, one finds that the total heat flux in the system is zero, that is a contradiction.
 However, when one applies the first order approximation, one finds the evolution operator expressed by the left expression from the Boltzmann transport equation.
-That evolution operator tends to zero when time tends to infinity, this is the equilibrium condition\cite{reif}.
+That evolution operator tends to zero when time tends to infinity, this is the equilibrium condition \cite{reif}.
 This result, that describes inhomogeneties, is not expressed in the $H$ theorem.\\
-In order to contribute to a consistent $H$ theorem, in this paper, the quantum and classical $H$ theorem will be proved proposing a new $H$ functional for both cases considering a non-homogeneous distribution function and a system divided into imaginary cells.
+%In order to contribute to a consistent $H$ theorem, in this paper, the quantum and classical $H$ theorem are proved proposing a new $H$ functional for both cases considering a non-homogeneous distribution function and a system divided into imaginary cells.
 With those assumptions, particles transport is allowed inside the system.
@@
 
-\section{Equilibrium}
-We begin by defining the form of the functional $H$
+\section{The equilibrium of dilute gases: a review}
+\subsection{Classical approach}
+In the classical version, the Boltzmann transport equation is used to shown the $H$-theorem. That equation is
+\begin{equation}
+    \left( \frac{\partial }{\partial t}+v_1 \cdot \nabla_r +\frac{F}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|v_1-v_2|(f_2'f_1'-f_2f_1),
+\end{equation}
+where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system and $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively.\\
+The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
+\begin{equation}
+    H=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,
+\end{equation} \label{hfunctional}
+where $f(v,t)$ is a homogeneous distribution function.
+The $H$ theorem establishes that if $f(v,t)$ satisfies the Boltzmann transport equation, then $\frac{dH}{dt}\leq 0$.
+The system arrives to the equilibrium state when $\frac{dH}{dt}=0$ and due to this, the distribution function to the system in equilibrium corresponds to the Maxwell-Boltzmann distribution
+\subsection{Quantum approach}
+The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
+\begin{equation}
+    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
+\end{equation}
+where $n_i$ is the occupation number and $g_i$ is a energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is obtained
 \begin{eqnarray}
-    H(t)&=&\sum_{M} \int_V \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]\mathrm{d}V \label{entropy}.
+    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
+    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
 \end{eqnarray}
+where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
+\begin{equation}
+  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
+\end{equation}
+where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expression are obtained by the perturbation theory.\\
+The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}).
+This prove is in the same way as in the classical version.
+That prove is developed taking account of the homogeneous distribution function, as in the classical version.\\
+In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high energy levels to
+\begin{equation}
+    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}).
+\end{equation}
+This expression can be obtained also from the Boltzmann $H$ functional by means defining 
+\begin{equation}
+    f=\frac{n_{\kappa}}{ g_{\kappa}}.
+\end{equation}
+\section{Ensemble approach}
+\subsection{Quantum scheme}
+
+\textcolor{red}{We begin by defining the form of the functional $H$
+\begin{eqnarray}
+    H(t)&=&\sum_{M} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \delta V_M \delta \epsilon_n \label{entropy}.
+\end{eqnarray}}
 This functional, in equilibrium should be a density of entropy, which we will call variational entropy.
@@
 quantum particles (which may be fermions for upper sing or bosons for lower sing) N, global volume V, and temperature T. The
-global system in turn consists of a set of $K$ small cells, whose local distribution functions, $f_{Mn}(\epsilon_{n},t)$, are
-identially zero outside the cell’s volume.
-The distribution function will depend on the position $\bar{r}$, the energy per cell $\epsilon_{n}$ and time $t$.
+global system in turn consists of a set of $K$ small cells with a constant volume $V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
+The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$ and time $t$ and the energy spectra will be divided into $g_n$ volumes.
 We will use the uppercase latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
@@
 \begin{eqnarray}
-    H_M&=& \int_V \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]\mathrm{d}V \label{entropycell},\\
-    {\mathcal{N}}_M&=&\int_V \sum_{n}f_{Mn}(\epsilon_{n} ,t)\mathrm{d}V, \nonumber \\
-{\mathcal{E}}_M&=&\int_V \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\mathrm{d}V.
+    H_M&=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \delta \epsilon_n \label{entropycell},\\
+    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \delta \epsilon_n, \nonumber \\
+{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \delta \epsilon_n.
 \end{eqnarray}
-When the system is in equilibrium, the quantities $\bar{\mathcal{N}}_M$ and $\bar{\mathcal{E}}_M$ will not depend on the cell number. This is
+When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number. This is
 \begin{equation}
@@
 \begin{eqnarray}
-    &&\sum_{M} \left[\int_V \sum_{n}f_{Mn}(\epsilon_{n} ,t)\mathrm{d}V\right]=\sum_{M} {\mathcal{N}}_{M}=N; \nonumber \\
-    &&\sum_{M}\left[\int_V \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{Mn}\mathrm{d}V\right]=\sum_M {\mathcal{E}}_M=E, \label{restriccions1}
+    &&\sum_{M} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\right]\delta V_M \delta \epsilon_n=\sum_{M} {\mathcal{N}}_{M} \delta \epsilon_n=N; \nonumber \\
+    &&\sum_{M}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{Mn}\right]\delta V_M \delta g_n=\sum_M {\mathcal{E}}_M \delta \epsilon_n=E, \label{restriccions1}
 \end{eqnarray}
 where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Inspired by the Hamilton principle, H will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining
+Inspired by the Hamilton principle, $H$ will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
+where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
+\begin{equation}
+    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
+\end{equation}
+where we have used that the $H$ functional in equilibrium is proportional to the entropy of the system.
+In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
+With this, we conclude
+\begin{equation}
+    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
+\end{equation}
+In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
 \begin{equation}
@@
 \end{eqnarray}{}
-It is necessary to remark that the microcanonical ensemble is used, but the Bose-Einstein and Fermi-Dirac distribution functions are obtained from the grant canonical ensemble.
-This result is a consequence of describing the system by its cells.
-The appearing of the chemical potential is a consequence of this description.
-Equiprobability assumption from the microcanonical ensemble is recovered because of the chemical potential and the temperature are equal and constants in equilibrium, then the total distribution function is a constant.\\
-From here, one can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
-\begin{equation}
-      H(t)=\sum_M \sum_n \int_V \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\mathrm{d}V \pm \int_V \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right]\mathrm{d}V\label{H-entropy}.
-  \end{equation}
-  This result lets us calculate the entropy of a quantum system with inhomogenities, this functional could be used as a thermodynamic variable to describe quantum systems.
-\section{The change of H for relaxation processes}
-
-The functional H has the following form
-\begin{equation}
-    H(t)=\sum_{M} \int_V\sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right]\mathrm{d}V \label{entropy2},
-\end{equation}
-where, for simplicity, we have defined $f(\bar{r},\epsilon_{n},t)\equiv f_{Mn}$.
-These functional has the following microcanonical restrictions
+We have proved that when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multiplier are homogeneous in the system.\\
+We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
 \begin{eqnarray}
-        \int_V\sum_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int_V\sum_{n}\epsilon_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}{}
-  where $\bar \mathcal{ N}$ and $\bar \mathcal{E}$ are the mean particle number and the mean energy by cell in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}=\int_V \sum_n \bar{f}_n; \ \ \ \ \bar{\mathcal{E}}=\int_V \sum_n \epsilon_n\bar{f}_n
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from the mean number of particles and mean energy by cell from the equilibrium respectively, with $\Delta_M(t)\ll N$ and $\delta_M(t) \ll E$.
-Those condition described a system out of equilibrium but no so far from it.
-\\
-If one performs the derivative of the functional H with respect to time
-\begin{equation}
-   \frac{dH(t)}{dt}= \sum_n \sum_M \int_{V}\dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right]\mathrm{d}V.\label{deltaH}
-\end{equation}{}
-Using the first-order approximation, that is
-\begin{eqnarray}
-   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}g_{Mn}, \label{firstorder}
-\end{eqnarray}{}
-one obtains
-
-\begin{eqnarray}
-    \frac{dH(t)}{dt}&=&\sum_n \sum_M \int_V\bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM}\mathrm{d}V \nonumber \\
-    &=&\sum_n \sum_M \int_V\bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\mathrm{d}V.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}{}
-Here the deviation $g_{Mn}=g_{Mn}(\bar{r},\epsilon_{n},t)$.\\
-It's necessary to remark that, in the case of fermions
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}.
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=0$, this is, all energy levels are occupied, then the system will not have inhomogenities because of the exclusion principle and as consequence of this,  $g_{nM}=0$.\\
-In the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, one can approximate the logarithm functions to their first-order Taylor series.
-Besides, this approximation is for both signs for $g_{nM}$.
-That approximation applied to logarithm functions in (\ref{cambioH1}) around $\bar f_n g_{nM}=0$ is
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}{}
-with the previous approximation, (\ref{cambioH1}) will be
-\begin{eqnarray}
-    \frac{dH}{dt}&=&\sum_n \sum_M \int_V\bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\}\mathrm{d}V \nonumber \\
-    &&-\sum_{n}\sum_{M}\int_V\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\mathrm{d}V.\label{cambioH2}
-\end{eqnarray}{}
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-\begin{eqnarray}
-    \frac{dH}{dt}&=&\sum_n \sum_M \int_V\bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\mathrm{d}V.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}{}
-On the other hand, one can obtain from the restrictions the following expressions
-\begin{eqnarray}
-    &&\int_V\sum_n \bar{f}_n g_{nM}\mathrm{d}V=\Delta_M(t) \ \  \Rightarrow \ \ \int_V \sum_n \bar{f}_n \dot{g}_{nM}\mathrm{d}V=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n \int_V \bar{f}_n g_{nM}\epsilon_n\mathrm{d}V=\delta_M(t) \ \  \Rightarrow \ \ \int_V \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n\mathrm{d}V=\dot{\delta}_M(t)
-\end{eqnarray}{}
-and as a consequence of $\sum_M \Delta_M(t)=\sum_M \delta_M(t)=0$, one finds
-\begin{equation}
-    \sum_M \dot{\Delta}_M(t)=\sum_M \dot{\delta}_{M}(t)=0.
-\end{equation}{}
-Substituting the previous expression to (\ref{cambioH3}) one obtains
-\begin{eqnarray}
-   \frac{dH}{dt}&=& \bar{\alpha}\sum_M \dot{\Delta}_M(t)+\bar{\beta}\sum_M \dot{\delta}_M(t)+\sum_n \sum_M \int_V\left \{ \bar{f}_{n}g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\mathrm{d}V \nonumber \\
-   &=&\sum_n  \bar{f}_{n}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V= \sum_n \left(\frac{1\pm e^{\bar{\alpha}+\bar{\beta}\epsilon_n}}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_n}\pm 1} \right)\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V.
-\nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over M will be expressed in two summations
-\begin{equation}
-    \frac{dH}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L}\int_V g^{+}_{nJ}\dot{g}^{+}_{nJ}\mathrm{d}V+\sum^{N}_J \int_V g^{-}_{nJ}\dot{g}^{-}_{nJ}\mathrm{d}V \right), \label{cambioH5}
-\end{equation}{}
-where $L+N=M$.
-Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or missing of particles or energy respectively.\\
-On the one hand, $\dot{g}^{+}_{nJ}<0$ as a consequence of lost of particles or energy and so , $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ as a consequence of gain of particles or energy and therefore $g^{-}_{nJ}<0$.
-As a result of the previous idea
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
-and consequently, (\ref{cambioH5}) obtains the following form
-\begin{equation}
-    \frac{dH}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L}\int_V |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\mathrm{d}V+\sum^{N}_J \int_V |g^{-}_{nJ}||\dot{g}^{-}_{nJ}|\mathrm{d}V \right).
-\label{cambioH6}
-\end{equation}{}
-In conclusion, $\frac{dH}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) that $\frac{dH}{dt}=0$.
-with the previous prove, in analogous to the classical scheme, the homogeneous distribution function hypothesis must be eliminated to be consistent with this prove.
-Now, in the following section, variational procedure and the relaxation processes will be used in the classical scheme to prove the $H$ theorem eliminating the homogeneous distribution function hypothesis.
-
-\section{Classical scheme}
-\subsection{Variational procedure}
+      H(t)&=&\sum_M \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
+      &&\pm  \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right]\delta V_M \delta \epsilon_n\label{H-entropy}.
+  \end{eqnarray}
+  This result lets us calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe quantum ideal gases.
+\subsection{Classical scheme: variational procedure}
 We proposed the following $H$ functional
@@
 \end{equation}{} 
+The constant $C$ can be omitted defining the following $H$ functional
+\begin{equation}
+   H(t)=\int_{}^{}\int_{}^{} \left[f(r,v,t) \ln f(r,v,t)-f(r,v,t)\right]\mathrm{d}^3v \mathrm{d}^3r \label{CH3},
+\end{equation}{}
+  but $\int \int f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N$ where $N$ is the total particle number.
+As $N$ is a constant and the functional $H$ has sense when one calculates the change of $H$, this constant can be omitted.
+\section{$H$-functional for relaxation processes in out of equilibrium systems}
 
-\subsubsection{Relaxation processes}
+\subsection{Quantum Scheme}
+
+The functional H \textcolor{red}{(variational entropy)} has the following form
+\begin{equation}
+    H(t)=\sum_{M} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right]\delta V_M \delta \epsilon_n \label{entropy2},
+\end{equation}
+where, for simplicity, we have defined $f(\bar{r},\epsilon_{n},t)\equiv f_{Mn}$.
+These functional has the following microcanonical restrictions
+\begin{eqnarray}
+        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
+  \end{eqnarray}
+  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
+  \begin{equation}
+      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
+  \end{equation}
+  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
+Those condition described a system out of equilibrium but no so far from it.
+\\
+If one performs the derivative of the functional H with respect to time, we obtain
+\begin{equation}
+   \frac{dH(t)}{dt}= \sum_n \sum_M \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right]\delta V_M \delta \epsilon_n.\label{deltaH}
+\end{equation}{}
+Using the first-order approximation
+\begin{eqnarray}
+   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}|g_{Mn}|, \label{firstorder}
+\end{eqnarray}{}
+we get
+
+\begin{eqnarray}
+    \frac{dH(t)}{dt}&=&\sum_n \sum_M \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM}\delta V_M \delta \epsilon_n \nonumber \\
+    &=&\sum_n \sum_M \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\delta V_M \delta \epsilon_n.\nonumber \\
+    \label{cambioH1}
+\end{eqnarray}{}
+It's necessary to remark that, in the case of fermions
+\begin{eqnarray}
+   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
+\end{eqnarray}{}
+The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
+Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogenities because of the exclusion principle and as consequence of this,  $g_{nM}=0$.
+This result corresponds to the system in the zero temperature condition.
+We can observe that in this condition, (\ref{fermionrestriction}) is violated.
+Due to this, the zero temperature condition is excluded.\\
+\\
+We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
+In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
+With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
+\begin{equation}
+    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
+\label{lnapproximation}
+\end{equation}{}
+With the previous approximation, (\ref{cambioH1}) will be
+\begin{eqnarray}
+    \frac{dH}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\}\delta V_M \delta \epsilon_n \nonumber \\
+    &&-\sum_{n}\sum_{M}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\delta V_M \delta \epsilon_n.\label{cambioH2}
+\end{eqnarray}{}
+Making use of the expression (\ref{relation}),
+(\ref{cambioH2}) casts into
+\begin{eqnarray}
+    \frac{dH}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\delta V_M \delta \epsilon_n.
+\nonumber \\
+    \label{cambioH3}
+\end{eqnarray}{}
+On the other hand, we obtain from the restrictions the following expressions
+\begin{eqnarray}
+    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
+    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
+\end{eqnarray}{}
+and as a consequence of $\sum_M \Delta_M(t) \delta V_M =\sum_M \delta_M(t)\delta V_M =0$, one finds
+\begin{equation}
+    \sum_M \dot{\Delta}_M(t)\delta V_M =\sum_M \dot{\delta}_{M}(t)\delta V_M =0.
+\end{equation}{}
+Substituting the previous expression to (\ref{cambioH3}) we obtain
+\begin{eqnarray}
+   \frac{dH}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM}\delta V_M \delta \epsilon_n. \nonumber \\ \label{cambioH4}
+\end{eqnarray}{}
+Now, the summation over M will be expressed in two summations
+\begin{equation}
+    \frac{dH}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\delta V_M \delta \epsilon_n+\sum^{N}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ}\delta V_M \delta \epsilon_n \right), \label{cambioH5}
+\end{equation}{}
+where $L+N=M$.
+Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or missing of particles or energy respectively.\\
+On the one hand, $\dot{g}^{+}_{nJ}<0$ describes the lost of particles and/or energy and so , $g^{+}_{nJ}>0$.
+On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
+\\
+We sort the previous ideas in the following form
+\begin{eqnarray}
+   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
+   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
+\end{eqnarray}{}
+\\
+and consequently, (\ref{cambioH5}) obtains the following form
+\begin{equation}
+    \frac{dH}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\delta V_M \delta \epsilon_n+\sum^{N}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}|\delta V_M \delta \epsilon_n \right), \label{cambioH6}
+\end{equation}{}
+and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{dH}{dt}<0$.\\
+When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{dH}{dt}=0$. 
+\\
+The next step in this model is to obtain the analogous transport equation through variation methods.
+In order to verify that transport equation can be obtained by applying the variation method, in the next section, the Boltzmann transport equation will be obtained applying the mentioned method to a modified Boltzmann $H$ functional.
+
+\subsection{Clasical scheme}
 Now one uses the restrictions (\ref{restrictionoutside}) but in the continuous case
@@
 \end{equation}{}
-where $\Delta(r,t)$ y $\Omega(r,t)$ could be seen as a deviation from the mean number of particles and mean energy by differential element $d^3r$.
-Performing the derivative of the functional $H$ (\ref{CH2})with respect to time
+where $\Delta(r,t)$ and $\Omega(r,t)$ could be seen as a deviation from the mean number of particles and mean energy by differential element $d^3r$.
+Performing the derivative of the functional $H$ (\ref{CH2}) with respect to time
 \begin{equation}
@@
 
-\section{Discussion}
-For quantum ideal gases, the H-theorem has been proved performing the variational method.
-In the classical scheme, the $H$-theorem is proved using the Boltzmann transport equation for a homogeneous distribution function.
-Classical $H$-theorem don't show the idea of particles transport in a system.
-That assumption in quantum mechanics presented by Tolman \cite{tolman} remains.
-This work eliminates that assumption defining the $H$ functional (\ref{entropy}) including a non-homogeneous distribution function.
-The prove of the quantum $H$-theorem in this work includes the idea of particles transport in relaxation processes using the first order approximation.
-This idea has not found in the literature.
-As consequence of this, all ideal gases placed in non-equilibrium state with inhomogeneities, when time tends to infinity, the system tends to the equilibrium state.
-This result complements the classical $H$-theorem due to the lack of the non-homogeneous distribution function.
-In addition, the classical $H$ theorem was analysed to include the particles transport and complete the classical $H$-theorem modifying the classical $H$ functional proposed by Boltzmann.
-Those proves only are applied to relaxation processes but the fact of including a non-homogeneous distribution function is a new way to complete the $H$ theorem for non-equilibrium systems.
-In other hand, the equation (\ref{H-entropy}) lets us calculate the entropy for a quantum system and describe the evolution of that system using the $H$ functional as a thermodynamic variable for a non-equilibrium system.
-Furthermore, thermodynamic variables to describe non-equilibrium systems is a problem in discussion.
-The previous idea adds new knowledge to this problem.
+\section{The Boltzmann transport equation}
+The Boltzmann $H$ functional is defined as
+\begin{equation}
+    H=\int_{}^{}f(\vec{v},t)\ln f(\vec{v},t)\mathrm{d}^{3}v,
+\end{equation}
+where $f$ is the homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space.
+In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous, that is, $f=f(\vec{r},\vec{v},t)$, where $r$ is the position of the particles defined in the $\mu$-space.
+Because of that, we defined a modified Boltzmann $H$ functional as
+\begin{equation}
+    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
+\end{equation}
+This equation involves the position dependence. This dependence is necessary to obtain the complete Boltzmann transport equation.\\
+Due to $H$ functional in equilibrium corresponds to a density of entropy and also, according to the second law of thermodynamics, the entropy is maximum, then the following relation must be hold in equilibrium
+\begin{equation}
+    \delta H'=0.\label{variation1}
+\end{equation}
+Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
+\begin{equation}
+    0=\left( 1+\ln f \right)\left( \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f \right).
+\end{equation}
+The previous relation holds if $(1+\ln f)=0$ or $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.\\
+Assuming that $(1+\ln f)=0$, we find that $f$ is a constant and consequently, $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.
+In other hand, if $f$ is not a constant, the Boltzmann transport equation will be
+\begin{equation}
+    \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f=0.\label{transport}
+\end{equation}
+The equation (\ref{transport}) is hold only in equilibrium.
+If the system is out of equilibrium, the variation is not equal to zero.
+As a consequence of the $H$ theorem, the variation is less to zero.
+In conclusion, is necessary to include in (\ref{transport}) a term that described this behavior, this term, as in \cite{huang}, will be the collision term and will be expressed in such form as
+\begin{equation}
+    \frac{\partial f}{\partial t}+\vec{v}\cdot \nabla f+\frac{\vec{F}}{m}\cdot \nabla_{\vec{v}} f=\left( \frac{df}{dt} \right)_{coll}, \label{classictransport}
+\end{equation}
+where $\vec{F}$ is a external force and $m$ is the mass of the particles.\\
+\\
+The equation (\ref{classictransport}) correspond to the Boltzmann transport equation.
+With this conclusion, in the next section, the analogous transport equation will be obtained applying the variation method to (\ref{entropy2}).
+
+\section{Analogous transport equation for ideal quantum gases}
+Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$.
+\begin{equation}
+    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
+\end{equation}
+where $\mathcal{N}_{M}^{n}$ is defined as
+\begin{equation}
+    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
+\end{equation}
+Besides, the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
+This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
+\\
+In orden to obtain the analogous transport equation to this system, we apply the variation method to the functional $H$.
+Due to $H$ is considered in equilibrium as a density of entropy, and therefore a maximum, the variation must be zero.\\
+Appliying the variation to the functional $H$, we obtain the following equation of motion
+\begin{equation}
+    0=\sum_M \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
+\label{transportequation1}
+\end{equation}
+This equation is valid in equilibrium.
+When the system is out of equilibrium, the variation is different from zero.
+Due to this, an expression has to be added to (\ref{transportequation1}).
+This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
+Moreover, this term contributes to the equation of motion per each cell and cell.
+Therefore, (\ref{transportequation1}) casts into
+\begin{eqnarray}
+    \sum_n \sum_M\left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_M \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
+    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
+\nonumber \\ \label{transportequation2}
+\end{eqnarray}
+\section{Classic limit}
+\textcolor{red}{The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
+Let us define the distribution function as the number of particles in a defined cell and a energy level $N_{nM}$ inside a volume $\delta V_M \delta \epsilon_n$ as
+\begin{equation}
+    f_{nM}=\frac{N_{nM}}{\delta V_M \delta g_{n} }.
+\end{equation}
+Using this definition on the variational entropy (\ref{entropy2}) we obtain
+\begin{eqnarray}
+    H&=& \sum_M \sum_n
+    \left[  
+           \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+           \right)\pm 
+           \left(  
+                  1\mp \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+           \right) \ln 
+           \left(  
+                   1\mp \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+           \right)
+    \right]\delta V_M \delta \epsilon_{n}. \nonumber \\
+    \label{h-quantic} 
+\end{eqnarray}
+In the classic limit, the particles are in the high energy levels and the number of particles  in most of the groups of state $n$ are small compared with the number of the elementary states $\epsilon_n$ of the group, that is
+\begin{equation}
+    \frac{N_{nM}}{\delta V_M \delta \epsilon_{n} } \approx 0,
+\end{equation}
+and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
+\begin{eqnarray}
+    H&=& \sum_M \sum_n
+    \left[  
+           \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+           \right)
+    \right]\delta V_M \delta \epsilon_{n}. \label{h-quantic2}
+\end{eqnarray}
+In the case when the system is spatial-homogeneous, the definition of the distribution function is 
+\begin{equation}
+    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
+\end{equation}
+and using the same limit we obtain
+\begin{eqnarray}
+    H&=& \sum
+    \left[  
+           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{n}}{ \delta \epsilon_{n}}
+           \right)
+    \right] \delta \epsilon_{n}. \label{h-quantic4}
+\end{eqnarray}
+This expression is equivalent to the
+Boltzmann's $H$ functional showed by Tolman
+\cite{tolman}.}
+\section{Conclusions}
+\textcolor{red}{
+In the literature, the Boltzmann's $H$ theorem is proved using a spatial-homogeneous distribution function.
+In spite of this method, we have used the spatial-homogeneous distribution function in each cell such that the distribution function in any cell is different to any other cell in the volume.
+This is the form how we have introduced inhomogeneities in the system.\\
+We have proved that the Maxwell-Boltzmann distribution can be obtained performing the Lagrange multipliers method for an ideal gas in the microcanonical conditions.
+In analogy to the previous method, we have proved that the Bose-Einstein and Fermi-Dirac distribution can be obtained performing the Lagrange multipliers method for ideal quantum gases in the microcanonical conditions.\\
+Another important idea in the classic $H$ theorem is the necessity of Boltzmann's transport equation.
+In comparison with this method, we have made use of the assumption of the behavior in time of the deviation $g$ and the assumption of that the system is out of equilibrium but close enough to the equilibrium state.
+Furthermore, we have introduced the inhomogeneities in the system dividing the volume $V$ in $P$ cells.\\
+We have also obtained the Boltzmann's transport equation performing the variational method to the Boltzmann's $H$ functional.
+It was necessary to include the in-homogeneous assumption in the distribution function to obtain the spatial dependency in the Boltzmann's transport equation.
+The collision term had to be introduced in the same way that was introduced by Boltzmann in \cite{huang}.\\
+In Quantum Mechanics, the spatial-homogeneous distribution function hyphotesis is not clear.
+Due to this, we have introduced inhomogeneities in the same way that the classical scheme and the $H$ theorem for quantum ideal gases was proved using the same assumptions that the classical scheme.
+Moreover the equation of motion (\ref{transportequation2}) could be identified as a quantum transport equation.
+The form of the collision term is not calculated in this work.
+\\
+Besides the variational entropy can be approximated to the Boltzmann's $H$ functional in the limit of high energy levels.\\
+In other hand, the equation (\ref{H-entropy}) lets us calculate the entropy for a quantum system and describe its evolution using the $H$ functional as a thermodynamic variable for a non-equilibrium system.
+Because of the use of thermodynamic variables to describe non-equilibrium systems is a problem in discussion, the previous idea adds new knowledge to this problem.
+}
+\clearpage
 
 \begin{thebibliography}{}
-\bibitem{goldstein}
-H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
+%\bibitem{goldstein}
+%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
 
-\bibitem{sakurai}
-J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
+%\bibitem{sakurai}
+%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
 
@@
 
+\bibitem{reif}
+F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
+
+\bibitem{patrick}
+R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
+
+\bibitem{kei}
+Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
+
+\bibitem{onsager}
+Onsager, L. (1931), Physical Review 37 (4): 405-426.
+
 \bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann’s H-theorem, its limitations, and thebirth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
+Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
 
@@
 
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
+\bibitem{hamiltonian}
+Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
+
+\bibitem{H-theorem-violation}
+Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
+
+\bibitem{H-theorem-and-entropy}
+Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
 
 \bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
+
+\bibitem{entropic-framework}
+Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
+
+\bibitem{thermal-harmonic}
+F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
+Phys. Rev. E 91, 042116 – 15 April 2015.
+
+\bibitem{quantum-transport}
+Robert Hussein and Sigmund Kohler,
+Phys. Rev. B 89, 205424, 21 May 2014
+
+\bibitem{quantum-transport-reservoirs}
+Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
+Phys. Rev. A 102, 022207, 11 August 2020.
+
+\bibitem{htheorem2}
+R. Silva {\it et al} EPL 89 10004 (2010).
+
+\bibitem{quantum1}
+J. Math. Phys. 47, 073303 (2006).
+
+\bibitem{quantum2}
+Z. Physik 268, 139-143 (1974).
+
+\bibitem{quantum-entropy}
+J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
+
+\bibitem{quantum-colapse}
+Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
+
+\bibitem{quantum3}
+Gring, M.
+and Kuhnert, M.
+and Langen, T.
+and Kitagawa, T.
+and Rauer, B.
+and Schreitl, M.
+and Mazets, I.
+and Smith, D.
+Adu and Demler, E.
+and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
+
+\bibitem{quantum4}
+Physical Review E 91, 062106 (2015).
+
+%\bibitem{relaxationphenomena}
+%Journal of Statistical Physics, 30; 2 (1983).
+
+%\bibitem{htheorem}
+%Review of Modern Physics; 74 (2002).
 
@@
 
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
+\bibitem{ergodic-theorem}
+Eur. Phys. J. H 35, 201–237 (2010).
+
+\bibitem{hydrodynamic-simulation}
+Review of Modern Physics 74(4):1203-1220.
 

### 2020-10-14 notasparaarticulo.tex
discussions and remarks were included, conclusions and introduction were rewritten and some details in the calculus were included

@@
 \usepackage{xcolor}
+\usepackage{calrsfs}
+\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
 
@@
 
-\textcolor{red}{In the literature, the basis for equilibrium systems in the classical scheme is well-established.
-From thermodynamics to statistical mechanics, the procedures for equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature \cite{huang, reif, patrick}.
-This is not the case of out of equilibrium systems.
-There are some well-established theoretical schemes to deal with out of equilibrium systems in the classical scheme, such as the Onsager formulation (called linear thermodynamics), some few other restricted developments proposed outstanding theoretical authors \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.
-The Kinetic Theory of dilute gases describes the behavior of a dilute classical gas by means the Boltzmann transport equation.
-Furthermore, the evolution to a equilibrium state for any system can be proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some some aspects in the Boltzmann's $H$ theorem that deserve some review in order to accomplish better consistency and is still in discussion [6-10].
-%\cite{paradox1, paradox2, hamiltonian, H-theorem-violation, H-theorem-and-entropy}.
+\textcolor{red}{In the literature, the theoretical basis to describe equilibrium systems in the classical scheme is well-established.
+From thermodynamics to statistical mechanics, the procedures to describe mathematically equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature \cite{huang, reif, patrick}.
+In the case of the out of equilibrium systems, there are many works that attempt to describe out of equilibrium systems (however those works do not work in some systems, for example, the internal behavior of a glass[], which behavior is too slow compare with the time observation).
+Examples of some theoretical schemes to describe out of equilibrium systems can be the Onsager formulation (called linear thermodynamics), some few other restricted developments that attempt to complete the Onsager formulation \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.\\
+In the case of the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann's by means the Boltzmann transport equation.
+Furthermore, the evolution to an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
+However, there are some some aspects in the Boltzmann's $H$ theorem and the Boltzmann's $H$ functional that deserve some review in order to accomplish better consistency, for example the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\ %\cite{paradox1, paradox2, hamiltonian, H-theorem-violation, H-theorem-and-entropy}.
+\\
 In order to obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the $H$ theorem and the Boltzmann transport equation are necessary such that in the appropriate limit, the  Boltzmann's $H$ theorem and the Boltzmann transport equation should be recovered.\\
 One of the first developments in this field was developed by Tolman \cite{tolman}.
-He establishes the quantum version of the $H$ theorem using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
-The equation of motion for the occupation number was obtained by performing the time perturbation theory.
-However, in order to obtain the flux relations in analogy to the Boltzmann's method, it is required a quantum version of the transport equation that includes a non-homogeneous distribution function.\\
-In another hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (the analogous to the Boltzmann transport equation) [12-15].
+The quantum version of the $H$ theorem was established by Tolman using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
+The equation of motion for the occupation number was obtained by applying the time perturbation theory.
+However, in order to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
+In other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (the analogous to the Boltzmann transport equation) [12-15].
 %\cite{entropic-framework, thermal-harmonic, quantum-transport, quantum-transport-reservoirs}.
 Besides some papers construct the $H$ functional and prove the $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous-hypothesis and the correspondence principle are not clear or discussed.
-Moreover the quantum $H$ theorem, the quantum entropy and the second law of thermodynamics in quantum mechanics is still in discussion [19-26].\\ %\cite{quantum-entropy, quantum-colapse, quantum3, quantum4, binary, contradictions,ergodic-theorem, hydrodynamic-simulation}.\\
-In order to solve those problems, in this work we define a new $H$ functional, which name is variational entropy.
-This variational entropy includes as main variable the distribution function of an ideal quantum gas (Bosons or Fermions particles) inside a volume $V$ divided into cells.
-Before developing the non-equilibrium description, the statistical mechanics results for equilibrium systems are recovered from the variational entropy.
-Furthermore we prove that the behavior of the variational entropy in time is less or equal to zero for systems that are close enough to the equilibrium state.
-This prove corresponds with the $H$ theorem.
-In addition, the Boltzmann transport equation is obtained from the Boltzmann's $H$ functional and, in analogy with this, a quantum version of the Boltzmann transport equation is obtained from the variational entropy.
-Finally, in order to be consistence with the correspondence principle, the Boltzmann's $H$ functional is recovered from the variational entropy in the limit of the high level energies.}
+However, the homogeneous distribution function hypothesis and the correspondence principle are not clear or discussed.
+Moreover the validity of the quantum $H$ theorem and the second low of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\ %\cite{quantum-entropy, quantum-colapse, quantum3, quantum4, binary, contradictions,ergodic-theorem, hydrodynamic-simulation}.\\
+\\
+In order to contribute to a better consistent $H$ theorem and a new formalism to describe out of equilibrium systems, we define the theoretical basis to describe out of equilibrium systems in order to be completed in the future.\\
+In this work we define a new $H$ functional (variational entropy) which depends on the distribution function of an ideal quantum gas (Bosons or Fermions particles).
+This gas is contained inside a volume $V$ divided into cells.
+Before developing the non-equilibrium description, we recover the Fermi-Dirac and Bose-Einstein distributions from the variational entropy when the system is in the equilibrium state.
+Furthermore we prove that the derivative of the variational entropy with respect to time is less or equal to zero for relaxation processes (the action of any perturbed system that returns to the equilibrium state) in any system.
+This prove corresponds with the statement of the $H$ theorem.
+In addition we obtain the Boltzmann transport equation from the Boltzmann's $H$ functional and, in analogy with this, we find a quantum version of the Boltzmann transport equation from the variational entropy.
+Finally, in order to be consistence with the correspondence principle, we prove that the Boltzmann's $H$ functional is proportional to the variational entropy in the limit of the high level energies.}
 
@@
 \end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system and $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively.\\
+where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
 The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
@@
 \end{equation}
-where $n_i$ is the occupation number and $g_i$ is a energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is obtained
+where $n_i$ is the occupation number and $g_i$ is a energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
 \begin{eqnarray}
@@
 where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expression are obtained by the perturbation theory.\\
-The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}).
-This prove is in the same way as in the classical version.
-That prove is developed taking account of the homogeneous distribution function, as in the classical version.\\
+The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}), taking account the spatial-homogeneous distribution function hypothesis, as in the classical version.\\
 In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high energy levels to
 \begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}).
+    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
 \end{equation}
@@
 \end{eqnarray}}
-This functional, in equilibrium should be a density of entropy, which we will call variational entropy.
+This functional, in equilibrium should be a density of entropy, which we will call \textit{variational entropy}.
 This functional describes a system with global energy E, global number of free
 quantum particles (which may be fermions for upper sing or bosons for lower sing) N, global volume V, and temperature T. The
-global system in turn consists of a set of $K$ small cells with a constant volume $V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$ and time $t$ and the energy spectra will be divided into $g_n$ volumes.
+global system in turn consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
+The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
+Besides the energy spectra will be divided into $\delta \epsilon_n$ volumes.
 We will use the uppercase latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
@@
 local identical volumes (\textit{i.e.} $V_M = V/K, \forall M$). Each cell has its own local chemical potential, $\mu_M$, local
-$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution.
+$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution
 function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
@@
 \end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number. This is
+When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
 \begin{equation}
@@
 \end{equation}
-where we have used that the $H$ functional in equilibrium is proportional to the entropy of the system.
+where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
 In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
@@
 \end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, one obtains
+If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
 \begin{equation}
@@
 \end{eqnarray}{}
-We have proved that when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multiplier are homogeneous in the system.\\
+We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multiplier are homogeneous in the system.\\
 We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
@@
 \subsection{Classical scheme: variational procedure}
-We proposed the following $H$ functional
+We propose the following $H$ functional
 
 \begin{equation}
-   H(t)=\int_{}^{}\int_{}^{} f(r,v,t) \ln f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{CH2}.
-\end{equation}{}
+   H'(t)=\int_{}^{}\int_{}^{} f(r,v,t) \ln f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{CH2}.
+\end{equation}
 The functional (\ref{CH2}) contains a no-homogeneous distribution function but keeps the same form of the Boltzman's $H$ functional.
-Excluding the homogeneous distribution function hypothesis, one uses the same procedure, that it was shown in the previous section, to prove the $H$-theorem for relaxation processes.\\
+Excluding the homogeneous distribution function hypothesis, we use the same procedure that it was shown in the previous section to prove the $H$-theorem.\\
 If one considers the microcanonical ensemble conditions 
@@
 \begin{eqnarray}
-\frac{\delta H}{\delta f(r',v')}&=&\int_{}^{}\frac{\delta}{\delta f(r',v')}\left[f(r,v)\ln f(r,v)  \right]\mathrm{d}^3v \mathrm{d}^3r-\int_{}^{}\alpha(r)\int_{}^{}\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
+\frac{\delta H'}{\delta f(r',v')}&=&\int\int_{}^{}\frac{\delta}{\delta f(r',v')}\left[f(r,v)\ln f(r,v)  \right]\mathrm{d}^3v \mathrm{d}^3r-\int_{}^{}\alpha(r)\int_{}^{}\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
 &&-\int_{}^{}\beta(r) \int_{}^{}\epsilon(v)\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
@@
 where $C$ is a constant. It is trivial that using the variational procedures, one obtains the form of the Maxwell-Boltzmann distribution. \\
-As $H$ in equilibrium is a density of entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
+As $H'$ in equilibrium \textcolor{red}{is proportional to} entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
 \begin{equation}
@@
 \begin{equation}
-   H(t)=\int_{}^{}\int_{}^{} \left[f(r,v,t) \ln f(r,v,t)-f(r,v,t)\right]\mathrm{d}^3v \mathrm{d}^3r \label{CH3},
+   H''(t)=\int_{}^{}\int_{}^{} \left[f(r,v,t) \ln f(r,v,t)-f(r,v,t)\right]\mathrm{d}^3v \mathrm{d}^3r \label{CH3},
 \end{equation}{}
   but $\int \int f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N$ where $N$ is the total particle number.
-As $N$ is a constant and the functional $H$ has sense when one calculates the change of $H$, this constant can be omitted.
+As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
 \section{$H$-functional for relaxation processes in out of equilibrium systems}
@@
 
-The functional H \textcolor{red}{(variational entropy)} has the following form
+The \textit{variational entropy} has the following form
 \begin{equation}
@@
 \end{equation}
-where, for simplicity, we have defined $f(\bar{r},\epsilon_{n},t)\equiv f_{Mn}$.
-These functional has the following microcanonical restrictions
+where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
+This distribution function \textcolor{red}{must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells)}, and the following microcanonical restrictions
 \begin{eqnarray}
@@
   and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-Those condition described a system out of equilibrium but no so far from it.
+Those condition described a system out of equilibrium but no so far from it, that is a perturbed system.
 \\
-If one performs the derivative of the functional H with respect to time, we obtain
+If we perform the derivative of the functional $H$ with respect to time, we obtain
 \begin{equation}
@@
 \end{eqnarray}{}
-and as a consequence of $\sum_M \Delta_M(t) \delta V_M =\sum_M \delta_M(t)\delta V_M =0$, one finds
+and as a consequence of $\sum_M \Delta_M(t) \delta V_M =\sum_M \delta_M(t)\delta V_M =0$, we find
 \begin{equation}
@@
 where $L+N=M$.
-Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or missing of particles or energy respectively.\\
+Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or/and missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or/and missing of particles or energy respectively.\\
 On the one hand, $\dot{g}^{+}_{nJ}<0$ describes the lost of particles and/or energy and so , $g^{+}_{nJ}>0$.
@@
 \end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{dH}{dt}<0$.\\
+and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{dH}{dt}<0$.
+With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{dH}{dt}<0$.\\
 When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{dH}{dt}=0$.
+Then, we can say that any quantum ideal gas perturbed, evolves always to the equilibrium state, such as the classical $H$ theorem.
 \\
-The next step in this model is to obtain the analogous transport equation through variation methods.
-In order to verify that transport equation can be obtained by applying the variation method, in the next section, the Boltzmann transport equation will be obtained applying the mentioned method to a modified Boltzmann $H$ functional.
+Now we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
 
 \subsection{Clasical scheme}
-Now one uses the restrictions (\ref{restrictionoutside}) but in the continuous case
+Now we use the restrictions (\ref{restrictionoutside}) but in the form of the continuous case
 \begin{equation}
@@
 \begin{equation}
-    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\left[ 1+\ln f(r,v,t) \right]\dot f(r,v,t) \mathrm{d}^3v \mathrm{d}^3r \label{dH1}.
+    \frac{dH'}{dt}=\int_{}^{}\int_{}^{}\left[ 1+\ln f(r,v,t) \right]\dot f(r,v,t) \mathrm{d}^3v \mathrm{d}^3r \label{dH1},
 \end{equation}{}
-Applying the first order approximation (\ref{firstorder}) to (\ref{dH1}), one obtains
+and applying the first order approximation (\ref{firstorder}) to (\ref{dH1}), we obtain
 \begin{equation}
-    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g(r,v,t) \right\} \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,1}.
+    \frac{dH'}{dt}=\int_{}^{}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g(r,v,t) \right\} \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,1}.
 \end{equation}{}
@@
 \begin{eqnarray}
-\frac{dH}{dt}&=&\int_{}^{} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g(r,v,t) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r,
+\frac{dH'}{dt}&=&\int_{}^{} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g(r,v,t) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r,
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r+\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,2}.
+\frac{dH'}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r+\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,2}.
 \nonumber \\
 \end{eqnarray}{}
-If one uses the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), one obtains the same result. Then, applying that result to (\ref{dH1,2}), one finds
+If we use the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), we obtain the same result. Then applying this to (\ref{dH1,2}), we find
 \begin{eqnarray}
-\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r.
+\frac{dH'}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r.
 \end{eqnarray}{}
-Using the same consideration in the previous section to $g$ and $\dot g$, one proves that $\frac{dH}{dt}<0$. If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH}{dt}=0$.
+Using the same consideration in the previous section to $g$ and $\dot g$, that is to divide the spatial integral in two, one integral for cells that have excess of particles and/or energy, and the second for cells that have missing of particles and/or energy, and then using (\ref{separacion}), we prove that $\frac{dH'}{dt}<0$.
+If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH'}{dt}=0$.
+\\
+\\
+The next step in this model is to obtain the analogous transport equation through variation methods.
+In order to verify that transport equation can be obtained by applying the variation method, in the next section the Boltzmann transport equation will be obtained applying the mentioned method to the modified Boltzmann $H$ functional (\ref{CH2}).
 
@@
 In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous, that is, $f=f(\vec{r},\vec{v},t)$, where $r$ is the position of the particles defined in the $\mu$-space.
-Because of that, we defined a modified Boltzmann $H$ functional as
-\begin{equation}
-    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
-\end{equation}
-This equation involves the position dependence. This dependence is necessary to obtain the complete Boltzmann transport equation.\\
-Due to $H$ functional in equilibrium corresponds to a density of entropy and also, according to the second law of thermodynamics, the entropy is maximum, then the following relation must be hold in equilibrium
+Because of that, we use the modified Boltzmann $H$ functional (\ref{CH2})
+%\begin{equation}
+%    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
+%\end{equation}
+that involves the position dependence. %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
+\\
+Due to $H$ functional in equilibrium is proportional to entropy and also, according to the second law of thermodynamics, the entropy is maximum, then the following relation must be hold in equilibrium
 \begin{equation}
@@
 As a consequence of the $H$ theorem, the variation is less to zero.
-In conclusion, is necessary to include in (\ref{transport}) a term that described this behavior, this term, as in \cite{huang}, will be the collision term and will be expressed in such form as
+In conclusion, is necessary to include in (\ref{transport}) a term that described this behavior.
+This term, as in \cite{huang}, will be the collision term and will be expressed in such form as
 \begin{equation}
@@
 \section{Analogous transport equation for ideal quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$.
+Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
 \begin{equation}
@@
 \end{equation}
-Besides, the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
+Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
 This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-\\
 In orden to obtain the analogous transport equation to this system, we apply the variation method to the functional $H$.
-Due to $H$ is considered in equilibrium as a density of entropy, and therefore a maximum, the variation must be zero.\\
+Due to $H$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.\\
 Appliying the variation to the functional $H$, we obtain the following equation of motion
@@
 This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, this term contributes to the equation of motion per each cell and cell.
+Moreover, we propose that this term contributes to the equation of motion per each cell \textcolor{blue}{(explanation about why this term contributes)}.
 Therefore, (\ref{transportequation1}) casts into
@@
 \end{eqnarray}
-\section{Classic limit}
-\textcolor{red}{The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-Let us define the distribution function as the number of particles in a defined cell and a energy level $N_{nM}$ inside a volume $\delta V_M \delta \epsilon_n$ as
+\section{Classical limit}
+\textcolor{red}{%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
+We begin defining the distribution function as the number of particles in a defined cell and a energy level $N_{nM}$ inside a volume $\delta V_M \delta \epsilon_n$ as
 \begin{equation}
@@
 \begin{eqnarray}
-    H&=& \sum
+    H&=& \sum_M \sum_n
     \left[  
@@
            \right)
-    \right] \delta \epsilon_{n}. \label{h-quantic4}
+    \right] \delta V_M \delta \epsilon_{n}=V\sum_n
+    \left[  
+           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{n}}{ \delta \epsilon_{n}}
+           \right)
+    \right] \delta \epsilon_{n} \nonumber \\
+    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=V H_{Boltz},\label{h-quantic4}
 \end{eqnarray}
-This expression is equivalent to the
-Boltzmann's $H$ functional showed by Tolman
-\cite{tolman}.}
+where $B_{boltz}$ is the Boltzmann's $H$ functional showed by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.} }
+\section{Discussions and remarks}
+\textcolor{red}{
+In the literature, the Boltzmann's $H$ theorem is proved using a spatial homogeneous distribution function.
+In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
+However we included inhomogeneities in the system through the division of the volume $V$ in cells and the %fact that the distribution function is different for each cell (
+non-homogeneous distribution function assumption in the total volume.
+Moreover the Boltzmann's $H$ theorem is still hold when the system includes inhomogeneities, only if the local equilibrium hypothesis is hold.
+We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
+It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
+Also we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through variational method.
+Nevertheless in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
+This is a consequence of the $H$ theorem.\\
+\\
+In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
+With the prove of the quantum version of the $H$ theorem we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis is hold.
+In addition to this, the system is in the equilibrium state when the derivative of the variational entropy respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
+In other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
+However in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
+This problem may be a future work.\\
+\\
+We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
+This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
+Moreover if we see the expression for the Boltzmann's $H$ functional in equilibrium
+\begin{equation}
+    H_{boltz}=-\frac{S}{Vk},
+\end{equation}
+where $S$ is the entropy of the system and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
+\begin{equation}
+    H\propto S.
+\end{equation}
+Finally we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.}
 \section{Conclusions}
-\textcolor{red}{
-In the literature, the Boltzmann's $H$ theorem is proved using a spatial-homogeneous distribution function.
-In spite of this method, we have used the spatial-homogeneous distribution function in each cell such that the distribution function in any cell is different to any other cell in the volume.
-This is the form how we have introduced inhomogeneities in the system.\\
-We have proved that the Maxwell-Boltzmann distribution can be obtained performing the Lagrange multipliers method for an ideal gas in the microcanonical conditions.
-In analogy to the previous method, we have proved that the Bose-Einstein and Fermi-Dirac distribution can be obtained performing the Lagrange multipliers method for ideal quantum gases in the microcanonical conditions.\\
-Another important idea in the classic $H$ theorem is the necessity of Boltzmann's transport equation.
-In comparison with this method, we have made use of the assumption of the behavior in time of the deviation $g$ and the assumption of that the system is out of equilibrium but close enough to the equilibrium state.
-Furthermore, we have introduced the inhomogeneities in the system dividing the volume $V$ in $P$ cells.\\
-We have also obtained the Boltzmann's transport equation performing the variational method to the Boltzmann's $H$ functional.
-It was necessary to include the in-homogeneous assumption in the distribution function to obtain the spatial dependency in the Boltzmann's transport equation.
-The collision term had to be introduced in the same way that was introduced by Boltzmann in \cite{huang}.\\
-In Quantum Mechanics, the spatial-homogeneous distribution function hyphotesis is not clear.
-Due to this, we have introduced inhomogeneities in the same way that the classical scheme and the $H$ theorem for quantum ideal gases was proved using the same assumptions that the classical scheme.
-Moreover the equation of motion (\ref{transportequation2}) could be identified as a quantum transport equation.
-The form of the collision term is not calculated in this work.
-\\
-Besides the variational entropy can be approximated to the Boltzmann's $H$ functional in the limit of high energy levels.\\
-In other hand, the equation (\ref{H-entropy}) lets us calculate the entropy for a quantum system and describe its evolution using the $H$ functional as a thermodynamic variable for a non-equilibrium system.
-Because of the use of thermodynamic variables to describe non-equilibrium systems is a problem in discussion, the previous idea adds new knowledge to this problem.
+\textcolor{red}{We conclude the following statements:\\
+\begin{enumerate}
+    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis is hold.
+    \item We can obtain the transport equation from the $H$ functional.
+    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
+    \item We find an expression for the entropy to quantum gases.
+    \end{enumerate}
   }
+%  \begin{equation}
+%      \mathcal{H}
+%  \end{equation}
+
 \clearpage

### 2020-10-16 notasparaarticulo.tex
discussions and remarks section was included, conclusions was rewritten, introduction was modified and some new ideas were included

@@
 \DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
+\newcommand{\Sm}{\Ss_M}
+\newcommand{\de}{\delta}
+\newcommand{\ep}{\epsilon}
+\newcommand{\Ss}{\mathcal{S}}
 
@@
 From thermodynamics to statistical mechanics, the procedures to describe mathematically equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature \cite{huang, reif, patrick}.
-In the case of the out of equilibrium systems, there are many works that attempt to describe out of equilibrium systems (however those works do not work in some systems, for example, the internal behavior of a glass[], which behavior is too slow compare with the time observation).
+In the case of the out of equilibrium systems, there are many works that attempt to describe out of equilibrium systems (however those works do not work in some systems, for example, the internal behavior of a glass due to the time relaxation is very big \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} as examples).
 Examples of some theoretical schemes to describe out of equilibrium systems can be the Onsager formulation (called linear thermodynamics), some few other restricted developments that attempt to complete the Onsager formulation \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.\\
@@
 In order to contribute to a better consistent $H$ theorem and a new formalism to describe out of equilibrium systems, we define the theoretical basis to describe out of equilibrium systems in order to be completed in the future.\\
-In this work we define a new $H$ functional (variational entropy) which depends on the distribution function of an ideal quantum gas (Bosons or Fermions particles).
+In this work we define a new functional (variational entropy) which depends on the distribution function of an ideal quantum gas (Bosons or Fermions particles).
 This gas is contained inside a volume $V$ divided into cells.
@@
 
-\textcolor{red}{We begin by defining the form of the functional $H$
+\textcolor{red}{We begin by defining the form of the functional 
 \begin{eqnarray}
-    H(t)&=&\sum_{M} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    \Ss (t)&=&\sum_{M} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
     &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
@@
 Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $V_M = V/K, \forall M$). Each cell has its own local chemical potential, $\mu_M$, local
-$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution
+local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its own local chemical potential, $\mu_M$, local
+variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
 function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
 \begin{eqnarray}
-    H_M&=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \delta \epsilon_n.
+    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
+    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
+{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
 \end{eqnarray}
@@
 \begin{eqnarray}
-    &&\sum_{M} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\right]\delta V_M \delta \epsilon_n=\sum_{M} {\mathcal{N}}_{M} \delta \epsilon_n=N; \nonumber \\
-    &&\sum_{M}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{Mn}\right]\delta V_M \delta g_n=\sum_M {\mathcal{E}}_M \delta \epsilon_n=E, \label{restriccions1}
+    &&\sum_{M} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
+    &&\sum_{M}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_M {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
 \end{eqnarray}
 where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Inspired by the Hamilton principle, $H$ will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining
+Inspired by the Hamilton principle, $\Ss$ will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining
 \begin{eqnarray}
@@
 \begin{eqnarray}
-      H(t)&=&\sum_M \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right]\delta V_M \delta \epsilon_n\label{H-entropy}.
+      \Ss(t)&=&\sum_M \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
+      &&\pm  \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
   \end{eqnarray}
@@
 \begin{equation}
-   H'(t)=\int_{}^{}\int_{}^{} f(r,v,t) \ln f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{CH2}.
+   H'(t)=\sum_M\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH2}.
 \end{equation}
@@
 \begin{equation}
-    \int_{}^{}\int_{}^{}f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N, \ \ \ \int_{}^{}\int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v \mathrm{d}^3r=E \label{micro},
+    \sum_M\int_{}^{}f_M(v,t)\mathrm{d}^3v =N, \ \ \ \sum_M\int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v =E \label{micro},
 \end{equation}{}
@@
 \begin{eqnarray}
-\frac{\delta H'}{\delta f(r',v')}&=&\int\int_{}^{}\frac{\delta}{\delta f(r',v')}\left[f(r,v)\ln f(r,v)  \right]\mathrm{d}^3v \mathrm{d}^3r-\int_{}^{}\alpha(r)\int_{}^{}\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
-&&-\int_{}^{}\beta(r) \int_{}^{}\epsilon(v)\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
-&=&\ln f(r',v')+1-\alpha(r')-\beta(r') \epsilon(v')=0.
+\frac{\delta H'}{\delta f_J(v')}&=&\sum_M\int_{}^{}\frac{\delta}{\delta f_J(v')}\left[f_M(v)\ln f_M(v)  \right]\mathrm{d}^3v -\sum_M\alpha_M\int_{}^{}\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
+&&-\sum_M\beta_M \int_{}^{}\epsilon(v)\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
+&=&\ln f_J(v')+1-\alpha_J-\beta_J \epsilon(v')=0.
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-\ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f(r',v')=e^{\alpha(r') +\beta(r') \epsilon(v')-1} \nonumber \\
-&=&Ce^{\alpha(r')+\beta(r') \epsilon(v') } \label{relacion1},
+\ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f_J(v')=e^{\alpha_J +\beta_J \epsilon(v')-1} \nonumber \\
+&=&Ce^{\alpha_J+\beta_J \epsilon(v') } \label{relacion1},
 \end{eqnarray}{}
@@
 \begin{equation}
-   H''(t)=\int_{}^{}\int_{}^{} \left[f(r,v,t) \ln f(r,v,t)-f(r,v,t)\right]\mathrm{d}^3v \mathrm{d}^3r \label{CH3},
+   H''(t)=\sum_M\int_{}^{} \left[f_M(v,t) \ln f_M(v,t)-f_M(v,t)\right]\mathrm{d}^3v  \label{CH3},
 \end{equation}{}
-  but $\int \int f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N$ where $N$ is the total particle number.
+  but $\sum_M \int f_M(v,t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
 As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\section{$H$-functional for relaxation processes in out of equilibrium systems}
+\\
+  \textcolor{green}{In other hand, if the distribution function is spatial-homogeneous, $f$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
+  \begin{equation}
+      H'(t)=\int \sum_M [f(v,t)\ln f(v,t)] \mathrm{d}^{3}v = P\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v= P H_{boltz}(t), 
+  \end{equation}
+  where
+  \begin{equation}
+      H_{boltz}(t)=\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v.
+  \end{equation}
+  Besides, if we analyse the following expression
+  \begin{eqnarray}
+      \int Pf(v,t) \ln [Pf(v,t)] \mathrm{d}^{3}v = \int [(P\ln P)f(v,t) + Pf(v,t) \ln f(v,t)]\mathrm{d}^{3}v, \label{sum-h}
+  \end{eqnarray}
+  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the the first term in (\ref{sum-h}) is negible and and consequently we conclude that
+  \begin{equation}
+      \int f'(v,t) \ln f'(v,t) \mathrm{d}^{3}v = P\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v; \ \ \ f'(v,t)= Pf(v,t). \label{aditive-property} 
+  \end{equation}
+  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.}
+
+%----------------------------------------------
+
+\section{Variational entropy for relaxation processes in out of equilibrium systems}
 
@@
 \begin{equation}
-    H(t)=\sum_{M} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right]\delta V_M \delta \epsilon_n \label{entropy2},
+    \Ss(t)=\sum_{M} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
 \end{equation}
@@
 \\
-If we perform the derivative of the functional $H$ with respect to time, we obtain
+If we perform the derivative of the variational entropy with respect to time, we obtain
 \begin{equation}
-   \frac{dH(t)}{dt}= \sum_n \sum_M \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right]\delta V_M \delta \epsilon_n.\label{deltaH}
+   \frac{d \Ss (t)}{dt}= \sum_n \sum_M \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
 \end{equation}{}
@@
 \begin{eqnarray}
-    \frac{dH(t)}{dt}&=&\sum_n \sum_M \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM}\delta V_M \delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_M \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\delta V_M \delta \epsilon_n.\nonumber \\
+    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_M \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
+    &=&\sum_n \sum_M \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
     \label{cambioH1}
@@
 \begin{eqnarray}
-    \frac{dH}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\}\delta V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\delta V_M \delta \epsilon_n.\label{cambioH2}
+    \frac{d\Ss}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
+    &&-\sum_{n}\sum_{M}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-    \frac{dH}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\delta V_M \delta \epsilon_n.
+    \frac{d\Ss}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
 \nonumber \\
@@
 \end{eqnarray}{}
-and as a consequence of $\sum_M \Delta_M(t) \delta V_M =\sum_M \delta_M(t)\delta V_M =0$, we find
+and as a consequence of $\sum_M \Delta_M(t) \de V_M  =\sum_M \delta_M(t) \de V_M =0$, we find
 \begin{equation}
-    \sum_M \dot{\Delta}_M(t)\delta V_M =\sum_M \dot{\delta}_{M}(t)\delta V_M =0.
+    \sum_M \dot{\Delta}_M(t) \de V_M =\sum_M \dot{\delta}_{M}(t) \de V_M=0.
 \end{equation}{}
@@
 \begin{eqnarray}
-   \frac{dH}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM}\delta V_M \delta \epsilon_n. \nonumber \\ \label{cambioH4}
+   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
 \end{eqnarray}{}
@@
 \begin{equation}
-    \frac{dH}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\delta V_M \delta \epsilon_n+\sum^{N}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ}\delta V_M \delta \epsilon_n \right), \label{cambioH5}
+    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_M \delta \epsilon_n+\sum^{N}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_M\delta \epsilon_n \right), \label{cambioH5}
 \end{equation}{}
@@
 \begin{equation}
-    \frac{dH}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\delta V_M \delta \epsilon_n+\sum^{N}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}|\delta V_M \delta \epsilon_n \right), \label{cambioH6}
+    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_M \delta \epsilon_n+\sum^{N}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_M \delta \epsilon_n \right), \label{cambioH6}
 \end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{dH}{dt}<0$.
+and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
 With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{dH}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{dH}{dt}=0$.
+When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
 Then, we can say that any quantum ideal gas perturbed, evolves always to the equilibrium state, such as the classical $H$ theorem.
@@
 \begin{equation}
-    \int_{}^{}f(r,v,t)\mathrm{d}^3v=\mathcal{N}+\Delta(r,t); \ \ \ \int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v=\mathcal{E}+\Omega(r,t)\label{restriccionescambio},
+    \int_{}^{}f_M(v,t)\mathrm{d}^3v=\mathcal{N}+\Delta_M(t); \ \ \ \int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v=\mathcal{E}+\Omega_M(t)\label{restriccionescambio},
 \end{equation}{}
-where $\Delta(r,t)$ and $\Omega(r,t)$ could be seen as a deviation from the mean number of particles and mean energy by differential element $d^3r$.
+where $\Delta_M(t)$ and $\Omega_M(t)$ could be seen as a deviation from the mean number of particles and mean energy.
+
 Performing the derivative of the functional $H$ (\ref{CH2}) with respect to time
 \begin{equation}
-    \frac{dH'}{dt}=\int_{}^{}\int_{}^{}\left[ 1+\ln f(r,v,t) \right]\dot f(r,v,t) \mathrm{d}^3v \mathrm{d}^3r \label{dH1},
+    \frac{dH'}{dt}=\sum_M\int_{}^{}\left[ 1+\ln f_M(v,t) \right]\dot f_M(v,t) \mathrm{d}^3v  \label{dH1},
 \end{equation}{}
@@
 \begin{equation}
-    \frac{dH'}{dt}=\int_{}^{}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g(r,v,t) \right\} \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,1}.
+    \frac{dH'}{dt}=\sum_M\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g_M(v,t) \right\} \right]\dot g_M(v,t)\mathrm{d}^3v  \label{dH1,1}.
 \end{equation}{}
@@
 \begin{eqnarray}
-\frac{dH'}{dt}&=&\int_{}^{} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g(r,v,t) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r,
+\frac{dH'}{dt}&=&\sum_M \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g_M(v,t) \right]\dot g_M(v,t)\mathrm{d}^3v,
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-\frac{dH'}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r+\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,2}.
-\nonumber \\
+\frac{dH'}{dt}&=&\sum_M\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g_M(v,t)\mathrm{d}^3v +\sum_M\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v \label{dH1,2}. \nonumber \\
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-\frac{dH'}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r.
+\frac{dH'}{dt}&=&\sum_M\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v.
 \end{eqnarray}{}
@@
 In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous, that is, $f=f(\vec{r},\vec{v},t)$, where $r$ is the position of the particles defined in the $\mu$-space.
-Because of that, we use the modified Boltzmann $H$ functional (\ref{CH2})
+Because of that, we use the following functional
+\begin{equation}
+    H_{mod}=\int f(\vec{r},\vec{v},t) \ln f(\vec{r},\vec{v},t) \mathrm{d}^{3}v,
+\end{equation}
+that involves the position dependence.
 %\begin{equation}
@@
 %\end{equation}
-that involves the position dependence. %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
+ %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
 \\
@@
 \begin{equation}
-    \delta H'=0.\label{variation1}
+    \delta H_{mod}=0.\label{variation1}
 \end{equation}
@@
 \end{eqnarray}
+
+%---------------------------------------------
+
 \section{Classical limit}
 \textcolor{red}{%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and a energy level $N_{nM}$ inside a volume $\delta V_M \delta \epsilon_n$ as
+We begin defining the distribution function as the number of particles in a defined cell and a energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
 \begin{equation}
-    f_{nM}=\frac{N_{nM}}{\delta V_M \delta g_{n} }.
+    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
 \end{equation}
@@
     \left[  
-           \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}} \ln 
+           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
            \left( 
-                  \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
            \right)\pm 
            \left(  
-                  1\mp \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
            \right) \ln 
            \left(  
-                   1\mp \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
            \right)
-    \right]\delta V_M \delta \epsilon_{n}. \nonumber \\
+    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
     \label{h-quantic} 
@@
 \begin{equation}
-    \frac{N_{nM}}{\delta V_M \delta \epsilon_{n} } \approx 0,
+    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
 \end{equation}
@@
     \left[  
-           \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}} \ln 
+           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
            \left( 
-                  \frac{N_{nM}}{\delta V_M \delta \epsilon_{n}}
+                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
            \right)
-    \right]\delta V_M \delta \epsilon_{n}. \label{h-quantic2}
+    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
 \end{eqnarray}
@@
            \right)
-    \right] \delta V_M \delta \epsilon_{n}=V\sum_n
+    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
     \left[  
@@
     \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=V H_{Boltz},\label{h-quantic4}
+    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M P H_{Boltz},\label{h-quantic4}
 \end{eqnarray}
-where $B_{boltz}$ is the Boltzmann's $H$ functional showed by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.} }
+where $B_{boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
+We can observe in the same way that in the classical scheme that if the number of particles is too big compared to the number of cells $P$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.}
+
+%--------------------------------------
+
 \section{Discussions and remarks}
@@
 \\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
+In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
 With the prove of the quantum version of the $H$ theorem we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis is hold.
@@
 \begin{equation}
-    H\propto S.
+    \Ss\propto S.
 \end{equation}
-Finally we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.}
+Finally We find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.}
+
+%-----------------------------------------------
+
 \section{Conclusions}
@@
 R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
+
+\bibitem{cristal1}
+Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
+
+\bibitem{cristal2}
+Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
+
+\bibitem{cristal3}
+Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
 

### 2020-10-21 notasparaarticulo.tex
the tex. file was checked on grammarly, some comments were included in chapter 7 as a footnote, the abstract is still unchanged

@@
 
-\textcolor{red}{In the literature, the theoretical basis to describe equilibrium systems in the classical scheme is well-established.
-From thermodynamics to statistical mechanics, the procedures to describe mathematically equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature \cite{huang, reif, patrick}.
-In the case of the out of equilibrium systems, there are many works that attempt to describe out of equilibrium systems (however those works do not work in some systems, for example, the internal behavior of a glass due to the time relaxation is very big \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} as examples).
-Examples of some theoretical schemes to describe out of equilibrium systems can be the Onsager formulation (called linear thermodynamics), some few other restricted developments that attempt to complete the Onsager formulation \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the case of the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann's by means the Boltzmann transport equation.
+\textcolor{red}{In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
+From thermodynamics to statistical mechanics, the procedures to describe mathematically equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature.
+In the case of the out of equilibrium systems, many physics attempt to describe out of equilibrium systems (however those works do not work in some systems, for example, the internal behavior of glasses due to the time relaxation is  \cite{huang, reif, patrick}.
+In the case of the out of equilibrium systems, many physics attempt to describe out of equilibrium systems creating new theoretical basis (however those works do not work in some systems, for example, the internal behavior of glasses due to the time relaxation is very big \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} as examples).
+Examples of some theoretical basis can be the Onsager formulation (called linear thermodynamics), some few other restricted developments that attempt to complete the Onsager formulation \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.\\
+In the case of the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann's through the Boltzmann transport equation.
 Furthermore, the evolution to an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some some aspects in the Boltzmann's $H$ theorem and the Boltzmann's $H$ functional that deserve some review in order to accomplish better consistency, for example the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\ %\cite{paradox1, paradox2, hamiltonian, H-theorem-violation, H-theorem-and-entropy}.
+However, there are some aspects in the Boltzmann's $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\ %\cite{paradox1, paradox2, hamiltonian, H-theorem-violation, H-theorem-and-entropy}.
 \\
-In order to obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the $H$ theorem and the Boltzmann transport equation are necessary such that in the appropriate limit, the  Boltzmann's $H$ theorem and the Boltzmann transport equation should be recovered.\\
+To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the $H$ theorem and the Boltzmann transport equation is necessary such that in the appropriate limit, the  Boltzmann's $H$ theorem and the Boltzmann transport equation should be recovered.\\
 One of the first developments in this field was developed by Tolman \cite{tolman}.
-The quantum version of the $H$ theorem was established by Tolman using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
-The equation of motion for the occupation number was obtained by applying the time perturbation theory.
-However, in order to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-In other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (the analogous to the Boltzmann transport equation) [12-15].
+He proposes one of the first quantum versions of the $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
+The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
+However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
+On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (the analogous to the Boltzmann transport equation) [12-15].
 %\cite{entropic-framework, thermal-harmonic, quantum-transport, quantum-transport-reservoirs}.
-Besides some papers construct the $H$ functional and prove the $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are not clear or discussed.
-Moreover the validity of the quantum $H$ theorem and the second low of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\ %\cite{quantum-entropy, quantum-colapse, quantum3, quantum4, binary, contradictions,ergodic-theorem, hydrodynamic-simulation}.\\
+Besides, some authors construct the $H$ functional and prove the $H$ theorem \cite{htheorem2, quantum1, quantum2}.
+However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or discussed.
+Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\ %\cite{quantum-entropy, quantum-colapse, quantum3, quantum4, binary, contradictions,ergodic-theorem, hydrodynamic-simulation}.\\
 \\
-In order to contribute to a better consistent $H$ theorem and a new formalism to describe out of equilibrium systems, we define the theoretical basis to describe out of equilibrium systems in order to be completed in the future.\\
-In this work we define a new functional (variational entropy) which depends on the distribution function of an ideal quantum gas (Bosons or Fermions particles).
+In order to contribute to a better consistent $H$ theorem and a new formalism to describe out of equilibrium systems, we define the theoretical basis to describe out of equilibrium systems.\\
+In this work, we define a new functional (variational entropy) which depends on the distribution function of an ideal quantum gas (Bosons or Fermions).
 This gas is contained inside a volume $V$ divided into cells.
 Before developing the non-equilibrium description, we recover the Fermi-Dirac and Bose-Einstein distributions from the variational entropy when the system is in the equilibrium state.
-Furthermore we prove that the derivative of the variational entropy with respect to time is less or equal to zero for relaxation processes (the action of any perturbed system that returns to the equilibrium state) in any system.
+Furthermore, we prove that the derivative of the variational entropy with respect to time is less or equal to zero for relaxation processes (the action of any perturbed system that returns to the equilibrium state) in any system.
 This prove corresponds with the statement of the $H$ theorem.
-In addition we obtain the Boltzmann transport equation from the Boltzmann's $H$ functional and, in analogy with this, we find a quantum version of the Boltzmann transport equation from the variational entropy.
-Finally, in order to be consistence with the correspondence principle, we prove that the Boltzmann's $H$ functional is proportional to the variational entropy in the limit of the high level energies.}
+Besides, we obtain the Boltzmann transport equation from the Boltzmann's $H$ functional and, in analogy with this, we find a quantum version of the Boltzmann transport equation from the variational entropy.
+Finally, to be consistent with the correspondence principle, we prove that the Boltzmann's $H$ functional is proportional to the variational entropy in the limit of the high-level energies.}
 
@@
 \subsection{Classical approach}
-In the classical version, the Boltzmann transport equation is used to shown the $H$-theorem. That equation is
+In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
 \begin{equation}
@@
 \begin{equation}
-    H=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,
-\end{equation} \label{hfunctional}
+    H=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,\label{hfunctional}
+\end{equation} 
 where $f(v,t)$ is a homogeneous distribution function.
 The $H$ theorem establishes that if $f(v,t)$ satisfies the Boltzmann transport equation, then $\frac{dH}{dt}\leq 0$.
-The system arrives to the equilibrium state when $\frac{dH}{dt}=0$ and due to this, the distribution function to the system in equilibrium corresponds to the Maxwell-Boltzmann distribution
+The system arrives to the equilibrium state when $\frac{dH}{dt}=0$, and due to this the distribution function to the system in equilibrium corresponds to the Maxwell-Boltzmann distribution
 \subsection{Quantum approach}
@@
 \end{equation}
-where $n_i$ is the occupation number and $g_i$ is a energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
+where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
 \begin{eqnarray}
@@
 \end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expression are obtained by the perturbation theory.\\
-The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}), taking account the spatial-homogeneous distribution function hypothesis, as in the classical version.\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high energy levels to
+where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
+The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}), taking account of the spatial-homogeneous distribution function hypothesis, as in the classical version.\\
+In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
 \begin{equation}
@@
 \end{equation}
-This expression can be obtained also from the Boltzmann $H$ functional by means defining 
+This expression also can be obtained also from the Boltzmann $H$ functional through defining 
 \begin{equation}
@@
 \end{eqnarray}}
-This functional, in equilibrium should be a density of entropy, which we will call \textit{variational entropy}.
-This functional describes a system with global energy E, global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) N, global volume V, and temperature T. The
-global system in turn consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
+This functional will be named \textit{variational entropy}.
+This functional describes a system with a global energy E, a global number of free
+quantum particles (which may be fermions for upper sing or bosons for lower sing) N, global volume V, and a temperature T. 
+Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
+global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
 The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Besides the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
+Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
+We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
 .
@@
 Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its own local chemical potential, $\mu_M$, local
+local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
 variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
@@
 therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-In addition, the system is subjected to the microcanonical restrictions
+Besides, the system is subjected to the microcanonical restrictions
 \begin{eqnarray}
@@
 \end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multiplier are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
+We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
+We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
 \begin{eqnarray}
@@
   \end{eqnarray}
-  This result lets us calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe quantum ideal gases.
+  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.
 \subsection{Classical scheme: variational procedure}
@@
 \end{eqnarray}{}
-where $C$ is a constant. It is trivial that using the variational procedures, one obtains the form of the Maxwell-Boltzmann distribution. \\
+where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
 As $H'$ in equilibrium \textcolor{red}{is proportional to} entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
@@
 \\
-  \textcolor{green}{In other hand, if the distribution function is spatial-homogeneous, $f$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
+  \textcolor{gray}{In other hand, if the distribution function is spatial-homogeneous, $f$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
   \begin{equation}
@@
   \end{equation}
-  Besides, if we analyse the following expression
+  Besides, if we analyze the following expression
   \begin{eqnarray}
@@
   \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the the first term in (\ref{sum-h}) is negible and and consequently we conclude that
+  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
   \begin{equation}
@@
   and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-Those condition described a system out of equilibrium but no so far from it, that is a perturbed system.
+Those condition described a system out of equilibrium but no so far from it, that is a perturbed system.\\
+  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system in an out of equilibrium state but not far from the equilibrium state.
 \\
@@
 The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogenities because of the exclusion principle and as consequence of this,  $g_{nM}=0$.
+Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
 This result corresponds to the system in the zero temperature condition.
@@
 Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or/and missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or/and missing of particles or energy respectively.\\
-On the one hand, $\dot{g}^{+}_{nJ}<0$ describes the lost of particles and/or energy and so , $g^{+}_{nJ}>0$.
+On the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
 On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
@@
 When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any quantum ideal gas perturbed, evolves always to the equilibrium state, such as the classical $H$ theorem.
+Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
 \\
-Now we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
+Now we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first-order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
 
@@
 \end{equation}{}
-and applying the first order approximation (\ref{firstorder}) to (\ref{dH1}), we obtain
+and applying the first-order approximation (\ref{firstorder}) to (\ref{dH1}), we obtain
 \begin{equation}
@@
 \end{eqnarray}{}
-If we use the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), we obtain the same result. Then applying this to (\ref{dH1,2}), we find
+If we use the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), we obtain the same result. Then apply this to (\ref{dH1,2}), we find
 \begin{eqnarray}
@@
 The next step in this model is to obtain the analogous transport equation through variation methods.
-In order to verify that transport equation can be obtained by applying the variation method, in the next section the Boltzmann transport equation will be obtained applying the mentioned method to the modified Boltzmann $H$ functional (\ref{CH2}).
+In order to verify that the transport equation can be obtained by applying the variation method, in the next section, the Boltzmann transport equation will be obtained applying the mentioned method to the modified Boltzmann $H$ functional (\ref{CH2}).
 
@@
 \\
-Due to $H$ functional in equilibrium is proportional to entropy and also, according to the second law of thermodynamics, the entropy is maximum, then the following relation must be hold in equilibrium
+The following relation must be held in equilibrium 
 \begin{equation}
-    \delta H_{mod}=0.\label{variation1}
+    \delta H_{mod}=0,\label{variation1}
 \end{equation}
+This due to $H$ functional in equilibrium is proportional to entropy and a maximum. 
 Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
@@
 As a consequence of the $H$ theorem, the variation is less to zero.
-In conclusion, is necessary to include in (\ref{transport}) a term that described this behavior.
+In conclusion, it is necessary to include a term in (\ref{transport}).
 This term, as in \cite{huang}, will be the collision term and will be expressed in such form as
@@
 \end{equation}
-where $\vec{F}$ is a external force and $m$ is the mass of the particles.\\
+where $\vec{F}$ is an external force, and $m$ is the mass of the particles.\\
 \\
-The equation (\ref{classictransport}) correspond to the Boltzmann transport equation.
+The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation.
 With this conclusion, in the next section, the analogous transport equation will be obtained applying the variation method to (\ref{entropy2}).
@@
 This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In orden to obtain the analogous transport equation to this system, we apply the variation method to the functional $H$.
+In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $H$.
 Due to $H$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.\\
-Appliying the variation to the functional $H$, we obtain the following equation of motion
+Applying the variation to the functional $H$, we obtain the following equation of motion
 \begin{equation}
@@
 When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be added to (\ref{transportequation1}).
+Due to this, an expression has to be included in (\ref{transportequation1}).
 This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell \textcolor{blue}{(explanation about why this term contributes)}.
-Therefore, (\ref{transportequation1}) casts into
+Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{\textcolor{red}{The collision term is only in a defined cell and local group of energy.
+If we are describing the total system, we need to include all contributions (each cell and energy group).}} Therefore, (\ref{transportequation1}) casts into
 \begin{eqnarray}
@@
 \textcolor{red}{%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and a energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
+We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
 \begin{equation}
@@
 \end{eqnarray}
-In the classic limit, the particles are in the high energy levels and the number of particles  in most of the groups of state $n$ are small compared with the number of the elementary states $\epsilon_n$ of the group, that is
+In the classic limit, the particles are in the high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
 \begin{equation}
@@
 \end{eqnarray}
-where $B_{boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe in the same way that in the classical scheme that if the number of particles is too big compared to the number of cells $P$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.}
+where $B_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
+We can observe that if the number of particles is too big compared to the number of cells $P$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.}
 
@@
 \textcolor{red}{
-In the literature, the Boltzmann's $H$ theorem is proved using a spatial homogeneous distribution function.
+In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
 In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However we included inhomogeneities in the system through the division of the volume $V$ in cells and the %fact that the distribution function is different for each cell (
-non-homogeneous distribution function assumption in the total volume.
-Moreover the Boltzmann's $H$ theorem is still hold when the system includes inhomogeneities, only if the local equilibrium hypothesis is hold.
+However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
+Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
 We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
 It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through variational method.
-Nevertheless in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
+Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
+Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
 This is a consequence of the $H$ theorem.\\
@@
 In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the prove of the quantum version of the $H$ theorem we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis is hold.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-In other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be a future work.\\
+With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
+In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
+On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
+However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
+This problem may be future work.\\
 \\
@@
 This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover if we see the expression for the Boltzmann's $H$ functional in equilibrium
+Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
 \begin{equation}
-    H_{boltz}=-\frac{S}{Vk},
+    H_{Boltz}=-\frac{S}{Vk},
 \end{equation}
-where $S$ is the entropy of the system and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
+where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
 \begin{equation}
@@
 \end{equation}
-Finally We find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.}
+Finally, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.}
 
@@
 \begin{enumerate}
-    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis is hold.
+    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
     \item We can obtain the transport equation from the $H$ functional.

### 2020-10-27 notasparaarticulo.tex
La introducción fue modificada, el orden de los temas fue modificado y se agregaron varias partes a algunas secciones para que tengan sentido. Lo de rojo es lo agregado o modificado, lo de gris es la parte nueva que no se ha discutido aún

@@
 
-\section{Abstract}
+\section*{Abstract}
 We review in detail the quantum formulation of the Boltzamnn's H theorem.
@@
 \textcolor{red}{In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-From thermodynamics to statistical mechanics, the procedures to describe mathematically equilibrium systems for both schemes (phenomenological for thermodynamics and microscopical for statistical mechanics) are well-known and works in almost every system in nature.
-In the case of the out of equilibrium systems, many physics attempt to describe out of equilibrium systems (however those works do not work in some systems, for example, the internal behavior of glasses due to the time relaxation is  \cite{huang, reif, patrick}.
-In the case of the out of equilibrium systems, many physics attempt to describe out of equilibrium systems creating new theoretical basis (however those works do not work in some systems, for example, the internal behavior of glasses due to the time relaxation is very big \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} as examples).
-Examples of some theoretical basis can be the Onsager formulation (called linear thermodynamics), some few other restricted developments that attempt to complete the Onsager formulation \cite{kei, onsager} and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the case of the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann's through the Boltzmann transport equation.
-Furthermore, the evolution to an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the Boltzmann's $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\ %\cite{paradox1, paradox2, hamiltonian, H-theorem-violation, H-theorem-and-entropy}.
+The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
+In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
+Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
+In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
+Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
+However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
 \\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the $H$ theorem and the Boltzmann transport equation is necessary such that in the appropriate limit, the  Boltzmann's $H$ theorem and the Boltzmann transport equation should be recovered.\\
+To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
 One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
+He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
 The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
 However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (the analogous to the Boltzmann transport equation) [12-15].
-%\cite{entropic-framework, thermal-harmonic, quantum-transport, quantum-transport-reservoirs}.
-Besides, some authors construct the $H$ functional and prove the $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\ %\cite{quantum-entropy, quantum-colapse, quantum3, quantum4, binary, contradictions,ergodic-theorem, hydrodynamic-simulation}.\\
+On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
+Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
+However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
+Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
 \\
-In order to contribute to a better consistent $H$ theorem and a new formalism to describe out of equilibrium systems, we define the theoretical basis to describe out of equilibrium systems.\\
-In this work, we define a new functional (variational entropy) which depends on the distribution function of an ideal quantum gas (Bosons or Fermions).
-This gas is contained inside a volume $V$ divided into cells.
-Before developing the non-equilibrium description, we recover the Fermi-Dirac and Bose-Einstein distributions from the variational entropy when the system is in the equilibrium state.
-Furthermore, we prove that the derivative of the variational entropy with respect to time is less or equal to zero for relaxation processes (the action of any perturbed system that returns to the equilibrium state) in any system.
-This prove corresponds with the statement of the $H$ theorem.
-Besides, we obtain the Boltzmann transport equation from the Boltzmann's $H$ functional and, in analogy with this, we find a quantum version of the Boltzmann transport equation from the variational entropy.
-Finally, to be consistent with the correspondence principle, we prove that the Boltzmann's $H$ functional is proportional to the variational entropy in the limit of the high-level energies.}
+In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
+In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
+Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
+The first part of this work consists of obtaining the classical equilibrium distribution function (Maxwell-Boltzmann distribution), the Boltzmann transport equation from the Boltzmann $H$ functional (through variational procedures), and the proof of the Boltzmann's $H$ theorem (without using the Boltzmann transport equation).
+The second part consists of obtaining the quantum equilibrium distribution functions (Bose-Einstein and Fermi-Dirac distributions), the quantum analogous to the Boltzmann transport equation from a new functional named as \textit{variational entropy}, and the proof of the quantum analogous Boltzmann's $H$ theorem.
+The third part consists of analyzing the corresponding principle between the \textit{variational entropy} and Boltzmann's $H$ functional.}
 
@@
 
-\section{The equilibrium of dilute gases: a review}
-\subsection{Classical approach}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
+%----------------------------------
+\section{Classical scheme}
+\subsection{Maxwell-Boltzmann distribution function}
+\textcolor{red}{The Boltzmann's $H$ functional is defined as
 \begin{equation}
-    \left( \frac{\partial }{\partial t}+v_1 \cdot \nabla_r +\frac{F}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|v_1-v_2|(f_2'f_1'-f_2f_1),
+    H(t)=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,
 \end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
+where $f(v,t)$ is the spatial-homogeneous distribution function of a system. This Boltzmann's $H$ functional describes a system with a global energy $E$, a global number of free
+classical particles $N$, global volume $V$, and a temperature $T$. 
+Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
+However, we want to consider a $H$ functional with a non-homogeneous distribution function. Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
 \begin{equation}
-    H=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,\label{hfunctional}
+   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH2}.
 \end{equation}
-where $f(v,t)$ is a homogeneous distribution function.
-The $H$ theorem establishes that if $f(v,t)$ satisfies the Boltzmann transport equation, then $\frac{dH}{dt}\leq 0$.
-The system arrives to the equilibrium state when $\frac{dH}{dt}=0$, and due to this the distribution function to the system in equilibrium corresponds to the Maxwell-Boltzmann distribution
-\subsection{Quantum approach}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
+This $H'$ describes the same system mention above.
+Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(v,t)$.
+The distribution function will depend on the position of the cells, the velocity $v$, and time $t$.
+Without loss of generality, we choose these cells to have
+local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
+$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
+function and the energy $\epsilon(v)$, the last three properties are given by:
+\begin{eqnarray}
+    H_M &=&  \int f_M(v,t) \ln f_{M}(v,t) \mathrm{d}^{3}v \label{Hcell},\\
+    {\mathcal{N}}_M&=& \int f_{M}(v ,t) \mathrm{d}^{3}v, \nonumber \\
+{\mathcal{E}}_M&=& \int f_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v.
+\end{eqnarray}
+When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
 \begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
+   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
 \end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
+The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
+Besides, the local distribution function must satisfy the following microcanonical restrictions
+\begin{equation}
+    \sum_{M=1}^{K}\int_{}^{}f_M(v,t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v =E \label{micro}.
+\end{equation}
+We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
+Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining}
 \begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
+    \frac{\delta H'}{\delta f_J(v')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(v')}\left[f_M(v)\ln f_M(v)  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
+    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(v)\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
+    &=&\ln f_J(v')+1-\alpha_J-\beta_J \epsilon(v')=0.
 \end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
+Isolating the distribution function 
+\begin{eqnarray}
+    \ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f_J(v')=e^{\alpha_J +\beta_J \epsilon(v')-1} \nonumber \\
+    &=&Ce^{\alpha_J+\beta_J \epsilon(v') } \label{relacion1},
+\end{eqnarray}
+where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
+As $H'$ in equilibrium \textcolor{red}{is proportional to} entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
 \begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
+    f(v)=Ce^{\alpha+\beta \epsilon(v)}\equiv \bar{f}(v).
+\end{equation}{} 
+The constant $C$ can be omitted defining the following $H$ functional
+\begin{equation}
+   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(v,t) \ln f_M(v,t)-f_M(v,t)\right]\mathrm{d}^3v  \label{CH3},
 \end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}), taking account of the spatial-homogeneous distribution function hypothesis, as in the classical version.\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
+  but $\sum_{M=1}^{K} \int f_M(v,t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
+As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
+\\
+  \textcolor{gray}{On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
   \begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
+      H'(t)=\int \sum_{M=1}^{K} [f(v,t)\ln f(v,t)] \mathrm{d}^{3}v = K\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v= K H_{boltz}(t), 
   \end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
+  where
   \begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
+      H_{boltz}(t)=\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v.
   \end{equation}
-\section{Ensemble approach}
-\subsection{Quantum scheme}
+  Besides, if we analyze the following expression
+  \begin{eqnarray}
+      \int Kf(v,t) \ln [Kf(v,t)] \mathrm{d}^{3}v = \int [(K\ln K)f(v,t) + Kf(v,t) \ln f(v,t)]\mathrm{d}^{3}v, \label{sum-h}
+  \end{eqnarray}
+  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
+  \begin{equation}
+      \int f'(v,t) \ln f'(v,t) \mathrm{d}^{3}v = K\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v; \ \ \ f'(v,t)= Kf(v,t). \label{aditive-property} 
+  \end{equation}
+  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.}\\
+  \textcolor{red}{In the next subsection, we prove the classical $H$ theorem without using the Boltzmann transport equation.
+It is important to remark that we will use the $H'$ functional.
+This implies that we are considering a non-homogeneous distribution function.}
+
+%----------------------------------------------
+
+%-----------------------------------------------
+
+\subsection{Proof of the Boltzmann's $H$ theorem}
+\textcolor{red}{To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(v,t)$ must satisfy the following assumptions:
+\begin{itemize}
+    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
+    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
+    \item The following restrictions
+    \begin{eqnarray}
+        \int f_{M}(v,t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(v,t) \epsilon(v) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
+  \end{eqnarray}
+  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
+  \begin{equation}
+      \bar{\mathcal{N}}= \int \bar{f}(v) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(v)\epsilon(v) \mathrm{d}^{3}v,
+  \end{equation}
+  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
+\end{itemize} 
+ Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
+  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.}  \\
+Performing the derivative of (\ref{CH2}) with respect to time, we obtain
+\begin{equation}
+    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(v,t) \right]\dot f_M(v,t) \mathrm{d}^3v  \label{dH1}.
+\end{equation}{}
+\textcolor{red}{Using the first-order approximation
+\begin{eqnarray}
+   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}|g_{Mn}|, \label{firstorder}
+\end{eqnarray}}
+(\ref{dH1}) yields
+\begin{equation}
+    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g_M(v,t) \right\} \right]\dot g_M(v,t)\mathrm{d}^3v  \label{dH1,1}.
+\end{equation}
+\textcolor{red}{We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n g_{nM}=0$ 
+\begin{equation}
+    \ln [\bar{f}(v)+\bar{f}(v) g_{M}(v,t)] \approx \ln [\bar{f}(v)]+ g_{M}(v,t). \label{lnapproximationclassical}
+\end{equation}}
+With the previous approximation, (\ref{dH1,1}) obtains the following form 
+\begin{eqnarray}
+    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g_M(v,t) \right]\dot g_M(v,t)\mathrm{d}^3v,
+\end{eqnarray}{}
+using (\ref{relacion1}) with homogeneous multipliers
+\begin{eqnarray}
+    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g_M(v,t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v \label{dH1,2}.
+\nonumber \\
+\end{eqnarray}{}
+\textcolor{red}{On the other hand, we obtain from the restrictions the following expressions
+\begin{eqnarray}
+    \int \bar{f}(v) g_{M}(v,t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(v) \dot{g}_{M}(v,t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
+    \int  \bar{f}(v) g_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(v) \dot{g}_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
+\end{eqnarray}{}
+and as a consequence of $\sum_{M=1}^{K}
+\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
+\begin{equation}
+    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
+\end{equation}}
+Then using the previous result to (\ref{dH1,2}), we find
+\begin{eqnarray}
+\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v.
+\end{eqnarray}
+Now, the summation over $M$ will be expressed in two summations
+\begin{equation}
+    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(v)g_J^{+}(v,t)\dot g_J^{+}(v,t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(v)g_J^{-}(v,t)\dot g_J^{-}(v,t)\mathrm{d}^3v. \label{classicalH3}
+\end{equation}
+where $L+P=K$.
+\textcolor{red}{The group of $L$ cells represents such cells that have an excess of particles or/and energy (from the mean value in equilibrium).
+In contrast, the group of $P$ cells represents such cells that have to miss particles or/and energy.
+Besides, $\dot{g}^{+}_{J}$ and $\dot{g}^{-}_{J}$  represent the change on the deviations on cells that have excess or/and missing of particles or energy respectively while $g^{+}_{J}$ and $g^{-}_{J}$ represent the deviation on cells that have excess or/and missing of particles or energy respectively.\\
+On the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
+On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
+\\
+We sort the previous ideas in the following form
+\begin{eqnarray}
+   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
+   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
+\end{eqnarray}}
+and consequently, (\ref{classicalH3}) obtains the following form
+\begin{equation}
+    \frac{dH'}{dt}=-\left[
+                               \sum_J^{L}\int_{}^{}\bar f(v)|g_J^{+}(v,t)||\dot g_J^{+}(v,t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(v)|g_J^{-}(v,t)||\dot g_J^{-}(v,t)|\mathrm{d}^3v 
+    \right]. \label{classicalH4}
+\end{equation}
+\textcolor{red}{We can observe that $\bar{f}(v)$ in (\ref{classicalH4}) is always positive.
+Also the deviation and its derivative are positive, then all the expression is positive.
+However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.} With this we proved that $\frac{dH'}{dt}<0$.
+If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH'}{dt}=0$.\\
+\textcolor{red}{Joining both results, we can say the following statement:\\
+Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
+Consider also the system has inhomogeneities and suffers a relaxation process.
+If we define the following functional
+\begin{equation}
+   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH3},
+\end{equation}
+where $f_M(v,t)$ is the local distribution function of each cell in the system, then in the first-order approximation
+\begin{equation}
+    \frac{dH'}{dt} \leq 0.
+\end{equation}
+This statement will be the classical $H$ theorem with inhomogeneities.}
+\\
+In the next subsection, we obtain the Boltzmann transport equation from the Boltzmann's $H$ functional.
+%-----------------------------------
+\subsection{The Boltzmann transport equation}
+The Boltzmann $H$ functional is defined as
+\begin{equation}
+    H=\int_{}^{}f(\vec{v},t)\ln f(\vec{v},t)\mathrm{d}^{3}v,
+\end{equation}
+where $f$ is the spatial-homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space.
+\textcolor{red}{In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous.
+To obtain the spatial dependence in Boltzmann's transport equation, we need to assume that cells are too small, such that they can be considered as differential volume elements.
+That is
+\begin{equation}
+    f_M(\vec{v},t)    \ \ \ \Rightarrow \ \ \ f(\vec{r},\vec{v},t),
+\end{equation}
+where $r$ is the position of the particles defined in the $\mu$-space.} Because of that, we use the following functional
+\begin{equation}
+    H_{mod}=\int f(\vec{r},\vec{v},t) \ln f(\vec{r},\vec{v},t) \mathrm{d}^{3}v,
+\end{equation}
+that involves the position dependence.
+%\begin{equation}
+%    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
+%\end{equation}
+ %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
+\\
+The following relation must be held in equilibrium 
+\begin{equation}
+    \delta H_{mod}=0.\label{variation1}
+\end{equation}
+This due to $H$ functional in equilibrium is proportional to entropy and a maximum.\\ 
+Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
+\begin{equation}
+    0=\left( 1+\ln f \right)\left( \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f \right).
+\end{equation}
+The previous relation holds if $(1+\ln f)=0$ or $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.\\
+Assuming that $(1+\ln f)=0$, we find that $f$ is a constant and consequently, $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.
+On the other hand, if $f$ is not a constant, the equation of motion will be
+\begin{equation}
+    \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f=0.\label{transport}
+\end{equation}
+The equation (\ref{transport}) is hold only in equilibrium.
+If the system is out of equilibrium, the variation is not equal to zero.
+As a consequence of the classical $H$ theorem, the variation is less to zero.
+In conclusion, it is necessary to include a term in (\ref{transport}).
+This term, as in \cite{huang}, will be the collision term and will be expressed in such form as
+\begin{equation}
+    \frac{\partial f}{\partial t}+\vec{v}\cdot \nabla f+\frac{\vec{F}}{m}\cdot \nabla_{\vec{v}} f=\left( \frac{df}{dt} \right)_{coll}, \label{classictransport}
+\end{equation}
+where $\vec{F}$ is an external force, and $m$ is the mass of the particles.\\
+\\
+\textcolor{red}{The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation. \\
+We observed that the variational method and the inclusion of a non-spatial homogeneous distribution function are consistent with the results obtained from Statistical Mechanics.
+In other words, we proved the classical $H$ theorem and we obtained the Boltzmann transport equation from the Boltzmann $H$ functional.\\
+In the next section, we will apply the procedure used in this section to a new functional that describes a quantum gas (bosons or fermions).}
+
+%------------------------------
+
+\section{Quantum scheme}
+\subsection{Variational entropy, Bose-Einstein, and Fermi-Dirac distribution}
 
@@
 \begin{eqnarray}
-    \Ss (t)&=&\sum_{M} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
     &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
@@
 This functional will be named \textit{variational entropy}.
-This functional describes a system with a global energy E, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) N, global volume V, and a temperature T. 
+This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
+quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
 Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
@@
 \begin{eqnarray}
-    &&\sum_{M} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_M {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
+    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
+    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
 \end{eqnarray}
 where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Inspired by the Hamilton principle, $\Ss$ will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining
+Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
 \begin{eqnarray}
@@
 \begin{eqnarray}
-      \Ss(t)&=&\sum_M \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
+      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
+      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
   \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.
-\subsection{Classical scheme: variational procedure}
-We propose the following $H$ functional
+  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
+  \textcolor{red}{We obtained some results about equilibrium systems.
+Those results are consistent with the information obtained from Statistical Mechanics.
+Now we proceed to analyze the behavior of derivative of the variational entropy $\Ss$ with respect to time in order to find the quantum analogous to Boltzmann's $H$ theorem.}
+  %-----------------------------------------
 
+\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
+
+We rewrite the \textit{variational entropy} (\ref{entropy}) as
 \begin{equation}
-   H'(t)=\sum_M\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-The functional (\ref{CH2}) contains a no-homogeneous distribution function but keeps the same form of the Boltzman's $H$ functional.
-Excluding the homogeneous distribution function hypothesis, we use the same procedure that it was shown in the previous section to prove the $H$-theorem.\\
-If one considers the microcanonical ensemble conditions 
-\begin{equation}
-    \sum_M\int_{}^{}f_M(v,t)\mathrm{d}^3v =N, \ \ \ \sum_M\int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v =E \label{micro},
-\end{equation}{}
-one obtains the distribution function that minimizes $H$
-\begin{eqnarray}
-\frac{\delta H'}{\delta f_J(v')}&=&\sum_M\int_{}^{}\frac{\delta}{\delta f_J(v')}\left[f_M(v)\ln f_M(v)  \right]\mathrm{d}^3v -\sum_M\alpha_M\int_{}^{}\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
-&&-\sum_M\beta_M \int_{}^{}\epsilon(v)\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
-&=&\ln f_J(v')+1-\alpha_J-\beta_J \epsilon(v')=0.
-\end{eqnarray}{}
-isolating the distribution function 
-\begin{eqnarray}
-\ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f_J(v')=e^{\alpha_J +\beta_J \epsilon(v')-1} \nonumber \\
-&=&Ce^{\alpha_J+\beta_J \epsilon(v') } \label{relacion1},
-\end{eqnarray}{}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium \textcolor{red}{is proportional to} entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(v)=Ce^{\alpha+\beta \epsilon(v)}.
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_M\int_{}^{} \left[f_M(v,t) \ln f_M(v,t)-f_M(v,t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}{}
-  but $\sum_M \int f_M(v,t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  \textcolor{gray}{In other hand, if the distribution function is spatial-homogeneous, $f$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_M [f(v,t)\ln f(v,t)] \mathrm{d}^{3}v = P\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v= P H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Pf(v,t) \ln [Pf(v,t)] \mathrm{d}^{3}v = \int [(P\ln P)f(v,t) + Pf(v,t) \ln f(v,t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(v,t) \ln f'(v,t) \mathrm{d}^{3}v = P\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v; \ \ \ f'(v,t)= Pf(v,t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.}
-
-%----------------------------------------------
-
-\section{Variational entropy for relaxation processes in out of equilibrium systems}
-
-\subsection{Quantum Scheme}
-
-The \textit{variational entropy} has the following form
-\begin{equation}
-    \Ss(t)=\sum_{M} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
+    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
 \end{equation}
 where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function \textcolor{red}{must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells)}, and the following microcanonical restrictions
+This distribution function \textcolor{red}{must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells)}, and the following restrictions
 \begin{eqnarray}
@@
   and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-Those condition described a system out of equilibrium but no so far from it, that is a perturbed system.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system in an out of equilibrium state but not far from the equilibrium state.
-\\
+\textcolor{red}{The previous conditions and restrictions are the discrete analogous case of the classical (continuous) case.
+Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
+  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.}  \\
 If we perform the derivative of the variational entropy with respect to time, we obtain
 \begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_M \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
+   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
 \end{equation}{}
-Using the first-order approximation
+Using the first-order approximation (\ref{firstorder}), we get
 \begin{eqnarray}
-   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}|g_{Mn}|, \label{firstorder}
-\end{eqnarray}{}
-we get
-
-\begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_M \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_M \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
+    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
+    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
     \label{cambioH1}
@@
 \begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
+    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
+    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_M \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
+    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
 \nonumber \\
@@
 \end{eqnarray}{}
-and as a consequence of $\sum_M \Delta_M(t) \de V_M  =\sum_M \delta_M(t) \de V_M =0$, we find
+and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
 \begin{equation}
-    \sum_M \dot{\Delta}_M(t) \de V_M =\sum_M \dot{\delta}_{M}(t) \de V_M=0.
+    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
 \end{equation}{}
@@
 \end{eqnarray}{}
-Now, the summation over M will be expressed in two summations
+Now, the summation over $M$ will be expressed in two summations, such as in the classical case
 \begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_M \delta \epsilon_n+\sum^{N}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_M\delta \epsilon_n \right), \label{cambioH5}
+    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
 \end{equation}{}
-where $L+N=M$.
+where $L+P=K$.
 Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or/and missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or/and missing of particles or energy respectively.\\
@@
 \begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_M \delta \epsilon_n+\sum^{N}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_M \delta \epsilon_n \right), \label{cambioH6}
+    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
 \end{equation}{}
 and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{dH}{dt}<0$.\\
+With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
 When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
 Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-\\
-Now we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first-order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
+\textcolor{red}{Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
+Now, we proceed to obtain the quantum analogous of the Boltzmann transport equation from the variational entropy $\Ss$ through variational method.}
+%ow we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first-order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
 
-\subsection{Clasical scheme}
-Now we use the restrictions (\ref{restrictionoutside}) but in the form of the continuous case
-\begin{equation}
-    \int_{}^{}f_M(v,t)\mathrm{d}^3v=\mathcal{N}+\Delta_M(t); \ \ \ \int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v=\mathcal{E}+\Omega_M(t)\label{restriccionescambio},
-\end{equation}{}
-where $\Delta_M(t)$ and $\Omega_M(t)$ could be seen as a deviation from the mean number of particles and mean energy.
+  %-----------------------------------------
 
-Performing the derivative of the functional $H$ (\ref{CH2}) with respect to time
-\begin{equation}
-    \frac{dH'}{dt}=\sum_M\int_{}^{}\left[ 1+\ln f_M(v,t) \right]\dot f_M(v,t) \mathrm{d}^3v  \label{dH1},
-\end{equation}{}
-and applying the first-order approximation (\ref{firstorder}) to (\ref{dH1}), we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_M\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g_M(v,t) \right\} \right]\dot g_M(v,t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}{}
-Using the same approximation (\ref{lnapproximation})
-to (\ref{dH1,1}) 
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_M \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g_M(v,t) \right]\dot g_M(v,t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_M\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g_M(v,t)\mathrm{d}^3v +\sum_M\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v \label{dH1,2}. \nonumber \\
-\end{eqnarray}{}
-If we use the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), we obtain the same result. Then apply this to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_M\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v.
-\end{eqnarray}{}
-Using the same consideration in the previous section to $g$ and $\dot g$, that is to divide the spatial integral in two, one integral for cells that have excess of particles and/or energy, and the second for cells that have missing of particles and/or energy, and then using (\ref{separacion}), we prove that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH'}{dt}=0$.
-\\
-\\
-The next step in this model is to obtain the analogous transport equation through variation methods.
-In order to verify that the transport equation can be obtained by applying the variation method, in the next section, the Boltzmann transport equation will be obtained applying the mentioned method to the modified Boltzmann $H$ functional (\ref{CH2}).
-
-\section{The Boltzmann transport equation}
-The Boltzmann $H$ functional is defined as
-\begin{equation}
-    H=\int_{}^{}f(\vec{v},t)\ln f(\vec{v},t)\mathrm{d}^{3}v,
-\end{equation}
-where $f$ is the homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space.
-In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous, that is, $f=f(\vec{r},\vec{v},t)$, where $r$ is the position of the particles defined in the $\mu$-space.
-Because of that, we use the following functional
-\begin{equation}
-    H_{mod}=\int f(\vec{r},\vec{v},t) \ln f(\vec{r},\vec{v},t) \mathrm{d}^{3}v,
-\end{equation}
-that involves the position dependence.
-%\begin{equation}
-%    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
-%\end{equation}
- %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
-\\
-The following relation must be held in equilibrium 
-\begin{equation}
-    \delta H_{mod}=0,\label{variation1}
-\end{equation}
-This due to $H$ functional in equilibrium is proportional to entropy and a maximum. 
-Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
-\begin{equation}
-    0=\left( 1+\ln f \right)\left( \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f \right).
-\end{equation}
-The previous relation holds if $(1+\ln f)=0$ or $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.\\
-Assuming that $(1+\ln f)=0$, we find that $f$ is a constant and consequently, $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.
-In other hand, if $f$ is not a constant, the Boltzmann transport equation will be
-\begin{equation}
-    \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f=0.\label{transport}
-\end{equation}
-The equation (\ref{transport}) is hold only in equilibrium.
-If the system is out of equilibrium, the variation is not equal to zero.
-As a consequence of the $H$ theorem, the variation is less to zero.
-In conclusion, it is necessary to include a term in (\ref{transport}).
-This term, as in \cite{huang}, will be the collision term and will be expressed in such form as
-\begin{equation}
-    \frac{\partial f}{\partial t}+\vec{v}\cdot \nabla f+\frac{\vec{F}}{m}\cdot \nabla_{\vec{v}} f=\left( \frac{df}{dt} \right)_{coll}, \label{classictransport}
-\end{equation}
-where $\vec{F}$ is an external force, and $m$ is the mass of the particles.\\
-\\
-The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation.
-With this conclusion, in the next section, the analogous transport equation will be obtained applying the variation method to (\ref{entropy2}).
-
-\section{Analogous transport equation for ideal quantum gases}
+  \subsection{The quantum analogous of the Boltzmann transport equation}
 Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
@@
 This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $H$.
-Due to $H$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.\\
-Applying the variation to the functional $H$, we obtain the following equation of motion
+In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
+Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
+Then the equation of motion is
 \begin{equation}
-    0=\sum_M \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
+    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
 \label{transportequation1}
@@
 \begin{eqnarray}
-    \sum_n \sum_M\left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_M \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
+    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
     &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
@@
 \end{eqnarray}
+The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas.
+
+  %----------------------------------------
+
+%-------------------------------------------------
 
@@
 \begin{eqnarray}
-    H&=& \sum_M \sum_n
+    H&=& \sum_{M=1}^{K} \sum_n
     \left[  
@@
 \end{eqnarray}
-In the classic limit, the particles are in the high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
+In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
 \begin{equation}
@@
     \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M P H_{Boltz},\label{h-quantic4}
+    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
 \end{eqnarray}
 where $B_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $P$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.}
+We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.}
 
@@
 \end{thebibliography}{}
+\clearpage
+
+\section{Suplementary material}
+\subsection{The equilibrium of dilute gases: a review}
+\subsubsection{Classical approach}
+In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
+\begin{equation}
+    \left( \frac{\partial }{\partial t}+v_1 \cdot \nabla_r +\frac{F}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|v_1-v_2|(f_2'f_1'-f_2f_1),
+\end{equation}
+where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
+The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
+\begin{equation}
+    H=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,\label{hfunctional}
+\end{equation} 
+where $f(v,t)$ is a homogeneous distribution function.
+The $H$ theorem establishes that if $f(v,t)$ satisfies the Boltzmann transport equation, then $\frac{dH}{dt}\leq 0$.
+The system arrives to the equilibrium state when $\frac{dH}{dt}=0$, and due to this the distribution function to the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
+
+\subsubsection{Quantum approach}
+The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
+\begin{equation}
+    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
+\end{equation}
+where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
+\begin{eqnarray}
+    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
+    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
+\end{eqnarray}
+where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
+\begin{equation}
+  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
+\end{equation}
+where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
+The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}), taking account of the spatial-homogeneous distribution function hypothesis, as in the classical version.\\
+In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
+\begin{equation}
+    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
+\end{equation}
+This expression also can be obtained also from the Boltzmann $H$ functional through defining 
+\begin{equation}
+    f=\frac{n_{\kappa}}{ g_{\kappa}}.
+\end{equation}
 \end{document}

### 2020-11-06 notasparaarticulo.tex
some parts were modified. Those parts are in red

@@
 
-\textcolor{red}{In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
+In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
 The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
@@
 Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-The first part of this work consists of obtaining the classical equilibrium distribution function (Maxwell-Boltzmann distribution), the Boltzmann transport equation from the Boltzmann $H$ functional (through variational procedures), and the proof of the Boltzmann's $H$ theorem (without using the Boltzmann transport equation).
-The second part consists of obtaining the quantum equilibrium distribution functions (Bose-Einstein and Fermi-Dirac distributions), the quantum analogous to the Boltzmann transport equation from a new functional named as \textit{variational entropy}, and the proof of the quantum analogous Boltzmann's $H$ theorem.
-The third part consists of analyzing the corresponding principle between the \textit{variational entropy} and Boltzmann's $H$ functional.}
+\textcolor{red}{The first section of this work consists of obtaining some classical results from Statistical mechanics (the equilibrium distribution function, the Boltzmann transport equation, and an alternative proof of the classical $H$ theorem).
+The second section consists of obtaining the same results from quantum mechanics.
+The third consists of analyzing the corresponding principle between the \textit{variational entropy} and Boltzmann's $H$ functional.}
 
@@
 \section{Classical scheme}
+\textcolor{red}{This section will be divided in three subsections, the first will be dedicated to obtain the equilibrium distribution function (Maxwell-Boltzmann distribution) from the Boltzmann's $H$ functional applying variational procedure.}
 \subsection{Maxwell-Boltzmann distribution function}
-\textcolor{red}{The Boltzmann's $H$ functional is defined as
+The Boltzmann's $H$ functional is defined as
 \begin{equation}
@@
 We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining}
+Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
 \begin{eqnarray}
@@
 where $L+P=K$.
-\textcolor{red}{The group of $L$ cells represents such cells that have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells represents such cells that have to miss particles or/and energy.
-Besides, $\dot{g}^{+}_{J}$ and $\dot{g}^{-}_{J}$  represent the change on the deviations on cells that have excess or/and missing of particles or energy respectively while $g^{+}_{J}$ and $g^{-}_{J}$ represent the deviation on cells that have excess or/and missing of particles or energy respectively.\\
-On the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
+\textcolor{red}{The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
+In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
+Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
+On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
+In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
+Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
 On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
@@
 \end{equation}
-\textcolor{red}{We can observe that $\bar{f}(v)$ in (\ref{classicalH4}) is always positive.
+We can observe that $\bar{f}(v)$ in (\ref{classicalH4}) is always positive.
 Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.} With this we proved that $\frac{dH'}{dt}<0$.
+However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
+With this we proved that $\frac{dH'}{dt}<0$.
 If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-\textcolor{red}{Joining both results, we can say the following statement:\\
+Joining both results, we can say the following statement:\\
 Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
@@
 \end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.}
+This statement will be the classical $H$ theorem with inhomogeneities.
 \\
-In the next subsection, we obtain the Boltzmann transport equation from the Boltzmann's $H$ functional.
+%\textcolor{red}{The last result we want to obtain is the Boltzmann transport equation from the Boltzmann $H$ functional. This procedure is developed in the next subsection.}
+
 %-----------------------------------
+
 \subsection{The Boltzmann transport equation}
@@
 where $f$ is the spatial-homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space.
-\textcolor{red}{In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous.
+In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous.
 To obtain the spatial dependence in Boltzmann's transport equation, we need to assume that cells are too small, such that they can be considered as differential volume elements.
@@
 \end{equation}
-where $r$ is the position of the particles defined in the $\mu$-space.} Because of that, we use the following functional
+where $r$ is the position of the particles defined in the $\mu$-space. Because of that, we use the following functional
 \begin{equation}
@@
 \\
-\textcolor{red}{The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation. \\
+The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation. \\
 We observed that the variational method and the inclusion of a non-spatial homogeneous distribution function are consistent with the results obtained from Statistical Mechanics.
 In other words, we proved the classical $H$ theorem and we obtained the Boltzmann transport equation from the Boltzmann $H$ functional.\\
-In the next section, we will apply the procedure used in this section to a new functional that describes a quantum gas (bosons or fermions).}
+\textcolor{red}{In contrast, there is not quantum analogous of the Boltzmann transport equation.
+Even more, the analogous quantum $H$ theorem is not clear on the spatial-homogeneous distribution hypothesis.
+To obtain those results in the quantum mechanic scheme, the next section will develop the same procedure used in this section.}
 
@@
 \section{Quantum scheme}
+\textcolor{red}{This section will be divided into three subsections (as in the previous section).
+The first part consists of obtaining the quantum equilibrium distribution functions (Bose-Einstein and Fermi-Dirac distributions) from a new functional named \textit{variational entropy}.
+In the second part, we prove the quantum analogous of the $H$ theorem using the same way as in the previous section.
+Finally, in the third part, we obtain an equation of motion for the distribution function from the \textit{variational entropy}.
+This equation can be treated as the quantum analogous of the Boltzmann transport equation.}
 \subsection{Variational entropy, Bose-Einstein, and Fermi-Dirac distribution}
 
-\textcolor{red}{We begin by defining the form of the functional 
+We begin by defining the form of the functional 
 \begin{eqnarray}
@@
     &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
-\end{eqnarray}}
+\end{eqnarray}
 This functional will be named \textit{variational entropy}.
@@
   This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
-  \textcolor{red}{We obtained some results about equilibrium systems.
+ % \textcolor{red}{We obtained some results about equilibrium systems.
 Those results are consistent with the information obtained from Statistical Mechanics.
@@
 where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function \textcolor{red}{must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells)}, and the following restrictions
+This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
 \begin{eqnarray}
@@
   and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\textcolor{red}{The previous conditions and restrictions are the discrete analogous case of the classical (continuous) case.
+The previous conditions and restrictions are the discrete analogous case of the classical (continuous) case.
 Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.}  \\
+  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
+\\
 If we perform the derivative of the variational entropy with respect to time, we obtain
@@
     \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}{}
+\end{equation}
 where $L+P=K$.
-Besides, $\dot{g}^{+}_{nJ}$ and $\dot{g}^{-}_{nJ}$  represent the change on the deviations on cells that have excess or/and missing of particles or energy respectively while $g^{+}_{nJ}$ and $g^{-}_{nJ}$ represent the deviation on cells that have excess or/and missing of particles or energy respectively.\\
-On the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
+\textcolor{red}{Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
+On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
+In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
+Also, on the one hand}, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
 On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
@@
 Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-\textcolor{red}{Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-Now, we proceed to obtain the quantum analogous of the Boltzmann transport equation from the variational entropy $\Ss$ through variational method.}
+Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
+%Now, we proceed to obtain the quantum analogous of the Boltzmann transport equation from the variational entropy $\Ss$ through variational method.
 %ow we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first-order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
@@
 This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{\textcolor{red}{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).}} Therefore, (\ref{transportequation1}) casts into
+Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
+If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
 \begin{eqnarray}
@@
 \end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas.
+The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
+\textcolor{red}{We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
+However, the correspondence principle is important to recover all results from Classical Mechanics.
+This correspondence principle will be discussed in the next section.}
 
@@
 
-\section{Classical limit}
-\textcolor{red}{%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
+\section{Correspondence principle}
+%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
 We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
@@
 where $B_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.}
+We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.
 
@@
 \section{Discussions and remarks}
-\textcolor{red}{
+
 In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
@@
 \end{equation}
-Finally, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.}
+Finally, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.
 
@@
 \section{Conclusions}
-\textcolor{red}{We conclude the following statements:\\
+We conclude the following statements:\\
 \begin{enumerate}
@@
     \end{enumerate}
-  }
+
 %  \begin{equation}

### 2020-12-15 notasparaarticulo.tex
The order of the sections were replaced and some modification were applied

@@
 \section*{Abstract}
-We review in detail the quantum formulation of the Boltzamnn's H theorem.
-Starting from a H functional defined in the energy space.
-By means a variational procedure, we demonstrate the analogous of the quantum H theorem.
-At difference of the Tolman's procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous system, and show that when $\frac{dH}{dt}=0$, the system reaches the equilibrium condition and then proposed H functional becomes the well known expression of the entropy of an ideal quantum gas.
-The time evolution of the system happens in such a way that $\frac{dH}{dt}<0$.
-Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium, Finally, in the limit of high energy levels, the Boltzmann's $H$ functional is recovered.
+Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
+This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
+The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
+The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
+In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
+Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
 
@@
 Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of obtaining some classical results from Statistical mechanics (the equilibrium distribution function, the Boltzmann transport equation, and an alternative proof of the classical $H$ theorem).
-The second section consists of obtaining the same results from quantum mechanics.
-The third consists of analyzing the corresponding principle between the \textit{variational entropy} and Boltzmann's $H$ functional.}
-
-%It is well known for the Physics community that the set of schemes, methods, techniques, and approximations of Statistical Physics to deal with equilibrium systems, are relatively well established.
-Liouvilles' theorem, Ergodic hypothesis and the main corollary of the H theorem, are some of the conceptual basis on which the theoretical structures of Statistical Thermodynamics have been grown and are sustained \cite{huang, reif, patrick}.
-This is not the case of the Statistical Physics of out of equilibrium systems, in which Poincares' theorem, Ergodic hypothesis, and H theorem establish the general conceptual basis of the theoretical developments, but these later are not as general and satisfactory as those for equilibrium [ref].
-There are of course some well stablished theoretical schemes to deal with out of equilibrium systems like those developed by Onsager known as linear thermodynamics, and some few other restricted developments proposed by some outstanding theoretical authors \cite{kei, onsager}.
-Nevertheless, even the very fundamentals of out of equilibrium Statistical physics like Boltzmanns' H Theorem, still presents some aspects that deserve some review in order to accomplish better consistency [ref].
-In the framework of classical mechanics, the prove of the Boltzmanns' H theorem in the scheme of Kinetic Theory of dilute gases, is founded on the Boltzmann Transport  Equation, which in turn is supported on the Molecular Chaos Hypothesis.
-%It is commonly accepted that BHT solves in a more complete way the Poicares' paradox and concilies the the theoretical conflict between a time reversible microscopic dynamics and an inherent macroscopic one \cite{paradox1, paradox2}.
-The quantum version of H Theorem, still currently is under analysis and discussion \cite{quantum1, quantum2, quantum3, quantum4}.
-
-%The behavior of the $H$ functional in times bigger than collisions, will be soft and decreasing.
-When the change of $H$ is zero, the system will be in equilibrium and the corresponding distribution function will be the Maxwell-Boltzmann distribution\cite{huang, patrick}.
-\\%
-
-%revisar la redaccion y coherencia de estas ideas% 
-%One of those discussions is the propose of Tolman.
-He established the quantum $H$ theorem \cite{tolman} using the probability transition relation, the random phases hypothesis and a homogeneous distribution function.
-The lack of the quantum Boltzmann transport equation in analogy to classical statistical mechanics has stopped the development of the quantum transport formulation and the quantum non-equilibrium formulation.
-Another discission is the prove of the quantum $H$ theorem using n-colissions.
-This idea completes the $H$ theorem in a classical and quantum scheme \cite{binary}.
-There are some special conditions in quantum mechanics where the second law of thermodynamics is violated \cite{contradictions}, making that the quantum $H$ theorem is not thoroughly understood.\\
-%In transport phenomena, the Boltzmann transport equation is solved using the zero and first order approximation.
-When one uses the zero order approximation, one finds that the total heat flux in the system is zero, that is a contradiction.
-However, when one applies the first order approximation, one finds the evolution operator expressed by the left expression from the Boltzmann transport equation.
-That evolution operator tends to zero when time tends to infinity, this is the equilibrium condition \cite{reif}.
-This result, that describes inhomogeneties, is not expressed in the $H$ theorem.\\
-%In order to contribute to a consistent $H$ theorem, in this paper, the quantum and classical $H$ theorem are proved proposing a new $H$ functional for both cases considering a non-homogeneous distribution function and a system divided into imaginary cells.
-With those assumptions, particles transport is allowed inside the system.
-Performing variational procedures, one finds that, in the quantum case, Bose-Einstein and Fermi-Dirac distributions minimize the $H$ functional and, in the classical case, Maxwell-Boltzmann distribution minimize the other $H$ functional.
-A expression of entropy to quantum systems will be obtained in the variational procedure.
-In other hand, making use of the hypothesis that the system is in a relaxation process and the non-homogeneous distribution function, the $H$ theorem will be proved for both cases.
+\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
+The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
+The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
+Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
 
@@
 \section{Classical scheme}
-\textcolor{red}{This section will be divided in three subsections, the first will be dedicated to obtain the equilibrium distribution function (Maxwell-Boltzmann distribution) from the Boltzmann's $H$ functional applying variational procedure.}
-\subsection{Maxwell-Boltzmann distribution function}
-The Boltzmann's $H$ functional is defined as
+\subsection{The Boltzmann transport equation}
+In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
 \begin{equation}
-    H(t)=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,
+    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
 \end{equation}
-where $f(v,t)$ is the spatial-homogeneous distribution function of a system. This Boltzmann's $H$ functional describes a system with a global energy $E$, a global number of free
+where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
+\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
+On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
+However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
+The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
+%-------------------------
+\subsection{H-theorem and the Maxwell-Boltzmann distribution}
+The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
+\begin{equation}
+   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
+\end{equation} 
+where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
 classical particles $N$, global volume $V$, and a temperature $T$. 
 Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-However, we want to consider a $H$ functional with a non-homogeneous distribution function. Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
+\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
+The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
+As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
+This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
 \begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH2}.
+    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
 \end{equation}
-This $H'$ describes the same system mention above.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(v,t)$.
-The distribution function will depend on the position of the cells, the velocity $v$, and time $t$.
+This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
+The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
+\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
+For more information, see \cite{huang}.}}
+%---------------------------
+\subsection{Out of equilibrium, non-homogeneous distributions}
+The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
+However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
+Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
+\begin{equation}
+   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
+\end{equation}
+This $H'$ describes a system with a global energy $E$, a global number of free
+classical particles $N$, global volume $V$, and a temperature $T$.
+Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
+The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
 Without loss of generality, we choose these cells to have
 local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and the energy $\epsilon(v)$, the last three properties are given by:
+$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
+function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
 \begin{eqnarray}
-    H_M &=&  \int f_M(v,t) \ln f_{M}(v,t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}_M&=& \int f_{M}(v ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}_M&=& \int f_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v.
+    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
+    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
+{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
 \end{eqnarray}
@@
 \begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
+   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
 \end{equation}
@@
 \begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(v,t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v =E \label{micro}.
+    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
 \end{equation}
@@
 \begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(v')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(v')}\left[f_M(v)\ln f_M(v)  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(v)\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(v')+1-\alpha_J-\beta_J \epsilon(v')=0.
+    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
+    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
+    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
 \end{eqnarray}
@@
 \begin{eqnarray}
-    \ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f_J(v')=e^{\alpha_J +\beta_J \epsilon(v')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(v') } \label{relacion1},
+    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
+    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
 \end{eqnarray}
 where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium \textcolor{red}{is proportional to} entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
+As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
 \begin{equation}
-    f(v)=Ce^{\alpha+\beta \epsilon(v)}\equiv \bar{f}(v).
+    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
 \end{equation}{} 
@@
 \begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(v,t) \ln f_M(v,t)-f_M(v,t)\right]\mathrm{d}^3v  \label{CH3},
+   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
 \end{equation}
-  but $\sum_{M=1}^{K} \int f_M(v,t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
+  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
 As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
 \\
-  \textcolor{gray}{On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
+  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
   \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(v,t)\ln f(v,t)] \mathrm{d}^{3}v = K\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v= K H_{boltz}(t), 
+      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
   \end{equation}
@@
   \begin{equation}
-      H_{boltz}(t)=\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v.
+      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
   \end{equation}
@@
   \begin{eqnarray}
-      \int Kf(v,t) \ln [Kf(v,t)] \mathrm{d}^{3}v = \int [(K\ln K)f(v,t) + Kf(v,t) \ln f(v,t)]\mathrm{d}^{3}v, \label{sum-h}
+      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
   \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
+  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
   \begin{equation}
-      \int f'(v,t) \ln f'(v,t) \mathrm{d}^{3}v = K\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v; \ \ \ f'(v,t)= Kf(v,t). \label{aditive-property} 
+      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
   \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.}\\
-  \textcolor{red}{In the next subsection, we prove the classical $H$ theorem without using the Boltzmann transport equation.
-It is important to remark that we will use the $H'$ functional.
-This implies that we are considering a non-homogeneous distribution function.}
-
-%----------------------------------------------
-
-%-----------------------------------------------
-
-\subsection{Proof of the Boltzmann's $H$ theorem}
-\textcolor{red}{To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(v,t)$ must satisfy the following assumptions:
+  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
+  \\
+To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
 \begin{itemize}
@@
     \begin{eqnarray}
-        \int f_{M}(v,t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(v,t) \epsilon(v) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
+        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
   \end{eqnarray}
@@
   \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(v) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(v)\epsilon(v) \mathrm{d}^{3}v,
+      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
   \end{equation}
@@
  Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.}  \\
+  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
+\\
 Performing the derivative of (\ref{CH2}) with respect to time, we obtain
 \begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(v,t) \right]\dot f_M(v,t) \mathrm{d}^3v  \label{dH1}.
+    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
 \end{equation}{}
-\textcolor{red}{Using the first-order approximation
+Using the first-order approximation
 \begin{eqnarray}
-   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}|g_{Mn}|, \label{firstorder}
-\end{eqnarray}}
+   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
+\end{eqnarray}
 (\ref{dH1}) yields
 \begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g_M(v,t) \right\} \right]\dot g_M(v,t)\mathrm{d}^3v  \label{dH1,1}.
+    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
 \end{equation}
-\textcolor{red}{We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n g_{nM}=0$ 
+We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
 \begin{equation}
-    \ln [\bar{f}(v)+\bar{f}(v) g_{M}(v,t)] \approx \ln [\bar{f}(v)]+ g_{M}(v,t). \label{lnapproximationclassical}
-\end{equation}}
+    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
+\end{equation}
 With the previous approximation, (\ref{dH1,1}) obtains the following form 
 \begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g_M(v,t) \right]\dot g_M(v,t)\mathrm{d}^3v,
+    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
 \end{eqnarray}{}
@@
 \begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g_M(v,t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v \label{dH1,2}.
+    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
 \nonumber \\
-\end{eqnarray}{}
-\textcolor{red}{On the other hand, we obtain from the restrictions the following expressions
+\end{eqnarray}
+On the other hand, we obtain from the restrictions the following expressions
 \begin{eqnarray}
-    \int \bar{f}(v) g_{M}(v,t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(v) \dot{g}_{M}(v,t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(v) g_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(v) \dot{g}_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
+    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
+    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
 \end{eqnarray}{}
@@
     \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}}
+\end{equation}
 Then using the previous result to (\ref{dH1,2}), we find
 \begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v.
+\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
 \end{eqnarray}
@@
 \begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(v)g_J^{+}(v,t)\dot g_J^{+}(v,t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(v)g_J^{-}(v,t)\dot g_J^{-}(v,t)\mathrm{d}^3v. \label{classicalH3}
+    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
+\label{classicalH3}
 \end{equation}
 where $L+P=K$.
-\textcolor{red}{The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
+The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
 In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
@@
    &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}}
+\end{eqnarray}
 and consequently, (\ref{classicalH3}) obtains the following form
@@
     \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(v)|g_J^{+}(v,t)||\dot g_J^{+}(v,t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(v)|g_J^{-}(v,t)||\dot g_J^{-}(v,t)|\mathrm{d}^3v 
+                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
     \right]. \label{classicalH4}
 \end{equation}
-We can observe that $\bar{f}(v)$ in (\ref{classicalH4}) is always positive.
+We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
 Also the deviation and its derivative are positive, then all the expression is positive.
@@
 With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH'}{dt}=0$.\\
+If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
 Joining both results, we can say the following statement:\\
@@
 \begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH3},
+   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
 \end{equation}
-where $f_M(v,t)$ is the local distribution function of each cell in the system, then in the first-order approximation
+where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
 \begin{equation}
@@
 \end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.
-\\
-%\textcolor{red}{The last result we want to obtain is the Boltzmann transport equation from the Boltzmann $H$ functional. This procedure is developed in the next subsection.}
-
-%-----------------------------------
-
-\subsection{The Boltzmann transport equation}
-The Boltzmann $H$ functional is defined as
-\begin{equation}
-    H=\int_{}^{}f(\vec{v},t)\ln f(\vec{v},t)\mathrm{d}^{3}v,
-\end{equation}
-where $f$ is the spatial-homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space.
-In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous.
-To obtain the spatial dependence in Boltzmann's transport equation, we need to assume that cells are too small, such that they can be considered as differential volume elements.
-That is
-\begin{equation}
-    f_M(\vec{v},t)    \ \ \ \Rightarrow \ \ \ f(\vec{r},\vec{v},t),
-\end{equation}
-where $r$ is the position of the particles defined in the $\mu$-space. Because of that, we use the following functional
-\begin{equation}
-    H_{mod}=\int f(\vec{r},\vec{v},t) \ln f(\vec{r},\vec{v},t) \mathrm{d}^{3}v,
-\end{equation}
-that involves the position dependence.
-%\begin{equation}
-%    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
-%\end{equation}
- %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
-\\
-The following relation must be held in equilibrium 
-\begin{equation}
-    \delta H_{mod}=0.\label{variation1}
-\end{equation}
-This due to $H$ functional in equilibrium is proportional to entropy and a maximum.\\ 
-Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
-\begin{equation}
-    0=\left( 1+\ln f \right)\left( \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f \right).
-\end{equation}
-The previous relation holds if $(1+\ln f)=0$ or $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.\\
-Assuming that $(1+\ln f)=0$, we find that $f$ is a constant and consequently, $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.
-On the other hand, if $f$ is not a constant, the equation of motion will be
-\begin{equation}
-    \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f=0.\label{transport}
-\end{equation}
-The equation (\ref{transport}) is hold only in equilibrium.
-If the system is out of equilibrium, the variation is not equal to zero.
-As a consequence of the classical $H$ theorem, the variation is less to zero.
-In conclusion, it is necessary to include a term in (\ref{transport}).
-This term, as in \cite{huang}, will be the collision term and will be expressed in such form as
-\begin{equation}
-    \frac{\partial f}{\partial t}+\vec{v}\cdot \nabla f+\frac{\vec{F}}{m}\cdot \nabla_{\vec{v}} f=\left( \frac{df}{dt} \right)_{coll}, \label{classictransport}
-\end{equation}
-where $\vec{F}$ is an external force, and $m$ is the mass of the particles.\\
-\\
-The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation. \\
-We observed that the variational method and the inclusion of a non-spatial homogeneous distribution function are consistent with the results obtained from Statistical Mechanics.
-In other words, we proved the classical $H$ theorem and we obtained the Boltzmann transport equation from the Boltzmann $H$ functional.\\
-\textcolor{red}{In contrast, there is not quantum analogous of the Boltzmann transport equation.
-Even more, the analogous quantum $H$ theorem is not clear on the spatial-homogeneous distribution hypothesis.
-To obtain those results in the quantum mechanic scheme, the next section will develop the same procedure used in this section.}
+This statement will be the classical $H$ theorem with inhomogeneities.\\
+\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
 
@@
 \section{Quantum scheme}
-\textcolor{red}{This section will be divided into three subsections (as in the previous section).
-The first part consists of obtaining the quantum equilibrium distribution functions (Bose-Einstein and Fermi-Dirac distributions) from a new functional named \textit{variational entropy}.
-In the second part, we prove the quantum analogous of the $H$ theorem using the same way as in the previous section.
-Finally, in the third part, we obtain an equation of motion for the distribution function from the \textit{variational entropy}.
-This equation can be treated as the quantum analogous of the Boltzmann transport equation.}
-\subsection{Variational entropy, Bose-Einstein, and Fermi-Dirac distribution}
+\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
+The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
+\begin{equation}
+    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
+\end{equation}
+where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
+\begin{eqnarray}
+    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
+    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
+\end{eqnarray}
+where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
+\begin{equation}
+  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
+\end{equation}
+where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
+\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
+\begin{equation}
+    \frac{dH}{dt}\leq 0,
+\end{equation}
+and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
+The proof of this theorem is in the same way as in the classical version.
+Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
+Those distributions can be obtained from a sufficient condition for the equilibrium state
+\begin{equation}
+    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
+\end{equation}
+where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
+In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
+\begin{equation}
+    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
+\end{equation}
+This expression also can be obtained also from the Boltzmann $H$ functional through defining 
+\begin{equation}
+    f=\frac{n_{\kappa}}{ g_{\kappa}}.
+\end{equation}
+%--------------------------
+\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
 
@@
   This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
- % \textcolor{red}{We obtained some results about equilibrium systems.
-Those results are consistent with the information obtained from Statistical Mechanics.
-Now we proceed to analyze the behavior of derivative of the variational entropy $\Ss$ with respect to time in order to find the quantum analogous to Boltzmann's $H$ theorem.}
-  %-----------------------------------------
-
-\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
-
+ %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
 We rewrite the \textit{variational entropy} (\ref{entropy}) as
@@
   and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the discrete analogous case of the classical (continuous) case.
+The previous conditions and restrictions are the quantum analogous case of the classical case.
 Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
@@
 where $L+P=K$.
-\textcolor{red}{Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
+Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
 On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
 In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand}, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
+Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
 On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
@@
 Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-%Now, we proceed to obtain the quantum analogous of the Boltzmann transport equation from the variational entropy $\Ss$ through variational method.
-%ow we will apply this method to the modified Boltzmann's $H$ functional (\ref{CH2}) to prove the $H$ theorem without using only the first-order approximation (\ref{firstorder}), and the behavior of the deviation $g_{nM}$ and its time derivative (\ref{separacion}).
+We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
+However, the correspondence principle is important to recover all results from Classical Mechanics.
+This correspondence principle will be discussed in the next section.
 
@@
 
-  \subsection{The quantum analogous of the Boltzmann transport equation}
+%---------------------------------------------
+
+\section{Quantum Classical correspondence}
+%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
+We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
+\begin{equation}
+    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
+\end{equation}
+Using this definition on the variational entropy (\ref{entropy2}) we obtain
+\begin{eqnarray}
+    \Ss&=& \sum_{M=1}^{K} \sum_n
+    \left[  
+           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
+           \right)\pm 
+           \left(  
+                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
+           \right) \ln 
+           \left(  
+                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
+           \right)
+    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
+    \label{h-quantic} 
+\end{eqnarray}
+In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
+\begin{equation}
+    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
+\end{equation}
+and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
+\begin{eqnarray}
+    \Ss&=& \sum_{M=1}^{K} \sum_n
+    \left[  
+           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
+           \right)
+    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
+\end{eqnarray}
+In the case when the system is spatial-homogeneous, the definition of the distribution function is 
+\begin{equation}
+    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
+\end{equation}
+and using the same limit we obtain
+\begin{eqnarray}
+    \Ss&=& \sum_{M=1}^{K} \sum_n
+    \left[  
+           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{n}}{ \delta \epsilon_{n}}
+           \right)
+    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
+    \left[  
+           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
+           \left( 
+                  \frac{N_{n}}{ \delta \epsilon_{n}}
+           \right)
+    \right] \delta \epsilon_{n} \nonumber \\
+    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
+\end{eqnarray}
+where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
+We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
+\textcolor{blue}{In the equilibrium case, we have the following variational entropy
+\begin{eqnarray}
+    \Ss&=& \sum_{M=1}^{K}\sum_{n}
+        \left[
+                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
+                    \left(
+                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
+                    \right)
+        \right. \nonumber \\
+          && \pm \left. \left(
+                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
+                  \right) \ln
+                \left(
+                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
+                \right) \right]. \label{equilibriumvariational}
+\end{eqnarray}
+To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
+\begin{equation}
+    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
+\end{equation}
+and then
+\begin{equation}
+    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
+\end{equation}
+If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
+\begin{eqnarray}
+    \Ss&=& \sum_{M=1}^{K}\sum_{n}
+        \left[
+                e^{\alpha_M+\beta_M \epsilon} \ln 
+                    \left(
+                            e^{\alpha_M+\beta_M \epsilon}
+                    \right)
+        \right. \nonumber \\
+          && \pm \left. \left(
+                        1 \mp e^{\alpha_M+\beta_M \epsilon}
+                  \right) \ln
+                \left(
+                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
+                \right) \right],
+\end{eqnarray}
+but if (\ref{classicalapproximation}) holds, then 
+\begin{equation}
+     e^{\alpha_M+\beta_M \epsilon}\ll 1,
+\end{equation}
+consequently 
+\begin{equation}
+    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
+\end{equation}
+and finally
+\begin{equation}
+    \Ss=\sum_{M=1}^{K}\sum_{n}
+        \left[
+                e^{\alpha_M+\beta_M \epsilon} \ln 
+                    \left(
+                            e^{\alpha_M+\beta_M \epsilon}
+                    \right)
+        \right],
+\end{equation}
+that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
+\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
+
+%--------------------------------------
+
+  \section{Relaxation processes in degenerated quantum gases}
 Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
@@
 The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
-\textcolor{red}{We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.}
 
@@
 %-------------------------------------------------
-
-%---------------------------------------------
-
-\section{Correspondence principle}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    H&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    H&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    H&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
-\end{eqnarray}
-where $B_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.
-
-%--------------------------------------
 
@@
 \end{equation}
-Finally, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of the entropy.
+\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
+It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
+This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
+Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
+Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
+According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
+In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
+Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
 
@@
 \end{thebibliography}{}
-\clearpage
 
-\section{Suplementary material}
-\subsection{The equilibrium of dilute gases: a review}
-\subsubsection{Classical approach}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+v_1 \cdot \nabla_r +\frac{F}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|v_1-v_2|(f_2'f_1'-f_2f_1),
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-    H=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(v,t)$ is a homogeneous distribution function.
-The $H$ theorem establishes that if $f(v,t)$ satisfies the Boltzmann transport equation, then $\frac{dH}{dt}\leq 0$.
-The system arrives to the equilibrium state when $\frac{dH}{dt}=0$, and due to this the distribution function to the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-
-\subsubsection{Quantum approach}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
-\end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
-\begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
-\end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
-\begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
-\end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-The quantum $H$-theorem is proved using (\ref{changen}) into (\ref{quantumh}), taking account of the spatial-homogeneous distribution function hypothesis, as in the classical version.\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
 \end{document}

### 2020-12-15 tex/classical.tex
Sources reformatted to match the template of Entropy journal.

%\section{Classical}
%\subsection{Boltzmann transport equation}
%\subsection{H theorem and the Maxwell-Boltzmann distribution}
%Todo esto de arriba en la forma tradicional y muy resumido. Solo remarcando lo fundamental.
%\subsection{Out of equilibrium, non-homogeneous distributions}.
%Entrar a la generalizacion, ahi si de manera muy detallada y con amplitud discutida
%fisicamente, del tratamiento con las celdas. Es decir, la demostracion del teorema H
%clasico pero con el enfoque de funciones de distribucion no homogeneas y por tanto
%con el tratamiento de celdas.

%----------------------------------

\section{Classical scheme}

\subsection{The Boltzmann transport equation}

In the classical version, the Boltzmann transport equation is required to show
the $H$-theorem. That equation is
%
\begin{equation}\label{eq:transport}
  \left(
    \frac{\partial}{\partial t}+\vec{v}_1 \cdot\nabla_r
    +\frac{\vec{F}}{m}\cdot\nabla_{v_1}
  \right)f_1=\int\mathrm{d}\Omega\int
    \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1),
\end{equation}
%
where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the
external force applied to the system, $f_1$, $f_2$ are the distribution
functions of the particles 1 and 2 respectively before the collision,
and $f'_1$, $f'_2$ are the distribution function of the particles 1 and 2
respectively after the collision.

On the one hand, dynamics, inhomogeneities, and external forces are described
by the left-hand side of the equation (\ref{eq:transport}). On the other hand, the
binary collisions among particles are involved in the right-hand side of the
previous equation. However, the \textit{molecular chaos hypothesis} was applied
to obtain the right-hand side of (\ref{eq:transport}). The \textit{molecular chaos
hypothesis} establishes that spatial and velocity coordinates of the particles
are not correlated.

%-------------------------
\subsection{H-theorem and the Maxwell-Boltzmann distribution}
The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
%
\begin{equation}\label{eq:hfunctional}
   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,
\end{equation} 
%
where $f(\vec{v},t)$ is a homogeneous distribution function.
This $H$ functional describes a system with a global energy $E$, a global
number of free classical particles $N$, global volume $V$, and a temperature $T$.
Besides, the Boltzmann's $H$ functional correspond to a density of entropy
when the system arrives at the equilibrium state.

The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
transport equation and the molecular chaos hypothesis, then
$dH/dt\leq0$ and the system arrives to the equilibrium state when
$dH/dt=0$.
The theorem can be proved by applying the time derivative
to Eq. (\ref{eq:hfunctional}) and using the Boltzmann transport equation
(\ref{eq:transport}). As a consequence of this theorem, the distribution function
that describes the system in equilibrium corresponds to the Maxwell-Boltzmann
distribution. This distribution can be obtained by solving a sufficient
condition for $f$ to obtain the equilibrium distribution function $f_0$:
%
\begin{equation}
    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
\end{equation}
%

This condition can be obtained from the binary collision term of the Boltzmann
transport equation (\ref{eq:transport}).

The $H$-theorem fixes the time evolution direction of systems in nature, such
as the second law of thermodynamics does. {\color{blue}footnote: }The molecular hypothesis is %
fundamental to determine the physical meaning of the $H$ theorem. For more %
information, see \cite{bib:huang}.

%---------------------------
\subsection{Out of equilibrium, non-homogeneous distributions}

The Boltzmann $H$-functional (\ref{eq:hfunctional}) is defined with a spatially
homogeneous distribution $f(\vec{v},t)$. However, we want to consider a $H$
functional with a non-homogeneous distribution function in order to generalize
the Boltzmann $H$ theorem. Because of this, we introduce a modified $H$
functional (denoted as $H'$) defined as
%
\begin{equation}\label{eq:CH2}
   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%
This $H'$ describes a system with a global energy $E$, a global number of free
classical particles $N$, global volume $V$, and a temperature $T$. Nevertheless,
the global system, in turn, consists of a set of $K$ small cells with a constant
volume $\delta V_M$, whose local homogeneous distribution functions is
$f_{M}(\vec{v},t)$.

The distribution function will depend on the position of the cells, the velocity
$\vec{v}$, and time $t$. Without loss of generality, we choose these cells to
have local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$).
Each cell has its local chemical potential, $\mu_M$, local
$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local
energy, $\mathcal{E}'_M$. In terms of the local distribution 
function and the energy $\epsilon(\vec{v})$, the last three properties are
given by:
%
\begin{eqnarray}
    H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
      \mathrm{d}^{3}v \label{Hcell},\nonumber\\
    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber\\
    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
\end{eqnarray}
%

When the system is in equilibrium, the local number of particles and the local
energy do not depend on the cell number, this is
%
\begin{equation}
   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}';
   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
\end{equation}
%
The particles in the system are considered to be free, \textit{i.e.} the
particles do not interact with each other. Besides, the local distribution
function must satisfy the following microcanonical restrictions 
%
\begin{equation}\label{micro}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N,
    \quad\sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
\end{equation}
%

We can see that the functional (\ref{eq:CH2}) contains a non-homogeneous
distribution function but keeps the same form of the Boltzmann's $H$ functional.

Inspired by the Hamilton principle, $H'$ will be maximized with those
restrictions to obtain the equilibrium distribution function for a classical gas
using the Lagrange multipliers method obtaining
%
\begin{eqnarray}
    \frac{\delta H'}{\delta f_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
      \frac{\delta}{\delta f_J(\vec{v}')}\left[
        f_M(\vec{v})\ln f_M(\vec{v})
        \right]
       \mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int
       	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
      \mathrm{d}^3v\nonumber\\
    & & -\sum_{M=1}^{K}\beta_M\int\epsilon(\vec{v})
    	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
      \mathrm{d}^3v \nonumber\\
    & = & \ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
\end{eqnarray}
%
Isolating the distribution function 
%
\begin{eqnarray}
    \ln f_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
       -1\quad\Rightarrow\quad f_J(\vec{v}')\\
    & = & e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1}\nonumber\\
    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
\end{eqnarray}
%
where $C$ is a constant. It is trivial that using the variational procedures we
obtain the form of the Maxwell-Boltzmann distribution.

As $H'$ in equilibrium is proportional to entropy, it is trivial to think that
Lagrange multipliers do not depend on the position because of the system is in
equilibrium and consequently, the distribution function is homogeneous, that is
%
\begin{equation}
    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
\end{equation} 
%
The constant $C$ can be omitted defining the following $H$ functional
%
\begin{equation}
   H''(t)=\sum_{M=1}^{K}\int \left[f_M(\vec{v},t)
    \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
\end{equation}
%
but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the
total particle number. As $N$ is a constant and any $H$ functional has sense
when we calculate its time derivative, this constant can be omitted.

On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{eq:CH2}) as
%
\begin{equation}
    H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v
         = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
\end{equation}
%
where
%
\begin{equation}
    H_{boltz}(t)=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
\end{equation}
%
Besides, if we analyze the following expression
%
\begin{equation}
    \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
\end{equation}
%
we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
%
\begin{equation}
    \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
\end{equation}
%
We obtained the additive property when we supposed the distribution function is
spatial-homogeneous.

To prove the Boltzmann's $H$ theorem (using (\ref{eq:CH2}) as functional), we need
to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the
following assumptions:
%
\begin{itemize}
  \item The local equilibrium hypothesis in each cell ($f_{M}$ must be
    spatial-homogeneous for each $M$).
  \item The non-homogeneous distribution function assumption in the total
    volume ($f_{M}$ must be different among all cells).
  \item The following restrictions
    %
    \begin{equation}\label{restrictionoutsideclassical}
      \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=
      \bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ 
      \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=
      \bar{\mathcal{E}}+ \delta_M(t),
    \end{equation}
    %
    where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are
    the local particle number and the local energy in equilibrium
    %
    \begin{equation}
      \bar{\mathcal{N}}=
      \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \quad
      \bar{\mathcal{E}}=
      \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
    \end{equation}
    %
    and $\Delta_M$, $\delta_M$ could be seen as a deviation from
    $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively,
    with $\Delta_M(t)\ll\bar{\mathcal{N}}$ and
    $\delta_M(t)\ll\bar{\mathcal{E}}$.
\end{itemize} 
%
Those conditions described a system out of equilibrium but no so far from it,
that is a perturbed system and will suffer a \textit{relaxation process, i.e.}
The time evolution of a perturbated system to the equilibrium state.

It is important to remark that the quantities $\Delta_M$, $\delta_M$ are
sufficiently big to be different from fluctuations in the system but too small
such that the system is not far from the equilibrium state.

Performing the derivative of (\ref{eq:CH2}) with respect to time, we obtain
%
\begin{equation}\label{dH1}
    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\left[
      1+\ln f_M(\vec{v},t)
    \right]\dot f_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%
Using the first-order approximation
%
\begin{equation}\label{firstorder}
   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ 
   \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|,
\end{equation}
%
(\ref{dH1}) yields
%
\begin{equation}\label{dH1-1}
    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\bar f(\vec{v}) \left[
      1+\ln \left\{
        \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t)
      \right\}
    \right]\dot g_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%

We can approximate the logarithm function in (\ref{dH1-1}) to its first-order
Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
%
\begin{equation}\label{lnapproximationclassical}
    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx
    \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t).
\end{equation}
%
With the previous approximation, (\ref{dH1-1}) obtains the following form 
%
\begin{eqnarray}
    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int \bar f(\vec{v})\left[
      1+\ln \bar f(\vec{v})+g_M(\vec{v},t)
    \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
\end{eqnarray}
%
using (\ref{relacion1}) with homogeneous multipliers
%
\begin{equation}\label{dH1-2}
    \frac{dH'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})\left[
      \alpha+\beta \epsilon(\vec{v})
    \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}
    \int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v . \nonumber 
\end{equation}
%
On the other hand, we obtain from the restrictions the following expressions
%
\begin{eqnarray}
    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow&
    \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow&
    \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber 
\end{eqnarray}
%
and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) =\sum_{M=1}^{K} \delta_M(t) =0$,
we find
%
\begin{equation}
    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
\end{equation}
%
Then using the previous result to (\ref{dH1-2}), we find
%
\begin{equation}
  \frac{dH'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%
Now, the summation over $M$ will be expressed in two summations
%
\begin{equation}\label{classicalH3}
    \frac{dH'}{dt}=\sum_J^{L}\int
      \bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v
      +\sum_J^{P}\int
        \bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%
where $L+P=K$. The group of $L$ cells is such that cells have an excess of
particles or/and energy (from the mean value in equilibrium). In contrast,
the group of $P$ cells is such that cells have to miss particles or/and energy
(from the mean value in equilibrium). Besides, $\dot{g}^{+}_{J}$ represents the
change on the deviation on cells that have an excess of particles or energy
while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that
have missing particles or energy. 
On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells
that have an excess of particles or energy. In contrast, $g^{-}_{J}$ represents
the value of the deviation on cells that have missing particles or energy.

Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles
and/or energy and so, $g^{+}_{J}>0$. On the other hand, $\dot{g}^{-}_{J}>0$
describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.

We sort the previous ideas in the following form
%
\begin{equation}\label{separacionclassical}
\begin{array}{rl}
  g^{+}_{J}=+|g^{+}_{J}|; & \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|,\\
  g^{-}_{J}=-|g^{-}_{J}|; & \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}|,
 \end{array}
\end{equation}
%
and consequently, (\ref{classicalH3}) obtains the following form
%
\begin{equation}
    \frac{dH'}{dt}=-\left[
      \sum_J^{L}\int\bar f(\vec{v})|g_J^{+}(\vec{v},t)|
        |\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v
      +\sum_J^{P}\int\bar f(\vec{v})|g_J^{-}(\vec{v},t)|
      |\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v 
    \right]. \label{classicalH4}
\end{equation}
%
We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always
positive. Also the deviation and its derivative are positive, then all the
expression is positive. However, the global sign makes the derivative of the
$H'$ with respect to time is always less to zero. With this we proved that
$\frac{dH'}{dt}<0$. If the system is in equilibrium,
$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.

Joining both results, we can say the following statement:
Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal
volume elements), total energy $E$ and total number of free particles $N$.
Consider also the system has inhomogeneities and suffers a relaxation process.
If we define the following functional
%
\begin{equation}\label{CH3}
   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v,
\end{equation}
%
where $f_M(\vec{v},t)$ is the local distribution function of each cell in the
system, then in the first-order approximation
%
\begin{equation}
    \frac{dH'}{dt} \leq 0.
\end{equation}
%
This statement will be the classical $H$ theorem with inhomogeneities.

In the next section, we present a proposal quantum version of the
$H$ theorem defined by Tolman and our proposal quantum version of the $H$
theorem applying the method of the volume divided into cells.

### 2020-12-15 tex/conclusions.tex
Sources reformatted to match the template of Entropy journal.

--- 2020-12-15 notasparaarticulo.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{xcolor}
-\usepackage{calrsfs}
-\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
-\newcommand{\Sm}{\Ss_M}
-\newcommand{\de}{\delta}
-\newcommand{\ep}{\epsilon}
-\newcommand{\Ss}{\mathcal{S}}
-
-\title{Quantum H theorem, variational entropy, and relaxation processes}
-\author{labfqot}
-\date{March 2020}
-
-\begin{document}
-
-\maketitle
-
-\section*{Abstract}
-Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
-This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
-The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
-The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
-In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
-Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
-
-\section{Introduction}
-
-In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
-In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
-\\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
-One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
-The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
-However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
-Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
-\\
-In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
-In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
-Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
-The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
-The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
-Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
-
-%----------------------------------
-\section{Classical scheme}
-\subsection{The Boltzmann transport equation}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
-On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
-However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
-The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
-%-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. 
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
-The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
-As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
-\begin{equation}
-    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
-\end{equation}
-This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
-The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
-\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
-For more information, see \cite{huang}.}}
-%---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
-The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
-However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
-Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
-The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
-\begin{eqnarray}
-    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Besides, the local distribution function must satisfy the following microcanonical restrictions
-\begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
-\end{equation}
-We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
-\begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
-\end{eqnarray}
-Isolating the distribution function 
-\begin{eqnarray}
-    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
-\end{eqnarray}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}
-  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
-  \\
-To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
-\begin{itemize}
-    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
-    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
-    \item The following restrictions
-    \begin{eqnarray}
-        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\end{itemize} 
- Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
-\\
-Performing the derivative of (\ref{CH2}) with respect to time, we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
-\end{equation}{}
-Using the first-order approximation
-\begin{eqnarray}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
-\end{eqnarray}
-(\ref{dH1}) yields
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}
-We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
-\begin{equation}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
-\end{equation}
-With the previous approximation, (\ref{dH1,1}) obtains the following form 
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
-\nonumber \\
-\end{eqnarray}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K}
-\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}
-Then using the previous result to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{eqnarray}
-Now, the summation over $M$ will be expressed in two summations
-\begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
-\label{classicalH3}
-\end{equation}
-where $L+P=K$.
-The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
-On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
-   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}
-and consequently, (\ref{classicalH3}) obtains the following form
-\begin{equation}
-    \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
-    \right]. \label{classicalH4}
-\end{equation}
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
-Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
-With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-Joining both results, we can say the following statement:\\
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
-\end{equation}
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
-\begin{equation}
-    \frac{dH'}{dt} \leq 0.
-\end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.\\
-\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
-
-%------------------------------
-
-\section{Quantum scheme}
-\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
-\end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
-\begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
-\end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
-\begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
-\end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
-\begin{equation}
-    \frac{dH}{dt}\leq 0,
-\end{equation}
-and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version.
-Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
-Those distributions can be obtained from a sufficient condition for the equilibrium state
-\begin{equation}
-    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
-\end{equation}
-where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
-%--------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
-
-We begin by defining the form of the functional 
-\begin{eqnarray}
-    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
-\end{eqnarray}
-This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
-Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
-.
-.
-, K$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
-\begin{eqnarray}
-    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-Besides, the system is subjected to the microcanonical restrictions
-\begin{eqnarray}
-    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
-\end{eqnarray}
-where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
-\begin{eqnarray}
-&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
-\end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
-\begin{equation}
-    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
-\end{equation}
-where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
-With this, we conclude
-\begin{equation}
-    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
-\end{equation}
-In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
-\begin{equation}
-    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
-\begin{equation}
-    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}{}
-and finally
-\begin{eqnarray}
-    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
-\begin{eqnarray}
-      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
-  \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
- %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
-We rewrite the \textit{variational entropy} (\ref{entropy}) as
-\begin{equation}
-    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
-\end{equation}
-where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
-\begin{eqnarray}
-        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the quantum analogous case of the classical case.
-Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
-\\
-If we perform the derivative of the variational entropy with respect to time, we obtain
-\begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}{}
-Using the first-order approximation (\ref{firstorder}), we get
-\begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}{}
-It's necessary to remark that, in the case of fermions
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
-This result corresponds to the system in the zero temperature condition.
-We can observe that in this condition, (\ref{fermionrestriction}) is violated.
-Due to this, the zero temperature condition is excluded.\\
-\\
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}{}
-With the previous approximation, (\ref{cambioH1}) will be
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}{}
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}{}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}{}
-Substituting the previous expression to (\ref{cambioH3}) we obtain
-\begin{eqnarray}
-   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over $M$ will be expressed in two summations, such as in the classical case
-\begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}
-where $L+P=K$.
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
-and consequently, (\ref{cambioH5}) obtains the following form
-\begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.
-
-  %-----------------------------------------
-
-%---------------------------------------------
-
-\section{Quantum Classical correspondence}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
-\end{eqnarray}
-where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
-\textcolor{blue}{In the equilibrium case, we have the following variational entropy
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
-                    \left(
-                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                  \right) \ln
-                \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
-                \right) \right]. \label{equilibriumvariational}
-\end{eqnarray}
-To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
-\begin{equation}
-    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
-\end{equation}
-and then
-\begin{equation}
-    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
-\end{equation}
-If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}
-                  \right) \ln
-                \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
-                \right) \right],
-\end{eqnarray}
-but if (\ref{classicalapproximation}) holds, then 
-\begin{equation}
-     e^{\alpha_M+\beta_M \epsilon}\ll 1,
-\end{equation}
-consequently 
-\begin{equation}
-    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
-\end{equation}
-and finally
-\begin{equation}
-    \Ss=\sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right],
-\end{equation}
-that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
-\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
-
-%--------------------------------------
-
-  \section{Relaxation processes in degenerated quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
-\begin{equation}
-    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
-\end{equation}
-where $\mathcal{N}_{M}^{n}$ is defined as
-\begin{equation}
-    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
-\end{equation}
-Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
-This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
-Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
-Then the equation of motion is
-\begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\label{transportequation1}
-\end{equation}
-This equation is valid in equilibrium.
-When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be included in (\ref{transportequation1}).
-This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
-\begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
-    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\nonumber \\ \label{transportequation2}
-\end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
-
-  %----------------------------------------
-
-%-------------------------------------------------
-
-\section{Discussions and remarks}
-
-In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
-In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
-It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$ theorem.\\
-\\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be future work.\\
-\\
-We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
-This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
-\begin{equation}
-    H_{Boltz}=-\frac{S}{Vk},
-\end{equation}
-where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
-\begin{equation}
-    \Ss\propto S.
-\end{equation}
-\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
-It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
-This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
-Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
-Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
-In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
-
-%-----------------------------------------------
-
 \section{Conclusions}
@@
 \clearpage
-
-\begin{thebibliography}{}
-%\bibitem{goldstein}
-%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
-
-%\bibitem{sakurai}
-%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
-
-\bibitem{huang}
-K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).
-
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
-
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
-
-\bibitem{cristal1}
-Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
-
-\bibitem{cristal2}
-Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
-
-\bibitem{cristal3}
-Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
-
-\bibitem{kei}
-Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
-
-\bibitem{onsager}
-Onsager, L. (1931), Physical Review 37 (4): 405-426.
-
-\bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
-
-\bibitem{paradox2}
-Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.
-
-\bibitem{hamiltonian}
-Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
-
-\bibitem{H-theorem-violation}
-Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
-
-\bibitem{H-theorem-and-entropy}
-Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
-
-\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
-
-\bibitem{entropic-framework}
-Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
-
-\bibitem{thermal-harmonic}
-F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
-Phys. Rev. E 91, 042116 – 15 April 2015.
-
-\bibitem{quantum-transport}
-Robert Hussein and Sigmund Kohler,
-Phys. Rev. B 89, 205424, 21 May 2014
-
-\bibitem{quantum-transport-reservoirs}
-Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
-Phys. Rev. A 102, 022207, 11 August 2020.
-
-\bibitem{htheorem2}
-R. Silva {\it et al} EPL 89 10004 (2010).
-
-\bibitem{quantum1}
-J. Math. Phys. 47, 073303 (2006).
-
-\bibitem{quantum2}
-Z. Physik 268, 139-143 (1974).
-
-\bibitem{quantum-entropy}
-J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
-
-\bibitem{quantum-colapse}
-Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
-
-\bibitem{quantum3}
-Gring, M.
-and Kuhnert, M.
-and Langen, T.
-and Kitagawa, T.
-and Rauer, B.
-and Schreitl, M.
-and Mazets, I.
-and Smith, D.
-Adu and Demler, E.
-and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
-
-\bibitem{quantum4}
-Physical Review E 91, 062106 (2015).
-
-%\bibitem{relaxationphenomena}
-%Journal of Statistical Physics, 30; 2 (1983).
-
-%\bibitem{htheorem}
-%Review of Modern Physics; 74 (2002).
-
-\bibitem{binary}
-Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.
-
-\bibitem{contradictions}
-Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.
-
-\bibitem{ergodic-theorem}
-Eur. Phys. J. H 35, 201–237 (2010).
-
-\bibitem{hydrodynamic-simulation}
-Review of Modern Physics 74(4):1203-1220.
-
-\end{thebibliography}{}
-
-\end{document}

### 2020-12-15 tex/disscussion.tex
Sources reformatted to match the template of Entropy journal.

--- 2020-12-15 notasparaarticulo.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{xcolor}
-\usepackage{calrsfs}
-\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
-\newcommand{\Sm}{\Ss_M}
-\newcommand{\de}{\delta}
-\newcommand{\ep}{\epsilon}
-\newcommand{\Ss}{\mathcal{S}}
-
-\title{Quantum H theorem, variational entropy, and relaxation processes}
-\author{labfqot}
-\date{March 2020}
-
-\begin{document}
-
-\maketitle
-
-\section*{Abstract}
-Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
-This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
-The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
-The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
-In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
-Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
-
-\section{Introduction}
-
-In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
-In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
-\\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
-One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
-The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
-However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
-Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
-\\
-In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
-In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
-Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
-The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
-The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
-Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
-
-%----------------------------------
-\section{Classical scheme}
-\subsection{The Boltzmann transport equation}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
-On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
-However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
-The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
-%-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. 
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
-The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
-As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
-\begin{equation}
-    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
-\end{equation}
-This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
-The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
-\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
-For more information, see \cite{huang}.}}
-%---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
-The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
-However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
-Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
-The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
-\begin{eqnarray}
-    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Besides, the local distribution function must satisfy the following microcanonical restrictions
-\begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
-\end{equation}
-We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
-\begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
-\end{eqnarray}
-Isolating the distribution function 
-\begin{eqnarray}
-    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
-\end{eqnarray}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}
-  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
-  \\
-To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
-\begin{itemize}
-    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
-    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
-    \item The following restrictions
-    \begin{eqnarray}
-        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\end{itemize} 
- Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
-\\
-Performing the derivative of (\ref{CH2}) with respect to time, we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
-\end{equation}{}
-Using the first-order approximation
-\begin{eqnarray}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
-\end{eqnarray}
-(\ref{dH1}) yields
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}
-We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
-\begin{equation}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
-\end{equation}
-With the previous approximation, (\ref{dH1,1}) obtains the following form 
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
-\nonumber \\
-\end{eqnarray}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K}
-\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}
-Then using the previous result to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{eqnarray}
-Now, the summation over $M$ will be expressed in two summations
-\begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
-\label{classicalH3}
-\end{equation}
-where $L+P=K$.
-The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
-On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
-   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}
-and consequently, (\ref{classicalH3}) obtains the following form
-\begin{equation}
-    \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
-    \right]. \label{classicalH4}
-\end{equation}
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
-Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
-With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-Joining both results, we can say the following statement:\\
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
-\end{equation}
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
-\begin{equation}
-    \frac{dH'}{dt} \leq 0.
-\end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.\\
-\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
-
-%------------------------------
-
-\section{Quantum scheme}
-\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
-\end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
-\begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
-\end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
-\begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
-\end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
-\begin{equation}
-    \frac{dH}{dt}\leq 0,
-\end{equation}
-and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version.
-Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
-Those distributions can be obtained from a sufficient condition for the equilibrium state
-\begin{equation}
-    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
-\end{equation}
-where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
-%--------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
-
-We begin by defining the form of the functional 
-\begin{eqnarray}
-    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
-\end{eqnarray}
-This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
-Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
-.
-.
-, K$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
-\begin{eqnarray}
-    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-Besides, the system is subjected to the microcanonical restrictions
-\begin{eqnarray}
-    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
-\end{eqnarray}
-where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
-\begin{eqnarray}
-&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
-\end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
-\begin{equation}
-    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
-\end{equation}
-where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
-With this, we conclude
-\begin{equation}
-    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
-\end{equation}
-In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
-\begin{equation}
-    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
-\begin{equation}
-    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}{}
-and finally
-\begin{eqnarray}
-    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
-\begin{eqnarray}
-      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
-  \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
- %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
-We rewrite the \textit{variational entropy} (\ref{entropy}) as
-\begin{equation}
-    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
-\end{equation}
-where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
-\begin{eqnarray}
-        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the quantum analogous case of the classical case.
-Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
-\\
-If we perform the derivative of the variational entropy with respect to time, we obtain
-\begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}{}
-Using the first-order approximation (\ref{firstorder}), we get
-\begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}{}
-It's necessary to remark that, in the case of fermions
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
-This result corresponds to the system in the zero temperature condition.
-We can observe that in this condition, (\ref{fermionrestriction}) is violated.
-Due to this, the zero temperature condition is excluded.\\
-\\
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}{}
-With the previous approximation, (\ref{cambioH1}) will be
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}{}
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}{}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}{}
-Substituting the previous expression to (\ref{cambioH3}) we obtain
-\begin{eqnarray}
-   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over $M$ will be expressed in two summations, such as in the classical case
-\begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}
-where $L+P=K$.
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
-and consequently, (\ref{cambioH5}) obtains the following form
-\begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.
-
-  %-----------------------------------------
-
-%---------------------------------------------
-
-\section{Quantum Classical correspondence}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
-\end{eqnarray}
-where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
-\textcolor{blue}{In the equilibrium case, we have the following variational entropy
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
-                    \left(
-                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                  \right) \ln
-                \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
-                \right) \right]. \label{equilibriumvariational}
-\end{eqnarray}
-To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
-\begin{equation}
-    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
-\end{equation}
-and then
-\begin{equation}
-    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
-\end{equation}
-If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}
-                  \right) \ln
-                \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
-                \right) \right],
-\end{eqnarray}
-but if (\ref{classicalapproximation}) holds, then 
-\begin{equation}
-     e^{\alpha_M+\beta_M \epsilon}\ll 1,
-\end{equation}
-consequently 
-\begin{equation}
-    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
-\end{equation}
-and finally
-\begin{equation}
-    \Ss=\sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right],
-\end{equation}
-that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
-\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
-
-%--------------------------------------
-
-  \section{Relaxation processes in degenerated quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
-\begin{equation}
-    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
-\end{equation}
-where $\mathcal{N}_{M}^{n}$ is defined as
-\begin{equation}
-    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
-\end{equation}
-Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
-This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
-Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
-Then the equation of motion is
-\begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\label{transportequation1}
-\end{equation}
-This equation is valid in equilibrium.
-When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be included in (\ref{transportequation1}).
-This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
-\begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
-    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\nonumber \\ \label{transportequation2}
-\end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
-
-  %----------------------------------------
-
-%-------------------------------------------------
-
 \section{Discussions and remarks}
@@
 %-----------------------------------------------
-
-\section{Conclusions}
-We conclude the following statements:\\
-\begin{enumerate}
-    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
-    \item We can obtain the transport equation from the $H$ functional.
-    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
-    \item We find an expression for the entropy to quantum gases.
-    \end{enumerate}
-
-%  \begin{equation}
-%      \mathcal{H}
-%  \end{equation}
-
-\clearpage
-
-\begin{thebibliography}{}
-%\bibitem{goldstein}
-%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
-
-%\bibitem{sakurai}
-%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
-
-\bibitem{huang}
-K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).
-
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
-
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
-
-\bibitem{cristal1}
-Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
-
-\bibitem{cristal2}
-Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
-
-\bibitem{cristal3}
-Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
-
-\bibitem{kei}
-Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
-
-\bibitem{onsager}
-Onsager, L. (1931), Physical Review 37 (4): 405-426.
-
-\bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
-
-\bibitem{paradox2}
-Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.
-
-\bibitem{hamiltonian}
-Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
-
-\bibitem{H-theorem-violation}
-Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
-
-\bibitem{H-theorem-and-entropy}
-Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
-
-\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
-
-\bibitem{entropic-framework}
-Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
-
-\bibitem{thermal-harmonic}
-F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
-Phys. Rev. E 91, 042116 – 15 April 2015.
-
-\bibitem{quantum-transport}
-Robert Hussein and Sigmund Kohler,
-Phys. Rev. B 89, 205424, 21 May 2014
-
-\bibitem{quantum-transport-reservoirs}
-Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
-Phys. Rev. A 102, 022207, 11 August 2020.
-
-\bibitem{htheorem2}
-R. Silva {\it et al} EPL 89 10004 (2010).
-
-\bibitem{quantum1}
-J. Math. Phys. 47, 073303 (2006).
-
-\bibitem{quantum2}
-Z. Physik 268, 139-143 (1974).
-
-\bibitem{quantum-entropy}
-J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
-
-\bibitem{quantum-colapse}
-Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
-
-\bibitem{quantum3}
-Gring, M.
-and Kuhnert, M.
-and Langen, T.
-and Kitagawa, T.
-and Rauer, B.
-and Schreitl, M.
-and Mazets, I.
-and Smith, D.
-Adu and Demler, E.
-and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
-
-\bibitem{quantum4}
-Physical Review E 91, 062106 (2015).
-
-%\bibitem{relaxationphenomena}
-%Journal of Statistical Physics, 30; 2 (1983).
-
-%\bibitem{htheorem}
-%Review of Modern Physics; 74 (2002).
-
-\bibitem{binary}
-Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.
-
-\bibitem{contradictions}
-Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.
-
-\bibitem{ergodic-theorem}
-Eur. Phys. J. H 35, 201–237 (2010).
-
-\bibitem{hydrodynamic-simulation}
-Review of Modern Physics 74(4):1203-1220.
-
-\end{thebibliography}{}
-
-\end{document}

### 2020-12-15 tex/introduction.tex
Sources reformatted to match the template of Entropy journal.

--- 2020-12-15 notasparaarticulo.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{xcolor}
-\usepackage{calrsfs}
-\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
-\newcommand{\Sm}{\Ss_M}
-\newcommand{\de}{\delta}
-\newcommand{\ep}{\epsilon}
-\newcommand{\Ss}{\mathcal{S}}
-
-\title{Quantum H theorem, variational entropy, and relaxation processes}
-\author{labfqot}
-\date{March 2020}
-
-\begin{document}
-
-\maketitle
-
-\section*{Abstract}
-Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
-This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
-The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
-The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
-In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
-Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
-
 \section{Introduction}
 
-In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
-In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
-\\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
-One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
-The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
-However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
-Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
-\\
-In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
-In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
-Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
-The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
-The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
-Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
+In the literature, the theoretical basis that allows us to describe equilibrium
+systems in the classical scheme is well-established.
+The procedures to describe
+equilibrium systems are well-known and work in almost every system in nature
+(Thermodynamics for a phenomenological description and statistical mechanics
+for a microscopic description). In contrast, a unified mathematical model to
+describe the evolution of any system to an equilibrium state has not developed.
+Nevertheless, developments can be found in the literature and describe the
+evolution to equilibrium to almost any system (the internal behavior of glasses
+\cite{bib:zanotto2018} and the entropy measurement
+\cite{bib:schmelzer2018,bib:nemilov2018} are some examples where the recent
+developments do not work.) Examples of those developments can be the Onsager
+formulation (called linear thermodynamics)
+\cite{bib:keizer1987,bib:onsager1931}, and the Kinetic Theory of dilute gases
+developed by Boltzmann.
 
-%----------------------------------
-\section{Classical scheme}
-\subsection{The Boltzmann transport equation}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
-On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
-However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
-The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
-%-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. 
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
-The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
-As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
-\begin{equation}
-    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
-\end{equation}
-This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
-The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
-\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
-For more information, see \cite{huang}.}}
-%---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
-The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
-However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
-Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
-The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
-\begin{eqnarray}
-    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Besides, the local distribution function must satisfy the following microcanonical restrictions
-\begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
-\end{equation}
-We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
-\begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
-\end{eqnarray}
-Isolating the distribution function 
-\begin{eqnarray}
-    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
-\end{eqnarray}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}
-  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
-  \\
-To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
-\begin{itemize}
-    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
-    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
-    \item The following restrictions
-    \begin{eqnarray}
-        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\end{itemize} 
- Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
-\\
-Performing the derivative of (\ref{CH2}) with respect to time, we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
-\end{equation}{}
-Using the first-order approximation
-\begin{eqnarray}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
-\end{eqnarray}
-(\ref{dH1}) yields
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}
-We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
-\begin{equation}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
-\end{equation}
-With the previous approximation, (\ref{dH1,1}) obtains the following form 
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
-\nonumber \\
-\end{eqnarray}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K}
-\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}
-Then using the previous result to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{eqnarray}
-Now, the summation over $M$ will be expressed in two summations
-\begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
-\label{classicalH3}
-\end{equation}
-where $L+P=K$.
-The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
-On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
-   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}
-and consequently, (\ref{classicalH3}) obtains the following form
-\begin{equation}
-    \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
-    \right]. \label{classicalH4}
-\end{equation}
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
-Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
-With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-Joining both results, we can say the following statement:\\
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
-\end{equation}
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
-\begin{equation}
-    \frac{dH'}{dt} \leq 0.
-\end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.\\
-\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
+In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas
+was described by Boltzmann through the Boltzmann transport equation.
+Furthermore, the evolution of an equilibrium state for any system was proved by
+means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the
+spatial-homogeneous distribution function hypothesis.
+However, there are some aspects in the classical $H$ theorem and the
+Boltzmann's $H$ functional that deserve some review to accomplish better
+consistency, for example, the modification of the $H$ theorem in stochastic
+trajectories, violations in the second law of thermodynamics, and the relation
+between Shannon's measure of information and the Boltzmann's entropy %[6-10].
+\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.
 
-%------------------------------
+To obtain the well-known correspondence principle between classical mechanics
+and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or
+quantum $H$ theorem) and the quantum analogous to the Boltzmann transport
+equation is necessary such that in the appropriate limit, the  classical $H$
+theorem and the Boltzmann transport equation should be recovered.
 
-\section{Quantum scheme}
-\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
-\end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
-\begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
-\end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
-\begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
-\end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
-\begin{equation}
-    \frac{dH}{dt}\leq 0,
-\end{equation}
-and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version.
-Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
-Those distributions can be obtained from a sufficient condition for the equilibrium state
-\begin{equation}
-    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
-\end{equation}
-where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
-%--------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
+One of the first developments in this field was developed by Tolman
+\cite{bib:tolman}. He proposes one of the first quantum versions of the
+Boltzmann's $H$ theorem using the probability transition relation, the random
+phases hypothesis, and a homogeneous distribution function. The equation of
+motion for the occupation number (necessary to prove the quantum $H$ theorem in
+the same way that Boltzmann's procedure) was obtained by applying the time
+perturbation theory. However, to obtain the flux relations in analogy to the
+Boltzmann's method, a quantum version of the transport equation that includes a
+non-homogeneous distribution function is required.
 
-We begin by defining the form of the functional 
-\begin{eqnarray}
-    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
-\end{eqnarray}
-This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
-Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
-.
-.
-, K$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
-\begin{eqnarray}
-    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-Besides, the system is subjected to the microcanonical restrictions
-\begin{eqnarray}
-    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
-\end{eqnarray}
-where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
-\begin{eqnarray}
-&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
-\end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
-\begin{equation}
-    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
-\end{equation}
-where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
-With this, we conclude
-\begin{equation}
-    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
-\end{equation}
-In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
-\begin{equation}
-    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
-\begin{equation}
-    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}{}
-and finally
-\begin{eqnarray}
-    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
-\begin{eqnarray}
-      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
-  \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
- %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
-We rewrite the \textit{variational entropy} (\ref{entropy}) as
-\begin{equation}
-    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
-\end{equation}
-where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
-\begin{eqnarray}
-        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the quantum analogous case of the classical case.
-Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
-\\
-If we perform the derivative of the variational entropy with respect to time, we obtain
-\begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}{}
-Using the first-order approximation (\ref{firstorder}), we get
-\begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}{}
-It's necessary to remark that, in the case of fermions
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
-This result corresponds to the system in the zero temperature condition.
-We can observe that in this condition, (\ref{fermionrestriction}) is violated.
-Due to this, the zero temperature condition is excluded.\\
-\\
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}{}
-With the previous approximation, (\ref{cambioH1}) will be
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}{}
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}{}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}{}
-Substituting the previous expression to (\ref{cambioH3}) we obtain
-\begin{eqnarray}
-   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over $M$ will be expressed in two summations, such as in the classical case
-\begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}
-where $L+P=K$.
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
-and consequently, (\ref{cambioH5}) obtains the following form
-\begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.
+On the other hand, some works in the quantum operator formalism attempt to
+describe quantum transport phenomena using the Hamiltonian of the system and
+the master equation (analogous to the Boltzmann transport equation)
+\cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. %[12-15]
+Besides, some authors construct the $H$ functional and develop the proof of the
+quantum $H$ theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
+However, the homogeneous distribution function hypothesis and the
+correspondence principle are uncleared or not discussed. Moreover, the validity
+of the quantum $H$ theorem and the second law of thermodynamics, and the
+interpretation of the quantum entropy is still in discussion %[19-26].
+\cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:acharya2019,%
+bib:kastner2017,bib:gring2012,bib:han2015,bib:das2018}.
 
-  %-----------------------------------------
+In order to contribute to the construction of a consistent classical and
+quantum $H$ theorem, and a new formalism to describe out of equilibrium
+systems, we develop a new theoretical basis to describe them. In particular, we
+are interested in including a non-homogeneous distribution function in the
+mathematical model, and so including non-homogeneous systems in the $H$
+theorem. Besides, we are interested in developing the quantum analogous of
+Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution
+function and solve the problem of the correspondence principle mentioned above.
 
-%---------------------------------------------
-
-\section{Quantum Classical correspondence}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
-\end{eqnarray}
-where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
-\textcolor{blue}{In the equilibrium case, we have the following variational entropy
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
-                    \left(
-                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                  \right) \ln
-                \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
-                \right) \right]. \label{equilibriumvariational}
-\end{eqnarray}
-To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
-\begin{equation}
-    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
-\end{equation}
-and then
-\begin{equation}
-    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
-\end{equation}
-If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}
-                  \right) \ln
-                \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
-                \right) \right],
-\end{eqnarray}
-but if (\ref{classicalapproximation}) holds, then 
-\begin{equation}
-     e^{\alpha_M+\beta_M \epsilon}\ll 1,
-\end{equation}
-consequently 
-\begin{equation}
-    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
-\end{equation}
-and finally
-\begin{equation}
-    \Ss=\sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right],
-\end{equation}
-that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
-\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
-
-%--------------------------------------
-
-  \section{Relaxation processes in degenerated quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
-\begin{equation}
-    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
-\end{equation}
-where $\mathcal{N}_{M}^{n}$ is defined as
-\begin{equation}
-    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
-\end{equation}
-Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
-This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
-Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
-Then the equation of motion is
-\begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\label{transportequation1}
-\end{equation}
-This equation is valid in equilibrium.
-When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be included in (\ref{transportequation1}).
-This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
-\begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
-    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\nonumber \\ \label{transportequation2}
-\end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
-
-  %----------------------------------------
-
-%-------------------------------------------------
-
-\section{Discussions and remarks}
-
-In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
-In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
-It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$ theorem.\\
-\\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be future work.\\
-\\
-We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
-This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
-\begin{equation}
-    H_{Boltz}=-\frac{S}{Vk},
-\end{equation}
-where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
-\begin{equation}
-    \Ss\propto S.
-\end{equation}
-\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
-It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
-This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
-Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
-Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
-In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
-
-%-----------------------------------------------
-
-\section{Conclusions}
-We conclude the following statements:\\
-\begin{enumerate}
-    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
-    \item We can obtain the transport equation from the $H$ functional.
-    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
-    \item We find an expression for the entropy to quantum gases.
-    \end{enumerate}
-
-%  \begin{equation}
-%      \mathcal{H}
-%  \end{equation}
-
-\clearpage
-
-\begin{thebibliography}{}
-%\bibitem{goldstein}
-%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
-
-%\bibitem{sakurai}
-%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
-
-\bibitem{huang}
-K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).
-
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
-
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
-
-\bibitem{cristal1}
-Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
-
-\bibitem{cristal2}
-Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
-
-\bibitem{cristal3}
-Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
-
-\bibitem{kei}
-Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
-
-\bibitem{onsager}
-Onsager, L. (1931), Physical Review 37 (4): 405-426.
-
-\bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
-
-\bibitem{paradox2}
-Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.
-
-\bibitem{hamiltonian}
-Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
-
-\bibitem{H-theorem-violation}
-Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
-
-\bibitem{H-theorem-and-entropy}
-Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
-
-\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
-
-\bibitem{entropic-framework}
-Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
-
-\bibitem{thermal-harmonic}
-F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
-Phys. Rev. E 91, 042116 – 15 April 2015.
-
-\bibitem{quantum-transport}
-Robert Hussein and Sigmund Kohler,
-Phys. Rev. B 89, 205424, 21 May 2014
-
-\bibitem{quantum-transport-reservoirs}
-Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
-Phys. Rev. A 102, 022207, 11 August 2020.
-
-\bibitem{htheorem2}
-R. Silva {\it et al} EPL 89 10004 (2010).
-
-\bibitem{quantum1}
-J. Math. Phys. 47, 073303 (2006).
-
-\bibitem{quantum2}
-Z. Physik 268, 139-143 (1974).
-
-\bibitem{quantum-entropy}
-J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
-
-\bibitem{quantum-colapse}
-Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
-
-\bibitem{quantum3}
-Gring, M.
-and Kuhnert, M.
-and Langen, T.
-and Kitagawa, T.
-and Rauer, B.
-and Schreitl, M.
-and Mazets, I.
-and Smith, D.
-Adu and Demler, E.
-and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
-
-\bibitem{quantum4}
-Physical Review E 91, 062106 (2015).
-
-%\bibitem{relaxationphenomena}
-%Journal of Statistical Physics, 30; 2 (1983).
-
-%\bibitem{htheorem}
-%Review of Modern Physics; 74 (2002).
-
-\bibitem{binary}
-Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.
-
-\bibitem{contradictions}
-Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.
-
-\bibitem{ergodic-theorem}
-Eur. Phys. J. H 35, 201–237 (2010).
-
-\bibitem{hydrodynamic-simulation}
-Review of Modern Physics 74(4):1203-1220.
-
-\end{thebibliography}{}
-
-\end{document}
+The first section of this work consists of a review of the
+Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann
+distribution, and our alternative form to prove the classical $H$ theorem. The
+second part consists of a review of the proposal from Tolman about the quantum
+version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions,
+and our alternative form to prove the quantum version of the $H$ theorem. The
+third section consists of a classical-quantum correspondence between the
+quantum $H$ functional and the classical $H$ functional. Finally, the last
+section consists of a proposal to explore relaxation processes in the quantum
+case obtaining the analogous quantum Boltzmann transport equation.

### 2020-12-15 tex/qccorrespondence.tex
Sources reformatted to match the template of Entropy journal.

--- 2020-12-15 notasparaarticulo.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{xcolor}
-\usepackage{calrsfs}
-\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
-\newcommand{\Sm}{\Ss_M}
-\newcommand{\de}{\delta}
-\newcommand{\ep}{\epsilon}
-\newcommand{\Ss}{\mathcal{S}}
-
-\title{Quantum H theorem, variational entropy, and relaxation processes}
-\author{labfqot}
-\date{March 2020}
-
-\begin{document}
-
-\maketitle
-
-\section*{Abstract}
-Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
-This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
-The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
-The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
-In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
-Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
-
-\section{Introduction}
-
-In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
-In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
-\\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
-One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
-The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
-However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
-Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
-\\
-In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
-In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
-Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
-The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
-The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
-Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
-
-%----------------------------------
-\section{Classical scheme}
-\subsection{The Boltzmann transport equation}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
-On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
-However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
-The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
-%-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. 
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
-The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
-As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
-\begin{equation}
-    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
-\end{equation}
-This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
-The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
-\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
-For more information, see \cite{huang}.}}
-%---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
-The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
-However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
-Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
-The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
-\begin{eqnarray}
-    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Besides, the local distribution function must satisfy the following microcanonical restrictions
-\begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
-\end{equation}
-We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
-\begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
-\end{eqnarray}
-Isolating the distribution function 
-\begin{eqnarray}
-    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
-\end{eqnarray}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}
-  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
-  \\
-To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
-\begin{itemize}
-    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
-    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
-    \item The following restrictions
-    \begin{eqnarray}
-        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\end{itemize} 
- Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
-\\
-Performing the derivative of (\ref{CH2}) with respect to time, we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
-\end{equation}{}
-Using the first-order approximation
-\begin{eqnarray}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
-\end{eqnarray}
-(\ref{dH1}) yields
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}
-We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
-\begin{equation}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
-\end{equation}
-With the previous approximation, (\ref{dH1,1}) obtains the following form 
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
-\nonumber \\
-\end{eqnarray}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K}
-\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}
-Then using the previous result to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{eqnarray}
-Now, the summation over $M$ will be expressed in two summations
-\begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
-\label{classicalH3}
-\end{equation}
-where $L+P=K$.
-The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
-On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
-   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}
-and consequently, (\ref{classicalH3}) obtains the following form
-\begin{equation}
-    \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
-    \right]. \label{classicalH4}
-\end{equation}
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
-Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
-With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-Joining both results, we can say the following statement:\\
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
-\end{equation}
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
-\begin{equation}
-    \frac{dH'}{dt} \leq 0.
-\end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.\\
-\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
-
-%------------------------------
-
-\section{Quantum scheme}
-\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
-\end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
-\begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
-\end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
-\begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
-\end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
-\begin{equation}
-    \frac{dH}{dt}\leq 0,
-\end{equation}
-and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version.
-Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
-Those distributions can be obtained from a sufficient condition for the equilibrium state
-\begin{equation}
-    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
-\end{equation}
-where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
-%--------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
-
-We begin by defining the form of the functional 
-\begin{eqnarray}
-    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
-\end{eqnarray}
-This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
-Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
-.
-.
-, K$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
-\begin{eqnarray}
-    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-Besides, the system is subjected to the microcanonical restrictions
-\begin{eqnarray}
-    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
-\end{eqnarray}
-where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
-\begin{eqnarray}
-&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
-\end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
-\begin{equation}
-    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
-\end{equation}
-where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
-With this, we conclude
-\begin{equation}
-    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
-\end{equation}
-In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
-\begin{equation}
-    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
-\begin{equation}
-    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}{}
-and finally
-\begin{eqnarray}
-    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
-\begin{eqnarray}
-      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
-  \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
- %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
-We rewrite the \textit{variational entropy} (\ref{entropy}) as
-\begin{equation}
-    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
-\end{equation}
-where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
-\begin{eqnarray}
-        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the quantum analogous case of the classical case.
-Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
-\\
-If we perform the derivative of the variational entropy with respect to time, we obtain
-\begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}{}
-Using the first-order approximation (\ref{firstorder}), we get
-\begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}{}
-It's necessary to remark that, in the case of fermions
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
-This result corresponds to the system in the zero temperature condition.
-We can observe that in this condition, (\ref{fermionrestriction}) is violated.
-Due to this, the zero temperature condition is excluded.\\
-\\
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}{}
-With the previous approximation, (\ref{cambioH1}) will be
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}{}
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}{}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}{}
-Substituting the previous expression to (\ref{cambioH3}) we obtain
-\begin{eqnarray}
-   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over $M$ will be expressed in two summations, such as in the classical case
-\begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}
-where $L+P=K$.
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
-and consequently, (\ref{cambioH5}) obtains the following form
-\begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.
-
-  %-----------------------------------------
-
-%---------------------------------------------
-
 \section{Quantum Classical correspondence}
@@
 %--------------------------------------
-
-  \section{Relaxation processes in degenerated quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
-\begin{equation}
-    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
-\end{equation}
-where $\mathcal{N}_{M}^{n}$ is defined as
-\begin{equation}
-    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
-\end{equation}
-Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
-This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
-Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
-Then the equation of motion is
-\begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\label{transportequation1}
-\end{equation}
-This equation is valid in equilibrium.
-When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be included in (\ref{transportequation1}).
-This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
-\begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
-    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\nonumber \\ \label{transportequation2}
-\end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
-
-  %----------------------------------------
-
-%-------------------------------------------------
-
-\section{Discussions and remarks}
-
-In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
-In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
-It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$ theorem.\\
-\\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be future work.\\
-\\
-We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
-This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
-\begin{equation}
-    H_{Boltz}=-\frac{S}{Vk},
-\end{equation}
-where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
-\begin{equation}
-    \Ss\propto S.
-\end{equation}
-\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
-It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
-This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
-Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
-Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
-In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
-
-%-----------------------------------------------
-
-\section{Conclusions}
-We conclude the following statements:\\
-\begin{enumerate}
-    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
-    \item We can obtain the transport equation from the $H$ functional.
-    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
-    \item We find an expression for the entropy to quantum gases.
-    \end{enumerate}
-
-%  \begin{equation}
-%      \mathcal{H}
-%  \end{equation}
-
-\clearpage
-
-\begin{thebibliography}{}
-%\bibitem{goldstein}
-%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
-
-%\bibitem{sakurai}
-%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
-
-\bibitem{huang}
-K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).
-
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
-
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
-
-\bibitem{cristal1}
-Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
-
-\bibitem{cristal2}
-Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
-
-\bibitem{cristal3}
-Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
-
-\bibitem{kei}
-Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
-
-\bibitem{onsager}
-Onsager, L. (1931), Physical Review 37 (4): 405-426.
-
-\bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
-
-\bibitem{paradox2}
-Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.
-
-\bibitem{hamiltonian}
-Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
-
-\bibitem{H-theorem-violation}
-Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
-
-\bibitem{H-theorem-and-entropy}
-Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
-
-\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
-
-\bibitem{entropic-framework}
-Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
-
-\bibitem{thermal-harmonic}
-F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
-Phys. Rev. E 91, 042116 – 15 April 2015.
-
-\bibitem{quantum-transport}
-Robert Hussein and Sigmund Kohler,
-Phys. Rev. B 89, 205424, 21 May 2014
-
-\bibitem{quantum-transport-reservoirs}
-Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
-Phys. Rev. A 102, 022207, 11 August 2020.
-
-\bibitem{htheorem2}
-R. Silva {\it et al} EPL 89 10004 (2010).
-
-\bibitem{quantum1}
-J. Math. Phys. 47, 073303 (2006).
-
-\bibitem{quantum2}
-Z. Physik 268, 139-143 (1974).
-
-\bibitem{quantum-entropy}
-J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
-
-\bibitem{quantum-colapse}
-Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
-
-\bibitem{quantum3}
-Gring, M.
-and Kuhnert, M.
-and Langen, T.
-and Kitagawa, T.
-and Rauer, B.
-and Schreitl, M.
-and Mazets, I.
-and Smith, D.
-Adu and Demler, E.
-and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
-
-\bibitem{quantum4}
-Physical Review E 91, 062106 (2015).
-
-%\bibitem{relaxationphenomena}
-%Journal of Statistical Physics, 30; 2 (1983).
-
-%\bibitem{htheorem}
-%Review of Modern Physics; 74 (2002).
-
-\bibitem{binary}
-Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.
-
-\bibitem{contradictions}
-Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.
-
-\bibitem{ergodic-theorem}
-Eur. Phys. J. H 35, 201–237 (2010).
-
-\bibitem{hydrodynamic-simulation}
-Review of Modern Physics 74(4):1203-1220.
-
-\end{thebibliography}{}
-
-\end{document}

### 2020-12-15 tex/quantum.tex
Sources reformatted to match the template of Entropy journal.

--- 2020-12-15 notasparaarticulo.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{xcolor}
-\usepackage{calrsfs}
-\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
-\newcommand{\Sm}{\Ss_M}
-\newcommand{\de}{\delta}
-\newcommand{\ep}{\epsilon}
-\newcommand{\Ss}{\mathcal{S}}
-
-\title{Quantum H theorem, variational entropy, and relaxation processes}
-\author{labfqot}
-\date{March 2020}
-
-\begin{document}
-
-\maketitle
-
-\section*{Abstract}
-Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
-This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
-The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
-The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
-In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
-Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
-
-\section{Introduction}
-
-In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
-In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
-\\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
-One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
-The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
-However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
-Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
-\\
-In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
-In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
-Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
-The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
-The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
-Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
-
-%----------------------------------
-\section{Classical scheme}
-\subsection{The Boltzmann transport equation}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
-On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
-However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
-The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
-%-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. 
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
-The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
-As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
-\begin{equation}
-    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
-\end{equation}
-This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
-The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
-\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
-For more information, see \cite{huang}.}}
-%---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
-The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
-However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
-Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
-The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
-\begin{eqnarray}
-    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Besides, the local distribution function must satisfy the following microcanonical restrictions
-\begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
-\end{equation}
-We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
-\begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
-\end{eqnarray}
-Isolating the distribution function 
-\begin{eqnarray}
-    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
-\end{eqnarray}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}
-  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
-  \\
-To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
-\begin{itemize}
-    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
-    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
-    \item The following restrictions
-    \begin{eqnarray}
-        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\end{itemize} 
- Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
-\\
-Performing the derivative of (\ref{CH2}) with respect to time, we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
-\end{equation}{}
-Using the first-order approximation
-\begin{eqnarray}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
-\end{eqnarray}
-(\ref{dH1}) yields
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}
-We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
-\begin{equation}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
-\end{equation}
-With the previous approximation, (\ref{dH1,1}) obtains the following form 
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
-\nonumber \\
-\end{eqnarray}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K}
-\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}
-Then using the previous result to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{eqnarray}
-Now, the summation over $M$ will be expressed in two summations
-\begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
-\label{classicalH3}
-\end{equation}
-where $L+P=K$.
-The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
-On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
-   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}
-and consequently, (\ref{classicalH3}) obtains the following form
-\begin{equation}
-    \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
-    \right]. \label{classicalH4}
-\end{equation}
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
-Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
-With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-Joining both results, we can say the following statement:\\
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
-\end{equation}
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
-\begin{equation}
-    \frac{dH'}{dt} \leq 0.
-\end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.\\
-\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
-
-%------------------------------
-
 \section{Quantum scheme}
@@
   \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
+  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.
+
 %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
@@
 where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
+
 This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
@@
 %---------------------------------------------
-
-\section{Quantum Classical correspondence}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
-\end{eqnarray}
-where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
-\textcolor{blue}{In the equilibrium case, we have the following variational entropy
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
-                    \left(
-                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                  \right) \ln
-                \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
-                \right) \right]. \label{equilibriumvariational}
-\end{eqnarray}
-To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
-\begin{equation}
-    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
-\end{equation}
-and then
-\begin{equation}
-    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
-\end{equation}
-If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}
-                  \right) \ln
-                \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
-                \right) \right],
-\end{eqnarray}
-but if (\ref{classicalapproximation}) holds, then 
-\begin{equation}
-     e^{\alpha_M+\beta_M \epsilon}\ll 1,
-\end{equation}
-consequently 
-\begin{equation}
-    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
-\end{equation}
-and finally
-\begin{equation}
-    \Ss=\sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right],
-\end{equation}
-that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
-\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
-
-%--------------------------------------
-
-  \section{Relaxation processes in degenerated quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
-\begin{equation}
-    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
-\end{equation}
-where $\mathcal{N}_{M}^{n}$ is defined as
-\begin{equation}
-    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
-\end{equation}
-Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
-This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
-Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
-Then the equation of motion is
-\begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\label{transportequation1}
-\end{equation}
-This equation is valid in equilibrium.
-When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be included in (\ref{transportequation1}).
-This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
-\begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
-    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
-\nonumber \\ \label{transportequation2}
-\end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
-
-  %----------------------------------------
-
-%-------------------------------------------------
-
-\section{Discussions and remarks}
-
-In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
-In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
-It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$ theorem.\\
-\\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be future work.\\
-\\
-We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
-This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
-\begin{equation}
-    H_{Boltz}=-\frac{S}{Vk},
-\end{equation}
-where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
-\begin{equation}
-    \Ss\propto S.
-\end{equation}
-\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
-It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
-This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
-Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
-Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
-In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
-
-%-----------------------------------------------
-
-\section{Conclusions}
-We conclude the following statements:\\
-\begin{enumerate}
-    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
-    \item We can obtain the transport equation from the $H$ functional.
-    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
-    \item We find an expression for the entropy to quantum gases.
-    \end{enumerate}
-
-%  \begin{equation}
-%      \mathcal{H}
-%  \end{equation}
-
-\clearpage
-
-\begin{thebibliography}{}
-%\bibitem{goldstein}
-%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
-
-%\bibitem{sakurai}
-%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
-
-\bibitem{huang}
-K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).
-
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
-
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
-
-\bibitem{cristal1}
-Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
-
-\bibitem{cristal2}
-Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
-
-\bibitem{cristal3}
-Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
-
-\bibitem{kei}
-Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
-
-\bibitem{onsager}
-Onsager, L. (1931), Physical Review 37 (4): 405-426.
-
-\bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
-
-\bibitem{paradox2}
-Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.
-
-\bibitem{hamiltonian}
-Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
-
-\bibitem{H-theorem-violation}
-Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
-
-\bibitem{H-theorem-and-entropy}
-Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
-
-\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
-
-\bibitem{entropic-framework}
-Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
-
-\bibitem{thermal-harmonic}
-F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
-Phys. Rev. E 91, 042116 – 15 April 2015.
-
-\bibitem{quantum-transport}
-Robert Hussein and Sigmund Kohler,
-Phys. Rev. B 89, 205424, 21 May 2014
-
-\bibitem{quantum-transport-reservoirs}
-Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
-Phys. Rev. A 102, 022207, 11 August 2020.
-
-\bibitem{htheorem2}
-R. Silva {\it et al} EPL 89 10004 (2010).
-
-\bibitem{quantum1}
-J. Math. Phys. 47, 073303 (2006).
-
-\bibitem{quantum2}
-Z. Physik 268, 139-143 (1974).
-
-\bibitem{quantum-entropy}
-J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
-
-\bibitem{quantum-colapse}
-Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
-
-\bibitem{quantum3}
-Gring, M.
-and Kuhnert, M.
-and Langen, T.
-and Kitagawa, T.
-and Rauer, B.
-and Schreitl, M.
-and Mazets, I.
-and Smith, D.
-Adu and Demler, E.
-and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
-
-\bibitem{quantum4}
-Physical Review E 91, 062106 (2015).
-
-%\bibitem{relaxationphenomena}
-%Journal of Statistical Physics, 30; 2 (1983).
-
-%\bibitem{htheorem}
-%Review of Modern Physics; 74 (2002).
-
-\bibitem{binary}
-Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.
-
-\bibitem{contradictions}
-Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.
-
-\bibitem{ergodic-theorem}
-Eur. Phys. J. H 35, 201–237 (2010).
-
-\bibitem{hydrodynamic-simulation}
-Review of Modern Physics 74(4):1203-1220.
-
-\end{thebibliography}{}
-
-\end{document}

### 2020-12-15 tex/relaxproc.tex
Sources reformatted to match the template of Entropy journal.

--- 2020-12-15 notasparaarticulo.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{xcolor}
-\usepackage{calrsfs}
-\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
-\newcommand{\Sm}{\Ss_M}
-\newcommand{\de}{\delta}
-\newcommand{\ep}{\epsilon}
-\newcommand{\Ss}{\mathcal{S}}
-
-\title{Quantum H theorem, variational entropy, and relaxation processes}
-\author{labfqot}
-\date{March 2020}
-
-\begin{document}
-
-\maketitle
-
-\section*{Abstract}
-Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure.
-This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$.
-The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas.
-The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle.
-In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained.
-Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.
-
-\section{Introduction}
-
-In the literature, the theoretical basis that allows us to describe equilibrium systems in the classical scheme is well-established.
-The procedures to describe equilibrium systems are well-known and work in almost every system in nature (Thermodynamics for a phenomenological description and statistical mechanics for a microscopic description).
-In contrast, a unified mathematical model to describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the evolution to equilibrium to almost any system (the internal behavior of glasses \cite{cristal1} and the entropy measurement \cite{cristal2, cristal3} are some examples where the recent developments do not work.) Examples of those developments can be the Onsager formulation (called linear thermodynamics) \cite{kei, onsager}, and the Kinetic Theory of dilute gases developed by Boltzmann.\\
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the Boltzmann's $H$ functional that deserve some review to accomplish better consistency, for example, the modification of the $H$ theorem in stochastic trajectories, violations in the second law of thermodynamics, and the relation between Shannon's measure of information and the Boltzmann's entropy  [6-10].\\
-\\
-To obtain the well-known correspondence principle between classical mechanics and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or quantum $H$ theorem) and the quantum analogous to the Boltzmann transport equation is necessary such that in the appropriate limit, the  classical $H$ theorem and the Boltzmann transport equation should be recovered.\\
-One of the first developments in this field was developed by Tolman \cite{tolman}.
-He proposes one of the first quantum versions of the Boltzmann's $H$ theorem using the probability transition relation, the random phases hypothesis, and a homogeneous distribution function.
-The equation of motion for the occupation number (necessary to prove the quantum $H$ theorem in the same way that Boltzmann's procedure) was obtained by applying the time perturbation theory.
-However, to obtain the flux relations in analogy to the Boltzmann's method, a quantum version of the transport equation that includes a non-homogeneous distribution function is required.\\
-On the other hand, some works in the quantum operator formalism attempt to describe quantum transport phenomena using the Hamiltonian of the system and the master equation (analogous to the Boltzmann transport equation) [12-15].
-Besides, some authors construct the $H$ functional and develop the proof of the quantum $H$ theorem \cite{htheorem2, quantum1, quantum2}.
-However, the homogeneous distribution function hypothesis and the correspondence principle are uncleared or not discussed.
-Moreover, the validity of the quantum $H$ theorem and the second law of thermodynamics, and the interpretation of the quantum entropy is still in discussion [19-26].\\
-\\
-In order to contribute to the construction of a consistent classical and quantum $H$ theorem, and a new formalism to describe out of equilibrium systems, we develop a new theoretical basis to describe them.
-In particular, we are interested in including a non-homogeneous distribution function in the mathematical model, and so including non-homogeneous systems in the $H$ theorem.
-Besides, we are interested in developing the quantum analogous of Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution function and solve the problem of the correspondence principle mentioned above.\\
-\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem.
-The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem.
-The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional.
-Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}
-
-%----------------------------------
-\section{Classical scheme}
-\subsection{The Boltzmann transport equation}
-In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is
-\begin{equation}
-    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}
-\end{equation}
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
-\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}).
-On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation.
-However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}).
-The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
-%-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
-\begin{equation}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
-\end{equation} 
-where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. 
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy when the system arrives at the equilibrium state.\\
-\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$.
-The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}).
-As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution.
-This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$
-\begin{equation}
-    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
-\end{equation}
-This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
-The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does.
-\footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem.
-For more information, see \cite{huang}.}}
-%---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
-The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$.
-However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem.
-Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
-\end{equation}
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$.
-Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$.
-The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are given by:
-\begin{eqnarray}
-    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
-    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
-{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Besides, the local distribution function must satisfy the following microcanonical restrictions
-\begin{equation}
-    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.
-\end{equation}
-We can see that the functional (\ref{CH2}) contains a non-homogeneous distribution function but keeps the same form of the Boltzmann's $H$ functional.\\
-Inspired by the Hamilton principle, $H'$ will be maximized with those restrictions to obtain the equilibrium distribution function for a classical gas using the Lagrange multipliers method obtaining
-\begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
-    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
-\end{eqnarray}
-Isolating the distribution function 
-\begin{eqnarray}
-    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
-    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
-\end{eqnarray}
-where $C$ is a constant. It is trivial that using the variational procedures we obtain the form of the Maxwell-Boltzmann distribution. \\
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
-\begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
-\end{equation}{} 
-The constant $C$ can be omitted defining the following $H$ functional
-\begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
-\end{equation}
-  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number.
-As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted.
-\\
-  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as
-  \begin{equation}
-      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-  \end{equation}
-  where
-  \begin{equation}
-      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-  \end{equation}
-  Besides, if we analyze the following expression
-  \begin{eqnarray}
-      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-  \end{eqnarray}
-  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-  \begin{equation}
-      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-  \end{equation}
-  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
-  \\
-To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:
-\begin{itemize}
-    \item The local equilibrium hypothesis in each cell ($f_{M}$ must be spatial-homogeneous for each $M$).
-    \item The non-homogeneous distribution function assumption in the total volume ($f_{M}$ must be different among all cells).
-    \item The following restrictions
-    \begin{eqnarray}
-        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-\end{itemize} 
- Those conditions described a system out of equilibrium but no so far from it, that is a perturbed system and will suffer a \textit{relaxation process}\footnote{The time evolution of a perturbated system to the equilibrium state.}.\\
-  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.
-\\
-Performing the derivative of (\ref{CH2}) with respect to time, we obtain
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.
-\end{equation}{}
-Using the first-order approximation
-\begin{eqnarray}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}
-\end{eqnarray}
-(\ref{dH1}) yields
-\begin{equation}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.
-\end{equation}
-We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
-\begin{equation}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}
-\end{equation}
-With the previous approximation, (\ref{dH1,1}) obtains the following form 
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
-\end{eqnarray}{}
-using (\ref{relacion1}) with homogeneous multipliers
-\begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}.
-\nonumber \\
-\end{eqnarray}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
-    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K}
-\Delta_M(t)  =\sum_{M=1}^{K} \delta_M(t) =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
-\end{equation}
-Then using the previous result to (\ref{dH1,2}), we find
-\begin{eqnarray}
-\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{eqnarray}
-Now, the summation over $M$ will be expressed in two summations
-\begin{equation}
-    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
-\label{classicalH3}
-\end{equation}
-where $L+P=K$.
-The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium).
-In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium).
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles and/or energy and so, $g^{+}_{J}>0$.
-On the other hand, $\dot{g}^{-}_{J}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{J}=+|g^{+}_{J}|; \ \ \  \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|, \nonumber \\
-   &&g^{-}_{J}=-|g^{-}_{J}|; \ \ \ \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}| \label{separacionclassical},
-\end{eqnarray}
-and consequently, (\ref{classicalH3}) obtains the following form
-\begin{equation}
-    \frac{dH'}{dt}=-\left[
-                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v
-    \right]. \label{classicalH4}
-\end{equation}
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive.
-Also the deviation and its derivative are positive, then all the expression is positive.
-However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero.
-With this we proved that $\frac{dH'}{dt}<0$.
-If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\
-Joining both results, we can say the following statement:\\
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-\begin{equation}
-   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},
-\end{equation}
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation
-\begin{equation}
-    \frac{dH'}{dt} \leq 0.
-\end{equation}
-This statement will be the classical $H$ theorem with inhomogeneities.\\
-\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}
-
-%------------------------------
-
-\section{Quantum scheme}
-\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
-\end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
-\begin{eqnarray}
-    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
-\end{eqnarray}
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
-\begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
-\end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
-\begin{equation}
-    \frac{dH}{dt}\leq 0,
-\end{equation}
-and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version.
-Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
-Those distributions can be obtained from a sufficient condition for the equilibrium state
-\begin{equation}
-    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
-\end{equation}
-where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-This expression also can be obtained also from the Boltzmann $H$ functional through defining 
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
-%--------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
-
-We begin by defining the form of the functional 
-\begin{eqnarray}
-    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
-\end{eqnarray}
-This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
-Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
-.
-.
-, K$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
-\begin{eqnarray}
-    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
-\end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
-\begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
-\end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-Besides, the system is subjected to the microcanonical restrictions
-\begin{eqnarray}
-    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
-\end{eqnarray}
-where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
-\begin{eqnarray}
-&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
-\end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
-\begin{equation}
-    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
-\end{equation}
-where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
-With this, we conclude
-\begin{equation}
-    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
-\end{equation}
-In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
-\begin{equation}
-    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
-\begin{equation}
-    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}{}
-and finally
-\begin{eqnarray}
-    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
-\begin{eqnarray}
-      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
-  \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.\\
- %\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
-We rewrite the \textit{variational entropy} (\ref{entropy}) as
-\begin{equation}
-    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
-\end{equation}
-where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
-\begin{eqnarray}
-        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
-  \begin{equation}
-      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
-  \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the quantum analogous case of the classical case.
-Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
-\\
-If we perform the derivative of the variational entropy with respect to time, we obtain
-\begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}{}
-Using the first-order approximation (\ref{firstorder}), we get
-\begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}{}
-It's necessary to remark that, in the case of fermions
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
-This result corresponds to the system in the zero temperature condition.
-We can observe that in this condition, (\ref{fermionrestriction}) is violated.
-Due to this, the zero temperature condition is excluded.\\
-\\
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}{}
-With the previous approximation, (\ref{cambioH1}) will be
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}{}
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-\begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}{}
-On the other hand, we obtain from the restrictions the following expressions
-\begin{eqnarray}
-    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}{}
-and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}{}
-Substituting the previous expression to (\ref{cambioH3}) we obtain
-\begin{eqnarray}
-   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over $M$ will be expressed in two summations, such as in the classical case
-\begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}
-where $L+P=K$.
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-\\
-We sort the previous ideas in the following form
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
-and consequently, (\ref{cambioH5}) obtains the following form
-\begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.
-
-  %-----------------------------------------
-
-%---------------------------------------------
-
-\section{Quantum Classical correspondence}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
-\end{eqnarray}
-where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
-\textcolor{blue}{In the equilibrium case, we have the following variational entropy
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
-                    \left(
-                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                  \right) \ln
-                \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
-                \right) \right]. \label{equilibriumvariational}
-\end{eqnarray}
-To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
-\begin{equation}
-    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
-\end{equation}
-and then
-\begin{equation}
-    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
-\end{equation}
-If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}
-                  \right) \ln
-                \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
-                \right) \right],
-\end{eqnarray}
-but if (\ref{classicalapproximation}) holds, then 
-\begin{equation}
-     e^{\alpha_M+\beta_M \epsilon}\ll 1,
-\end{equation}
-consequently 
-\begin{equation}
-    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
-\end{equation}
-and finally
-\begin{equation}
-    \Ss=\sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right],
-\end{equation}
-that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
-\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
-
-%--------------------------------------
-
   \section{Relaxation processes in degenerated quantum gases}
@@
 %-------------------------------------------------
-
-\section{Discussions and remarks}
-
-In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
-In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
-It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$ theorem.\\
-\\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be future work.\\
-\\
-We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
-This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
-\begin{equation}
-    H_{Boltz}=-\frac{S}{Vk},
-\end{equation}
-where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
-\begin{equation}
-    \Ss\propto S.
-\end{equation}
-\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
-It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
-This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
-Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
-Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
-In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
-
-%-----------------------------------------------
-
-\section{Conclusions}
-We conclude the following statements:\\
-\begin{enumerate}
-    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
-    \item We can obtain the transport equation from the $H$ functional.
-    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
-    \item We find an expression for the entropy to quantum gases.
-    \end{enumerate}
-
-%  \begin{equation}
-%      \mathcal{H}
-%  \end{equation}
-
-\clearpage
-
-\begin{thebibliography}{}
-%\bibitem{goldstein}
-%H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).
-
-%\bibitem{sakurai}
-%J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.
-
-\bibitem{huang}
-K. Huang, \emph{Statistical mechanics}, John Wiley \& Sons, Inc., (1963).
-
-\bibitem{reif}
-F. Reif {\it Fundamentals of Statistical and Thermal Physics.},Waveland Press Inc, 2009.
-
-\bibitem{patrick}
-R. Fitzpatrick, {\it Thermodynamics and Statistical Mechanics: An intermediate level.},University Reprints 2012 (2012).
-
-\bibitem{cristal1}
-Zanotto, E.D.; Mauro, J.C., {\it Comment on “Glass Transition, Crystallization of Glass-Forming Melts, and Entropy”} Entropy 2018, 20, 103.. Entropy 2018, 20, 703.
-
-\bibitem{cristal2}
-Schmelzer, J.W.P.; Tropin, T.V. {\it Glass Transition, Crystallization of Glass-Forming Melts, and Entropy}, Entropy 2018, 20, 103.
-
-\bibitem{cristal3}
-Nemilov, S.V. {\it On the Possibility of Calculating Entropy, Free Energy, and Enthalpy of Vitreous Substances}, Entropy 2018, 20, 187.
-
-\bibitem{kei}
-Keizer J., \emph{Statistical thermodynamics of nonequilibrium processes}, Springer-Verlag, New York, 1987.
-
-\bibitem{onsager}
-Onsager, L. (1931), Physical Review 37 (4): 405-426.
-
-\bibitem{paradox1}
-Harvey R. Brown and Wayne Myrvold, {\it Boltzmann's H-theorem, its limitations, and the birth of (fully) statistical mechanics}, physics.hist-ph,0809.1304,2008.
-
-\bibitem{paradox2}
-Dragoljub A. Cucic, {\it Paradoxes of Thermodynamics and Statistical Physics}, physics.gen-ph,0912.1756,2009.
-
-\bibitem{hamiltonian}
-Wang, Q.A.; El Kaabouchiu, A. {\it From Random Motion of Hamiltonian Systems to Boltzmann’s H Theorem and Second Law of Thermodynamics: a Pathway by Path Probability}, Entropy 2014, 16, 885-894.
-
-\bibitem{H-theorem-violation}
-Gorban, A.N. {\it General H-theorem and Entropies that Violate the Second Law}, Entropy 2014, 16, 2408-2432.
-
-\bibitem{H-theorem-and-entropy}
-Ben-Naim, A. {\it Entropy, Shannon’s Measure of Information and Boltzmann’s H-Theorem}, Entropy 2017, 19, 48.
-
-\bibitem{tolman} Richard C. Tolman, \emph{The principles of statistical mechanics}, Oxford, 1938.
-
-\bibitem{entropic-framework}
-Li, S.-N.; Cao, B.-Y. {\it On Entropic Framework Based on Standard and Fractional Phonon Boltzmann Transport Equations}, Entropy 2019, 21, 204.
-
-\bibitem{thermal-harmonic}
-F. Nicacio, A. Ferraro, A. Imparato, M. Paternostro, and F. L. Semião,
-Phys. Rev. E 91, 042116 – 15 April 2015.
-
-\bibitem{quantum-transport}
-Robert Hussein and Sigmund Kohler,
-Phys. Rev. B 89, 205424, 21 May 2014
-
-\bibitem{quantum-transport-reservoirs}
-Giulio Amato, Heinz-Peter Breuer, Sandro Wimberger, Alberto Rodríguez, and Andreas Buchleitner
-Phys. Rev. A 102, 022207, 11 August 2020.
-
-\bibitem{htheorem2}
-R. Silva {\it et al} EPL 89 10004 (2010).
-
-\bibitem{quantum1}
-J. Math. Phys. 47, 073303 (2006).
-
-\bibitem{quantum2}
-Z. Physik 268, 139-143 (1974).
-
-\bibitem{quantum-entropy}
-J. Acharya, I. Issa, N. V. Shende and A. B. Wagner, {\it Measuring Quantum Entropy}, 2019 IEEE International Symposium on Information Theory (ISIT), Paris, France, 2019, pp. 3012-3016
-
-\bibitem{quantum-colapse}
-Kastner, R.E. {\it On Quantum Collapse as a Basis for the Second Law of Thermodynamics}, Entropy 2017, 19, 106.
-
-\bibitem{quantum3}
-Gring, M.
-and Kuhnert, M.
-and Langen, T.
-and Kitagawa, T.
-and Rauer, B.
-and Schreitl, M.
-and Mazets, I.
-and Smith, D.
-Adu and Demler, E.
-and Schmiedmayer, J.; {\it Relaxation and Prethermalization in an Isolated Quantum System}; 337; 6100; 1318-1322 (2012).
-
-\bibitem{quantum4}
-Physical Review E 91, 062106 (2015).
-
-%\bibitem{relaxationphenomena}
-%Journal of Statistical Physics, 30; 2 (1983).
-
-%\bibitem{htheorem}
-%Review of Modern Physics; 74 (2002).
-
-\bibitem{binary}
-Das, Bandita and Biswas, Shyamal; Journal of Statistical Mechanics: Theory and Experiment;2018, 10, 1742-5468.
-
-\bibitem{contradictions}
-Lesovik, G. B. and Lebedev, A. V. and Sadovskyy, I. A. and Suslov, M. V. and Vinokur, V. M., Scientific Reports, Vol. 6, \# 1, 2016.
-
-\bibitem{ergodic-theorem}
-Eur. Phys. J. H 35, 201–237 (2010).
-
-\bibitem{hydrodynamic-simulation}
-Review of Modern Physics 74(4):1203-1220.
-
-\end{thebibliography}{}
-
-\end{document}

### 2020-12-15 tex/templatedocs/template.tex
Sources reformatted to match the template of Entropy journal.

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, appliedchem, applmech, applmicrobiol, applnano, applsci, arts, asi, atmosphere, atoms, audiolres, automation, axioms, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomechanics, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, brainsci, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coatings, colloids, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, crops, cryptography, crystals, curroncol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forensicsci, forests, fractalfract, fuels, futureinternet, futuretransp, futurepharmacol, futurephys, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, land, languages, laws, life, liquids, literature, livers, logistics, lubricants, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, metabolites, metals, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, minerals, mining, modelling, molbank, molecules, mps, mti, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photochem, photonics, physchem, physics, physiolsci, plants, plasma, pollutants, polymers, polysaccharides, proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, systems, taxonomy, technologies, telecom, textiles, thermo, tourismhosp, toxics, toxins, transplantology, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wevj, women, world

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2021}
\copyrightyear{2020}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{} 
\dateaccepted{} 
\datepublished{}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusion: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{Instead of the abstract}
%\entrylink{The Link to this entry published on the encyclopedia platform.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}

The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions on the authors' page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX-related questions please contact latex@mdpi.com.
%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Histories, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item; 
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}

% The MDPI table float is called specialtable
\begin{specialtable}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
%%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabular}
\end{specialtable}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{equation}
a = 1,
\end{equation}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\end{paracol}
\nointerlineskip
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}

% Example of a figure that spans the whole page width (the commands \widefigure and \begin{paracol}{2}, \linenumbers, and\switchcolumn need to be present). The same concept works for tables, too.
\begin{figure}[H]	
\widefigure
\includegraphics[width=15 cm]{Definitions/logo-mdpi}
\caption{This is a wide figure.\label{fig2}}
\end{figure}  
\begin{paracol}{2}
\linenumbers
\switchcolumn

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals.}

\informedconsent{Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.}

\dataavailability{Please refer to suggested Data Availability Statements in section “MDPI Research Data Policies” at \href{https://www.mdpi.com/ethics}{https://www.mdpi.com/ethics}.}

\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the~results''.}

%% Optional
\sampleavailability{Samples of the compounds ... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendix
\section{}
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{paracol}
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book1}
Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
\bibitem[Author3(year)]{ref-book2}
Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
\bibitem[Author4(year)]{ref-unpublish}
Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
% Reference 5
\bibitem[Author5(year)]{ref-communication}
Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
\bibitem[Author6(year)]{ref-proceeding}
Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of Presentation. In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds. (if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
% Reference 7
\bibitem[Author7(year)]{ref-thesis}
Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
\bibitem[Author8(year)]{ref-url}
Title of Site. Available online: URL (accessed on Day Month Year).
\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2020-12-16 tex/classical.tex
quantum.tex reformatted: 80chars or less

@@
 The $H$-theorem fixes the time evolution direction of systems in nature, such
-as the second law of thermodynamics does. {\color{blue}footnote: }The molecular hypothesis is %
+as the second law of thermodynamics does.\footnote{The molecular hypothesis is %
 fundamental to determine the physical meaning of the $H$ theorem. For more %
-information, see \cite{bib:huang}.
+information, see \cite{bib:huang}.}
 

### 2020-12-16 tex/quantum.tex
quantum.tex reformatted: 80chars or less

@@
 \section{Quantum scheme}
+
 \subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
-The $H$ functional in quantum mechanics presented by Tolman \cite{tolman} is defined as 
+
+The $H$ functional in quantum mechanics presented by Tolman \cite{bib:tolman}
+is defined as
+%
 \begin{equation}
@@
 \end{equation}
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an equation for the rate of change in the number of particles in group $\kappa$ is
+%
+where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an
+equation for the rate of change in the number of particles in group $\kappa$ is
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
+%
 where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
+%
 \begin{equation}
@@
 \end{equation}
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where $V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions are obtained by the perturbation theory.\\
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that if we have a quantum gas that satisfies the rate of change equation (\ref{changen})\footnote{It is important to remark that the equal a priori probabilities and random a priori phases hypothesis are implicit in the rate of change equation (\ref{changen}) and the random a priori phases hypothesis is considered as the quantum analogous to the molecular chaos hypothesis \cite{binary}}, then the $H$ functional (\ref{quantumh}) satisfies that
+%
+where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where
+$V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions
+are obtained by the perturbation theory.
+
+\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that
+if we have a quantum gas that satisfies the rate of change equation
+(\ref{changen})\footnote{It is important to remark that the equal a priori
+probabilities and random a priori phases hypothesis are implicit in the rate of
+change equation (\ref{changen}) and the random a priori phases hypothesis is
+considered as the quantum analogous to the molecular chaos hypothesis
+\cite{bib:das2018}}, then the $H$ functional (\ref{quantumh}) satisfies that
+%
 \begin{equation}
@@
 \end{equation}
-and as a consequence of this theorem, when the system is in equilibrium, $\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version.
-Also, when the system is in equilibrium, the distribution functions that correspond to the equilibrium state are the Fermi-Dirac and the Bose-Einstein distribution.
-Those distributions can be obtained from a sufficient condition for the equilibrium state
+%
+and as a consequence of this theorem, when the system is in equilibrium,
+$\frac{dH}{dt}=0$.\\
+The proof of this theorem is in the same way as in the classical version. Also,
+when the system is in equilibrium, the distribution functions that correspond
+to the equilibrium state are the Fermi-Dirac and the Bose-Einstein
+distribution. Those distributions can be obtained from a sufficient condition
+for the equilibrium state
+%
 \begin{equation}
@@
 \end{equation}
-where the upper sign refers to the Bose-Einstein distribution and the lower sign refers to the Fermi-Dirac distribution.}\\
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the limit of high-energy levels to
+%
+where the upper sign refers to the Bose-Einstein distribution and the lower
+sign refers to the Fermi-Dirac distribution.}
+
+In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the
+limit of high-energy levels to
+%
 \begin{equation}
@@
 \end{equation}
+%
 This expression also can be obtained also from the Boltzmann $H$ functional through defining
+%
 \begin{equation}
@@
 \end{equation}
+%
 %--------------------------
+
 \subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
@@
 We begin by defining the form of the functional
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
+%
 This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower sing) $N$, global volume $V$, and a temperature $T$. 
-Besides, the variational entropy should correspond to a density of entropy when the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{Mn}(\epsilon_{n},t)$.
-The distribution function will depend on the position of the cells, the energy per cell $\epsilon_{n}$, and time $t$.
-Also, the energy spectra will be divided into $\delta \epsilon_n$ volumes.
-We will use the uppercase Latin index to denote cell numbers, and unless stated otherwise, $M = 1, 2 .
-.
-.
-, K$.
-Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three properties are given by:
+This functional describes, as in the classical case, a system with a global
+energy $E$, a global number of free
+quantum particles (which may be fermions for upper sing or bosons for lower
+sing) $N$, global volume $V$, and a temperature $T$.
+Besides, the variational entropy should correspond to a density of entropy when
+the system arrives at the equilibrium state. The
+global system, in turn, consists of a set of $K$ small cells with a constant
+volume $\delta V_M$, whose local homogeneous distribution functions is
+$f_{Mn}(\epsilon_{n},t)$. The distribution function will depend on the position
+of the cells, the energy per cell $\epsilon_{n}$, and time $t$. Also, the
+energy spectra will be divided into $\delta \epsilon_n$ volumes. We will use
+the uppercase Latin index to denote cell numbers, and unless stated otherwise,
+$M = 1, 2 . . . , K$. Without loss of generality, we choose these cells to have
+local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each
+cell has its local chemical potential, $\mu_M$, local
+variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and
+local energy, $\mathcal{E}_M$. In terms of the local distribution
+function and local energy spectra, $\{\epsilon_{n}\}$, the last three
+properties are given by:
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number, this is
+%
+When the system is in equilibrium, the local number of particles and the local
+energy don't depend on the cell number, this is
+%
 \begin{equation}
@@
 \end{equation}
-The particles in the system are considered to be free, \textit{i.e.} the particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$.
-Besides, the system is subjected to the microcanonical restrictions
+%
+The particles in the system are considered to be free, \textit{i.e.} the
+particles do not interact with each other.
+Consequently, we can assume that the available quantum states do not depend on
+the cell features,
+therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all
+cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$. Besides, the
+system is subjected to the microcanonical restrictions
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $N$ and $E$ are the total particle number and the total energy of the system respectively.\\ 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method, we obtain
+%
+where $N$ and $E$ are the total particle number and the total energy of the
+system respectively.
+
+Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method,
+we obtain
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain from the variational procedure that 
+%
+where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain
+from the variational procedure that
+%
 \begin{equation}
@@
 \end{equation}
-where we used that the $H$ functional in equilibrium is proportional to the entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell.
-With this, we conclude
+%
+where we used that the $H$ functional in equilibrium is proportional to the
+entropy of the system.
+In other hand, in the expression (\ref{multipliers}), we expect that the
+variation of the logarithm with the number of particles by cells is indifferent
+from a particular cell. With this, we conclude 
+%
 \begin{equation}
@@
 \end{equation}
-In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore
+%
+In equilibrium, the energy and the Lagrange multiplier do not depend on the
+cell number and time, therefore
+%
 \begin{equation}
     \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}{}
-If one compares those multipliers with the mean energy and the mean particle number from the statistical mechanic's results, we obtain
+\end{equation}
+%
+If one compares those multipliers with the mean energy and the mean particle
+number from the statistical mechanic's results, we obtain
+%
 \begin{equation}
     \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}{}
+\end{equation}
+%
 and finally
+%
 \begin{eqnarray}
     \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}{}
-We proved, when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multipliers are homogeneous in the system.\\
-We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
+\end{eqnarray}
+%
+We proved, when the system is in equilibrium, the Lagrange multipliers don't
+depend on the cell number. In other words, Lagrange multipliers are homogeneous
+in the system.
+
+We can calculate the entropy from an ideal system, in other words, if one
+substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
+%
 \begin{eqnarray}
@@
   \end{eqnarray}
-  This result allows us to calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe those systems.
+%
+This result allows us to calculate the entropy of a quantum ideal gases. This
+entropy could be used as a thermodynamic variable to describe those systems.
 
@@
 We rewrite the \textit{variational entropy} (\ref{entropy}) as
+%
 \begin{equation}
@@
 \end{equation}
+%
 where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
 
-This distribution function must satisfy the local equilibrium hypothesis in each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the non-homogeneous distribution function assumption in the total volume ($f_{nM}$ must be different among all cells), and the following restrictions
+This distribution function must satisfy the local equilibrium hypothesis in
+each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the
+non-homogeneous distribution function assumption in the total volume ($f_{nM}$
+must be different among all cells), and the following restrictions
+%
 \begin{eqnarray}
@@
   \end{eqnarray}
-  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
+%
+where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle
+number and the local energy in equilibrium
+%
 \begin{equation}
@@
   \end{equation}
-  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$.
-The previous conditions and restrictions are the quantum analogous case of the classical case.
-Also, this system will suffer, as in the classical case, a \textit{relaxation process}.\\
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but sufficiently small to consider the system not far from the equilibrium state, as in the classical case.
-\\
-If we perform the derivative of the variational entropy with respect to time, we obtain
+%
+and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$
+and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$
+and $\delta_M(t) \ll \bar{\mathcal{E}}$. The previous conditions and
+restrictions are the quantum analogous case of the classical case. Also, this
+system will suffer, as in the classical case, a \textit{relaxation process}.
+
+  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from
+fluctuations in the system but sufficiently small to consider the system not
+far from the equilibrium state, as in the classical case.
+
+If we perform the derivative of the variational entropy with respect to time,
+we obtain
+%
 \begin{equation}
    \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}{}
+\end{equation}
+%
 Using the first-order approximation (\ref{firstorder}), we get
+%
 \begin{eqnarray}
@@
     \label{cambioH1}
-\end{eqnarray}{}
+\end{eqnarray}
+%
 It's necessary to remark that, in the case of fermions
+%
 \begin{eqnarray}
    1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}{}
-The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$.
-Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogeneities because of the exclusion principle, and as a consequence of this,  $g_{nM}=0$.
-This result corresponds to the system in the zero temperature condition.
-We can observe that in this condition, (\ref{fermionrestriction}) is violated.
-Due to this, the zero temperature condition is excluded.\\
-\\
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
-With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
+\end{eqnarray}
+%
+The previous expression establishes that the value of $g_{nM}$ is determined by
+$\bar f_n$. Furthermore, whether $\bar f_{n}=1$, this is, all lower energy
+levels are occupied, then the system will not have inhomogeneities because of
+the exclusion principle, and as a consequence of this,  $g_{nM}=0$. This result
+corresponds to the system in the zero temperature condition. We can observe
+that in this condition, (\ref{fermionrestriction}) is violated. Due to this,
+the zero temperature condition is excluded.
+
+We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$
+then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$. In fermions, in the case of a
+non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n
+|g_{nM}|$. With those relations, we can approximate the logarithm functions in
+(\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
+%
 \begin{equation}
@@
 \label{lnapproximation}
-\end{equation}{}
+\end{equation}
+%
 With the previous approximation, (\ref{cambioH1}) will be
+%
 \begin{eqnarray}
@@
     &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}{}
+\end{eqnarray}
+%
 Making use of the expression (\ref{relation}),
 (\ref{cambioH2}) casts into
+%
 \begin{eqnarray}
@@
     \label{cambioH3}
-\end{eqnarray}{}
+\end{eqnarray}
+%
 On the other hand, we obtain from the restrictions the following expressions
+%
 \begin{eqnarray}
@@
     &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}{}
+\end{eqnarray}
+%
 and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
+%
 \begin{equation}
     \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}{}
+\end{equation}
+%
 Substituting the previous expression to (\ref{cambioH3}) we obtain
+%
 \begin{eqnarray}
    \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}{}
-Now, the summation over $M$ will be expressed in two summations, such as in the classical case
+\end{eqnarray}
+%
+Now, the summation over $M$ will be expressed in two summations, such as in the
+classical case
+%
 \begin{equation}
@@
 \end{equation}
-where $L+P=K$.
-Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells that have an excess of particles or energy.
-In contrast, $g^{-}_{J}$ represents the value of the deviation on cells that have missing particles or energy.\\
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles and/or energy and so, $g^{+}_{nJ}>0$.
-On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-\\
+%
+where $L+P=K$. Besides, $\dot{g}^{+}_{J}$ represents the change on the
+deviation on cells that have an excess of particles or energy while
+$\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have
+missing particles or energy.
+On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells
+that have an excess of particles or energy. In contrast, $g^{-}_{J}$ represents
+the value of the deviation on cells that have missing particles or energy.
+
+Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles
+and/or energy and so, $g^{+}_{nJ}>0$. On the other hand, $\dot{g}^{-}_{nJ}>0$
+describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
+
 We sort the previous ideas in the following form
+%
 \begin{eqnarray}
@@
    &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
-\end{eqnarray}{}
-\\
+\end{eqnarray}
+%
+
 and consequently, (\ref{cambioH5}) obtains the following form
+%
 \begin{equation}
     \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}{}
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{d\Ss}{dt}<0$.
-With this, we proved that any quantum ideal gas perturbed (out of equilibrium but not so far from it), evolves such that $\frac{d\Ss}{dt}<0$.\\
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$.
-Then, we can say that any ideal quantum gas perturbed always evolves to the equilibrium state, such as the classical $H$ theorem.
-Then, we obtained the quantum version of the Boltzmann's $H$ theorem including inhomogeneities in the system.\\
-We obtained some results from the Statistical Mechanic in Quantum Mechanics scheme.
-However, the correspondence principle is important to recover all results from Classical Mechanics.
-This correspondence principle will be discussed in the next section.
+\end{equation}
+%
+and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then
+$\frac{d\Ss}{dt}<0$. With this, we proved that any quantum ideal gas perturbed
+(out of equilibrium but not so far from it), evolves such that
+$\frac{d\Ss}{dt}<0$.
+
+When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot
+g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$. Then, we can
+say that any ideal quantum gas perturbed always evolves to the equilibrium
+state, such as the classical $H$ theorem. Then, we obtained the quantum version
+of the Boltzmann's $H$ theorem including inhomogeneities in the system.
+
+We obtained some results from the Statistical Mechanic in Quantum Mechanics
+scheme. However, the correspondence principle is important to recover all
+results from Classical Mechanics. This correspondence principle will be
+discussed in the next section.
 

### 2020-12-16 tex/classical.tex
Typos, 1st intro paragraph, secs. 2.0, 2.1

@@
 
-\section{Classical scheme}
+\section{Classical scheme}\label{sec:classical}
+
+Classical statistical mechanics is theoretically founded on the Boltzmann transport
+equation and the hypo of molecular chaos. The fundamental connection
+between the microscopic nature of matter and the phenomenological laws of thermodynamics
+is the $H$-theorem.
 
@@
 
-In the classical version, the Boltzmann transport equation is required to show
-the $H$-theorem. That equation is
+From the Boltzmann kinetic theory of gases, the probability for a particle to be in a given
+state of position and velocity (\textit{aka} the distribution function) must satify
+the transport equation (Eq.~(\ref{eq:transport}), below).
+This equation is traditionally obtained by exploiting the volume-invariance of the $\mu$-space
+(\textit{i.e.} the space formed by the direct product of position and velocity)
+and considering that elastic collisions constitute the predominant mechanism
+for exchanging energy among particles \cite{bib:huang}.
 %
 \begin{equation}\label{eq:transport}
+	%\left(\frac{\partial f_1}{\partial t}\right)_{\mathrm{coll}}=
   \left(
-    \frac{\partial}{\partial t}+\vec{v}_1 \cdot\nabla_r
-    +\frac{\vec{F}}{m}\cdot\nabla_{v_1}
+    \frac{\partial}{\partial t}+\vec{v}_1 \cdot\nabla_{\vec r}
+    +\frac{\vec{F}}{m}\cdot\nabla_{\vec v_1}
   \right)f_1=\int\mathrm{d}\Omega\int
-    \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1),
+    \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1).
 \end{equation}
 %
-where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the
-external force applied to the system, $f_1$, $f_2$ are the distribution
-functions of the particles 1 and 2 respectively before the collision,
-and $f'_1$, $f'_2$ are the distribution function of the particles 1 and 2
-respectively after the collision.
+In Eq. (\ref{eq:transport}), $\Omega$ is the solid angle, $\sigma$ is the
+scattering cross section, $\vec F$ the
+external force applied to the system, and $f_1$ and $f_2$ ($f'_1$ and $f'_2$) are the distribution
+functions of particles 1 and 2, respectively, before (after) the collision.
 
-On the one hand, dynamics, inhomogeneities, and external forces are described
-by the left-hand side of the equation (\ref{eq:transport}). On the other hand, the
-binary collisions among particles are involved in the right-hand side of the
-previous equation. However, the \textit{molecular chaos hypothesis} was applied
-to obtain the right-hand side of (\ref{eq:transport}). The \textit{molecular chaos
-hypothesis} establishes that spatial and velocity coordinates of the particles
-are not correlated.
+Particle dynamics and external forces are described
+by the left-hand side of Eq.~(\ref{eq:transport}), and the right-hand side
+is derived considering binary collisions between particles and
+the \textit{molecular chaos hypothesis}, \textit{i.e.} it is assumed that
+the coordinates and velocities of the particles are not time-correlated.
 
@@
 \subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
+The $H$-theorem can be proved using the $H$ functional defined by Boltzmann 
 %
@@
 
-The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
+The $H$-theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
 transport equation and the molecular chaos hypothesis, then
@@
 as the second law of thermodynamics does.\footnote{The molecular hypothesis is %
-fundamental to determine the physical meaning of the $H$ theorem. For more %
+fundamental to determine the physical meaning of the $H$-theorem. For more %
 information, see \cite{bib:huang}.}
@@
 functional with a non-homogeneous distribution function in order to generalize
-the Boltzmann $H$ theorem. Because of this, we introduce a modified $H$
+the Boltzmann $H$-theorem. Because of this, we introduce a modified $H$
 functional (denoted as $H'$) defined as
@@
 
-To prove the Boltzmann's $H$ theorem (using (\ref{eq:CH2}) as functional), we need
+To prove the Boltzmann's $H$-theorem (using (\ref{eq:CH2}) as functional), we need
 to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the
@@
 %
-This statement will be the classical $H$ theorem with inhomogeneities.
+This statement will be the classical $H$-theorem with inhomogeneities.
 
 In the next section, we present a proposal quantum version of the
-$H$ theorem defined by Tolman and our proposal quantum version of the $H$
+$H$-theorem defined by Tolman and our proposal quantum version of the $H$
 theorem applying the method of the volume divided into cells.

### 2020-12-16 tex/disscussion.tex
Typos, 1st intro paragraph, secs. 2.0, 2.1

@@
-\section{Discussions and remarks}
+\section{Discussions and remarks}\label{sec:disscussion}
 
-In the literature, the Boltzmann's $H$ theorem is proved using a spatially homogeneous distribution function.
+In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function.
 In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
 However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$ theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$ theorem without using the Boltzmann's transport equation.
+Moreover, the Boltzmann's $H$-theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
+We can include the fact that we proved the Boltzmann's $H$-theorem without using the Boltzmann's transport equation.
 It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
@@
 Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$ theorem.\\
+This is a consequence of the $H$-theorem.\\
 \\
-In the quantum version, we used the same assumptions to prove the $H$ theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$ theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
+In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
+With the proof of the quantum version of the $H$-theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
 In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
@@
 Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$ theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
+According to the classical analysis of the $H$-theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
 In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.

### 2020-12-16 tex/introduction.tex
Typos, 1st intro paragraph, secs. 2.0, 2.1

@@
 Furthermore, the evolution of an equilibrium state for any system was proved by
-means the Boltzmann's $H$ theorem (or classical $H$ theorem) and the
+means the Boltzmann's $H$-theorem (or classical $H$-theorem) and the
 spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$ theorem and the
+However, there are some aspects in the classical $H$-theorem and the
 Boltzmann's $H$ functional that deserve some review to accomplish better
-consistency, for example, the modification of the $H$ theorem in stochastic
+consistency, for example, the modification of the $H$-theorem in stochastic
 trajectories, violations in the second law of thermodynamics, and the relation
@@
 To obtain the well-known correspondence principle between classical mechanics
-and quantum mechanics, a quantum version of the Boltzmann's $H$ theorem (or
-quantum $H$ theorem) and the quantum analogous to the Boltzmann transport
+and quantum mechanics, a quantum version of the Boltzmann's $H$-theorem (or
+quantum $H$-theorem) and the quantum analogous to the Boltzmann transport
 equation is necessary such that in the appropriate limit, the  classical $H$
@@
 \cite{bib:tolman}. He proposes one of the first quantum versions of the
-Boltzmann's $H$ theorem using the probability transition relation, the random
+Boltzmann's $H$-theorem using the probability transition relation, the random
 phases hypothesis, and a homogeneous distribution function. The equation of
-motion for the occupation number (necessary to prove the quantum $H$ theorem in
+motion for the occupation number (necessary to prove the quantum $H$-theorem in
 the same way that Boltzmann's procedure) was obtained by applying the time
@@
 Besides, some authors construct the $H$ functional and develop the proof of the
-quantum $H$ theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
+quantum $H$-theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
 However, the homogeneous distribution function hypothesis and the
 correspondence principle are uncleared or not discussed. Moreover, the validity
-of the quantum $H$ theorem and the second law of thermodynamics, and the
+of the quantum $H$-theorem and the second law of thermodynamics, and the
 interpretation of the quantum entropy is still in discussion %[19-26].
@@
 In order to contribute to the construction of a consistent classical and
-quantum $H$ theorem, and a new formalism to describe out of equilibrium
+quantum $H$-theorem, and a new formalism to describe out of equilibrium
 systems, we develop a new theoretical basis to describe them. In particular, we
@@
 theorem. Besides, we are interested in developing the quantum analogous of
-Boltzmann's $H$ theorem considering a non-spatial-homogeneous distribution
+Boltzmann's $H$-theorem considering a non-spatial-homogeneous distribution
 function and solve the problem of the correspondence principle mentioned above.
 
-The first section of this work consists of a review of the
-Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann
-distribution, and our alternative form to prove the classical $H$ theorem. The
-second part consists of a review of the proposal from Tolman about the quantum
-version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions,
-and our alternative form to prove the quantum version of the $H$ theorem. The
-third section consists of a classical-quantum correspondence between the
-quantum $H$ functional and the classical $H$ functional. Finally, the last
-section consists of a proposal to explore relaxation processes in the quantum
-case obtaining the analogous quantum Boltzmann transport equation.
+The organization of this article is as follows.
+In Section~\ref{sec:classical}, we highlight, to our purposes, the most
+fundamental assumptions required to proof the Boltzmann $H$-theorem,
+we provide an alternative method to obtain the Maxwell-Boltzmann
+distribution using the variational method, and an alternative demonstration of
+the $H$-theorem for classical systems. In Section~\ref{sec:quantum}
+we review the Tolman proposal for the quantum
+version of the $H$-theorem (quantum $H$-theorem)
+\commr{and how the Bose-Einstein and Fermi-Dirac distributions are treated within
+this framework},
+and our alternative proof of the quantum $H$-theorem. In 
+Section~\ref{sec:qccorrespondence}
+the classical-quantum correspondence between the
+quantum $H$ functional and the classical $H$ functional is analyzed.
+In Section~\ref{sec:relaxproc} we explore how relaxation processes occur
+in a quantum ideal gas; here we provide a time-evolution equation that could be
+considered the analogous of the classical Boltzmann transport equation. Finally,
+we discuss some key ideas derived from our approach in Section~\ref{sec:disscussion},
+and provide some conclusions in Section~\ref{sec:conclusions}.

### 2020-12-16 tex/quantum.tex
ssh test
Typos, 1st intro paragraph, secs. 2.0, 2.1

@@
-\section{Quantum scheme}
+\section{Quantum scheme}\label{sec:quantum}
 
@@
 %
-\begin{equation}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, \label{quantumh}
+\begin{equation}\label{quantumh}
+    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, 
 \end{equation}
@@
 
-%\subsection{The quantum analogous of the Boltzmann's $H$ theorem}
+%\subsection{The quantum analogous of the Boltzmann's $H$-theorem}
 We rewrite the \textit{variational entropy} (\ref{entropy}) as
@@
 say that any ideal quantum gas perturbed always evolves to the equilibrium
-state, such as the classical $H$ theorem. Then, we obtained the quantum version
-of the Boltzmann's $H$ theorem including inhomogeneities in the system.
+state, such as the classical $H$-theorem. Then, we obtained the quantum version
+of the Boltzmann's $H$-theorem including inhomogeneities in the system.
 

### 2020-12-16 tex/relaxproc.tex
Typos, 1st intro paragraph, secs. 2.0, 2.1

@@
-  \section{Relaxation processes in degenerated quantum gases}
+  \section{Relaxation processes in degenerated quantum gases}\label{sec:relaxproc}
 Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow

### 2020-12-17 tex/classical.tex
Reworded sec. 2.2

@@
 From the Boltzmann kinetic theory of gases, the probability for a particle to be in a given
-state of position and velocity (\textit{aka} the distribution function) must satify
+state of position and velocity at a time $t$,
+$f(\vec r,\vec v,t)$ (\textit{aka} the distribution function), must satify
 the transport equation (Eq.~(\ref{eq:transport}), below).
@@
 \subsection{H-theorem and the Maxwell-Boltzmann distribution}
-The $H$-theorem can be proved using the $H$ functional defined by Boltzmann 
+
+The evolution of a dilute gas towards thermodynamic equilibrium is frequently
+addressed by first defining the functional \cite{bib:tolman,bib:huang}
 %
 \begin{equation}\label{eq:hfunctional}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,
+   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
 \end{equation} 
 %
-where $f(\vec{v},t)$ is a homogeneous distribution function.
-This $H$ functional describes a system with a global energy $E$, a global
-number of free classical particles $N$, global volume $V$, and a temperature $T$.
-Besides, the Boltzmann's $H$ functional correspond to a density of entropy
-when the system arrives at the equilibrium state.
+Here, $f(\vec{v},t)$ is a homogeneous distribution function, \textit{i.e.}
+$f$ is assumed to be independent of $\vec r$.
+The functional $H$, originally introduced by Boltzmann in 1872,
+describes a diluted gas, at temperature $T$, with a total energy $E$, a total
+number of free classical particles $N$, and total volume $V$.
 
-The $H$-theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
-transport equation and the molecular chaos hypothesis, then
-$dH/dt\leq0$ and the system arrives to the equilibrium state when
-$dH/dt=0$.
-The theorem can be proved by applying the time derivative
-to Eq. (\ref{eq:hfunctional}) and using the Boltzmann transport equation
-(\ref{eq:transport}). As a consequence of this theorem, the distribution function
-that describes the system in equilibrium corresponds to the Maxwell-Boltzmann
-distribution. This distribution can be obtained by solving a sufficient
-condition for $f$ to obtain the equilibrium distribution function $f_0$:
+Subsequently, the time-evolution of the gas towards equilibrium is obtained through the
+$H$-theorem. This theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
+transport equation and the molecular chaos hypothesis is valid, then
+$dH/dt\leq0$ in general, and if $dH/dt=0$, then the system is in the equilibrium state.
+The $H$-theorem is proven by taking the time derivative
+of Eq. (\ref{eq:hfunctional}) and using the Boltzmann transport equation
+(\ref{eq:transport}) \cite{bib:huang}. The $H$-theorem fixes the direction of time-evolution
+of systems in nature, which is consistent with the second law of thermodynamics; in fact,
+$H$ can be associated with an entropy density.
+
+On the other hand, considering a diluted gas in equilibrium (\textit{i.e.}
+$(\partial f/\partial t)=0$),
+with no applied external forces (\textit{i.e.} $f$ is independent of $\vec r$), 
+it can be proven that the equilibrium distribution function obtained from
+Eq.~(\ref{eq:transport}) is precisely the Maxwell-Boltzmann
+distribution function. The proof of the above commences by finding the sufficient
+condition for $f$ to render a null r.h.s. of Eq. (\ref{eq:transport}). Such an
+$f$, which we denote here as $f_0$, must satisfy:
 %
-\begin{equation}
+\begin{equation}\label{eq:neccondf0}
     f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.
@@
 %
+Subsequently, the Maxwell-Boltzmann distrubution function can be obtained by
+taking the logarithm to Eq.~(\ref{eq:neccondf0}), and exploiting mechanical
+conserved quantities (see \cite[ch. 4.2]{bib:huang}).
 
-This condition can be obtained from the binary collision term of the Boltzmann
-transport equation (\ref{eq:transport}).
-
-The $H$-theorem fixes the time evolution direction of systems in nature, such
-as the second law of thermodynamics does.\footnote{The molecular hypothesis is %
-fundamental to determine the physical meaning of the $H$-theorem. For more %
-information, see \cite{bib:huang}.}
+We emphasize that the validity of the $H$-theorem relies significantly on assuming
+that the distribution function is homogeneous and that the hypothesis of
+molecular chaos is surmised.
 

### 2020-12-18 tex/classical.tex
Rewording: start of section 2.3

@@
 
-We emphasize that the validity of the $H$-theorem relies significantly on assuming
-that the distribution function is homogeneous and that the hypothesis of
-molecular chaos is surmised.
-
 %---------------------------
@@
 
-The Boltzmann $H$-functional (\ref{eq:hfunctional}) is defined with a spatially
-homogeneous distribution $f(\vec{v},t)$. However, we want to consider a $H$
-functional with a non-homogeneous distribution function in order to generalize
-the Boltzmann $H$-theorem. Because of this, we introduce a modified $H$
-functional (denoted as $H'$) defined as
+As we saw in the previous section,
+the validity of the $H$-theorem relies significantly on assuming
+that the distribution function is homogeneous and that the hypothesis of
+molecular chaos is surmised. However, in order to consider systems whose
+distribution functions are not necessarily homogeneous, which
+allows for studying systems out of equilibrium,
+we introduce a modified $H$ functional as follows.
+
+The proposed functional, denoted as $H'$, describes a diluted gas occupying a volume, $V$,
+divided into $K$ cells, which, without loss of generality,
+have local identical volumes (\textit{i.e.} $\delta V_M = V/K,\ M=1,\,\dots\,,K$).
+Each cell has its own local
+$H$ functional, $H'_M$, and local homogeneous distribution function,
+$f_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
+local temperature, $T_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
+global energy $E$, and a global number of free
+classical particles $N$. The functional takes the form
 %
@@
 %
-This $H'$ describes a system with a global energy $E$, a global number of free
-classical particles $N$, global volume $V$, and a temperature $T$. Nevertheless,
-the global system, in turn, consists of a set of $K$ small cells with a constant
-volume $\delta V_M$, whose local homogeneous distribution functions is
-$f_{M}(\vec{v},t)$.
 
-The distribution function will depend on the position of the cells, the velocity
-$\vec{v}$, and time $t$. Without loss of generality, we choose these cells to
-have local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$).
-Each cell has its local chemical potential, $\mu_M$, local
-$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local
-energy, $\mathcal{E}'_M$. In terms of the local distribution 
-function and the energy $\epsilon(\vec{v})$, the last three properties are
-given by:
+The distribution functions, $\{f_M(\vec{v},t)\}$, depend implicitly on the position of the cells,
+relative to the global system, and on the velocity
+$\vec{v}$ and time $t$. In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
 %
+\begin{subequations}
 \begin{eqnarray}
     H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
-      \mathrm{d}^{3}v \label{Hcell},\nonumber\\
-    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber\\
+      \mathrm{d}^{3}v \label{Hcell},\\
+    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\\
     \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
 \end{eqnarray}
+\end{subequations}
+%
+Notice that assuming local homogeneous distribution functions means that we are surmising the 
+local equilibrium hypothesis. In addition, $f_{M}(\vec{v},t)$ must satisfy the following
+microcanonical restrictions:
+%
+\begin{subequations}\label{eq:micro}
+\begin{equation}\label{eq:micron}
+    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N
+\end{equation}
+and
+\begin{equation}\label{eq:microe}
+    \sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
+\end{equation}
+\end{subequations}
 %
 
-When the system is in equilibrium, the local number of particles and the local
+If the complete system is in equilibrium, then the local number
+of particles and the local
 energy do not depend on the cell number, this is
 %
+\begin{subequations}\label{eq:eqNandE}
 \begin{equation}
-   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}';
+   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'
+\end{equation}
+%
+and
+%
+\begin{equation}
    \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
 \end{equation}
+\end{subequations}
 %
-The particles in the system are considered to be free, \textit{i.e.} the
-particles do not interact with each other. Besides, the local distribution
-function must satisfy the following microcanonical restrictions 
-%
-\begin{equation}\label{micro}
-    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N,
-    \quad\sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
-\end{equation}
-%
+In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
 

### 2020-12-19 tex/classical.tex
Rewording: up to p. 4 of sec. 2.3

@@
 describes a diluted gas, at temperature $T$, with a total energy $E$, a total
-number of free classical particles $N$, and total volume $V$.
+number of free classical particles $N$, and occupying a volume $V$.
 
@@
 %
-\begin{subequations}
+\begin{subequations}\label{eq:cellrestrictions}
 \begin{eqnarray}
@@
 
-We can see that the functional (\ref{eq:CH2}) contains a non-homogeneous
-distribution function but keeps the same form of the Boltzmann's $H$ functional.
-
-Inspired by the Hamilton principle, $H'$ will be maximized with those
-restrictions to obtain the equilibrium distribution function for a classical gas
-using the Lagrange multipliers method obtaining
+We now use the variational method in order to maximize $H'$, applying the restrictions
+given in Eqs.~(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) through Lagrange
+multipliers ($\{\alpha_M\}$ and $\{\beta_M\}$). This yields
 %
@@
 %
-Isolating the distribution function 
+Solving the last line for $f_J(\vec{v}')$ renders
 %
+\begin{comment}
 \begin{eqnarray}
@@
     & = & e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1}\nonumber\\
-    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},
+    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{eq:relacion1},
 \end{eqnarray}
+\end{comment}
+\begin{equation}\label{eq:relacion1}
+	f_J(\vec{v}') = Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') }
+\end{equation}
 %
-where $C$ is a constant. It is trivial that using the variational procedures we
-obtain the form of the Maxwell-Boltzmann distribution.
+where $C$ is a constant. We notice that using the variational procedure on
+$H'$ predicts that the
+distribution function of each cell has the form of the Maxwell-Boltzmann
+distribution function.
 
@@
 %
-using (\ref{relacion1}) with homogeneous multipliers
+using (\ref{eq:relacion1}) with homogeneous multipliers
 %

### 2020-12-21 tex/classical.tex
Rewording: up tp sec. 2.3.1

@@
 %-------------------------
-\subsection{H-theorem and the Maxwell-Boltzmann distribution}
+\subsection{A brief summary of the H-theorem and the Maxwell-Boltzmann distribution}
 
@@
 %
-\begin{equation}\label{eq:hfunctional}
-   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
+\begin{equation}\label{eq:hbfunctional}
+   H_{B}=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
 \end{equation} 
@@
 $f$ is assumed to be independent of $\vec r$.
-The functional $H$, originally introduced by Boltzmann in 1872,
+The functional $H_B$,
+originally introduced by Boltzmann in 1872,
 describes a diluted gas, at temperature $T$, with a total energy $E$, a total
 number of free classical particles $N$, and occupying a volume $V$.
+(We deviate from the custommary notation in order to
+clearly distinguish the Boltzmann functional $H$ from our proposals, and
+$f_B$ to denote the Maxwell-Boltzmann distribution function.)
 
-Subsequently, the time-evolution of the gas towards equilibrium is obtained through the
+Under this scheme, the time-evolution of the gas towards equilibrium is obtained through the
 $H$-theorem. This theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
 transport equation and the molecular chaos hypothesis is valid, then
-$dH/dt\leq0$ in general, and if $dH/dt=0$, then the system is in the equilibrium state.
+$dH_B/dt\leq0$ in general, and if $dH_B/dt=0$, then the system is in the equilibrium state.
 The $H$-theorem is proven by taking the time derivative
-of Eq. (\ref{eq:hfunctional}) and using the Boltzmann transport equation
+of Eq. (\ref{eq:hbfunctional}) and using the Boltzmann transport equation
 (\ref{eq:transport}) \cite{bib:huang}. The $H$-theorem fixes the direction of time-evolution
 of systems in nature, which is consistent with the second law of thermodynamics; in fact,
-$H$ can be associated with an entropy density.
+$H_B$ can be associated with an entropy density.
 
@@
 %---------------------------
-\subsection{Out of equilibrium, non-homogeneous distributions}
+\subsection{Alternative classical $H$-functional}
 
@@
 allows for studying systems out of equilibrium,
-we introduce a modified $H$ functional as follows.
+we introduce a modified $H$-functional as follows. Hereafter, we will
+use the term homogeneous distribution function to refer to a distribution
+function thad does not depend on the coordinates.
 
-The proposed functional, denoted as $H'$, describes a diluted gas occupying a volume, $V$,
+The proposed functional, denoted as $H'$, describes a diluted classical
+gas occupying a volume, $V$,
 divided into $K$ cells, which, without loss of generality,
@@
 Each cell has its own local
-$H$ functional, $H'_M$, and local homogeneous distribution function,
+$H$-functional, $H'_M$, and local homogeneous distribution function,
 $f_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
@@
 global energy $E$, and a global number of free
-classical particles $N$. The functional takes the form
+classical particles $N$, and it is considered to be perfectly isolated (\textit{i.e.} the
+global system is a microcanonical ensemble). We also assume that the number of particles
+in each cell is sufficiently large, so as to obtain accurate averages.
+The functional takes the form
 %
 \begin{equation}\label{eq:CH2}
-   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
+   H'(t)=\sum_{M=1}^{K}\int_{\delta V_M} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
 \end{equation}
@@
 relative to the global system, and on the velocity
-$\vec{v}$ and time $t$. In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
+$\vec{v}$ and time $t$. Each $f_M$ can be formally extended to the complete coordinate space
+by defining each $f_M$ to be zero outside the $M$-th cell, in such a manner that
+the distribution function of the complete system is a piece-wise sum
+of $\{f_M(\vec v,t)\}$:
+%
+\begin{equation}
+   f(\vec r,\vec v,t)=\sum_{M=1}^Kf(\vec r_M,\vec v,t)=\sum_{M=1}^Kf_M(\vec v,t).
+\end{equation}
+%
+Here $\vec r_M$ is the center of the cell of index $M$. This extended definition
+allows us to omit the symbol $\delta V_M$ on all integrals performed over the cell volume.
+In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
 %
@@
     H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
-      \mathrm{d}^{3}v \label{Hcell},\\
-    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\\
-    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
+      \mathrm{d}^{3}v \label{eq:Hcell},\\
+    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\label{eq:Ncell}\\
+    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v\label{eq:Ecell}.
 \end{eqnarray}
@@
 %
-Notice that assuming local homogeneous distribution functions means that we are surmising the 
-local equilibrium hypothesis. In addition, $f_{M}(\vec{v},t)$ must satisfy the following
+Notice that assuming homogeneous distribution functions per cell means that we are surmising the 
+local equilibrium hypothesis. In addition, the set $\{f_{M}(\vec{v},t)\}$ must satisfy the following
 microcanonical restrictions:
@@
 %
-
-If the complete system is in equilibrium, then the local number
-of particles and the local
-energy do not depend on the cell number, this is
-%
-\begin{subequations}\label{eq:eqNandE}
-\begin{equation}
-   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'
-\end{equation}
-%
-and
-%
-\begin{equation}
-   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
-\end{equation}
-\end{subequations}
-%
-In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
 
@@
 \begin{equation}\label{eq:relacion1}
-	f_J(\vec{v}') = Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') }
+	f_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
 \end{equation}
@@
 distribution function of each cell has the form of the Maxwell-Boltzmann
-distribution function.
+distribution function, which is consistent with the local equilibrium assumption.
 
-As $H'$ in equilibrium is proportional to entropy, it is trivial to think that
-Lagrange multipliers do not depend on the position because of the system is in
-equilibrium and consequently, the distribution function is homogeneous, that is
+\subsubsection{Properties of $H'$ for systems in equilibrium}
+
+If the complete system is in equilibrium without external forces applied to
+the gas, from classical thermodynamics of
+systems in equilibrium, we ascertain that the local number
+of particles and the local
+energy do not depend on the cell number, this is
+%
+\begin{subequations}\label{eq:eqNandE}
+\begin{equation}
+   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'
+\end{equation}
+%
+and
 %
 \begin{equation}
-    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
+   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
+\end{equation}
+\end{subequations}
+%
+In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
+Moreover, the distribution function is homogeneous, hence $f_M$
+does not depend on the cell number $M$ (\textit{i.e.} $f_M(\vec v,t)=f(\vec v,t)$,
+$\forall M$). Several properties follow directly from this, \textit{e.g.}
+from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$ (here we have used Eqs. (\ref{eq:Ecell})
+and (\ref{eq:Ncell})),
+and Eq.~(\ref{eq:CH2}), in terms of Eq.~(\ref{eq:hbfunctional}), reduces to
+%
+\begin{equation}
+    H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v
+         = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{B}(t).
 \end{equation}
 %
-The constant $C$ can be omitted defining the following $H$ functional
+In order to identify $H'$ with the entropy, it remains to show that $H'(t)$ is extensive,
+with respect to $f'(\vec v,t)\big(\equiv Kf_(\vec v,t)\big)$.
+This is shown by analyzing the following expression
+%
+\begin{eqnarray}\label{eq:sum-h}
+    \int Kf(\vec{v},t) \ln \big[Kf(\vec{v},t)\big] \mathrm{d}^{3}v & = & 
+        \int \big[(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)\big]\mathrm{d}^{3}v\nonumber\\
+        & = & K\int\Big\{ f(\vec{v},t)\big[\ln K + \ln f(\vec{v},t)\big] \Big\}\mathrm{d}^{3}v.
+\end{eqnarray}
+%
+We observe that if the number of particles in the $\mu$-space, $f(\vec{v},t)$, is much greater than
+the number of cells $K$, then the first term of Eq.~(\ref{eq:sum-h}) is negligible, and consequently:
+%
+\begin{equation}\label{aditive-property}
+    \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v \simeq
+    K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,
+\end{equation}
+%
+\textit{i.e.} $H_M'$ is extensive, and the sum $\sum_MH'_M$ is the $H$-functional of the complete
+system and reduces to the Boltzmann $H$-functional. Hence, $H'$ can also be identified with the
+system's entropy.
+
+Furthermore, in equilibrium the Lagrange multipliers are position- and time-independent,
+thus $f_M(\vec{v})$
+reduces to
+%
+\begin{equation}
+    f_M(\vec{v})=C\exp\left(\alpha+\beta \epsilon(\vec{v})\right)
+    \equiv \bar{f}(\vec{v}),\quad M=1,\,\dots\,,K.
+\end{equation} 
+%
+The constant $C$ can be omitted, which is shown by defining the following $H''$
+functional:
 %
@@
 %
-but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the
-total particle number. As $N$ is a constant and any $H$ functional has sense
-when we calculate its time derivative, this constant can be omitted.
+however $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$.
+Since $N$ is a constant and we are mainly intersted in the time-derivative of $H''$,
+$C$ can be conveniently omitted. In other words, $H'$ renders the Maxwell-Boltzmann
+distribution function for systems in equilibrium.
 
-On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{eq:CH2}) as
-%
-\begin{equation}
-    H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v
-         = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
-\end{equation}
-%
-where
-%
-\begin{equation}
-    H_{boltz}(t)=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
-\end{equation}
-%
-Besides, if we analyze the following expression
-%
-\begin{equation}
-    \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
-\end{equation}
-%
-we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
-%
-\begin{equation}
-    \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
-\end{equation}
-%
-We obtained the additive property when we supposed the distribution function is
-spatial-homogeneous.
+\subsubsection{Alternative $H$-theorem}
 

### 2020-12-21 tex/conclusions.tex
Typos, 1st intro paragraph, secs. 2.0, 2.1
Rewording: up tp sec. 2.3.1

@@
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 We conclude the following statements:\\
@@
     \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
-    \item We can obtain the transport equation from the $H$ functional.
-    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
+    \item We can obtain the transport equation from the $H$-functional.
+    \item The variational entropy is proportional to the Boltzmann's $H$-functional in the limit of the high energy levels.
     \item We find an expression for the entropy to quantum gases.

### 2020-12-21 tex/disscussion.tex
Rewording: start of section 2.3
Rewording: up tp sec. 2.3.1

@@
+%--------------------------------------
+% Me parece que sera necesario incluir la discusin sobre los procesos de relajacin.
+%Sin esa parte el contenido original del artculo podria parecer insuficiente para
+%una buena publicacion%
+
 \section{Discussions and remarks}\label{sec:disscussion}
@@
 It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium.
+Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$-functional through the variational method.
+Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$-functional is not zero when the system is out of equilibrium.
 This is a consequence of the $H$-theorem.\\
@@
 \\
-We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels.
+We also find that variational entropy is proportional to the Boltzmann's $H$-functional in the limit of high energy levels.
 This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium
+Moreover, if we see the expression for the Boltzmann's $H$-functional in equilibrium
 \begin{equation}
-    H_{Boltz}=-\frac{S}{Vk},
+    H_{B}=-\frac{S}{Vk},
 \end{equation}

### 2020-12-21 tex/introduction.tex
Rewording: up tp sec. 2.3.1

@@
 However, there are some aspects in the classical $H$-theorem and the
-Boltzmann's $H$ functional that deserve some review to accomplish better
+Boltzmann's $H$-functional that deserve some review to accomplish better
 consistency, for example, the modification of the $H$-theorem in stochastic
@@
 \cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. %[12-15]
-Besides, some authors construct the $H$ functional and develop the proof of the
+Besides, some authors construct the $H$-functional and develop the proof of the
 quantum $H$-theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
@@
 the classical-quantum correspondence between the
-quantum $H$ functional and the classical $H$ functional is analyzed.
+quantum $H$-functional and the classical $H$-functional is analyzed.
 In Section~\ref{sec:relaxproc} we explore how relaxation processes occur

### 2020-12-21 tex/qccorrespondence.tex
quantum.tex reformatted: 80chars or less
Typos, 1st intro paragraph, secs. 2.0, 2.1
Rewording: start of section 2.3
Rewording: up tp sec. 2.3.1

@@
-\section{Quantum Classical correspondence}
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\
+\section{Quantum Classical correspondence}\label{sec:qccorrespondence}
+%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$-functional in the appropriated limit.\\
 We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
@@
     \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}
+    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{B},\label{h-quantic4}
 \end{eqnarray}
-where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\
+where $H_{B}$ now is (\ref{eq:CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{bib:tolman} eq. (102.6) for more information.}\\
 We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
@@
 \end{equation}
-that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\
+that corresponds to the classical $H$-functional in equilibrium described in the energy space.}\\
 \textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}

### 2020-12-21 tex/quantum.tex
Rewording: up tp sec. 2.3.1

@@
 
-The $H$ functional in quantum mechanics presented by Tolman \cite{bib:tolman}
+The $H$-functional in quantum mechanics presented by Tolman \cite{bib:tolman}
 is defined as
@@
 considered as the quantum analogous to the molecular chaos hypothesis
-\cite{bib:das2018}}, then the $H$ functional (\ref{quantumh}) satisfies that
+\cite{bib:das2018}}, then the $H$-functional (\ref{quantumh}) satisfies that
 %
@@
 %
-This expression also can be obtained also from the Boltzmann $H$ functional through defining
+This expression also can be obtained also from the Boltzmann $H$-functional through defining
 %
@@
 %
-where we used that the $H$ functional in equilibrium is proportional to the
+where we used that the $H$-functional in equilibrium is proportional to the
 entropy of the system.

### 2020-12-21 tex/classical.tex
Rewording: fist part of sec. 2.3.2

@@
 %
-Here $\vec r_M$ is the center of the cell of index $M$. This extended definition
+Here $\vec r_M$ is the center of the cell of index $M$, and $f_M(\vec v,t)\neq f_N(\vec v,t)$
+for $M\neq N$. This extended definition
 allows us to omit the symbol $\delta V_M$ on all integrals performed over the cell volume.
@@
 \textit{i.e.} $H_M'$ is extensive, and the sum $\sum_MH'_M$ is the $H$-functional of the complete
-system and reduces to the Boltzmann $H$-functional. Hence, $H'$ can also be identified with the
+system, which reduces to the Boltzmann $H$-functional. Hence, $H'$ can also be identified with the
 system's entropy.
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:fbardef}
     f_M(\vec{v})=C\exp\left(\alpha+\beta \epsilon(\vec{v})\right)
@@
 
-\subsubsection{Alternative $H$-theorem}
+\subsubsection{Alternative proof of the classical $H$-theorem}
 
-To prove the Boltzmann's $H$-theorem (using (\ref{eq:CH2}) as functional), we need
-to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the
-following assumptions:
+Throughout this section, we will consider that the classical gas
+starts at some time as an out-of-equilibrium system,
+but in such a manner that it is inteuitively not too far from equilibrium,
+which ensures that the local equilibrium hypothesis remains valid during
+the time evolution of the system.
+Also we will use the following definitions:
+\begin{subequations}\label{eq:restrictionoutsideclassical}
 %
-\begin{itemize}
-  \item The local equilibrium hypothesis in each cell ($f_{M}$ must be
-    spatial-homogeneous for each $M$).
-  \item The non-homogeneous distribution function assumption in the total
-    volume ($f_{M}$ must be different among all cells).
-  \item The following restrictions
-    %
-    \begin{equation}\label{restrictionoutsideclassical}
-      \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=
-      \bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ 
-      \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=
-      \bar{\mathcal{E}}+ \delta_M(t),
+\begin{equation}
+      \mathcal{N}'_M(t)=\int f'_{M}(\vec{v},t) \mathrm{d}^{3}v=
+      \bar{\mathcal{N}}'+\Delta'_M(t)
 \end{equation}
 %
-    where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are
-    the local particle number and the local energy in equilibrium
+and
 %
 \begin{equation}
-      \bar{\mathcal{N}}=
-      \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \quad
-      \bar{\mathcal{E}}=
-      \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
+      \mathcal{E}'_M(t)=\int f'_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=
+      \bar{\mathcal{E}}'+ \delta'_M(t).
+\end{equation}
+\end{subequations}
+%
+Here $\bar {\mathcal{N}}'\ (=N/K)$ and $\bar{\mathcal{E}}'\ (=E/K)$ are
+the local particle number and the local energy in equilibrium, respectively, which are given by:
+%
+    \begin{equation}
+      \bar{\mathcal{N}}'=
+      \int \bar{f}(\vec{v}) \mathrm{d}^{3}v\quad\textrm{and} \quad
+      \bar{\mathcal{E}}'=
+      \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v.
     \end{equation}
     %
-    and $\Delta_M$, $\delta_M$ could be seen as a deviation from
-    $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively,
-    with $\Delta_M(t)\ll\bar{\mathcal{N}}$ and
-    $\delta_M(t)\ll\bar{\mathcal{E}}$.
-\end{itemize} 
-%
-Those conditions described a system out of equilibrium but no so far from it,
-that is a perturbed system and will suffer a \textit{relaxation process, i.e.}
-The time evolution of a perturbated system to the equilibrium state.
+Here, we have used $\bar{f}(\vec{v})$ as defined in Eq.~(\ref{eq:fbardef}).
+In Eqs.~(\ref{eq:restrictionoutsideclassical}), $\Delta'_M$ and $\delta'_M$
+are considered deviations relative to
+$\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$, respectively, and for systems that are not
+too far away from equilbrium, it is reasonable to expect on the one hand that
+$\Delta'_M(t)\ll\bar{\mathcal{N}}'$ and
+$\delta'_M(t)\ll\bar{\mathcal{E}}'$, and on the other hand that $\Delta'_M$ and
+$\delta'_M$ are sufficiently large compared with fluctuations of $\bar{\mathcal{N}}'$
+and $\bar{\mathcal{E}}'$.
 
-It is important to remark that the quantities $\Delta_M$, $\delta_M$ are
-sufficiently big to be different from fluctuations in the system but too small
-such that the system is not far from the equilibrium state.
-
+\commr{CHECKPOINT}\\
 Performing the derivative of (\ref{eq:CH2}) with respect to time, we obtain

### 2020-12-22 tex/classical.tex
Reword: middle part of 2.3.2. Typos: primed classical quantities.

@@
   \right)f_1=\int\mathrm{d}\Omega\int
-    \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1).
+    \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(\tilde f_2\tilde f_1-f_2f_1).
 \end{equation}
@@
 scattering cross section, $\vec F$ the
-external force applied to the system, and $f_1$ and $f_2$ ($f'_1$ and $f'_2$) are the distribution
+external force applied to the system, and $f_1$ and $f_2$ ($\tilde f_1$ and $\tilde f_2$) are the distribution
 functions of particles 1 and 2, respectively, before (after) the collision.
@@
 use the term homogeneous distribution function to refer to a distribution
-function thad does not depend on the coordinates.
+function thad does not depend on the coordinates, and will use primed functions
+and quantities to differentiate the classical case (this Section) \textit{vs} the quantum
+case (below, Section \ref{sec:quantum}).
 
-The proposed functional, denoted as $H'$, describes a diluted classical
+Our proposed functional, denoted as $\mathcal{H}'$, describes a diluted classical
 gas occupying a volume, $V$,
@@
 Each cell has its own local
-$H$-functional, $H'_M$, and local homogeneous distribution function,
-$f_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
-local temperature, $T_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
+$H$-functional, $\mathcal{H}'_M$, and local homogeneous distribution function,
+$f'_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
+local temperature, $T'_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
 global energy $E$, and a global number of free
@@
 \begin{equation}\label{eq:CH2}
-   H'(t)=\sum_{M=1}^{K}\int_{\delta V_M} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
+   \mathcal{H}'(t)=\sum_{M=1}^{K}\int_{\delta V_M} f'_M(\vec{v},t) \ln f'_M(\vec{v},t)\mathrm{d}^3v.
 \end{equation}
@@
 
-The distribution functions, $\{f_M(\vec{v},t)\}$, depend implicitly on the position of the cells,
+The distribution functions, $\{f'_M(\vec{v},t)\}$, depend implicitly on the position of the cells,
 relative to the global system, and on the velocity
-$\vec{v}$ and time $t$. Each $f_M$ can be formally extended to the complete coordinate space
-by defining each $f_M$ to be zero outside the $M$-th cell, in such a manner that
+$\vec{v}$ and time $t$. Each $f'_M$ can be formally extended to the complete coordinate space
+by defining each $f'_M$ to be zero outside the $M$-th cell, in such a manner that
 the distribution function of the complete system is a piece-wise sum
-of $\{f_M(\vec v,t)\}$:
+of $\{f'_M(\vec v,t)\}$:
 %
 \begin{equation}
-   f(\vec r,\vec v,t)=\sum_{M=1}^Kf(\vec r_M,\vec v,t)=\sum_{M=1}^Kf_M(\vec v,t).
+   f'(\vec r,\vec v,t)=\sum_{M=1}^Kf'(\vec r_M,\vec v,t)=\sum_{M=1}^Kf'_M(\vec v,t).
 \end{equation}
 %
-Here $\vec r_M$ is the center of the cell of index $M$, and $f_M(\vec v,t)\neq f_N(\vec v,t)$
+Here $\vec r_M$ is the center of the cell of index $M$, and $f'_M(\vec v,t)\neq f'_N(\vec v,t)$
 for $M\neq N$. This extended definition
 allows us to omit the symbol $\delta V_M$ on all integrals performed over the cell volume.
-In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
+In terms of $f'_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
 %
@@
 \begin{eqnarray}
-    H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
+    H_M' & = &  \int f'_M(\vec{v},t) \ln f'_{M}(\vec{v},t)
       \mathrm{d}^{3}v \label{eq:Hcell},\\
-    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\label{eq:Ncell}\\
-    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v\label{eq:Ecell}.
+    \mathcal{N}_M' & = & \int f'_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\label{eq:Ncell}\\
+    \mathcal{E}_M' & = & \int f'_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v\label{eq:Ecell}.
 \end{eqnarray}
@@
 Notice that assuming homogeneous distribution functions per cell means that we are surmising the 
-local equilibrium hypothesis. In addition, the set $\{f_{M}(\vec{v},t)\}$ must satisfy the following
+local equilibrium hypothesis. In addition, the set $\{f'_{M}(\vec{v},t)\}$ must satisfy the following
 microcanonical restrictions:
@@
 \begin{equation}\label{eq:micron}
-    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N
+    \sum_{M=1}^{K}\int f'_M(\vec{v},t)\mathrm{d}^3v =N
 \end{equation}
@@
 \begin{equation}\label{eq:microe}
-    \sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
+    \sum_{M=1}^{K}\int f'_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
 \end{equation}
@@
 \begin{eqnarray}
-    \frac{\delta H'}{\delta f_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
-      \frac{\delta}{\delta f_J(\vec{v}')}\left[
-        f_M(\vec{v})\ln f_M(\vec{v})
+    \frac{\delta H'}{\delta f'_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
+      \frac{\delta}{\delta f'_J(\vec{v}')}\left[
+        f'_M(\vec{v})\ln f'_M(\vec{v})
         \right]
        \mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int
-       	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
+       	\frac{\delta f'_M(\vec{v})}{\delta f'_J(\vec{v}')}
       \mathrm{d}^3v\nonumber\\
     & & -\sum_{M=1}^{K}\beta_M\int\epsilon(\vec{v})
-    	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
+    	\frac{\delta f'_M(\vec{v})}{\delta f'_J(\vec{v}')}
       \mathrm{d}^3v \nonumber\\
-    & = & \ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
+    & = & \ln f'_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
 \end{eqnarray}
 %
-Solving the last line for $f_J(\vec{v}')$ renders
+Solving the last line for $f'_J(\vec{v}')$ renders
 %
@@
 \begin{eqnarray}
-    \ln f_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
+    \ln f'_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
        -1\quad\Rightarrow\quad f_J(\vec{v}')\\
@@
 \begin{equation}\label{eq:relacion1}
-	f_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
+	f'_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
 \end{equation}
@@
 In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
-Moreover, the distribution function is homogeneous, hence $f_M$
-does not depend on the cell number $M$ (\textit{i.e.} $f_M(\vec v,t)=f(\vec v,t)$,
+Moreover, the distribution function is homogeneous, hence $f'_M$
+does not depend on the cell number $M$ (\textit{i.e.} $f'_M(\vec v,t)=f'(\vec v,t)$,
 $\forall M$). Several properties follow directly from this, \textit{e.g.}
@@
 \begin{equation}
-    H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v
-         = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{B}(t).
+    H'(t)=\int \sum_{M=1}^{K} [f'(\vec{v},t)\ln f'(\vec{v},t)] \mathrm{d}^{3}v
+         = K\int  f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v= K H_{B}(t).
 \end{equation}
@@
 In order to identify $H'$ with the entropy, it remains to show that $H'(t)$ is extensive,
-with respect to $f'(\vec v,t)\big(\equiv Kf_(\vec v,t)\big)$.
+with respect to $Kf'_(\vec v,t)$.
 This is shown by analyzing the following expression
@@
 \begin{eqnarray}\label{eq:sum-h}
-    \int Kf(\vec{v},t) \ln \big[Kf(\vec{v},t)\big] \mathrm{d}^{3}v & = & 
-        \int \big[(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)\big]\mathrm{d}^{3}v\nonumber\\
-        & = & K\int\Big\{ f(\vec{v},t)\big[\ln K + \ln f(\vec{v},t)\big] \Big\}\mathrm{d}^{3}v.
+    \int \big[Kf'(\vec{v},t)\big] \ln \big[Kf'(\vec{v},t)\big] \mathrm{d}^{3}v & = & 
+        \int \big[(K\ln K)f'(\vec{v},t) + Kf'(\vec{v},t) \ln f'(\vec{v},t)\big]\mathrm{d}^{3}v\nonumber\\
+        & = & K\int\Big\{ f'(\vec{v},t)\big[\ln K + \ln f'(\vec{v},t)\big] \Big\}\mathrm{d}^{3}v.
 \end{eqnarray}
 %
-We observe that if the number of particles in the $\mu$-space, $f(\vec{v},t)$, is much greater than
+We observe that if the number of particles in the $\mu$-space, $f'(\vec{v},t)$, is much greater than
 the number of cells $K$, then the first term of Eq.~(\ref{eq:sum-h}) is negligible, and consequently:
@@
 \begin{equation}\label{aditive-property}
-    \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v \simeq
-    K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,
+    \int \big[Kf'(\vec{v},t)\big] \ln \big[Kf'(\vec{v},t)\big] \mathrm{d}^{3}v \simeq
+    K\int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v,
 \end{equation}
@@
 Furthermore, in equilibrium the Lagrange multipliers are position- and time-independent,
-thus $f_M(\vec{v})$
+thus $f'_M(\vec{v})$
 reduces to
@@
 \begin{equation}\label{eq:fbardef}
-    f_M(\vec{v})=C\exp\left(\alpha+\beta \epsilon(\vec{v})\right)
-    \equiv \bar{f}(\vec{v}),\quad M=1,\,\dots\,,K.
+    f'_M(\vec{v})=C\exp\left(\alpha+\beta \epsilon(\vec{v})\right)
+    \equiv \bar{f}'(\vec{v}),\quad M=1,\,\dots\,,K.
 \end{equation} 
@@
 \begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int \left[f_M(\vec{v},t)
-    \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
+   H''(t)=\sum_{M=1}^{K}\int \left[f'_M(\vec{v},t)
+    \ln f'_M(\vec{v},t)-f'_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
 \end{equation}
 %
-however $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$.
+however $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$.
 Since $N$ is a constant and we are mainly intersted in the time-derivative of $H''$,
@@
 starts at some time as an out-of-equilibrium system,
-but in such a manner that it is inteuitively not too far from equilibrium,
+but in such a manner that it is intuitively not too far from equilibrium,
 which ensures that the local equilibrium hypothesis remains valid during
@@
       \bar{\mathcal{N}}'=
-      \int \bar{f}(\vec{v}) \mathrm{d}^{3}v\quad\textrm{and} \quad
+      \int \bar{f}'(\vec{v}) \mathrm{d}^{3}v\quad\textrm{and} \quad
       \bar{\mathcal{E}}'=
-      \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v.
+      \int \bar{f}'(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,
     \end{equation}
     %
-Here, we have used $\bar{f}(\vec{v})$ as defined in Eq.~(\ref{eq:fbardef}).
+where we have used $\bar{f}'(\vec{v})$ as defined in Eq.~(\ref{eq:fbardef}).
 In Eqs.~(\ref{eq:restrictionoutsideclassical}), $\Delta'_M$ and $\delta'_M$
 are considered deviations relative to
-$\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$, respectively, and for systems that are not
-too far away from equilbrium, it is reasonable to expect on the one hand that
+$\bar{\mathcal{N}}'$ and $\bar{\mathcal{E}}'$, respectively, and for systems that are not
+too far away from equilibrium, it is reasonable to expect on the one hand that
 $\Delta'_M(t)\ll\bar{\mathcal{N}}'$ and
@@
 $\delta'_M$ are sufficiently large compared with fluctuations of $\bar{\mathcal{N}}'$
-and $\bar{\mathcal{E}}'$.
-
-\commr{CHECKPOINT}\\
-Performing the derivative of (\ref{eq:CH2}) with respect to time, we obtain
-%
-\begin{equation}\label{dH1}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\left[
-      1+\ln f_M(\vec{v},t)
-    \right]\dot f_M(\vec{v},t)\mathrm{d}^3v.
-\end{equation}
-%
-Using the first-order approximation
-%
-\begin{equation}\label{firstorder}
-   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ 
-   \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|,
-\end{equation}
-%
-(\ref{dH1}) yields
-%
-\begin{equation}\label{dH1-1}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\bar f(\vec{v}) \left[
-      1+\ln \left\{
-        \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t)
-      \right\}
-    \right]\dot g_M(\vec{v},t)\mathrm{d}^3v.
+and $\bar{\mathcal{E}}'$. Under the same perspective, we can assume that every local
+distribution function can be written as
+\begin{equation}\label{eq:firstorder}
+   f'_{Mn}(\vec{v},t)=\bar{f}'_{n}(\vec{v})(1+g'_{Mn}(\vec{v},t)),\quad
+   1\gg|g'_{Mn}(\vec{v},t)|.
 \end{equation}
@@
 
-We can approximate the logarithm function in (\ref{dH1-1}) to its first-order
-Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 
+With the previous considerations, in the following we provide an alternative demonstration
+of the $H$-theorem, considering the $H$-functional defined by Eq.~(\ref{eq:CH2}).
+
+We commence by differentiating Eq.~(\ref{eq:CH2}) with respect to time,
+%
+\begin{equation}\label{eq:dH1}
+    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\left[
+      1+\ln f'_M(\vec{v},t)
+    \right]\dot f'_M(\vec{v},t)\mathrm{d}^3v.
+\end{equation}
+%
+(Here and in what follows we use the standard notation $(d/dt)h\equiv\dot h$
+for an arbitrary function $h$.) Substituting Eq.~(\ref{eq:firstorder}) into Eq.~(\ref{eq:dH1}) yields
+%
+\begin{equation}\label{eq:dH1-1}
+    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\bar f'(\vec{v}) \left[
+      1+\ln \left\{
+        \bar f'(\vec{v})+\bar f'(\vec{v})g'_M(\vec{v},t)
+      \right\}
+    \right]\dot g'_M(\vec{v},t)\mathrm{d}^3v.
+\end{equation}
+%
+The logarithmic term of Eq.~(\ref{eq:dH1-1}) expanded up to the first-order term
+of its Taylor series, around $g'_{nM}(\vec{v},t)=0$, is
 %
 \begin{equation}\label{lnapproximationclassical}
-    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx
-    \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t).
+    \ln [\bar{f}'(\vec{v})+\bar{f}'(\vec{v}) g'_{M}(\vec{v},t)] \approx
+    \ln [\bar{f}'(\vec{v})]+ g'_{M}(\vec{v},t),
 \end{equation}
 %
-With the previous approximation, (\ref{dH1-1}) obtains the following form 
+and and substituting this into Eq.~(\ref{eq:dH1-1}) gives
 %
 \begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int \bar f(\vec{v})\left[
-      1+\ln \bar f(\vec{v})+g_M(\vec{v},t)
-    \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,
+    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int \bar f'(\vec{v})\left[
+      1+\ln \bar f'(\vec{v})+g'_M(\vec{v},t)
+    \right]\dot g'_M(\vec{v},t)\mathrm{d}^3v.
 \end{eqnarray}
 %
+
+\commr{ CHECKPOINT So far, we have not restricted the initial condition of the system to be
+of any particular form.}
+
 using (\ref{eq:relacion1}) with homogeneous multipliers

### 2020-12-22 tex/classical.tex
Notation: using mathcal{H}, not H.

@@
 \begin{eqnarray}
-    H_M' & = &  \int f'_M(\vec{v},t) \ln f'_{M}(\vec{v},t)
+    \mathcal{H}_M' & = &  \int f'_M(\vec{v},t) \ln f'_{M}(\vec{v},t)
       \mathrm{d}^{3}v \label{eq:Hcell},\\
@@
 
-We now use the variational method in order to maximize $H'$, applying the restrictions
+We now use the variational method in order to maximize $\mathcal{H}'$, applying the restrictions
 given in Eqs.~(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) through Lagrange
@@
 \begin{eqnarray}
-    \frac{\delta H'}{\delta f'_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
+    \frac{\delta \mathcal{H}'}{\delta f'_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
       \frac{\delta}{\delta f'_J(\vec{v}')}\left[
@@
 where $C$ is a constant. We notice that using the variational procedure on
-$H'$ predicts that the
+$\mathcal{H}'$ predicts that the
 distribution function of each cell has the form of the Maxwell-Boltzmann
@@
 
-\subsubsection{Properties of $H'$ for systems in equilibrium}
+\subsubsection{Properties of $\mathcal{H}'$ for systems in equilibrium}
 
@@
 \begin{equation}
-    H'(t)=\int \sum_{M=1}^{K} [f'(\vec{v},t)\ln f'(\vec{v},t)] \mathrm{d}^{3}v
+    \mathcal{H}'(t)=\int \sum_{M=1}^{K} [f'(\vec{v},t)\ln f'(\vec{v},t)] \mathrm{d}^{3}v
          = K\int  f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v= K H_{B}(t).
@@
 %
-In order to identify $H'$ with the entropy, it remains to show that $H'(t)$ is extensive,
+In order to identify $\mathcal{H}'$ with the entropy, it remains to show that $\mathcal{H}'(t)$ is extensive,
 with respect to $Kf'_(\vec v,t)$.
@@
 %
-\textit{i.e.} $H_M'$ is extensive, and the sum $\sum_MH'_M$ is the $H$-functional of the complete
-system, which reduces to the Boltzmann $H$-functional. Hence, $H'$ can also be identified with the
+\textit{i.e.} $\mathcal{H}'_M$ is extensive, and the sum $\sum_M\mathcal{H}'_M$ is the $H$-functional of the complete
+system, which reduces to the Boltzmann $H$-functional. Hence, $\mathcal{H}'$ can also be identified with the
 system's entropy.
@@
 %
-The constant $C$ can be omitted, which is shown by defining the following $H''$
+The constant $C$ can be omitted, which is shown by defining the following $\mathcal{H}''$
 functional:
@@
 \begin{equation}
-   H''(t)=\sum_{M=1}^{K}\int \left[f'_M(\vec{v},t)
+   \mathcal{H}''(t)=\sum_{M=1}^{K}\int \left[f'_M(\vec{v},t)
     \ln f'_M(\vec{v},t)-f'_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
@@
 however $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$.
-Since $N$ is a constant and we are mainly intersted in the time-derivative of $H''$,
-$C$ can be conveniently omitted. In other words, $H'$ renders the Maxwell-Boltzmann
+Since $N$ is a constant and we are mainly intersted in the time-derivative of $\mathcal{H}''$,
+$C$ can be conveniently omitted. In other words, $\mathcal{H}'$ renders the Maxwell-Boltzmann
 distribution function for systems in equilibrium.
@@
 With the previous considerations, in the following we provide an alternative demonstration
-of the $H$-theorem, considering the $H$-functional defined by Eq.~(\ref{eq:CH2}).
+of the $H$-theorem, considering the $H$-functional, $\mathcal{H}'$, defined by Eq.~(\ref{eq:CH2}).
 
@@
 \begin{equation}\label{eq:dH1}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\left[
+    \frac{d\mathcal{H}'}{dt}=\sum_{M=1}^{K}\int\left[
       1+\ln f'_M(\vec{v},t)
@@
 %
-(Here and in what follows we use the standard notation $(d/dt)h\equiv\dot h$
-for an arbitrary function $h$.) Substituting Eq.~(\ref{eq:firstorder}) into Eq.~(\ref{eq:dH1}) yields
+(Here and throughout the article, we use the standard notation $\dot h\equiv(dh/dt)$.)
+Substituting Eq.~(\ref{eq:firstorder}) into Eq.~(\ref{eq:dH1}) yields
 %
 \begin{equation}\label{eq:dH1-1}
-    \frac{dH'}{dt}=\sum_{M=1}^{K}\int\bar f'(\vec{v}) \left[
+    \frac{d\mathcal{H}'}{dt}=\sum_{M=1}^{K}\int\bar f'(\vec{v}) \left[
       1+\ln \left\{
@@
 %
-and and substituting this into Eq.~(\ref{eq:dH1-1}) gives
+and substituting this into Eq.~(\ref{eq:dH1-1}) gives
 %
 \begin{eqnarray}
-    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int \bar f'(\vec{v})\left[
+    \frac{d\mathcal{H}'}{dt}&=&\sum_{M=1}^{K} \int \bar f'(\vec{v})\left[
       1+\ln \bar f'(\vec{v})+g'_M(\vec{v},t)
@@
 \begin{equation}\label{dH1-2}
-    \frac{dH'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})\left[
+    \frac{d\mathcal{H}'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})\left[
       \alpha+\beta \epsilon(\vec{v})
@@
 \begin{equation}
-  \frac{dH'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
+  \frac{d\mathcal{H}'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
 \end{equation}
@@
 \begin{equation}\label{classicalH3}
-    \frac{dH'}{dt}=\sum_J^{L}\int
+    \frac{d\mathcal{H}'}{dt}=\sum_J^{L}\int
       \bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v
@@
 \begin{equation}
-    \frac{dH'}{dt}=-\left[
+    \frac{d\mathcal{H}'}{dt}=-\left[
       \sum_J^{L}\int\bar f(\vec{v})|g_J^{+}(\vec{v},t)|
@@
 expression is positive. However, the global sign makes the derivative of the
-$H'$ with respect to time is always less to zero. With this we proved that
-$\frac{dH'}{dt}<0$. If the system is in equilibrium,
-$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.
+$\mathcal{H}'$ with respect to time is always less to zero. With this we proved that
+$\frac{d\mathcal{H}'}{dt}<0$. If the system is in equilibrium,
+$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{d\mathcal{H}'}{dt}=0$.
 
@@
 \begin{equation}\label{CH3}
-   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v,
+   \mathcal{H}'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v,
 \end{equation}
@@
 \begin{equation}
-    \frac{dH'}{dt} \leq 0.
+    \frac{d\mathcal{H}'}{dt} \leq 0.
 \end{equation}

### 2020-12-22 tex/classical.tex
Reworded late middle part of 2.3.2.

@@
 
-Throughout this section, we will consider that the classical gas
-starts at some time as an out-of-equilibrium system,
-but in such a manner that it is intuitively not too far from equilibrium,
+Throughout this section, we will consider a classical gas
+whose initial condition is a state close to equilibrium,
 which ensures that the local equilibrium hypothesis remains valid during
-the time evolution of the system.
+the time-evolution of the system.
 Also we will use the following definitions:
@@
 are considered deviations relative to
-$\bar{\mathcal{N}}'$ and $\bar{\mathcal{E}}'$, respectively, and for systems that are not
-too far away from equilibrium, it is reasonable to expect on the one hand that
+$\bar{\mathcal{N}}'$ and $\bar{\mathcal{E}}'$, respectively, and for systems that are
+sufficiently close to equilibrium, it is reasonable to expect on the one hand that
 $\Delta'_M(t)\ll\bar{\mathcal{N}}'$ and
@@
 \begin{equation}\label{eq:firstorder}
-   f'_{Mn}(\vec{v},t)=\bar{f}'_{n}(\vec{v})(1+g'_{Mn}(\vec{v},t)),\quad
-   1\gg|g'_{Mn}(\vec{v},t)|.
+   f'_{M}(\vec{v},t)=\bar{f}'(\vec{v})(1+g'_{M}(\vec{v},t)),\quad
+   1\gg|g'_{M}(\vec{v},t)|.
 \end{equation}
@@
 The logarithmic term of Eq.~(\ref{eq:dH1-1}) expanded up to the first-order term
-of its Taylor series, around $g'_{nM}(\vec{v},t)=0$, is
+of its Taylor series, around $g'_{M}(\vec{v},t)=0$, is
 %
@@
 
-\commr{ CHECKPOINT So far, we have not restricted the initial condition of the system to be
-of any particular form.}
-
-using (\ref{eq:relacion1}) with homogeneous multipliers
+So far, we have not restricted the initial condition of the system to be
+of any particular form, other than being close to equilibrium. Assuming
+that the initial state of the system is also such that the Lagrange multipliers
+are homogeneous (\textit{i.e.} $\alpha_M=\alpha$ and $\beta_M=\beta$, for all $M$),
+we have from Eq.~(\ref{eq:relacion1}):
+\commr{[Prof. Carrillo: A esto no le encuentro razón física, salvo que es lo que
+hace Huang\dots]}
 %
-\begin{equation}\label{dH1-2}
+\begin{equation}\label{eq:dH1-2}
     \frac{d\mathcal{H}'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})\left[
@@
     \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}
-    \int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v . \nonumber 
+    \int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v .
 \end{equation}
 %
-On the other hand, we obtain from the restrictions the following expressions
+In the first term of Eq.~(\ref{eq:dH1-2}), we have used Eq.~(\ref{eq:fbardef}).
+From the definitions of $\mathcal{N}'_M$ and $\mathcal{E}'_M$ 
+---Eqs.~(\ref{eq:restrictionoutsideclassical})---, and $f'_{M}(\vec{v},t)$
+---Eq.~(\ref{eq:firstorder})---, it is straightforward to show that
 %
+\begin{subequations}\label{eq:classsumdotdeltaseq0}
 \begin{eqnarray}
     \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow&
-    \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
+    \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \\
     \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow&
-    \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber 
+    \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t),
 \end{eqnarray}
+\end{subequations}
 %
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:sumderdeltaseq0}
     \sum_{M=1}^{K} \dot{\Delta}_M(t)  =\sum_{M=1}^{K} \dot{\delta}_{M}(t) =0.
@@
 %
-Then using the previous result to (\ref{dH1-2}), we find
+Therefore the first term of Eq.~(\ref{eq:dH1-2}) is null because of
+Eqs.~(\ref{eq:classsumdotdeltaseq0}) and (\ref{eq:sumderdeltaseq0}), which renders
 %
-\begin{equation}
+\begin{equation}\label{eq:dHpdtbefgdecclass}
   \frac{d\mathcal{H}'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
@@
 %
-Now, the summation over $M$ will be expressed in two summations
+In order to clearly determine the time-evolution of Eq.~(\ref{eq:dHpdtbefgdecclass}),
+we split the summation over $M$ into two terms:
 %
-\begin{equation}\label{classicalH3}
+\begin{equation}\label{eq:classicalH3}
     \frac{d\mathcal{H}'}{dt}=\sum_J^{L}\int
-      \bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v
+      \bar f'(\vec{v}){g'_I}^{+}(\vec{v},t){{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)\mathrm{d}^3v
         +\sum_J^{P}\int
-        \bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v.
+      \bar f'(\vec{v}){g'_J}^{-}(\vec{v},t){{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)\mathrm{d}^3v,
 \end{equation}
 %
-where $L+P=K$. The group of $L$ cells is such that cells have an excess of
-particles or/and energy (from the mean value in equilibrium). In contrast,
-the group of $P$ cells is such that cells have to miss particles or/and energy
-(from the mean value in equilibrium). Besides, $\dot{g}^{+}_{J}$ represents the
+where $L+P=K$. The above split is made based on the assumption that for any given initial
+state of the system, at $t_0$, some cells will have either a $g'_I(\vec{v},t_0)\geq0$
+or a $g'_J(\vec{v},t_0)<0$,
+which we denote as ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)$
+or ${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)$, respectively.
+%
+%---------------------------------------
+%The text below is not needed, also it may be wrong. $g(t_0)>0$ indicates $f(t_0)>\bar f$, but this does
+%not necessarily imply excess of particles and/or energy. E.g., a particular combination of a
+%few missing particles and a considerable energy excess might render the corresponding
+% $g(t_0)$ to be greater than zero.
+\begin{comment}
+Cells with ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)$, $I=1,\,\dots\,,L$, 
+have an excess of particles and/or energy (relative to the mean value in equilibrium).
+In contrast, cells with ${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)$, $J=1,\,\dots\,,P$,
+have a scarcity of particles and/or energy
+(relative to the mean value in equilibrium).
+In addition, $\dot{g}^{+}_{J}$ represents the
 change on the deviation on cells that have an excess of particles or energy
@@
 the value of the deviation on cells that have missing particles or energy.
-
 Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles
@@
 and consequently, (\ref{classicalH3}) obtains the following form
+\end{comment}
+%-----------------------------
 %
-\begin{equation}
+If the system's initial state is sufficiently close to equilibrium, it is
+physically suitable
+to assume that $\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$ in a monotonous manner,
+thus ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)\leq0$ and
+${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)>0$, for $t\geq t_0$. Consequently,
+Eq.~(\ref{eq:classicalH3}) can be re-written as
+%
+\begin{equation}\label{eq:classicalH4}
     \frac{d\mathcal{H}'}{dt}=-\left[
@@
       |\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v 
-    \right]. \label{classicalH4}
+    \right].
 \end{equation}
 %
-We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always
+Since every integrand in Eq.~(\ref{eq:classicalH4}) is positive,
+for all $t$ and $\vec v$, and
+$\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$, it follows that
+%
+\begin{equation}\label{eq:dHpdtleq0}
+ \frac{d\mathcal{H}'}{dt}\leq0.
+\end{equation}
+%
+
+We can observe that $\bar{f}(\vec{v})$ in (\ref{eq:classicalH4}) is always
 positive. Also the deviation and its derivative are positive, then all the
@@
 $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{d\mathcal{H}'}{dt}=0$.
+
+\commr{CHECKPOINT}
 

### 2020-12-23 tex/classical.tex
Reworded: classical section finished (1st round)

@@
 
-Classical statistical mechanics is theoretically founded on the Boltzmann transport
-equation and the hypo of molecular chaos. The fundamental connection
+Classical statistical mechanics is theoretically built upon the Boltzmann transport
+equation and the hypothesis of molecular chaos. The fundamental connection
 between the microscopic nature of matter and the phenomenological laws of thermodynamics
-is the $H$-theorem.
+is the $H$-theorem. In this article we propose an alternative approach, which is 
+developed by means of the variational method, applied to a functional that may include
+systems out-of-equilibrium (or non-homogeneous). Our proposal is limited to systems
+whose initial state is physically close to the equilibrium, in such a manner that the
+local equilibrium hypothesis is satisfied.
+
+We begin this section by quite briefly stating the Boltzmann transport equation, followed
+by an also quite brief review of the $H$-theorem demonstration, as performed in
+standard textbooks.
 
@@
 
+Before introducing our proposal, we bring the reader's attention to the following. In defining
+Eq.~(\ref{eq:hbfunctional}), it is assumed that the distribution function does not
+depend on the coordinates, which implies that the initial state of the system may not in
+equilibrium, but the difference between the equilibrium and non-equilibrium is such
+that it does not cause inhomogeneities to the system.
+
 %---------------------------
@@
 use the term homogeneous distribution function to refer to a distribution
-function thad does not depend on the coordinates, and will use primed functions
+function thad does not depend on the coordinates. In addition,
+in this section we will use primed functions
 and quantities to differentiate the classical case (this Section) \textit{vs} the quantum
@@
 of particles and the local
-energy do not depend on the cell number, this is
+energy do not depend on the cell number, on a statistical sense, this is
 %
@@
 %
-
-So far, we have not restricted the initial condition of the system to be
-of any particular form, other than being close to equilibrium. Assuming
-that the initial state of the system is also such that the Lagrange multipliers
-are homogeneous (\textit{i.e.} $\alpha_M=\alpha$ and $\beta_M=\beta$, for all $M$),
-we have from Eq.~(\ref{eq:relacion1}):
-\commr{[Prof. Carrillo: A esto no le encuentro razón física, salvo que es lo que
-hace Huang\dots]}
+Substituting $\ln \bar f'(\vec{v})=\exp\left(\alpha+\beta \epsilon(\vec{v})\right)$,
+see Eq.~(\ref{eq:fbardef}) and text following it, and omitting $C$ yields
 %
@@
 %
-In the first term of Eq.~(\ref{eq:dH1-2}), we have used Eq.~(\ref{eq:fbardef}).
 From the definitions of $\mathcal{N}'_M$ and $\mathcal{E}'_M$ 
@@
 %
-Therefore the first term of Eq.~(\ref{eq:dH1-2}) is null because of
-Eqs.~(\ref{eq:classsumdotdeltaseq0}) and (\ref{eq:sumderdeltaseq0}), which renders
+Therefore, because of
+Eqs.~(\ref{eq:classsumdotdeltaseq0}) and (\ref{eq:sumderdeltaseq0}), Eq.~(\ref{eq:dH1-2}) simplifies to
 %
@@
 and consequently, (\ref{classicalH3}) obtains the following form
+
+We can observe that $\bar{f}(\vec{v})$ in (\ref{eq:classicalH4}) is always
+positive. Also the deviation and its derivative are positive, then all the
+expression is positive. However, the global sign makes the derivative of the
+$\mathcal{H}'$ with respect to time is always less to zero. With this we proved that
+$\frac{d\mathcal{H}'}{dt}<0$. If the system is in equilibrium,
+$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{d\mathcal{H}'}{dt}=0$.
 \end{comment}
@@
 
-We can observe that $\bar{f}(\vec{v})$ in (\ref{eq:classicalH4}) is always
-positive. Also the deviation and its derivative are positive, then all the
-expression is positive. However, the global sign makes the derivative of the
-$\mathcal{H}'$ with respect to time is always less to zero. With this we proved that
-$\frac{d\mathcal{H}'}{dt}<0$. If the system is in equilibrium,
-$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{d\mathcal{H}'}{dt}=0$.
+In summary, if a gas occupying a volume V (which is divided into $K$ small cells), with a
+total energy $E$ and $N$ classical free particles, whose initial state is not in equilibrium,
+but sufficiently close to equilibrium ---such that the local equilibrium hypothesis apply---, then the funcional
+%
+\begin{equation}\label{eq:CH3}
+   \mathcal{H}'(t)=\sum_{M=1}^{K}\int f'_M(\vec{v},t) \ln f'_M(\vec{v},t)\mathrm{d}^3v
+\end{equation}
+%
+---where $f'_M(\vec{v},t)$ is the cell distribution function---,
+satisfies $d\mathcal{H}'/dt\leq0$, and the equality relation is attained at $t\to\infty$.
+In Eq.~(\ref{eq:CH3}), $f'_M(\vec{v},t)$ is the Maxwell-Boltzmann distribution
+function, which in general is different for different cells (\textit{i.e.} the complete system
+can be non-homogeneous), and each $f'_M(\vec v,t)$ is compatible
+with the cell number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,
+$T_M$, and Legendre multipliers $\alpha_M$ and $\beta_M$.
 
-\commr{CHECKPOINT}
-
+\begin{comment}
 Joining both results, we can say the following statement:
@@
 theorem applying the method of the volume divided into cells.
+\end{comment}

### 2020-12-27 tex/classical.tex
Rewording: up to section 3.1

@@
 \textit{i.e.} $\mathcal{H}'_M$ is extensive, and the sum $\sum_M\mathcal{H}'_M$ is the $H$-functional of the complete
-system, which reduces to the Boltzmann $H$-functional. Hence, $\mathcal{H}'$ can also be identified with the
-system's entropy.
+system, which reduces to the Boltzmann $H$-functional.
+\commb{Hence, $\mathcal{H}'$ can also be identified with the
+system's entropy.}
 
@@
 can be non-homogeneous), and each $f'_M(\vec v,t)$ is compatible
-with the cell number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,
-$T_M$, and Legendre multipliers $\alpha_M$ and $\beta_M$.
+with the cell properties (number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,
+$T_M$, and Legendre multipliers $\alpha_M$ and $\beta_M$).
 

### 2020-12-27 tex/disscussion.tex
Rewording: up to section 3.1

@@
 \end{equation}
-\textcolor{red}{Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
+Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
 It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
@@
 In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.}
+Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.
 

### 2020-12-27 tex/qccorrespondence.tex
Rewording: up to section 3.1

@@
 where $H_{B}$ now is (\ref{eq:CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{bib:tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.\\
-\textcolor{blue}{In the equilibrium case, we have the following variational entropy
+We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.
+
+In the equilibrium case, we have the following variational entropy
 \begin{eqnarray}
@@
 \end{equation}
-that corresponds to the classical $H$-functional in equilibrium described in the energy space.}\\
-\textcolor{red}{Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.}
+that corresponds to the classical $H$-functional in equilibrium described in the energy space.
+
+Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.
 

### 2020-12-27 tex/quantum.tex
Rewording: up to section 3.1

@@
 
-\subsection{H theorem and the Fermi-Dirac and Bose-Einstein distribution}
+The classical $H$-theorem is even to these days one of the pillars whereon the classical statistical
+mechanics is founded. Unfortunately, regardless of several attempts have been made
+\commr{[CITE Q. H-THEOREM EXAMPLES]}, 
+the generality of the classical $H$-theorem has no quantum match. In this section, we shall
+propose and analize an alternative quantum $H$-functional using the variational method.
+We start by briefly outline a typical textbook demonstration of the quantum $H$-theorem
+\cite{bib:tolman}, followed by the analysis of our proposed $H$-functional.
 
-The $H$-functional in quantum mechanics presented by Tolman \cite{bib:tolman}
-is defined as
+\subsection{$H$-theorem and the Fermi-Dirac and Bose-Einstein distribution functions}
+
+Consider a diluted gas of $N$ non-interacting quantum particles (either bosons or fermions),
+contained by a vessel of volume
+$V$, at some temperature $T$, and total energy E. Starting from the Boltzmann definition
+of entropy, the quantum $H$ functional is
 %
-\begin{equation}\label{quantumh}
-    H=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, 
+\begin{equation}
+   H_T=-\ln G,
 \end{equation}
 %
-where $n_i$ is the occupation number and $g_i$ is an energy cell. Also, an
-equation for the rate of change in the number of particles in group $\kappa$ is
+where $G$ is the total number of possible quantum states of the gas that satisfy the
+above conditions \cite{bib:tolman}.
+The quantum $H$-theorem can be demostrated as follows.  $G$ can be divided into groups of neighboring states,
+$g_k$, and a certain occupation numbers, $n_k$, can be associated with each of these
+groups. Thus, the above functional takes the form:
 %
-\begin{eqnarray}
+\begin{equation}\label{eq:quantumh}
+    H_T=\sum_i n_i \ln n_i -(n_i\pm g_i)\ln (g_i \pm n_i)\pm g_i\ln g_i, 
+\end{equation}
+%
+where the upper and lower signs are for bosons and fermions, respectively.
+Thus the time derivative of Eq.~(\ref{eq:quantumh}) is
+%
+\begin{equation}\label{eq:quantumdHdt}
+\frac{dH}{dt}=\sum_{\kappa}\left[\ln n_{\kappa}-\ln\left(g_{\kappa}\pm n_{\kappa}\right)\right]
+\frac{dn_{\kappa}}{dt}.
+\end{equation}
+%
+
+Arguing that the energy exchange between particles is dominated by collisions between particles,
+and using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
+%
+\begin{eqnarray}\label{eq:changen}
     \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
-    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}),\label{changen}
+    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}).
 \end{eqnarray}
 %
-where $A_{\kappa\lambda,\mu\nu}$ is a tensor expressed as
+Here $A_{\kappa\lambda,\mu\nu}n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})$ is the
+expected number of collisions per unit time, in which two particles will be moved from groups
+$\kappa$, $\lambda$ to $\mu$, $\nu$, and the tensor $A_{\kappa\lambda,\mu\nu}n_{\kappa}$ is given by
+%
+\begin{equation}\label{eq:amnkldef}
+  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon}.
+\end{equation}
+%
+In Eq.~(\ref{eq:amnkldef}), $\Delta\epsilon$ is the net energy change occuring during the
+collision and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where
+$V_{mn,kl}$ is the element of the transition matrix of a binary collision.
+It is important to remark that in deriving Eq.~(\ref{eq:changen}), the equal a priori
+probabilities and the random a priori phases hypotheses were surmised, and that 
+the random a priori phases hypothesis can be
+considered an analogous to the molecular chaos hypothesis
+\cite{bib:das2018}.
+
+Substituting Eq.~(\ref{eq:changen}) into Eq.~(\ref{eq:quantumdHdt}), and performing
+the algebra, it can be proven that
 %
 \begin{equation}
-  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon},
+    \frac{dH_T}{dt}\leq 0.
 \end{equation}
 %
-where $\epsilon$ is the energy and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where
-$V_{mn,kl}$ is the perturbation matrix of binary particles. Those expressions
-are obtained by the perturbation theory.
 
-\textcolor{red}{Tolman's proposal for the quantum $H$-theorem establishes that
-if we have a quantum gas that satisfies the rate of change equation
-(\ref{changen})\footnote{It is important to remark that the equal a priori
-probabilities and random a priori phases hypothesis are implicit in the rate of
-change equation (\ref{changen}) and the random a priori phases hypothesis is
-considered as the quantum analogous to the molecular chaos hypothesis
-\cite{bib:das2018}}, then the $H$-functional (\ref{quantumh}) satisfies that
+In equilibrium (at $t\to\infty$), $dn_{\kappa}/dt=0$, hence from Eq.~(\ref{eq:changen})
+%
+\begin{equation}\label{eq:boseiferdirdistfnctsrc}
+    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}}.
+\end{equation}
+%
+Considering that the energy is not lost during the collsion, the Bose-Einstein or Fermi-Dirac distribution
+functions can be recovered from Eq.~(\ref{eq:boseiferdirdistfnctsrc}):
 %
 \begin{equation}
-    \frac{dH}{dt}\leq 0,
+   n_{\kappa}=\frac{g_{\kappa}}{\exp(\alpha+\beta\epsilon_{\kappa}\mp1)}.
 \end{equation}
 %
-and as a consequence of this theorem, when the system is in equilibrium,
-$\frac{dH}{dt}=0$.\\
-The proof of this theorem is in the same way as in the classical version. Also,
-when the system is in equilibrium, the distribution functions that correspond
-to the equilibrium state are the Fermi-Dirac and the Bose-Einstein
-distribution. Those distributions can be obtained from a sufficient condition
-for the equilibrium state
+
+\textit{I.e.}, in equilibrium $dH_T/dt=0$ and the distribution function obtained from
+Eq.~(\ref{eq:quantumh}) is the expected distribution function.
+
+%----------------------------------------------
+\begin{comment}
+In addition, Tolman also showed that (\ref{eq:quantumh}) can be reduced, in the
+high-energy limit, to
 %
 \begin{equation}
-    \ln \frac{n_{\kappa}}{g_{\kappa}\pm n_{\kappa}}+\ln \frac{n_{\lambda}}{g_{\lambda}\pm n_{\lambda}}=\ln \frac{n_{\mu}}{g_{\mu}\pm n_{\nu}}+\ln \frac{n_{\nu}}{g_{\nu}\pm n_{\nu}},
-\end{equation}
-%
-where the upper sign refers to the Bose-Einstein distribution and the lower
-sign refers to the Fermi-Dirac distribution.}
-
-In addition to this, Tolman showed that (\ref{quantumh}) can be reduced in the
-limit of high-energy levels to
-%
-\begin{equation}
-    H = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
+    H_T = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
 \end{equation}
@@
 \end{equation}
+\end{comment}
 %

### 2020-12-28 tex/quantum.tex
Rewording: start of sec. 3.2

@@
 \begin{equation}\label{eq:quantumdHdt}
-\frac{dH}{dt}=\sum_{\kappa}\left[\ln n_{\kappa}-\ln\left(g_{\kappa}\pm n_{\kappa}\right)\right]
+\frac{dH_T}{dt}=\sum_{\kappa}\left[\ln n_{\kappa}-\ln\left(g_{\kappa}\pm n_{\kappa}\right)\right]
 \frac{dn_{\kappa}}{dt}.
@@
 
-\subsection{Out of equilibrium, non-homogeneous distributions for degenerated systems}
+\subsection{Out of equilibrium, non-homogeneous quantum systems}
 
-We begin by defining the form of the functional
+Consider a diluted gas enclosed by a perfectly isolated vessel of volume $V$, with
+total energy $E$, and total number of quantum particles $N$, which can be fermions or
+bosons. For our purposes, the volume $V$ is divided into $K$ small cells, each of which
+has a constant volume $\delta V_M=V/K$, a temperature $T_M$, a local energy $\epsilon_M$,
+a local number of particles $\mathcal{N}_M$, and a local set of distribution functions,
+$\{f_{Mn}(t)\}$. Each $f_{Mn}$ $\big(\equiv f(\vec r_M,\epsilon_{n},t)\big)$ represent the number
+of particles contained in the $M$-th cell that occupy the energy level $\epsilon_n$.
+We propose the following functional as an alternative $H$-functional
+for quantum non-homogeneous diluted gases:
 %
 \begin{eqnarray}
-    \Ss (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
     &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
@@
 %
-This functional will be named \textit{variational entropy}.
-This functional describes, as in the classical case, a system with a global
-energy $E$, a global number of free
-quantum particles (which may be fermions for upper sing or bosons for lower
-sing) $N$, global volume $V$, and a temperature $T$.
-Besides, the variational entropy should correspond to a density of entropy when
-the system arrives at the equilibrium state. The
-global system, in turn, consists of a set of $K$ small cells with a constant
-volume $\delta V_M$, whose local homogeneous distribution functions is
-$f_{Mn}(\epsilon_{n},t)$. The distribution function will depend on the position
-of the cells, the energy per cell $\epsilon_{n}$, and time $t$. Also, the
-energy spectra will be divided into $\delta \epsilon_n$ volumes. We will use
+Here, the upper and lower sign refer to bosons and fermions respectively.
+\color{blue}\commr{Also, the energy spectrum will be divided into $\delta \epsilon_n$ volumes}. We will use
 the uppercase Latin index to denote cell numbers, and unless stated otherwise,
@@
 cell has its local chemical potential, $\mu_M$, local
-variational functional, $\Sm$, local number of particles, $\mathcal{N}_M$, and
+variational functional, $\mathcal{H}_M$, local number of particles, $\mathcal{N}_M$, and
 local energy, $\mathcal{E}_M$. In terms of the local distribution
@@
 \begin{eqnarray}
-    \Sm &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
+    \mathcal{H}_M &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
     &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
@@
 
-Maximizing the variational entropy $\Ss$ using the Lagrange multipliers method,
+Maximizing the variational entropy $\mathcal{H}$ using the Lagrange multipliers method,
 we obtain
@@
 \begin{eqnarray}
-      \Ss(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
+      \mathcal{H}(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
       &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
@@
 \begin{equation}
-    \Ss(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
+    \mathcal{H}(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
 \end{equation}
@@
 \begin{equation}
-   \frac{d \Ss (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
+   \frac{d \mathcal{H} (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
 \end{equation}
@@
 \begin{eqnarray}
-    \frac{d\Ss (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
+    \frac{d\mathcal{H} (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
     &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
@@
 \begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
+    \frac{d\mathcal{H}}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
     &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
@@
 \begin{eqnarray}
-    \frac{d\Ss}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
+    \frac{d\mathcal{H}}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
 \nonumber \\
@@
 \begin{eqnarray}
-   \frac{d\Ss}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
+   \frac{d\mathcal{H}}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
 \end{eqnarray}
@@
 \begin{equation}
-    \frac{d\Ss}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
+    \frac{d\mathcal{H}}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
 \end{equation}
@@
 \begin{equation}
-    \frac{d\Ss}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
+    \frac{d\mathcal{H}}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
 \end{equation}
@@
 and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then
-$\frac{d\Ss}{dt}<0$. With this, we proved that any quantum ideal gas perturbed
+$\frac{d\mathcal{H}}{dt}<0$. With this, we proved that any quantum ideal gas perturbed
 (out of equilibrium but not so far from it), evolves such that
-$\frac{d\Ss}{dt}<0$.
+$\frac{d\mathcal{H}}{dt}<0$.
 
 When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot
-g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\Ss}{dt}=0$. Then, we can
+g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\mathcal{H}}{dt}=0$. Then, we can
 say that any ideal quantum gas perturbed always evolves to the equilibrium

### 2021-01-06 tex/quantum.tex
Rewording: start of sec. 3.2.1

@@
 %
-Considering that the energy is not lost during the collsion, the Bose-Einstein or Fermi-Dirac distribution
+Considering that the energy is conserved during the collsion, the Bose-Einstein or Fermi-Dirac distribution
 functions can be recovered from Eq.~(\ref{eq:boseiferdirdistfnctsrc}):
@@
 Consider a diluted gas enclosed by a perfectly isolated vessel of volume $V$, with
-total energy $E$, and total number of quantum particles $N$, which can be fermions or
+total energy $E$, and total number of quantum particles $N$, which can be free fermions or
 bosons. For our purposes, the volume $V$ is divided into $K$ small cells, each of which
-has a constant volume $\delta V_M=V/K$, a temperature $T_M$, a local energy $\epsilon_M$,
+has a constant volume $\delta V_M=V/K$ ($M=1,\,\dots\,,K$), a
+local temperature $T_M$, a local energy $\epsilon_M$,
 a local number of particles $\mathcal{N}_M$, and a local set of distribution functions,
-$\{f_{Mn}(t)\}$. Each $f_{Mn}$ $\big(\equiv f(\vec r_M,\epsilon_{n},t)\big)$ represent the number
-of particles contained in the $M$-th cell that occupy the energy level $\epsilon_n$.
+$\{f_{Mn}(t)\}$. Hereafter we will use the short-hand notation:
+%
+\begin{equation}
+   f_{Mn}(t)\equiv f(\vec r_M,\epsilon_{n},t),
+\end{equation}
+%
+where $\vec r_M$ is the radius vector pointing at the center of the $M$-th cell.
+$f_{Mn}(t)$ represent the number
+of particles contained in the $M$-th cell that occupy the energy level $\epsilon_n$ at
+time $t$. Since the particles are considered to be free, the energy levels should not
+depend on the cell properties, \textit{i.e.} the energy spectrum, $\{\epsilon_n\}$, is the same for all
+cells; thus there is no need to label $\epsilon_n$ with an index $M$.
+
 We propose the following functional as an alternative $H$-functional
@@
 \begin{eqnarray}
-    \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ]   \delta V_M \delta \epsilon_n \label{entropy}.
+    \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} \bigg[ f_{Mn}(t) \ln f_{Mn}(t)\bigg.\nonumber \\
+    &&\bigg.\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M\label{entropy}.
 \end{eqnarray}
@@
 Here, the upper and lower sign refer to bosons and fermions respectively.
-\color{blue}\commr{Also, the energy spectrum will be divided into $\delta \epsilon_n$ volumes}. We will use
-the uppercase Latin index to denote cell numbers, and unless stated otherwise,
-$M = 1, 2 . . . , K$. Without loss of generality, we choose these cells to have
-local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$). Each
-cell has its local chemical potential, $\mu_M$, local
-variational functional, $\mathcal{H}_M$, local number of particles, $\mathcal{N}_M$, and
-local energy, $\mathcal{E}_M$. In terms of the local distribution
-function and local energy spectra, $\{\epsilon_{n}\}$, the last three
-properties are given by:
+In addition, when needed, each cell will have an associated local chemical potential,
+$\alpha_M$, and its local $H$-functional, which is defined by:
 %
-\begin{eqnarray}
-    \mathcal{H}_M &=&  \sum_{n} [ f_{Mn}(\epsilon_{n},t) \ln f_{Mn}(\epsilon_{n},t)\nonumber \\
-    &&\pm (1 \mp f_{Mn}(\epsilon_{n},t)) \ln (1 \mp f_{Mn}(\epsilon_{n},t)) ] \de V_M \delta \epsilon_n \label{entropycell},\\
-    {\mathcal{N}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n} ,t) \de V_M \delta \epsilon_n, \nonumber \\
-{\mathcal{E}}_M&=& \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n} \de V_M \delta \epsilon_n.
-\end{eqnarray}
+\begin{equation}\label{eq:qHMdef}
+   \mathcal{H}_M(t)=\sum_{n} \bigg[ f_{Mn}(t) \ln f_{Mn}(t)
+   \pm \Big(1 \mp f_{Mn}(t)\Big) \ln \Big(1 \mp f_{Mn}(t)\Big) \bigg] \de V_M.
+\end{equation}
 %
-When the system is in equilibrium, the local number of particles and the local
-energy don't depend on the cell number, this is
+Therefore, $\mathcal{N}_M$ and $\mathcal{E}_M$ as functions of time are given by:
+%
+\begin{subequations}
+\begin{equation}
+    {\mathcal{N}}_M(t) = \sum_{n}f_{Mn}(t) \delta V_M
+\end{equation}
+%
+and
 %
 \begin{equation}
-   {\mathcal{N}}_M=\mathcal{N}\equiv \bar{\mathcal{N}}; \ \ \ \  {\mathcal{E}}_M=\mathcal{E}\equiv \bar{\mathcal{E}}.
+\mathcal{E}_M(t) = \sum_{n}f_{Mn}(t)\epsilon_{n} \de V_M
+\end{equation}
+\end{subequations}
+%
+which are, in addtion, constrained by the global microcanonical restrictions
+%
+\begin{subequations}
+\begin{equation}\label{eq:qGlobRestN}
+    \sum_{M=1}^{K}\left[\sum_{n}f_{Mn}(t)\right]\delta  V_M
+    =\sum_{M=1}^{K}\mathcal{N}_{M}(t)\delta V_M=N
 \end{equation}
 %
-The particles in the system are considered to be free, \textit{i.e.} the
-particles do not interact with each other.
-Consequently, we can assume that the available quantum states do not depend on
-the cell features,
-therefore the energy spectrum, denoted by $\epsilon_n$, is the same for all
-cells, \textit{i.e.} $\epsilon_{Mn} = \epsilon_n, \forall M$. Besides, the
-system is subjected to the microcanonical restrictions
+and
 %
-\begin{eqnarray}
-    &&\sum_{M=1}^{K} \left[ \sum_{n}f_{Mn}(\epsilon_{n} ,t)\de \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{N}}_{M} \de V_M=N; \nonumber \\
-    &&\sum_{M=1}^{K}\left[ \sum_{n}f_{Mn}(\epsilon_{n},t)\epsilon_{n}\delta \epsilon_n\right] \de V_M=\sum_{M=1}^{K} {\mathcal{E}}_M \delta V_M=E, \label{restriccions1}
-\end{eqnarray}
+\begin{equation}\label{eq:qGlobRestE}
+    \sum_{M=1}^{K}\left[\sum_{n}f_{Mn}(t)\epsilon_{n}\right]\delta V_M
+    =\sum_{M=1}^{K}\mathcal{E}_M(t)\delta V_M=E. 
+\end{equation}
+\end{subequations}
 %
-where $N$ and $E$ are the total particle number and the total energy of the
-system respectively.
 
-Maximizing the variational entropy $\mathcal{H}$ using the Lagrange multipliers method,
-we obtain
+Applying the variational method to $\mathcal{H}$, and using the using the
+Lagrange multipliers
+$\{\alpha_M\}$ and $\{\beta_M\}$ it is straightforward to obtain
+(see also text around Eq.~(\ref{eq:deltaHpdeltafpj})):
 %
-\begin{eqnarray}
-&&\ln \left(\frac{1\mp f_{Mn}(\epsilon_{n},t)}{f_{Mn}(\epsilon_{n},t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n}, \label{relation}\\ &&\Rightarrow \ \ f_{Mn}(\epsilon_{n},t)=\frac{1}{e^{-\alpha_M(t)-\beta_M(t) \epsilon_{n}}\pm 1} \equiv \bar{f}_{Mn}(\epsilon_{n},t) \label{distributionequilibrium}.
-\end{eqnarray}
+\begin{equation}\label{eq:relation}
+\ln \left(\frac{1\mp f_{Mn}(t)}{f_{Mn}(t)} \right)=-\alpha_M(t)-\beta_M(t) \epsilon_{n},
+\end{equation}
 %
-where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. We can obtain
+and solving for $f_{Mn}(t)$ yields
+%
+\begin{equation}\label{eq:qfMn}
+f_{Mn}(t)=\frac{1}{\exp\big(-\alpha_M(t)-\beta_M(t) \epsilon_{n}\big)\pm 1}.
+\end{equation}
+
+%============================================================
+\subsubsection{Properties of $\mathcal{H}$ in equilibrium}
+%============================================================
+If the system is in equilibrium (at $t\to\infty$), the local number of particles and the local
+energy do not depend on the cell number and they should be homogeneous, this is \commr{[This needs a rationale\dots]}
+%
+\begin{subequations}
+\begin{equation}
+   {\mathcal{N}}_M(t\to\infty)\equiv \bar{\mathcal{N}}=N/K
+\end{equation}
+%
+and
+%
+\begin{equation}
+	{\mathcal{E}}_M(t\to\infty)\equiv \bar{\mathcal{E}}=E/K.
+\end{equation}
+%
+\end{subequations}
+%
+
+\color{blue}
+
+%
+\commr{We can obtain
 from the variational procedure that
@@
 %
+[What is $\Omega_M$?]
+}
+
 where we used that the $H$-functional in equilibrium is proportional to the

### 2021-01-06 tex/quantum.tex
Typos: sec. 3

@@
 expected number of collisions per unit time, in which two particles will be moved from groups
-$\kappa$, $\lambda$ to $\mu$, $\nu$, and the tensor $A_{\kappa\lambda,\mu\nu}n_{\kappa}$ is given by
+($\kappa$, $\lambda$) to ($\mu$, $\nu$), and the tensor $A_{\kappa\lambda,\mu\nu}$ is given by
 %

### 2021-01-09 tex/quantum.tex
Rewording: up to sec. 3.2.1

@@
 \begin{equation}
-   n_{\kappa}=\frac{g_{\kappa}}{\exp(\alpha+\beta\epsilon_{\kappa}\mp1)}.
+   n_{\kappa}=\frac{g_{\kappa}}{\exp(\alpha+\beta\epsilon_{\kappa})\mp1}.
 \end{equation}
@@
 %
-\begin{eqnarray}
+\begin{eqnarray}\label{eq:qHdef}
     \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} \bigg[ f_{Mn}(t) \ln f_{Mn}(t)\bigg.\nonumber \\
-    &&\bigg.\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M\label{entropy}.
+    &&\bigg.\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M.
 \end{eqnarray}
 %
-Here, the upper and lower sign refer to bosons and fermions respectively.
+Here, the upper and lower sign refer to \commr{bosons and fermions respectively}.
 In addition, when needed, each cell will have an associated local chemical potential,
@@
 %============================================================
-If the system is in equilibrium (at $t\to\infty$), the local number of particles and the local
-energy do not depend on the cell number and they should be homogeneous, this is \commr{[This needs a rationale\dots]}
+If the system is in equilibrium (at $t\to\infty$), the temperature becomes homogeneous throughout
+the complete system. Also, the local number of particles, the local
+energy, and the Lagrange multipliers do not depend on the cell number and they should be homogeneous,
+this is \commr{[This needs a rationale\dots]}
 %
-\begin{subequations}
+\begin{subequations}\label{eq:qEqRestrictions}
 \begin{equation}
-   {\mathcal{N}}_M(t\to\infty)\equiv \bar{\mathcal{N}}=N/K
+   {\mathcal{N}}_M(t\to\infty)\equiv \bar{\mathcal{N}}=N/K,
+\end{equation}
+%
+%
+\begin{equation}
+	{\mathcal{E}}_M(t\to\infty)\equiv \bar{\mathcal{E}}=E/K,
+\end{equation}
+\end{subequations}
+%
+%
+\begin{subequations}\label{eq:qEqAlphaBeta}
+\begin{equation}
+	\alpha_M=\bar\alpha,\quad\forall M,
 \end{equation}
@@
 and
+\begin{equation}
+	\beta_M=\bar\beta, \quad\forall M.
+\end{equation}
+\end{subequations}
+%
+Substituting Eqs.~(\ref{eq:qEqAlphaBeta}) into Eq.~(\ref{eq:qfMn}) yields the distribution
+function of each cell in equilibrium to be
+%
+\begin{equation}\label{eq:qfneq}
+    \bar f_{Mn}=\bar f_n =\frac{1}{\exp\big(-\bar\alpha-\bar\beta \epsilon_n\big)\pm 1},\quad\forall M.
+\end{equation}
+%
+By means of the above equation, we can recover the distribution function and the entropy of a diluted quantum gas in equilibrium as follows. Setting $\bar\alpha=\mu/kT$ and $\bar\beta=-1/kT$,
+and substituting them in Eq.~(\ref{eq:qfneq}) renders the Fermi-Dirac and Bose-Einstein
+distribution functions:
 %
 \begin{equation}
-	{\mathcal{E}}_M(t\to\infty)\equiv \bar{\mathcal{E}}=E/K.
+    \bar{f}_{n}=\frac{1}{\exp\big(\frac{{\epsilon_n}-\mu}{kT}\big)\pm 1},
 \end{equation}
 %
-\end{subequations}
+and substituting Eq.~(\ref{eq:qfneq}) into the negative of Eq.(\ref{eq:qHdef}) provides the entropy:
 %
+\begin{eqnarray}\label{eq:varEntropyDef}
+      S&=&\sum_{M=1}^{K}
+        \sum_n  \left[
+        		\left(\frac{1}{\exp\big(-\bar{\alpha}-\bar{\beta}\epsilon_{n}\big)\pm 1} \right)
+           \commr{\ln \left(\frac{1}{\exp\big(-\bar{\alpha}-\bar{\beta}\epsilon_{n}\big)\pm 1} \right)}
+         \right]\nonumber \\
+      &&\quad\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}
+         \left(1 \mp \frac{1}{\exp\big(-\bar{\alpha}-\bar{\beta}\epsilon_{n}\big)\pm 1} \right)
+         \right] \delta V_M.
+\end{eqnarray}
+%
+For obtaining the last result, we have used the well-known fact that the $H$-functional
+is the negative of the entropy. The property provided by Eq.(\ref{eq:varEntropyDef})
+is precisely what we call the ``variational entropy'',
+and this name reflects the fact that it was obtained \textit{via}
+the variational method.
+
+%=============================================
+\subsubsection{Proof of the quantum $H$-theorem for non-homogeneous systems}
+%=============================================
 
@@
 
-%
-\commr{We can obtain
-from the variational procedure that
-%
-\begin{equation}
-    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
-\end{equation}
-%
-[What is $\Omega_M$?]
-}
-
-where we used that the $H$-functional in equilibrium is proportional to the
-entropy of the system.
-In other hand, in the expression (\ref{multipliers}), we expect that the
-variation of the logarithm with the number of particles by cells is indifferent
-from a particular cell. With this, we conclude 
-%
-\begin{equation}
-    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
-\end{equation}
-%
-In equilibrium, the energy and the Lagrange multiplier do not depend on the
-cell number and time, therefore
-%
-\begin{equation}
-    \bar f_{Mn}(\epsilon_{n},t)=\bar f_n(\epsilon_{n}) =\frac{1}{e^{-\alpha-\beta \epsilon_n}\pm 1}.
-\end{equation}
-%
-If one compares those multipliers with the mean energy and the mean particle
-number from the statistical mechanic's results, we obtain
-%
-\begin{equation}
-    \alpha=\frac{\mu}{kT}\equiv \bar{\alpha}; \ \ \ \ \beta=-\frac{1}{kT}\equiv \bar{\beta},
-\end{equation}
-%
-and finally
-%
-\begin{eqnarray}
-    \bar{f}_{n}(\epsilon_{n})&=&\frac{1}{e^{(\frac{{\epsilon_n}-\bar{\mu}}{kT})}\pm 1}\equiv \bar{f}_{n}.
-\end{eqnarray}
-%
-We proved, when the system is in equilibrium, the Lagrange multipliers don't
-depend on the cell number. In other words, Lagrange multipliers are homogeneous
-in the system.
-
-We can calculate the entropy from an ideal system, in other words, if one
-substitute (\ref{distributionequilibrium}) to (\ref{entropy}) we obtain
-%
-\begin{eqnarray}
-      \mathcal{H}(t)&=&\sum_{M=1}^{K} \sum_n  \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\nonumber \\
-      &&\pm  \ln \left[\prod_{M=1}^{K} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right] \de V_M \delta \epsilon_n\label{H-entropy}.
-  \end{eqnarray}
-%
-This result allows us to calculate the entropy of a quantum ideal gases. This
-entropy could be used as a thermodynamic variable to describe those systems.
-
-%\subsection{The quantum analogous of the Boltzmann's $H$-theorem}
-We rewrite the \textit{variational entropy} (\ref{entropy}) as
+We rewrite the \textit{variational entropy} (\ref{eq:qHdef}) as
 %

### 2021-01-10 tex/classical.tex
Rewording: start of sec. 3.2.1
Los los cambios que hizo el Dr. Carrillo hasta a página 10 fueron agregados

--- 2020-12-23 tex/classical.tex
@@
 
-Classical statistical mechanics is theoretically built upon the Boltzmann transport
-equation and the hypothesis of molecular chaos. The fundamental connection
-between the microscopic nature of matter and the phenomenological laws of thermodynamics
-is the $H$-theorem. In this article we propose an alternative approach, which is 
-developed by means of the variational method, applied to a functional that may include
-systems out-of-equilibrium (or non-homogeneous). Our proposal is limited to systems
-whose initial state is physically close to the equilibrium, in such a manner that the
-local equilibrium hypothesis is satisfied.
-
-We begin this section by quite briefly stating the Boltzmann transport equation, followed
-by an also quite brief review of the $H$-theorem demonstration, as performed in
-standard textbooks.
+Boltzmann Kinetic theory of gases is one of the theoretical basese of our present 
+knowledge of out equilibrium systems. This represent a fundamental connection
+between the microscopic nature of matter and the phenomenological macroscopic  laws of classical thermodynamics.
+The stochasticity introduced in the, in other way deterministic, kinetics of the
+particles of the gas, via the molecular chaos hypothesis, allows the demostration 
+of the celebreted Boltzmann $H$ theorem. In this article we propose an alternative approach, which is 
+developed by means of a variational method, applied to a functional.
+Let us start this section by briefly remaind of the important elements of the 
+standard derivation and demostration of the Boltzmann transport equation and the $H$ theorem as they are presented in classical textbooks \cite{bib:huang}.
 
@@
 
-From the Boltzmann kinetic theory of gases, the probability for a particle to be in a given
-state of position and velocity at a time $t$,
-$f(\vec r,\vec v,t)$ (\textit{aka} the distribution function), must satify
-the transport equation (Eq.~(\ref{eq:transport}), below).
-This equation is traditionally obtained by exploiting the volume-invariance of the $\mu$-space
-(\textit{i.e.} the space formed by the direct product of position and velocity)
-and considering that elastic collisions constitute the predominant mechanism
-for exchanging energy among particles \cite{bib:huang}.
+Boltzmann kinetic theory of gases starts by defining the distribution function 
+$f(\vec{r},\vec{v},t)$ as the average number of molecules that at time $t$, has the 
+values of position and velocity in the element of volumen in the $\mu$ space.
+Assuming a deterministic Newtonian description of the movement of the molecules 
+and the invariance of the volume measure of the $\mu$ space, one arrives to the Boltzmann rate equation.
+%
+\begin{equation}
+	f(\vec{r}+\vec{v}\delta t, \vec{v}+\vec{F} \delta t, t+\delta t)=f(\vec{r},\vec{v},t)+\left( \frac{\partial f}{\partial t} \right)_{coll} \delta t.
+\end{equation}
+%
+The collision term $\left( \frac{\partial f}{\partial t}\right)$ of this equation 
+describes the fuxes in and out from the volume element due to the collisions.
+The integro-differential Boltzmann transpor equation can be cast into.
 %
@@
 
-Particle dynamics and external forces are described
+Particle dynamics and the effects of external forces are described
 by the left-hand side of Eq.~(\ref{eq:transport}), and the right-hand side
-is derived considering binary collisions between particles and
+is derived considering binary collisions between particles and accepting
 the \textit{molecular chaos hypothesis}, \textit{i.e.} it is assumed that
-the coordinates and velocities of the particles are not time-correlated.
+the positions and velocities of the particles are not time-correlated.
 
@@
 %
-Here, $f(\vec{v},t)$ is a homogeneous distribution function, \textit{i.e.}
-$f$ is assumed to be independent of $\vec r$.
+Notice that $f(\vec{v},t)$ is a spacially homogeneous distribution function.
 The functional $H_B$,
@@
 number of free classical particles $N$, and occupying a volume $V$.
-(We deviate from the custommary notation in order to
-clearly distinguish the Boltzmann functional $H$ from our proposals, and
-$f_B$ to denote the Maxwell-Boltzmann distribution function.)
+In order to clearly distinguish the Boltzmann functional $H_B$, we denote here
+the Maxwell-Boltzmann distribution function as $f_B$.
 
-Under this scheme, the time-evolution of the gas towards equilibrium is obtained through the
-$H$-theorem. This theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
-transport equation and the molecular chaos hypothesis is valid, then
-$dH_B/dt\leq0$ in general, and if $dH_B/dt=0$, then the system is in the equilibrium state.
-The $H$-theorem is proven by taking the time derivative
-of Eq. (\ref{eq:hbfunctional}) and using the Boltzmann transport equation
-(\ref{eq:transport}) \cite{bib:huang}. The $H$-theorem fixes the direction of time-evolution
-of systems in nature, which is consistent with the second law of thermodynamics; in fact,
+The physically correct espontaneous time evolution of an out of equilibrium dilute gas 
+is corroborated by the $H$ theorem.
+This theorem establishes that if the homogeneous function $f(\vec{v},t)$ satisfies the Boltzmann
+transport equation and consequently, if the molecular chaos hypothesis is valid, then
+the system evolves in such a way that $dH_B/dt\leq0$, and if when $dH_B/dt=0$, 
+the system is in the equilibrium state. $H$-theorem is easily proven by 
+using the Boltzmann transport equation (\ref{eq:transport}) \cite{bib:huang}. 
+$H$-theorem assures that our microscopic approach to describe the spontaneous 
+time evolution of the system will be consistent with the phenomenological observations
+established in the 2nd-law of Classical; in fact,
 $H_B$ can be associated with an entropy density.
 
-On the other hand, considering a diluted gas in equilibrium (\textit{i.e.}
-$(\partial f/\partial t)=0$),
-with no applied external forces (\textit{i.e.} $f$ is independent of $\vec r$), 
-it can be proven that the equilibrium distribution function obtained from
+On the other hand, considering a diluted gas in equilibrium, \textit{i.e.}
+$(\partial f/\partial t)=0$,
+with no applied external forces $f$ is independent of $\vec r$, 
+it can be directly proven that the equilibrium distribution function obtained from
 Eq.~(\ref{eq:transport}) is precisely the Maxwell-Boltzmann
@@
 
-Before introducing our proposal, we bring the reader's attention to the following. In defining
-Eq.~(\ref{eq:hbfunctional}), it is assumed that the distribution function does not
-depend on the coordinates, which implies that the initial state of the system may not in
-equilibrium, but the difference between the equilibrium and non-equilibrium is such
-that it does not cause inhomogeneities to the system.
+Before introducing our proposal, we call the reader's attention to the following. In defining
+$H_B$ Eq.~(\ref{eq:hbfunctional}), it is assumed that the distribution function 
+$f_B$ is spattially homogeneous. Clearly this assumption simplifies the demostration 
+of the $H$ theorem, nevertheless, it introduces as well a limitate-conception
+of an out of equilibrium condition in the gas. Given the relatively simple nature of a dilute 
+gas, one of the main characteristics of an out of equilibrium condition would be the existence of
+inhomogeneities in the system.
 
 %---------------------------
-\subsection{Alternative classical $H$-functional}
+\subsection{Non-homogeneous classical $H$-functional}
 
@@
 that the distribution function is homogeneous and that the hypothesis of
-molecular chaos is surmised. However, in order to consider systems whose
-distribution functions are not necessarily homogeneous, which
-allows for studying systems out of equilibrium,
-we introduce a modified $H$-functional as follows. Hereafter, we will
-use the term homogeneous distribution function to refer to a distribution
-function thad does not depend on the coordinates. In addition,
-in this section we will use primed functions
-and quantities to differentiate the classical case (this Section) \textit{vs} the quantum
-case (below, Section \ref{sec:quantum}).
+molecular chaos is fullfiled. In order to extend the procedure for systems whose
+distribution functions are not necessarily homogeneous, would possibilities allows 
+for studying systems been in a more general out of equilibrium condition, we introduce a modified $H$-functional 
+as follows. For the sake of clearity and simplicity we will use primed functions
+and quantities to denotate the classical case and differentiate it from the quantum
+case.
 
-Our proposed functional, denoted as $\mathcal{H}'$, describes a diluted classical
-gas occupying a volume, $V$,
+Our proposed functional, denoted as $\mathcal{H}'$, describes a dilute classical
+gas occupying a volume $V$. For our theoretical treatment we will consider this volume
 divided into $K$ cells, which, without loss of generality,
-have local identical volumes (\textit{i.e.} $\delta V_M = V/K,\ M=1,\,\dots\,,K$).
-Each cell has its own local
-$H$-functional, $\mathcal{H}'_M$, and local homogeneous distribution function,
-$f'_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
-local temperature, $T'_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
-global energy $E$, and a global number of free
-classical particles $N$, and it is considered to be perfectly isolated (\textit{i.e.} the
-global system is a microcanonical ensemble). We also assume that the number of particles
+have identical volumes $\delta V_M = V/K,\ M=1,\,\dots\,,K$.
+Locally the $\mathcal{H}'$-functional is denoted by $\mathcal{H}'_M$, and 
+the local homogeneous distribution function, $f'_{M}(\vec{v},t)$; local 
+number of particles, $\mathcal{N}'_M$; local temperature, $T'_M$; and a local 
+energy, $\mathcal{E}'_M$. As a whole, the system has a
+energy $E$, and a global number of free classical particles $N$. Let us assume that the system is 
+to be isolated. We also assume that the number of particles
 in each cell is sufficiently large, so as to obtain accurate averages.
-The functional takes the form
+We start our analysis by proposing the following inhomogeneous functional
 %
@@
 relative to the global system, and on the velocity
-$\vec{v}$ and time $t$. Each $f'_M$ can be formally extended to the complete coordinate space
+$\vec{v}$ and time $t$. Notice that each $f'_M$ can be formally extended to the complete coordinate space
 by defining each $f'_M$ to be zero outside the $M$-th cell, in such a manner that
@@
 allows us to omit the symbol $\delta V_M$ on all integrals performed over the cell volume.
-In terms of $f'_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
+In terms of $f'_M(\vec{v},t)$ and a local variable of energy, $\epsilon(\vec{v})$, we have:
 %
@@
 %
-Notice that assuming homogeneous distribution functions per cell means that we are surmising the 
-local equilibrium hypothesis. In addition, the set $\{f'_{M}(\vec{v},t)\}$ must satisfy the following
-microcanonical restrictions:
+Notice that assuming homogeneous distribution functions per cell means that we are accepting valid the 
+local equilibrium hypothesis. In addition, the set $\{f'_{M}(\vec{v},t)\}$ must satisfy the following restrictions:
 %
@@
 
-We now use the variational method in order to maximize $\mathcal{H}'$, applying the restrictions
-given in Eqs.~(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) through Lagrange
-multipliers ($\{\alpha_M\}$ and $\{\beta_M\}$). This yields
+We now use the variational method in order to find the extremal of $\mathcal{H}'$, consistent with the restrictions
+(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) by using the corresponding Lagrange
+multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields
 %
-\begin{eqnarray}
+\begin{eqnarray}\label{eq:deltaHpdeltafpj}
     \frac{\delta \mathcal{H}'}{\delta f'_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
@@
 %
-where $C$ is a constant. We notice that using the variational procedure on
-$\mathcal{H}'$ predicts that the
+where $C$ is a constant. We notice that the use of the variational procedure on
+$\mathcal{H}'$ predicts that when the equilibrium is reached, the
 distribution function of each cell has the form of the Maxwell-Boltzmann
@@
 $\forall M$). Several properties follow directly from this, \textit{e.g.}
-from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$ (here we have used Eqs. (\ref{eq:Ecell})
-and (\ref{eq:Ncell})),
-and Eq.~(\ref{eq:CH2}), in terms of Eq.~(\ref{eq:hbfunctional}), reduces to
+from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$. Here we have used Eqs. (\ref{eq:Ecell}) and (\ref{eq:Ncell}).
+Eq.~(\ref{eq:CH2}), in terms of Eq.~(\ref{eq:hbfunctional}), can be cast into
 %
@@
 %
-In order to identify $\mathcal{H}'$ with the entropy, it remains to show that $\mathcal{H}'(t)$ is extensive,
+In order to identify $\mathcal{H}'$ with the entropy, still we need to show that $\mathcal{H}'(t)$ is extensive,
 with respect to $Kf'_(\vec v,t)$.
@@
 %
-We observe that if the number of particles in the $\mu$-space, $f'(\vec{v},t)$, is much greater than
+We observe that if the number of particles in the $\mu$-space, $f'(\vec{v},t)$, is much large than
 the number of cells $K$, then the first term of Eq.~(\ref{eq:sum-h}) is negligible, and consequently:
@@
 \textit{i.e.} $\mathcal{H}'_M$ is extensive, and the sum $\sum_M\mathcal{H}'_M$ is the $H$-functional of the complete
-system, which reduces to the Boltzmann $H$-functional. Hence, $\mathcal{H}'$ can also be identified with the
-system's entropy.
+system, which reduces to the Boltzmann $H$-functional.
+\commb{Hence, $\mathcal{H}'$ can also be identified with a density of entropy for the system.}
 
@@
 %
-however $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$.
-Since $N$ is a constant and we are mainly intersted in the time-derivative of $\mathcal{H}''$,
-$C$ can be conveniently omitted. In other words, $\mathcal{H}'$ renders the Maxwell-Boltzmann
+since $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$.
+$N$ is a constant and we are mainly intersted in the time-derivative of $\mathcal{H}''$, thus
+$C$ can be conveniently omitted. In other words, $\mathcal{H}'$ leads the Maxwell-Boltzmann
 distribution function for systems in equilibrium.
 
-\subsubsection{Alternative proof of the classical $H$-theorem}
+\subsubsection{Proof of the $H$-theorem for non-homogeneous distributions}
 
-Throughout this section, we will consider a classical gas
-whose initial condition is a state close to equilibrium,
+Along this section, we consider a classical gas
+whose initial condition is close to the equilibrium,
 which ensures that the local equilibrium hypothesis remains valid during
 the time-evolution of the system.
-Also we will use the following definitions:
+Also we will use the following definitions for the deviations of concentration 
+and energy from the equilibrium values:
 \begin{subequations}\label{eq:restrictionoutsideclassical}
@@
 %
-Here $\bar {\mathcal{N}}'\ (=N/K)$ and $\bar{\mathcal{E}}'\ (=E/K)$ are
-the local particle number and the local energy in equilibrium, respectively, which are given by:
+Here $\bar {\mathcal{N}}' =N/K$ and $\bar{\mathcal{E}}'=E/K$ are
+the cell particle number and the cell energy in equilibrium, respectively, they are given by:
 %
@@
 Substituting $\ln \bar f'(\vec{v})=\exp\left(\alpha+\beta \epsilon(\vec{v})\right)$,
-see Eq.~(\ref{eq:fbardef}) and text following it, and omitting $C$ yields
+see Eq.~(\ref{eq:fbardef}) and text following it, and omitting $C$ we obtain
 %
@@
 In summary, if a gas occupying a volume V (which is divided into $K$ small cells), with a
-total energy $E$ and $N$ classical free particles, whose initial state is not in equilibrium,
-but sufficiently close to equilibrium ---such that the local equilibrium hypothesis apply---, then the funcional
+total energy $E$ and $N$ classical free particles, whose initial state is not in the equilibrium,
+but sufficiently close to equilibrium, then the funcional
 %
@@
 In Eq.~(\ref{eq:CH3}), $f'_M(\vec{v},t)$ is the Maxwell-Boltzmann distribution
-function, which in general is different for different cells (\textit{i.e.} the complete system
-can be non-homogeneous), and each $f'_M(\vec v,t)$ is compatible
-with the cell number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,
+function, which in general is different for different cells \textit{i.e.} the complete system
+can be non-homogeneous, and each $f'_M(\vec v,t)$ is compatible
+with the cell properties, such as number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,
 $T_M$, and Legendre multipliers $\alpha_M$ and $\beta_M$.

### 2021-01-10 tex/quantum.tex
Los los cambios que hizo el Dr. Carrillo hasta a página 10 fueron agregados

--- 2021-01-06 tex/quantum.tex
@@
 The classical $H$-theorem is even to these days one of the pillars whereon the classical statistical
-mechanics is founded. Unfortunately, regardless of several attempts have been made
+physics is founded. Unfortunately, regardless of several attempts have been made
 \commr{[CITE Q. H-THEOREM EXAMPLES]}, 
-the generality of the classical $H$-theorem has no quantum match. In this section, we shall
+the generality of the classical $H$-theorem has no an equally robust quantum match. In this section, we shall
 propose and analize an alternative quantum $H$-functional using the variational method.
 We start by briefly outline a typical textbook demonstration of the quantum $H$-theorem
-\cite{bib:tolman}, followed by the analysis of our proposed $H$-functional.
+\cite{bib:tolman}, then, we present the analysis of our proposed $H$-functional.
 
@@
 %
-where $G$ is the total number of possible quantum states of the gas that satisfy the
-above conditions \cite{bib:tolman}.
+where $G$ describes the total number of accessible quantum states of the gas that satisfy the
+above mentioned conditions \cite{bib:tolman}.
 The quantum $H$-theorem can be demostrated as follows.  $G$ can be divided into groups of neighboring states,
@@
 
-Arguing that the energy exchange between particles is dominated by collisions between particles,
-and using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
+Assuming that the energy exchange between particles is produced by collisions between particles.
+By using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
 %
@@
 It is important to remark that in deriving Eq.~(\ref{eq:changen}), the equal a priori
-probabilities and the random a priori phases hypotheses were surmised, and that 
+probabilities and the random a priori phases hypotheses were taken valid. The 
 the random a priori phases hypothesis can be
 considered an analogous to the molecular chaos hypothesis
-\cite{bib:das2018}.
+\cite{bib:das2018} as the mechanism to introduce stochasticity into the system.
 
-Substituting Eq.~(\ref{eq:changen}) into Eq.~(\ref{eq:quantumdHdt}), and performing
-the algebra, it can be proven that
+Substituting Eq.~(\ref{eq:changen}) into Eq.~(\ref{eq:quantumdHdt}) after some algebra, it can be proven that
 %
 \begin{equation}
-    \frac{dH_T}{dt}\leq 0.
+    \frac{dH_T}{dt}\leq 0. \label{eq:H-theorem-tolman}
 \end{equation}
@@
 bosons. For our purposes, the volume $V$ is divided into $K$ small cells, each of which
-has a constant volume $\delta V_M=V/K$ ($M=1,\,\dots\,,K$), a
-local temperature $T_M$, a local energy $\epsilon_M$,
-a local number of particles $\mathcal{N}_M$, and a local set of distribution functions,
-$\{f_{Mn}(t)\}$. Hereafter we will use the short-hand notation:
+has a constant volume $\delta V_M=V/K$ ($M=1,\,\dots\,,K$), a temperature $T_M$, an energy $\epsilon_M$,
+a number of particles $\mathcal{N}_M$, and a distribution function,
+$\{f_{Mn}(t)\}$. Hereafter we will use a short-hand notation:
 %
@@
 Here, the upper and lower sign refer to bosons and fermions respectively.
-In addition, when needed, each cell will have an associated local chemical potential,
+In addition, when needed, each cell will have an associated chemical potential,
 $\alpha_M$, and its local $H$-functional, which is defined by:
@@
 
-Applying the variational method to $\mathcal{H}$, and using the using the
+Applying the variational method to $\mathcal{H}$, and using the
 Lagrange multipliers
-$\{\alpha_M\}$ and $\{\beta_M\}$ it is straightforward to obtain
-(see also text around Eq.~(\ref{eq:deltaHpdeltafpj})):
+$\{\alpha_M\}$ and $\{\beta_M\}$, it is straightforward to obtain
+(see also the discussion of Eq.~(\ref{eq:deltaHpdeltafpj})):
 %
@@
 \end{equation}
+%
+Thus, in this zero-order approximation the form of the distribution function is conserved.
 
@@
 %============================================================
-If the system is in equilibrium (at $t\to\infty$), the local number of particles and the local
-energy do not depend on the cell number and they should be homogeneous, this is \commr{[This needs a rationale\dots]}
+According to Eq.~(\ref{eq:H-theorem-tolman}) the system is in equilibrium at $t\to\infty$, then 
+the cell number of particles and the cell energy do not depend on the cell number, this is \commr{[This needs a rationale\dots]}
 %

### 2021-01-11 tex/quantum.tex
Rewording: up to quantum H-theorem proof.

--- 2021-01-09 tex/quantum.tex
@@
     \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} \bigg[ f_{Mn}(t) \ln f_{Mn}(t)\bigg.\nonumber \\
-    &&\bigg.\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M.
+    &&\qquad\qquad\bigg.\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M.
 \end{eqnarray}
@@
 In addition, when needed, each cell will have an associated local chemical potential,
-$\alpha_M$, and its local $H$-functional, which is defined by:
+$\alpha_M$, and a local $H$-functional, which is defined by:
 %
@@
 %
-\begin{subequations}
+\begin{subequations}\label{eq:qGlobRest}
 \begin{equation}\label{eq:qGlobRestN}
@@
 %============================================================
-\subsubsection{Properties of $\mathcal{H}$ in equilibrium}
+\subsubsection{Properties of $\mathcal{H}$ for systems in equilibrium}
 %============================================================
@@
 is the negative of the entropy. The property provided by Eq.(\ref{eq:varEntropyDef})
-is precisely what we call the ``variational entropy'',
+is precisely what we refer to as ``variational entropy'',
 and this name reflects the fact that it was obtained \textit{via}
@@
 
+For quantum systems, we also accept the validity of the local equilibrium hypothesis
+for every cell in the system. This allows us to define non-homogeneous systems,
+wherein thermodynamic quantities are well-defined on a cell-per-cell basis,
+and in terms of the equilibrium properties, we have
+
+%
+\begin{subequations}\label{eq:restrictionoutside}
+\begin{equation}
+        \mathcal{N}_M(t)=\sum_{n}f_{Mn}(t)=\bar{\mathcal{N}}+\Delta_M(t) 
+\end{equation}
+  and
+\begin{equation}
+        \mathcal{E}_M(t)=\sum_{n}\epsilon_{n}f_{Mn}(t)=\bar{\mathcal{E}}+ \delta_M(t).
+\end{equation}
+\end{subequations}
+%
+In Eqs.~(\ref{eq:restrictionoutside}) $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle
+number and the local energy of cells in equilibrium, which are given by Eqs.~(\ref{eq:qEqRestrictions}),
+and $\Delta_M$ and $\delta_M$ are deviations from $\bar{\mathcal{N}}$
+and $\bar{\mathcal{E}}$, respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$
+and $\delta_M(t) \ll \bar{\mathcal{E}}$.
+
+In the present context, $|\Delta_M|$ and $|\delta_M|$ are sufficiently large so they are not
+fluctuations of the system, but sufficiently small so that the local equilibrium hypothesis is
+valid for $t>0$ (we set $t_0=0$, and $t_0$ is the initial time at which the system is prepared).
+Therefore, it is reasonable to re-write the distribution functions as:
+%
+\begin{equation}\label{eq:qFirstOrd}
+   f_{Mn}(t)=\bar{f}_n(1+g_{Mn}(t)),\quad
+   1\gg|g_{Mn}(t)|,
+\end{equation}
+%
+from which it follows, by substituting Eq.~(\ref{eq:qFirstOrd}) into
+Eq.~(\ref{eq:restrictionoutside}), that $\Delta_M$ and $\delta_M$ satisfy:
+%
+\begin{subequations}
+\begin{equation}
+    \Delta_M(t)=\sum_n \bar{f}_n g_{nM}
+\end{equation}
+%
+and
+%
+\begin{equation}
+	\delta_M(t)=\sum_n  \bar{f}_n g_{nM}\epsilon_n.
+\end{equation}
+\end{subequations}
+%
+
+\commr{
+An additional consideration is necessary for treating gases composed of fermions,
+which must obey the relation $f_{Mn}(t)\leq1$, hence
+%
+\begin{eqnarray}\label{eq:fermionrestriction}
+   1-\bar f_n -\bar f_n g_{nM}\geq0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}\geq1+g_{nM}.
+\end{eqnarray}
+%
+$\bar f_{n}=1$ is certainly satisfied if the system is at zero temperature. In this state,
+all energy levels below and including the Fermi energy are occupied because of
+the exclusion principle, therefore the system will
+necessarily be homogeneous , and consequently $g_{nM}=0$. In this article, we will
+omit this scenario and will only discuss Fermi gases whose temperature is non zero.
+}
+
+To proof the quantum $H$-theorem, start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
+%
+\begin{equation}\label{eq:deltaH}
+   \frac{d \mathcal{H} (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M.
+\end{equation}
+%
+Subsequently, substitute Eq.~(\ref{eq:qFirstOrd}) in the above equation to obtain:
+%
+\begin{eqnarray}\label{eq:cambioH1}
+    \frac{d\mathcal{H} (t)}{dt}&=&
+      \sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[
+        \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})}
+      \right]\dot{g}_{nM} \delta V_M \nonumber \\
+    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{
+      \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}
+      -\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}
+    \right \}\delta V_M,
+\end{eqnarray}
+%
+Approximate the logarithmic terms, corresponding to
+Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$, and
+substitute this into Eq.~(\ref{eq:cambioH1}), which gives:
+%
+\begin{subequations}\label{eq:qlnApprox}
+\begin{equation}\label{eq:qlnApproxFermions}
+	\ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]
+    	\approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}
+\end{equation}
+%
+and
+%
+\begin{equation}\label{eq:qlnApproxBosons}
+    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM},
+\end{equation}
+\end{subequations}
+%
+respectively. Eq.~(\ref{eq:qlnApproxFermions}) is valid because, for non-extremely degenerated Fermi gases, 
+$1-\bar{f}_n \gg \bar{f}_n|g_{nM}|$, and Eq.~(\ref{eq:qlnApproxBosons}) is fulfilled because,
+for Boson gases,
+$1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$ when $\bar{f}_n \gg \bar{f}_n |g_{nM}|$.
+
+Combine Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}):
+%
+\begin{eqnarray}\label{eq:cambioH2}
+    \frac{d\mathcal{H}}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
+    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M,
+\end{eqnarray}
+%
+and substitute Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
+%
+\begin{equation}\label{eq:cambioH3}
+    \frac{d\mathcal{H}}{dt}=
+       \sum_n \sum_{M=1}^{K} \bar{f}_n\left \{
+          (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}
+           +g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM}
+       \right \} \de V_M.
+\end{equation}
+%
+
+Since both the total number of particles and the total energy of the system are constant, it follows
+from Eqs.~(\ref{eq:qGlobRest}) and (\ref{eq:restrictionoutside}) that:
+%
+\begin{subequations}
+\begin{equation}
+\frac{dN}{dt}=\sum_{M=1}^K\dot{\mathcal{N}}_M\delta V_M
+   =\sum_{M=1}^K\sum_n \bar{f}_n \dot{g}_{nM}\delta V_M=\sum_{M=1}^K\dot{\Delta}_M(t)\delta V_M=0
+\end{equation}
+%
+and
+%
+\begin{equation}
+\frac{dE}{dt}=\sum_{M=1}^K\dot{\mathcal{E}}_M\delta V_M
+   =\sum_{M=1}^K\sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n\delta V_M=\sum_{M=1}^K\dot{\delta}_M(t)\delta V_M=0.
+\end{equation}
+\end{subequations}
+%
+
+Substitute the previous expression in Eq.~(\ref{eq:cambioH3}) to obtain:
+%
+\begin{equation}\label{eq:cambioH4}
+   \frac{d\mathcal{H}}{dt}=\sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}
+   \sum_M  g_{nM}\dot{g}_{nM} \delta V_M \leq 0.
+\end{equation}
+%
+
+For obtaining the right-most hand side of Eq.~(\ref{eq:cambioH4}), 
+we have exploited the relation $g_{nM}\dot{g}_{nM}\leq0$ for $t>0$. This can be proven 
+by simply arguing that, in the initial
+state, if a cell is described by $g_{nM}(t_0)>0$ then $g_{nM}(t)\geq0$ and $\dot g_{nM}(t)\leq0$, and if
+$g_{nM}(t_0)<0$ then $g_{nM}(t)\leq0$ and $\dot g_{nM}(t)\geq0$. Here we have used the facts that
+the system in equilibrium is homogeneous, and that, by accepting the local equilibrium hypothesis, $g_{Mn}(t)$
+is a monotonic function and $g_{nM}\to0$ as $t\to\infty$ as the system aproaches to the equilibrium state.
+Another approach to prove Eq.~(\ref{eq:cambioH4}) consists of splitting the cells into two subsets,
+such as we did in the classical scenario, however, we will ommit this approach in this section.
+
 \color{blue}
 
-We rewrite the \textit{variational entropy} (\ref{eq:qHdef}) as
-%
-\begin{equation}
-    \mathcal{H}(t)=\sum_{M=1}^{K} \sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right] \de V_M \delta \epsilon_n \label{entropy2},
-\end{equation}
-%
-where, for simplicity, we have defined $f(\epsilon_{n},t)\equiv f_{Mn}$.
-
-This distribution function must satisfy the local equilibrium hypothesis in
-each cell ($f_{nM}$ must be spatial-homogeneous for each $M$), the
-non-homogeneous distribution function assumption in the total volume ($f_{nM}$
-must be different among all cells), and the following restrictions
-%
-\begin{eqnarray}
-        \sum_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \sum_{n}\epsilon_{n}f_{Mn} \delta \epsilon_n=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
-  \end{eqnarray}
-%
-where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle
-number and the local energy in equilibrium
-%
-\begin{equation}
-      \bar{\mathcal{N}}= \sum_n \bar{f}_n \delta \epsilon_n; \ \ \ \ \bar{\mathcal{E}}= \sum_n \epsilon_n\bar{f}_n \delta \epsilon_n,
-  \end{equation}
-%
-and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$
-and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$
-and $\delta_M(t) \ll \bar{\mathcal{E}}$. The previous conditions and
-restrictions are the quantum analogous case of the classical case. Also, this
-system will suffer, as in the classical case, a \textit{relaxation process}.
-
-  The quantities $\Delta_M,\delta_M$ are sufficiently big to be different from
-fluctuations in the system but sufficiently small to consider the system not
-far from the equilibrium state, as in the classical case.
-
-If we perform the derivative of the variational entropy with respect to time,
-we obtain
-%
-\begin{equation}
-   \frac{d \mathcal{H} (t)}{dt}= \sum_n \sum_{M=1}^{K} \dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right] \de V_M \delta \epsilon_n.\label{deltaH}
-\end{equation}
-%
-Using the first-order approximation (\ref{firstorder}), we get
-%
-\begin{eqnarray}
-    \frac{d\mathcal{H} (t)}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_{n}\ln \left[ \frac{\bar{f}_{n}(1+g_{nM})}{1\mp \bar{f}_{n} (1+ g_{nM})} \right]\dot{g}_{nM} \de V_M\delta \epsilon_n \nonumber \\
-    &=&\sum_n \sum_{M=1}^{K} \bar{f}_n \left \{ \ln [\bar{f}_n+\bar{f}_n g_{nM}]\dot{g}_{nM}-\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}  \right \}\de V_M \delta \epsilon_n.\nonumber \\
-    \label{cambioH1}
-\end{eqnarray}
-%
-It's necessary to remark that, in the case of fermions
-%
-\begin{eqnarray}
-   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}
-\end{eqnarray}
-%
-The previous expression establishes that the value of $g_{nM}$ is determined by
-$\bar f_n$. Furthermore, whether $\bar f_{n}=1$, this is, all lower energy
-levels are occupied, then the system will not have inhomogeneities because of
-the exclusion principle, and as a consequence of this,  $g_{nM}=0$. This result
-corresponds to the system in the zero temperature condition. We can observe
-that in this condition, (\ref{fermionrestriction}) is violated. Due to this,
-the zero temperature condition is excluded.
-
-We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$
-then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$. In fermions, in the case of a
-non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n
-|g_{nM}|$. With those relations, we can approximate the logarithm functions in
-(\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$
-%
-\begin{equation}
-    \ln [\bar{f}_n+\bar{f}_n g_{nM}] \approx \ln [\bar{f}_n]+ g_{nM}; \ \ \ \ \ln[1\mp\bar{f}_n\mp\bar{f}_n g_{nM}] \approx \ln[1\mp\bar{f}_n]\mp\frac{\bar{f}_n}{1\mp\bar{f}_{n}} g_{nM}.
-\label{lnapproximation}
-\end{equation}
-%
-With the previous approximation, (\ref{cambioH1}) will be
-%
-\begin{eqnarray}
-    \frac{d\mathcal{H}}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M \delta \epsilon_n.\label{cambioH2}
-\end{eqnarray}
-%
-Making use of the expression (\ref{relation}),
-(\ref{cambioH2}) casts into
-%
-\begin{eqnarray}
-    \frac{d\mathcal{H}}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\bar{\alpha}+\bar{\beta}{\epsilon}_n)\dot{g}_{nM}+ g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \} \de V_M \delta \epsilon_n.
-\nonumber \\
-    \label{cambioH3}
-\end{eqnarray}
-%
-On the other hand, we obtain from the restrictions the following expressions
-%
-\begin{eqnarray}
-    &&\sum_n \bar{f}_n g_{nM} \delta \epsilon_n=\Delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM} \delta \epsilon_n=\dot{\Delta}_M(t), \nonumber \\
-    &&\sum_n  \bar{f}_n g_{nM}\epsilon_n \delta \epsilon_n=\delta_M(t) \ \  \Rightarrow \ \  \sum_n \bar{f}_n \dot{g}_{nM}\epsilon_n \delta \epsilon_n=\dot{\delta}_M(t)
-\end{eqnarray}
-%
-and as a consequence of $\sum_{M=1}^{K} \Delta_M(t) \de V_M  =\sum_{M=1}^{K} \delta_M(t) \de V_M =0$, we find
-%
-\begin{equation}
-    \sum_{M=1}^{K} \dot{\Delta}_M(t) \de V_M =\sum_{M=1}^{K} \dot{\delta}_{M}(t) \de V_M=0.
-\end{equation}
-%
-Substituting the previous expression to (\ref{cambioH3}) we obtain
-%
-\begin{eqnarray}
-   \frac{d\mathcal{H}}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M  g_{nM}\dot{g}_{nM} \de V_M\delta \epsilon_n. \nonumber \\ \label{cambioH4}
-\end{eqnarray}
 %
@@
 When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot
-g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{d\mathcal{H}}{dt}=0$. Then, we can
+g_{nM}=0$ and finally from (\ref{eq:cambioH1}) $\frac{d\mathcal{H}}{dt}=0$. Then, we can
 say that any ideal quantum gas perturbed always evolves to the equilibrium

### 2021-01-11 tex/quantum.tex
Rewording: finished 1st round of sec. 3.

@@
 
-\color{blue}
-
+Succinctly, considering a diluted quantum gas contained in a vessel of volume $V$ (divided into
+$K$ small cells), with total energy $E$ and $N$ quantum free particles, which initially
+is slightly out of equilibrium ---so as to the local equilibrium hypothesis is valid---,
+the functional
 %
-Now, the summation over $M$ will be expressed in two summations, such as in the
-classical case
-%
-\begin{equation}
-    \frac{d\mathcal{H}}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
-\end{equation}
-%
-where $L+P=K$. Besides, $\dot{g}^{+}_{J}$ represents the change on the
-deviation on cells that have an excess of particles or energy while
-$\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have
-missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells
-that have an excess of particles or energy. In contrast, $g^{-}_{J}$ represents
-the value of the deviation on cells that have missing particles or energy.
-
-Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles
-and/or energy and so, $g^{+}_{nJ}>0$. On the other hand, $\dot{g}^{-}_{nJ}>0$
-describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.
-
-We sort the previous ideas in the following form
-%
-\begin{eqnarray}
-   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
-   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
+\begin{eqnarray}\label{eq:qHfunctTheo}
+    \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} \bigg[ f_{Mn}(t) \ln f_{Mn}(t)\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M,
 \end{eqnarray}
 %
-
-and consequently, (\ref{cambioH5}) obtains the following form
-%
-\begin{equation}
-    \frac{d\mathcal{H}}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
-\end{equation}
-%
-and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then
-$\frac{d\mathcal{H}}{dt}<0$. With this, we proved that any quantum ideal gas perturbed
-(out of equilibrium but not so far from it), evolves such that
-$\frac{d\mathcal{H}}{dt}<0$.
-
-When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot
-g_{nM}=0$ and finally from (\ref{eq:cambioH1}) $\frac{d\mathcal{H}}{dt}=0$. Then, we can
-say that any ideal quantum gas perturbed always evolves to the equilibrium
-state, such as the classical $H$-theorem. Then, we obtained the quantum version
-of the Boltzmann's $H$-theorem including inhomogeneities in the system.
-
-We obtained some results from the Statistical Mechanic in Quantum Mechanics
-scheme. However, the correspondence principle is important to recover all
-results from Classical Mechanics. This correspondence principle will be
-discussed in the next section.
-
-  %-----------------------------------------
-
-%---------------------------------------------
+where $f_{Mn}$ is the $M$-th cell distribution function,
+satisfies $d\mathcal{H}/dt\leq0$, and the equality condition is attained when the system approaches
+the equilibrium state at $t\to\infty$. In Eq.~(\ref{eq:qHfunctTheo}), and for a Fermi (Bose)
+gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distrubution function for each
+cell and locally each cell is in equilibrium, although the complete system may be non-homogeneous,
+and is caracterized by the respective $f_{Mn}$, number of particles $\mathcal{N}_M$, energy $\mathcal{E}_M$,
+temperature $T_M$, and Legendre Multipliers $\alpha_M$ and $\beta_M$.

### 2021-01-11 tex/introduction.tex

@@
 In the literature, the theoretical basis that allows us to describe equilibrium
-systems in the classical scheme is well-established.
+systems in the classical scheme are well-established.
 The procedures to describe
-equilibrium systems are well-known and work in almost every system in nature
+equilibrium systems are well-known and work properly for the treatment of almost every system in nature
 (Thermodynamics for a phenomenological description and statistical mechanics
 for a microscopic description). In contrast, a unified mathematical model to
-describe the evolution of any system to an equilibrium state has not developed.
+describe the evolution of any system from an arbitrary out of equilibrium condition to an equilibrium state has not been developed yet.
 Nevertheless, developments can be found in the literature and describe the
@@
 
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas
+In the Kinetic Theory of gases, the behavior of a dilute classical gas
 was described by Boltzmann through the Boltzmann transport equation.

### 2021-01-14 tex/classical.tex
Rewording: first paragraph of sec. 4

@@
 %
-\begin{equation}\label{eq:CH2}
+\begin{equation}\label{eq:cHdef}
    \mathcal{H}'(t)=\sum_{M=1}^{K}\int_{\delta V_M} f'_M(\vec{v},t) \ln f'_M(\vec{v},t)\mathrm{d}^3v.
@@
 from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$. Here we have used Eqs. (\ref{eq:Ecell}) and (\ref{eq:Ncell}).
-Eq.~(\ref{eq:CH2}), in terms of Eq.~(\ref{eq:hbfunctional}), can be cast into
+Eq.~(\ref{eq:cHdef}), in terms of Eq.~(\ref{eq:hbfunctional}), can be cast into
 %
@@
 With the previous considerations, in the following we provide an alternative demonstration
-of the $H$-theorem, considering the $H$-functional, $\mathcal{H}'$, defined by Eq.~(\ref{eq:CH2}).
+of the $H$-theorem, considering the $H$-functional, $\mathcal{H}'$, defined by Eq.~(\ref{eq:cHdef}).
 
-We commence by differentiating Eq.~(\ref{eq:CH2}) with respect to time,
+We commence by differentiating Eq.~(\ref{eq:cHdef}) with respect to time,
 %

### 2021-01-14 tex/qccorrespondence.tex
Rewording: finished 1st round of sec. 3.
Rewording: first paragraph of sec. 4

@@
 \section{Quantum Classical correspondence}\label{sec:qccorrespondence}
+
+In sections \ref{sec:classical} and \ref{sec:quantum}, we have seen that the variational
+method can be applied to $H$-functionals, which correctly describes the behavior of
+classical and quantum diluted gases, as to their respective time-evolution.
+Both $H$-functionals defined in Eqs.~(\ref{eq:cHdef}) and
+(\ref{eq:qHdef}) also recovers the well-known distribution functions, either Maxwell-Boltzmann
+for a classical gas, Fermi-Dirac for a Fermi gas, or Bose-Einstein for a Bose gas. Nevertheless,
+the functionals (\ref{eq:cHdef}) and
+(\ref{eq:qHdef}) are seemingly different, and in this section we will show that both are
+related by the correspondence principle. To this end,
+
+\color{blue}
+
 %The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$-functional in the appropriated limit.\\
@@
 \end{eqnarray}
-where $H_{B}$ now is (\ref{eq:CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{bib:tolman} eq. (102.6) for more information.}\\
+where $H_{B}$ now is (\ref{eq:cHdef}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{bib:tolman} eq. (102.6) for more information.}\\
 We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.

### 2021-01-14 tex/quantum.tex
se agregaron los cambios en la sección de quantum, agregué un paquete en el manuscript que lo dejé como comentario para incluir el cuadro de la demostración de un teorema

@@
 \begin{equation}
-\mathcal{E}_M(t) = \sum_{n}f_{Mn}(t)\epsilon_{n} \de V_M
+\mathcal{E}_M(t) = \sum_{n}f_{Mn}(t)\epsilon_{n} \de V_M,
 \end{equation}
@@
 %
-which are, in addtion, constrained by the global microcanonical restrictions
+which are, for the whole system, constrained by the microcanonical restrictions
 %
@@
     \sum_{M=1}^{K}\left[\sum_{n}f_{Mn}(t)\right]\delta  V_M
-    =\sum_{M=1}^{K}\mathcal{N}_{M}(t)\delta V_M=N
+    =\sum_{M=1}^{K}\mathcal{N}_{M}(t)\delta V_M=N,
 \end{equation}
@@
 %
-Thus, in this zero-order approximation the form of the distribution function is conserved.
+Thus, in this zero-order approximation the form of equilibrium distribution functions are conserved.
 
@@
 %============================================================
-If the system is in equilibrium (at $t\to\infty$), the temperature becomes homogeneous throughout
+If the system is in equilibrium (we will show that it will also be valid at $t\to\infty$), the temperature becomes homogeneous throughout
 the complete system. Also, the local number of particles, the local
 energy, and the Lagrange multipliers do not depend on the cell number and they should be homogeneous,
-this is \commr{[This needs a rationale\dots]}
+this is 
 %
@@
 By means of the above equation, we can recover the distribution function and the entropy of a diluted quantum gas in equilibrium as follows. Setting $\bar\alpha=\mu/kT$ and $\bar\beta=-1/kT$,
-and substituting them in Eq.~(\ref{eq:qfneq}) renders the Fermi-Dirac and Bose-Einstein
+and substituting them in Eq.~(\ref{eq:qfneq}), it renders the Fermi-Dirac and Bose-Einstein
 distribution functions:
@@
 %
-and substituting Eq.~(\ref{eq:qfneq}) into the negative of Eq.(\ref{eq:qHdef}) provides the entropy:
+and substituting Eq.~(\ref{eq:qfneq}) into the negative of Eq.(\ref{eq:qHdef}), 
+the entropy of a quantum ideal gas is obtained
 %
@@
 %
-For obtaining the last result, we have used the well-known fact that the $H$-functional
-is the negative of the entropy. The property provided by Eq.(\ref{eq:varEntropyDef})
-is precisely what we refer to as ``variational entropy'',
+This quantity is what we refer to as ``variational entropy'',
 and this name reflects the fact that it was obtained \textit{via}
-the variational method.
+the variational method. We will use it at the end of this article to analyze relaxation processes.
 
@@
 %
-In Eqs.~(\ref{eq:restrictionoutside}) $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle
-number and the local energy of cells in equilibrium, which are given by Eqs.~(\ref{eq:qEqRestrictions}),
+In Eqs.~(\ref{eq:restrictionoutside}) $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the cell particle
+number and the cell energy in equilibrium, which are given by Eqs.~(\ref{eq:qEqRestrictions}),
 and $\Delta_M$ and $\delta_M$ are deviations from $\bar{\mathcal{N}}$
@@
 In the present context, $|\Delta_M|$ and $|\delta_M|$ are sufficiently large so they are not
-fluctuations of the system, but sufficiently small so that the local equilibrium hypothesis is
+fluctuations of the system, but enough small so that the local equilibrium hypothesis is
 valid for $t>0$ (we set $t_0=0$, and $t_0$ is the initial time at which the system is prepared).
@@
 
-To proof the quantum $H$-theorem, start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
+\begin{proof}
+To proof the quantum $H$-theorem, we start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
 %
@@
       -\ln [1\mp\bar{f}_n\mp\bar{f}_n g_{nM}]\dot{g}_{nM}
-    \right \}\delta V_M,
+    \right \}\delta V_M.
 \end{eqnarray}
 %
-Approximate the logarithmic terms, corresponding to
+By representing the logarithmic terms, corresponding to
 Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$, and
-substitute this into Eq.~(\ref{eq:cambioH1}), which gives:
+substituting this into Eq.~(\ref{eq:cambioH1}), one obtains:
 %
@@
 
-Combine Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}):
+Combining Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}) yields:
 %
@@
 %
-and substitute Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
+and substituting Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
 %
@@
 
-Since both the total number of particles and the total energy of the system are constant, it follows
+Since both the total number of particles and the total energy 
+of the system are constant, it follows,
 from Eqs.~(\ref{eq:qGlobRest}) and (\ref{eq:restrictionoutside}) that:
@@
 \end{equation}
+\end{proof}
+Thus, the time evolution of the system occurs in such a way that 
+$\frac{d\mathcal{H}}{dt}$ is a decreasing function toward zero.
 %
@@
 For obtaining the right-most hand side of Eq.~(\ref{eq:cambioH4}), 
-we have exploited the relation $g_{nM}\dot{g}_{nM}\leq0$ for $t>0$. This can be proven 
+we have used the relation $g_{nM}\dot{g}_{nM}\leq0$ for $t>0$. This can be proven 
 by simply arguing that, in the initial
@@
 Another approach to prove Eq.~(\ref{eq:cambioH4}) consists of splitting the cells into two subsets,
-such as we did in the classical scenario, however, we will ommit this approach in this section.
+just as we did in the classical scenario.
 
@@
 $K$ small cells), with total energy $E$ and $N$ quantum free particles, which initially
-is slightly out of equilibrium ---so as to the local equilibrium hypothesis is valid---,
+is out of equilibrium ---but still the local equilibrium hypothesis is valid---,
 the functional
@@
 where $f_{Mn}$ is the $M$-th cell distribution function,
-satisfies $d\mathcal{H}/dt\leq0$, and the equality condition is attained when the system approaches
-the equilibrium state at $t\to\infty$. In Eq.~(\ref{eq:qHfunctTheo}), and for a Fermi (Bose)
+evolves in time such that $d\mathcal{H}/dt\leq0$, and the equality 
+condition is attained when the system reaches
+the equilibrium state. In Eq.~(\ref{eq:qHfunctTheo}), and for a Fermi (Bose)
 gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distrubution function for each

### 2021-01-17 tex/qccorrespondence.tex
Rewording: middle of sec. 4

@@
 (\ref{eq:qHdef}) are seemingly different, and in this section we will show that both are
-related by the correspondence principle. To this end,
+related by the correspondence principle.
+
+We start by arguing that, in the classic limit, $f_{nM}\approx0$ for the following reasons.
+Systems at very low temperatures, where the quantum effects cannot be ignored, are obviously excluded
+from the current analysis. For systems at sufficently large temperature, the particles
+occupy almost exclusively high-energy levels. 
+Furthermore, the energy spectrum approaches to be a continuum, as is expected by taking the limit
+$\hbar\to0$, and the number of particles per level are very close to zero.
+
+Subsequently, we substitute $f_{nM}\approx0$ into Eq.~(\ref{eq:qHdef}) to obtain:
+%
+\begin{equation}\label{h-quantic2}
+    \mathcal{H}= \sum_{M=1}^{K} \sum_n
+    \left[ f_{nM}(t)\ln \left(f_{nM}(t)\right)\right] \delta V_M
+    =\sum_{M=1}^{K} \sum_n
+    \left[ f_{M}(\epsilon_n,t) \ln \left( f_{M}(\epsilon_n,t)\right)\right] \delta V_M.
+\end{equation}
+%
 
@@
 
-%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$-functional in the appropriated limit.\\
-We begin defining the distribution function as the number of particles in a defined cell and an energy level $N_{nM}$ inside a volume $\de V_M \delta \epsilon_n$ as
-\begin{equation}
-    f_{nM}=\frac{N_{nM}}{ \de V_M \delta \epsilon_{n} }.
-\end{equation}
-Using this definition on the variational entropy (\ref{entropy2}) we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M\delta \epsilon_{n}}
-           \right)\pm 
-           \left(  
-                  1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right) \ln 
-           \left(  
-                   1\mp \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \nonumber \\
-    \label{h-quantic} 
-\end{eqnarray}
-In the classic limit, the particles occupy high energy levels, and the number of particles in most of the groups of state $n$ is small compared with the number of the elementary states $\epsilon_n$ of the group, that is
-\begin{equation}
-    \frac{N_{nM}}{ \de V_M \delta \epsilon_{n} } \approx 0,
-\end{equation}
-and with the previous approximation, the second term in (\ref{h-quantic}) is negligible, then (\ref{h-quantic}) yields
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{nM}}{ \de V_M \delta \epsilon_{n}}
-           \right)
-    \right] \de V_M \delta \epsilon_{n}. \label{h-quantic2}
-\end{eqnarray}
 In the case when the system is spatial-homogeneous, the definition of the distribution function is 
@@
            \right)
-    \right]  \de V_M \delta \epsilon_{n} = V \sum_n
+    \right]  \delta V_M \delta \epsilon_{n} = V \sum_n
     \left[  
@@
     \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{B},\label{h-quantic4}
+    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\delta V_M K H_{B},\label{h-quantic4}
 \end{eqnarray}

### 2021-01-17 tex/quantum.tex

@@
 The classical $H$-theorem is even to these days one of the pillars whereon the classical statistical
-physics is founded. Unfortunately, regardless of several attempts have been made
-\commr{[CITE Q. H-THEOREM EXAMPLES]}, 
+physics is founded. Unfortunately, regardless of several attempts have been made \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:han2015,bib:das2018,bib:vonneumann2010}, 
 the generality of the classical $H$-theorem has no an equally robust quantum match. In this section, we shall

### 2021-01-17 tex/relaxproc.tex

--- 2020-12-15 tex/relaxproc.tex
@@
-  \section{Relaxation processes in degenerated quantum gases}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
+  \section{Relaxation processes in degenerated quantum gases}\label{sec:relaxproc}
+Consider that the distribution function $f_{nM}$ is a function of the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
 \begin{equation}
-    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
+    f=f(\mathcal{N}_M^{n},t),
 \end{equation}
@@
 \end{equation}
-Besides the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$.
-This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
+Besides the local number of particles in a quantum energy state 
+is a function of time $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(t)$.\\
 In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
@@
 \begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
+    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}++\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}} \frac{\partial \mathcal{N}_J^{j}}{\partial t} \right].
 \label{transportequation1}
@@
 \begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
-    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right].
+    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}} \frac{\partial \mathcal{N}_J^{j}}{\partial t} \right].
 \nonumber \\ \label{transportequation2}

### 2021-01-17 tex/qccorrespondence.tex
Version: 20210117

@@
 
-\color{blue}
+Finally, the sum over the quatum energy levels can be replaced by an integral over the velocities
+by invoking both the uncertainty principle and the fact that, for free particles, the continuum
+energy spectrum
+can be written as a function of the velocity. Hence the quantum $H$-functional
+transforms, in the classical limit, to:
+%
+\begin{equation}\label{eq:qHclassLim}
+    \mathcal{H}= \sum_{M=1}^{K}\int C'f_M(\vec v,t)\ln\big[C' f_M(\vec v,t)\big] d\vec v,
+\end{equation}
+%
+where $C'$ collects the appropriate constants stemming from writing the energy spectrum as
+a function of $\vec v$.
 
-In the case when the system is spatial-homogeneous, the definition of the distribution function is 
-\begin{equation}
-    f_{n}=\frac{N_{n}}{ \delta \epsilon_{n} },
-\end{equation}
-and using the same limit we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K} \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right]  \delta V_M \delta \epsilon_{n} = V \sum_n
-    \left[  
-           \frac{N_{n}}{ \delta \epsilon_{n}} \ln 
-           \left( 
-                  \frac{N_{n}}{ \delta \epsilon_{n}}
-           \right)
-    \right] \delta \epsilon_{n} \nonumber \\
-    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\delta V_M K H_{B},\label{h-quantic4}
-\end{eqnarray}
-where $H_{B}$ now is (\ref{eq:cHdef}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{bib:tolman} eq. (102.6) for more information.}\\
-We can observe that if the number of particles is too big compared to the number of cells $K$, then we can conclude that the sum of the local variational entropy is equal to the variational entropy of the entire system.
-
-In the equilibrium case, we have the following variational entropy
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \ln 
-                    \left(
-                            \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}
-                  \right) \ln
-                \left(
-                        1 \mp \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1}            
-                \right) \right]. \label{equilibriumvariational}
-\end{eqnarray}
-To obtain the classical limit, that is, to recover the Maxwell-Boltzmann distribution from the Fermi-Dirac and Bose-Einstein distribution, it is necessary to hold the following approximation
-\begin{equation}
-    e^{-\alpha_M-\beta_M \epsilon}\gg 1, \label{classicalapproximation}
-\end{equation}
-and then
-\begin{equation}
-    \frac{1}{e^{-\alpha_M-\beta_M \epsilon}\pm 1} \approx e^{\alpha_M+\beta_M \epsilon}.
-\end{equation}
-If we apply (\ref{classicalapproximation}) to (\ref{equilibriumvariational}), we obtain
-\begin{eqnarray}
-    \Ss&=& \sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right. \nonumber \\
-          && \pm \left. \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}
-                  \right) \ln
-                \left(
-                        1 \mp e^{\alpha_M+\beta_M \epsilon}            
-                \right) \right],
-\end{eqnarray}
-but if (\ref{classicalapproximation}) holds, then 
-\begin{equation}
-     e^{\alpha_M+\beta_M \epsilon}\ll 1,
-\end{equation}
-consequently 
-\begin{equation}
-    \ln(1\mp e^{\alpha_M+\beta_M \epsilon}) \approx \ln 1 \approx 0,
-\end{equation}
-and finally
-\begin{equation}
-    \Ss=\sum_{M=1}^{K}\sum_{n}
-        \left[
-                e^{\alpha_M+\beta_M \epsilon} \ln 
-                    \left(
-                            e^{\alpha_M+\beta_M \epsilon}
-                    \right)
-        \right],
-\end{equation}
-that corresponds to the classical $H$-functional in equilibrium described in the energy space.
-
-Finally, as a complementary topic, in the next section, we propose a method to obtain the quantum analogous from the Boltzmann transport equation for relaxation processes.
+Because we have shown that both classical and quantum $H$-functionals, Eqs.~(\ref{eq:cHdef}) and
+(\ref{eq:qHdef}) recover the Maxwell-Boltzmann, Bose-Einstein, or Fermi-Dirac distribution
+functions in equilibrium, we will omit the quantum classical correspondence of the distribution functions,
+as it has been thoroughly discussed in textbooks.
 

### 2021-01-17 tex/quantum.tex
Version: 20210117

@@
 
-\begin{proof}
+%\begin{proof}
 To proof the quantum $H$-theorem, we start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
@@
 \end{equation}
-\end{proof}
+%\end{proof}
 Thus, the time evolution of the system occurs in such a way that 

### 2021-01-19 tex/disscussion.tex
Rewording: up to sec. 3.2.1
partial commit, rew: sec5

@@
 
+\color{black}
 \section{Discussions and remarks}\label{sec:disscussion}
 
+Traditionally, the proof of the classical $H$-theorem involves assuming that the
+distribution function is spatially homogeneous, and the proof starts with
+considering particular systems that are out of equilibrium, but in such a manner
+that their distribution functions are still homogeneous. This out-of-equilibrium systems
+are conceived considering the micro-states of the system, and the $H$-theorem determines
+the microscopic evolution of the system towards a state with the most-probable distribution function,
+given a thermodynamic state. 
+This approach does not cover systems with spatial inhomogeneities.
+To address this issue, we
+proposed a framework that may be useful to describe the time-evolution of systems
+initially out of equilibrium and spatially non-homogeneous. For this, we divided the
+system into small cells, so as to conceive a system wherein the local equilibrium
+hypothesis is valid for each cell and the global system is not homogeneous. Systems
+that satisfy the previous conditions will evolve towards global equilibrium, and the evolution is
+determined by the relations $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef}),
+and $d\mathcal{H}/dt\leq0$, Eq.~(\ref{eq:qHdef}), for classical and quantum gases, respectively.
+
+The $H$-functionals $\mathcal{H}'$ and $\mathcal{H}$ correctly recover the most-probable distribution
+functions both in out-of-equilibrium states (locally) and when the system is in the global
+equilibrium state. This enables the use of all the machinery of thermodynamics, including the
+proper definition of state variables for each cell and
+
+\color{blue}
 In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function.
@@
 \end{equation}
-Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
+Besides, we find an expression for the entropy density (\ref{eq:varEntropyDef}). This expression will be useful to describe quantum gases in terms of entropy.\\
 It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.

### 2021-01-22 tex/disscussion.tex
se agregaron los ultimos cambios, se unió la sección de discusión con conclusiones. La sección de relajación no se ha modificado hasta haber discutido lo que propone el Dr. Carrillo.

--- 2020-12-27 tex/disscussion.tex
@@
 
-\section{Discussions and remarks}\label{sec:disscussion}
+\section{Remarks and conclusions}\label{sec:disscussion}
 
 In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. 
-In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis.
-However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume.
-Moreover, the Boltzmann's $H$-theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds.
-We can include the fact that we proved the Boltzmann's $H$-theorem without using the Boltzmann's transport equation.
-It was only necessary to specify the behavior of the deviation $g$ and its derivative in time.
-Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$-functional through the variational method.
-Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$-functional is not zero when the system is out of equilibrium.
-This is a consequence of the $H$-theorem.\\
+Here, we considered inhomogeneities in the system through the division of the volume $V$ in cells 
+and the non-homogeneous distribution function assumption in the total volume. 
+We have shown that the Boltzmann's $H$-theorem still holds when the system includes 
+inhomogeneities, of course assuming that the local equilibrium hypothesis holds. 
+We demostrated the classical Boltzmann's $H$-theorem without using explicity the Boltzmann's transport equation. 
+Instead of that, it needed to specify the behavior of the deviation $g$ and its time derivative. 
 \\
-In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption.
-With the proof of the quantum version of the $H$-theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds.
-In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
-On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation.
-However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term.
-This problem may be future work.\\
+In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the 
+behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the 
+non-homogeneous distribution function assumption. 
+We have corroborated the correctness of the $f$ expression in equilibrium recovering 
+the MB, FD and BE distributions and in recovering the expression of the entropy of an ideal gas. 
+We demostrated as well that the functional $H$ leads to the proper fulfilling of the quantum classical correspondence.
 \\
-We also find that variational entropy is proportional to the Boltzmann's $H$-functional in the limit of high energy levels.
-This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}).
-Moreover, if we see the expression for the Boltzmann's $H$-functional in equilibrium
+\\
+We also have shown that variational entropy is proportional to the Boltzmann's $H$-functional in the non-degenerated limit. 
+This is a straightfowars conclusion. One may infer from the expression for the Boltzmann's $H$-functional in equilibrium that
 \begin{equation}
@@
 \end{equation}
-Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
-It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$.
-This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case.
-Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$.
-Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\
-According to the classical analysis of the $H$-theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
-In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time.
-Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.
+It is important to remark that we find that $\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$ 
+setting the expecting behavior of the deviation $g$. 
+This correct behavior results from the molecular chaos hypothesis in the classical case and 
+the random a priori phases hypothesis in the quantum case. 
+Specifically those hypothesis are included in the assumed behavior of $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$. 
+Those time derivatives correspond to the expected behavior in a meantime, that 
+is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\ 
+According to the classical $H$-theorem, if the molecular chaos hypothesis holds, 
+although within small times of the order of the \textit{relaxation time}, as a consequence of the 
+stochastic nature of the collisions, $H$ function evolves through a sequential series of 
+molecular chaos and non-molecular chaos, but a larger times $\frac{dH}{dt}\leq 0$.\\
+In analogy with this, we see that in our procedure this behavior is introduced by 
+$\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ to obtain the expected behavior in a relaxation time. 
+So, molecular chaos hypothesis and a random a priori phases hypothesis are implicit
+ in the expected behavior of the deviation from the equilibrium of the $H$ function such that 
+$\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$.
+\\
+\\
+In summaty, we propose a variational procedure to demostrate the classical and quantum $H$ theorem, that
+make possible to describe at a mesoscopic local view (cell-scale), the time evolution 
+of an out-of-equilibrium, spatially non-homogeneous, system to the equilibrium condition.
+In principle, this approach would make possible investigate the transport phenomena inherent to the
+equilibrium process, starting from an arbitrarily far-from-equilibrium initial condition.
 

### 2021-01-22 tex/qccorrespondence.tex
se agregaron los ultimos cambios, se unió la sección de discusión con conclusiones. La sección de relajación no se ha modificado hasta haber discutido lo que propone el Dr. Carrillo.

@@
 
-We start by arguing that, in the classic limit, $f_{nM}\approx0$ for the following reasons.
+We start by arguing that, in equilibrium, it is a trivial task to proof that expression (\ref{eq:qfneqtwo}) collapses
+into expression (\ref{eq:fbardef}) by simply letting that the 
+degeneration parameter $\xi=e^{-\frac{1}{K_B T}(\epsilon-\mu)} \gg 1$.
+A more general approach to see the quantum classical correspondence would be if the quantum functional 
+$H$ collapsed  into the classical one in the appropiated limit.
+This limit in our case will be $f_{nM} \approx 0$ because the following reasons.
 Systems at very low temperatures, where the quantum effects cannot be ignored, are obviously excluded
@@
 
-Because we have shown that both classical and quantum $H$-functionals, Eqs.~(\ref{eq:cHdef}) and
-(\ref{eq:qHdef}) recover the Maxwell-Boltzmann, Bose-Einstein, or Fermi-Dirac distribution
-functions in equilibrium, we will omit the quantum classical correspondence of the distribution functions,
-as it has been thoroughly discussed in textbooks.
-
 %--------------------------------------

### 2021-01-22 tex/disscussion.tex
Partial commit: rewording remarksanddiscussion

@@
 
-\section{Remarks and conclusions}\label{sec:disscussion}
+\color{black}
+\section{Discussions and remarks}\label{sec:disscussion}
 
-In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. 
-Here, we considered inhomogeneities in the system through the division of the volume $V$ in cells 
-and the non-homogeneous distribution function assumption in the total volume. 
-We have shown that the Boltzmann's $H$-theorem still holds when the system includes 
-inhomogeneities, of course assuming that the local equilibrium hypothesis holds. 
-We demostrated the classical Boltzmann's $H$-theorem without using explicity the Boltzmann's transport equation. 
-Instead of that, it needed to specify the behavior of the deviation $g$ and its time derivative. 
-\\
-In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the 
-behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the 
-non-homogeneous distribution function assumption. 
-We have corroborated the correctness of the $f$ expression in equilibrium recovering 
-the MB, FD and BE distributions and in recovering the expression of the entropy of an ideal gas. 
-We demostrated as well that the functional $H$ leads to the proper fulfilling of the quantum classical correspondence.
-\\
-\\
-We also have shown that variational entropy is proportional to the Boltzmann's $H$-functional in the non-degenerated limit. 
-This is a straightfowars conclusion. One may infer from the expression for the Boltzmann's $H$-functional in equilibrium that
-\begin{equation}
-    H_{B}=-\frac{S}{Vk},
-\end{equation}
-where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
-\begin{equation}
-    \Ss\propto S.
-\end{equation}
-It is important to remark that we find that $\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$ 
-setting the expecting behavior of the deviation $g$. 
-This correct behavior results from the molecular chaos hypothesis in the classical case and 
-the random a priori phases hypothesis in the quantum case. 
-Specifically those hypothesis are included in the assumed behavior of $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$. 
-Those time derivatives correspond to the expected behavior in a meantime, that 
-is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\ 
-According to the classical $H$-theorem, if the molecular chaos hypothesis holds, 
-although within small times of the order of the \textit{relaxation time}, as a consequence of the 
-stochastic nature of the collisions, $H$ function evolves through a sequential series of 
-molecular chaos and non-molecular chaos, but a larger times $\frac{dH}{dt}\leq 0$.\\
-In analogy with this, we see that in our procedure this behavior is introduced by 
-$\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ to obtain the expected behavior in a relaxation time. 
-So, molecular chaos hypothesis and a random a priori phases hypothesis are implicit
- in the expected behavior of the deviation from the equilibrium of the $H$ function such that 
-$\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$.
-\\
-\\
-In summaty, we propose a variational procedure to demostrate the classical and quantum $H$ theorem, that
-make possible to describe at a mesoscopic local view (cell-scale), the time evolution 
-of an out-of-equilibrium, spatially non-homogeneous, system to the equilibrium condition.
-In principle, this approach would make possible investigate the transport phenomena inherent to the
-equilibrium process, starting from an arbitrarily far-from-equilibrium initial condition.
+Traditionally, the proof of the classical $H$-theorem involves assuming that the
+distribution function is spatially homogeneous, and the proof starts with
+considering particular systems that are out of equilibrium, but in such a manner
+that their distribution functions are still homogeneous. This out-of-equilibrium systems
+are conceived considering the micro-states of the system, and the $H$-theorem determines
+the microscopic evolution of the system towards a state with the most-probable distribution function,
+given a thermodynamic state. 
+This approach does not cover systems with spatial inhomogeneities.
+To address this issue, we
+proposed a framework that may be useful to describe the time-evolution of systems
+initially non-homogeneous. For this, we divided the
+system into small cells, so as to conceive a system wherein we accept the local equilibrium
+hypothesis to be valid for each cell, but in such a manner that the global system is not homogeneous. Systems
+that satisfy the previous conditions will evolve towards global equilibrium, and the evolution is
+determined by the relations $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef}),
+and $d\mathcal{H}/dt\leq0$, Eq.~(\ref{eq:qHdef}), for classical and quantum gases, respectively.
+In some aspects, this can be considered as an extension or as a potential equivalent of the $H$ theorem
+for systems out-of-equilibrium in the spatial sense, \textit{i.e.} systems whose thermodynamic
+variables are not spatially homogeneous.
+
+The classical and quantum $H$-functionals, $\mathcal{H}'$ and $\mathcal{H}$, respectively,
+correctly recover the most-probable distribution
+functions both in out-of-equilibrium states (locally) and when the system attains the global
+equilibrium state. This enables the use of all the machinery of thermodynamics, including the
+proper definition of state variables as functions of time for each cell. In addition, our approach
+does not require to know the specific mechanisms by which energy or particles are transferred from one
+cell to its neighbors. Instead, the time-evolution from the out-of-equilibrium to equilibrium states are
+governed by monotonic functions that account for deviations from the global equilibrium.
+
+The specific forms of the functions $g'_M$ and $g_{nM}$ for classical and quantum systems will be the
+subject of our future work. However, some of their properties can be foreseen, \textit{e.g.} they must
+be consistent with the system relaxation times, they must be consistent with the mechanisms of 
+energy transfer between open systems, there must be related to the cell thermodynamic variables, etc.
+
+An important aspect of the framework proposed in this work is related to the entropy of systems
+out-of-equlibrium. Because the functionals $\mathcal{H}'$ and $\mathcal{H}$ can be related to
+the entropy of diluted gases, either classical or quantum, the fact that these functionals are defined
+over a system divided into cells enables their use for defining the entropy
+of other out-of-equilibrium systems. In particular, and 
+derived from our previous work (\textit{e.g.} \cite{bib:nicolas2020,bib:nicolas2016}), the
+$\mathcal H'$ and $\mathcal H$ functionals may serve to describe the entropy,
+as well as the entropy generation, occurring during the
+growth of complex physical systems, such as fractals. Possibly, studying these systems might also
+shed light on the functional form of $g'_M$ and $g_{nM}$.
+
+In conclusion, we have proposed a framework for studying classical and quantum out-of-equilibrium
+diluted gases, whose initial
+state is inhomogeneous in nature. This framework consists, briefly, of dividing the system into
+a set of small cells, and assuming that each cell is in equilibrium. Subsequently,
+$H$-functionals, $\mathcal H$, can be defined
+such that the time evolution towards equilibrium resembles the Boltzmann $H$-theorem: $d\mathcal H/dt\leq0$,
+and the equality condition is satisfied when the system attains the global equlibrium.
+The distribution functions,
+stemming from the $\mathcal H$ functionals, at any time, $t$, are the respective equilibrium distribution functions,
+which can be Maxwell-Boltzmann, Fermi-Dirac or Bose-Einstein, depending on the classical or quantum nature
+of the gas under consideration. Furthermore,
+the distribution function of each cell evolves so as to match the global distribution function in equilibrium.
+A proof of the relation $d\mathcal H/dt\leq0$ is performed through the variational method,
+the functional form of the distribution
+functions, in terms of a monotonic function $g$, are defined, and some of the properties of $g$ are discussed.
+The functionals $\mathcal H$ can be associated with the entropy of the system under consideration, hence we
+refer to this entropy as the variational entropy of the system. Finally,
+a potential application of this framework is outlined briefly.
 

### 2021-01-22 tex/disscussion.tex

@@
 
-\color{black}
+\color{red}
 \section{Discussions and remarks}\label{sec:disscussion}
@@
 
+\section{Remarks and conclusions}\label{sec:disscussion}
+
+In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. 
+Here, we considered inhomogeneities in the system through the division of the volume $V$ in cells 
+and the non-homogeneous distribution function assumption in the total volume. 
+We have shown that the Boltzmann's $H$-theorem still holds when the system includes 
+inhomogeneities, of course assuming that the local equilibrium hypothesis holds. 
+We demostrated the classical Boltzmann's $H$-theorem without using explicity the Boltzmann's transport equation. 
+Instead of that, it needed to specify the behavior of the deviation $g$ and its time derivative. 
+\\
+In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the 
+behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the 
+non-homogeneous distribution function assumption. 
+We have corroborated the correctness of the $f$ expression in equilibrium recovering 
+the MB, FD and BE distributions and in recovering the expression of the entropy of an ideal gas. 
+We demostrated as well that the functional $H$ leads to the proper fulfilling of the quantum classical correspondence.
+\\
+\\
+We also have shown that variational entropy is proportional to the Boltzmann's $H$-functional in the non-degenerated limit. 
+This is a straightfowars conclusion. One may infer from the expression for the Boltzmann's $H$-functional in equilibrium that
+\begin{equation}
+    H_{B}=-\frac{S}{Vk},
+\end{equation}
+where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
+\begin{equation}
+    \Ss\propto S.
+\end{equation}
+It is important to remark that we find that $\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$ 
+setting the expecting behavior of the deviation $g$. 
+This correct behavior results from the molecular chaos hypothesis in the classical case and 
+the random a priori phases hypothesis in the quantum case. 
+Specifically those hypothesis are included in the assumed behavior of $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$. 
+Those time derivatives correspond to the expected behavior in a meantime, that 
+is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\ 
+According to the classical $H$-theorem, if the molecular chaos hypothesis holds, 
+although within small times of the order of the \textit{relaxation time}, as a consequence of the 
+stochastic nature of the collisions, $H$ function evolves through a sequential series of 
+molecular chaos and non-molecular chaos, but a larger times $\frac{dH}{dt}\leq 0$.\\
+In analogy with this, we see that in our procedure this behavior is introduced by 
+$\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ to obtain the expected behavior in a relaxation time. 
+So, molecular chaos hypothesis and a random a priori phases hypothesis are implicit
+ in the expected behavior of the deviation from the equilibrium of the $H$ function such that 
+$\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$.
+\\
+\\
+In summaty, we propose a variational procedure to demostrate the classical and quantum $H$ theorem, that
+make possible to describe at a mesoscopic local view (cell-scale), the time evolution 
+of an out-of-equilibrium, spatially non-homogeneous, system to the equilibrium condition.
+In principle, this approach would make possible investigate the transport phenomena inherent to the
+equilibrium process, starting from an arbitrarily far-from-equilibrium initial condition.
+
 %-----------------------------------------------

### 2021-01-22 tex/disscussion.tex
Rewording: sec. 6 (disc. and concl).

--- 2021-01-22 tex/disscussion.tex
@@
 
-\color{black}
-\section{Discussions and remarks}\label{sec:disscussion}
+\section{Discussion and conclusion}\label{sec:disscussion}
 
@@
 
-The specific forms of the functions $g'_M$ and $g_{nM}$ for classical and quantum systems will be the
+The specific forms of the monotonic functions $g'_M$ and $g_{nM}$ for classical and quantum systems will be the
 subject of our future work. However, some of their properties can be foreseen, \textit{e.g.} they must
 be consistent with the system relaxation times, they must be consistent with the mechanisms of 
-energy transfer between open systems, there must be related to the cell thermodynamic variables, etc.
+energy transfer between open systems, they must be related to the cell thermodynamic variables, etc.
 
@@
 the functional form of the distribution
-functions, in terms of a monotonic function $g$, are defined, and some of the properties of $g$ are discussed.
+functions, in terms of a monotonic functions $g$, are defined, and some of the properties of $g$ are discussed.
 The functionals $\mathcal H$ can be associated with the entropy of the system under consideration, hence we

### 2021-01-22 tex/qccorrespondence.tex
Rewording: sec. 6 (disc. and concl).

--- 2021-01-17 tex/qccorrespondence.tex
@@
 
-We start by arguing that, in the classic limit, $f_{nM}\approx0$ for the following reasons.
+We start by arguing that, in equilibrium, it is straightforward to proof that
+Eq.~(\ref{eq:qfneqtwo}) collapses
+into Eq.~(\ref{eq:fbardef}) by taking the limit wherein the 
+degeneration parameter $\xi\equiv\exp\big(-(\epsilon-\mu)/(k_BT)\big) \gg 1$.
+Alternatively, a more general approach to show the quantum classical correspondence
+consists of analyzing the collapse from the quantum to the classical $H$-functionals,
+within the appropiated limit.
+For the case treated here, this limit is $f_{nM} \approx 0$ because of the following reasons.
 Systems at very low temperatures, where the quantum effects cannot be ignored, are obviously excluded

### 2021-01-22 tex/quantum.tex
se agregaron los ultimos cambios, se unió la sección de discusión con conclusiones. La sección de relajación no se ha modificado hasta haber discutido lo que propone el Dr. Carrillo.
Partial commit: rewording remarksanddiscussion
Rewording: sec. 6 (disc. and concl).

@@
 %
-\begin{equation}
+\begin{equation}\label{eq:qfneqtwo}
     \bar{f}_{n}=\frac{1}{\exp\big(\frac{{\epsilon_n}-\mu}{kT}\big)\pm 1},
@@
 and this name reflects the fact that it was obtained \textit{via}
-the variational method. We will use it at the end of this article to analyze relaxation processes.
+the variational method. {\color{blue} We will use it at the end of this article to analyze relaxation processes.}
 
@@
 %\begin{proof}
-To proof the quantum $H$-theorem, we start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
+To proof the quantum $H$-theorem, start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
 %
@@
 %
-By representing the logarithmic terms, corresponding to
-Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$, and
-substituting this into Eq.~(\ref{eq:cambioH1}), one obtains:
+Approximate the logarithmic terms, corresponding to
+Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$:
 %
@@
 
-Combining Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}) yields:
+Combine Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}):
 %
@@
 %
-and substituting Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
+and substitute Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
 %
@@
    \frac{d\mathcal{H}}{dt}=\sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}
-   \sum_M  g_{nM}\dot{g}_{nM} \delta V_M \leq 0.
+   \sum_M  g_{nM}\dot{g}_{nM} \delta V_M \leq 0. \qquad\qquad\textrm{QED.}
 \end{equation}
 %\end{proof}
-Thus, the time evolution of the system occurs in such a way that 
-$\frac{d\mathcal{H}}{dt}$ is a decreasing function toward zero.
 %
@@
 $K$ small cells), with total energy $E$ and $N$ quantum free particles, which initially
-is out of equilibrium ---but still the local equilibrium hypothesis is valid---,
+is out of equilibrium ---but in such a manner that the local equilibrium hypothesis is valid---,
 the functional
@@
 the equilibrium state. In Eq.~(\ref{eq:qHfunctTheo}), and for a Fermi (Bose)
-gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distrubution function for each
-cell and locally each cell is in equilibrium, although the complete system may be non-homogeneous,
+gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distrubution function for each cell.
+Locally, each cell is in equilibrium, although the complete system may be non-homogeneous,
 and is caracterized by the respective $f_{Mn}$, number of particles $\mathcal{N}_M$, energy $\mathcal{E}_M$,

### 2021-01-23 tex/relaxproc.tex
Se agregó un breve desarrolo para la sección de relajación en términos de la funcional H evaluada en diferentes puntos. Corregí un texto en rojo en la sección quantum donde estaban mal identificados los signos para bosones y fermiones. También agregué en rojo la subsección OUT OF EQUILIBRIUM, NON-HOMOGENEOUS SYSTEMS debido a que ahí no se hace mención sobre sistemas fuera de equilibrio sino hasta que demostramos el teorema H.

@@
   \section{Relaxation processes in degenerated quantum gases}\label{sec:relaxproc}
-Consider that the distribution function $f_{nM}$ is a function of the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$ as follow
-\begin{equation}
-    f=f(\mathcal{N}_M^{n},t),
+In order to obtain an equation of motion for the quantum gas, we evaluate 
+$\Delta H=H(t_2)-H(t_1)$ where the $H$ functional is evaluated in diferent times such that $t_2 > t_1$.
+We begin using the following notation $f_{nM}(t_2)\equiv f'_{nM}$ and 
+$f_{nM}(t_1) \equiv f''_{nM}$. 
+Now, using (\ref{eq:qHdef}) to calculate $\Delta H$ with the previous notation, we obtain
+\begin{eqnarray} \label{eq:deltaHdef}
+	\Delta H &=& \sum_{M}^{K}\sum_n [f'_{nM} \ln f'_{nM} - f''_{nM}	\ln f''_{nM}\nonumber \\ 
+			&&\pm (1\mp f'_{nM}) \ln(1\mp f'_{nM}) \mp (1\mp f''_{nM}) \ln (1\mp f''_{nM})]\delta V.  
+\end{eqnarray}
+In order to obtain a general equation of motion for the distribution function 
+$f_{nM}$, we only take the limit when $(t_2-t_1) \rightarrow 0 $. 
+However, our current work is focused on the deviation $g_{nM}$. So, keeping the same workline
+we apply the first order approximation (\ref{eq:qFirstOrd}) to (\ref{eq:deltaHdef}) to obtain the following expression
+\begin{eqnarray}\label{eq:deltaH1}
+	\Delta H &=& \sum_{M}^{K} \sum_n [\bar{f}_n(1+g'_{nM}) \ln \bar{f}_{n}(1+g'_{nM})  -  
+			\bar{f}_n(1+g''_{nM}) \ln \bar{f}_n(1+g''_{nM}) \nonumber \\
+		    &&  \pm (1 \mp \bar{f}_n \{1+g'_{nM}\}) \ln (1\mp \bar{f}_n\{1+g'_{nM} \})\nonumber \\
+		    &&  \mp (1\mp \bar{f}_n \{ 1+g''_{nM} \}) \ln (1\mp \bar{f}_n \{1+g''_{nM} \})]\delta V.\nonumber \\
+\end{eqnarray} 
+The equation (\ref{eq:deltaH1}) can be rewritten, after some approximations on the logarithm functions, as
+\begin{equation}\label{eq:deltaH2}
+\Delta H = \sum_{M}^{K}\sum_n [\bar{f}_n(1+\ln \bar{f}_n)-\bar{f}_n \{\ln(1\mp \bar{f}_n)\mp 1 \}](g'_{nM}-g''_{nM})\delta V.
 \end{equation}
-where $\mathcal{N}_{M}^{n}$ is defined as
-\begin{equation}
-    \mathcal{N}_{M}^{n}= f_{nM} \delta \epsilon_n.
-\end{equation}
-Besides the local number of particles in a quantum energy state 
-is a function of time $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(t)$.\\
-In order to obtain the analogous transport equation to this system, we apply the variation method to the functional $\Ss$.
-Due to $\Ss$ in equilibrium is proportional to entropy, and therefore a maximum, the variation must be zero.
-Then the equation of motion is
-\begin{equation}
-    0=\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}++\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}} \frac{\partial \mathcal{N}_J^{j}}{\partial t} \right].
-\label{transportequation1}
-\end{equation}
-This equation is valid in equilibrium.
-When the system is out of equilibrium, the variation is different from zero.
-Due to this, an expression has to be included in (\ref{transportequation1}).
-This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)_{coll}$.
-Moreover, we propose that this term contributes to the equation of motion per each cell.\footnote{The collision term is only in a defined cell and local group of energy.
-If we are describing the total system, we need to include all contributions (each cell and energy group).} Therefore, (\ref{transportequation1}) casts into
-\begin{eqnarray}
-    \sum_{M=1}^{K}\sum_n \left( \frac{df_{nM}}{dt} \right)_{coll}&=&\sum_{M=1}^{K} \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}} \frac{\partial \mathcal{N}_J^{j}}{\partial t} \right].
-\nonumber \\ \label{transportequation2}
-\end{eqnarray}
-The equation (\ref{transportequation2}) is the time evolution equation to a quantum ideal gas. \\
+The equation (\ref{eq:deltaH2}) must be analyzed taking the limit $(t_2-t_1) \rightarrow 0$ 
+to obtain the equation of motion for the deviation $g_{nM}$.
+This analysis will not be treated in this article.
 

### 2021-01-25 tex/classical.tex
Rewording: sec. 6 (disc. and concl).
Rewording: All sections finished but 5.

@@
 
-Boltzmann Kinetic theory of gases is one of the theoretical basese of our present 
-knowledge of out equilibrium systems. This represent a fundamental connection
+Boltzmann Kinetic theory of gases represents a fundamental connection
 between the microscopic nature of matter and the phenomenological macroscopic  laws of classical thermodynamics.
-The stochasticity introduced in the, in other way deterministic, kinetics of the
-particles of the gas, via the molecular chaos hypothesis, allows the demostration 
-of the celebreted Boltzmann $H$ theorem. In this article we propose an alternative approach, which is 
-developed by means of a variational method, applied to a functional.
-Let us start this section by briefly remaind of the important elements of the 
-standard derivation and demostration of the Boltzmann transport equation and the $H$ theorem as they are presented in classical textbooks \cite{bib:huang}.
+The stochasticity introduced by the molecular chaos hypothesis in the, otherwise deterministic, kinetics of the
+particles allows the demostration 
+of the celebrated Boltzmann $H$-theorem. In contrast, in this article we propose an alternative approach, which is 
+developed by means of a variational procedure applied to an $H$-functional.
+Let us start this section by briefly accounting for the important elements of the 
+standard derivation of the Boltzmann transport equation and demostration the $H$-theorem,
+such as they are presented in classical textbooks \cite{bib:huang}.
 
@@
 
-Boltzmann kinetic theory of gases starts by defining the distribution function 
-$f(\vec{r},\vec{v},t)$ as the average number of molecules that at time $t$, has the 
-values of position and velocity in the element of volumen in the $\mu$ space.
-Assuming a deterministic Newtonian description of the movement of the molecules 
-and the invariance of the volume measure of the $\mu$ space, one arrives to the Boltzmann rate equation.
+In Boltzmann kinetic theory of gases, one starts by defining the distribution function, 
+$f(\vec{r},\vec{v},t)$, as the average number of molecules that, at time $t$, have 
+position $\vec r$ and velocity $\vec v$, and are contained in a $\mu$-space volume element $d^3rd^3v$.
+Assuming a deterministic Newtonian description of the molecules motion, 
+as well as the invariance of the $\mu$-space volume measure, one arrives to the Boltzmann rate equation:
 %
 \begin{equation}
-	f(\vec{r}+\vec{v}\delta t, \vec{v}+\vec{F} \delta t, t+\delta t)=f(\vec{r},\vec{v},t)+\left( \frac{\partial f}{\partial t} \right)_{coll} \delta t.
+	f(\vec{r}+\vec{v}\delta t, \vec{v}+\vec{F} \delta t, t+\delta t)=f(\vec{r},\vec{v},t)+\left( \frac{\partial f}{\partial t} \right)_{\textrm{coll}} \delta t.
 \end{equation}
 %
-The collision term $\left( \frac{\partial f}{\partial t}\right)$ of this equation 
-describes the fuxes in and out from the volume element due to the collisions.
-The integro-differential Boltzmann transpor equation can be cast into.
+Here, the term $\left(\partial f/\partial t\right)_{\textrm{coll}}$
+describes the in and out fluxes from and towards the volume element, due to the collisions.
+Subsequently, from the previous equation, the integro-differential Boltzmann transport equation is obtained:
 %
@@
 The evolution of a dilute gas towards thermodynamic equilibrium is frequently
-addressed by first defining the functional \cite{bib:tolman,bib:huang}
+addressed by first defining the $H$-functional \cite{bib:tolman,bib:huang}:
 %
@@
 %
-Notice that $f(\vec{v},t)$ is a spacially homogeneous distribution function.
+Notice that $f_B(\vec{v},t)$ is a spatially homogeneous distribution function.
 The functional $H_B$,
 originally introduced by Boltzmann in 1872,
-describes a diluted gas, at temperature $T$, with a total energy $E$, a total
-number of free classical particles $N$, and occupying a volume $V$.
-In order to clearly distinguish the Boltzmann functional $H_B$, we denote here
+describes a diluted gas occupying a volume $V$, at temperature $T$, with total energy $E$, and total
+number of free classical particles $N$.
+In order to clearly distinguish the Boltzmann functional $H_B$, we denote hereafter
 the Maxwell-Boltzmann distribution function as $f_B$.
 
-The physically correct espontaneous time evolution of an out of equilibrium dilute gas 
-is corroborated by the $H$ theorem.
-This theorem establishes that if the homogeneous function $f(\vec{v},t)$ satisfies the Boltzmann
-transport equation and consequently, if the molecular chaos hypothesis is valid, then
-the system evolves in such a way that $dH_B/dt\leq0$, and if when $dH_B/dt=0$, 
-the system is in the equilibrium state. $H$-theorem is easily proven by 
-using the Boltzmann transport equation (\ref{eq:transport}) \cite{bib:huang}. 
-$H$-theorem assures that our microscopic approach to describe the spontaneous 
-time evolution of the system will be consistent with the phenomenological observations
-established in the 2nd-law of Classical; in fact,
+The physically correct espontaneous time-evolution of an out of equilibrium diluted gas 
+is corroborated by the $H$-theorem.
+This theorem establishes that if the two conditions, a) the homogeneous
+function $f(\vec{v},t)$ satisfies the Boltzmann
+transport equation and b) the molecular chaos hypothesis is valid, then
+the system evolves in such a manner that $dH_B/dt\leq0$, and if $dH_B/dt=0$, 
+then the system is in the equilibrium state. 
+The $H$-theorem is straightforward to prove through Eq.~(\ref{eq:transport}) \cite{bib:huang}, 
+and it assures the consistency between our microscopic approach to describe the system's spontaneous 
+time-evolution and the phenomenological observations
+established by the $2^{\textrm{nd}}$ law of classical thermodynamics; in fact,
 $H_B$ can be associated with an entropy density.
 
-On the other hand, considering a diluted gas in equilibrium, \textit{i.e.}
-$(\partial f/\partial t)=0$,
-with no applied external forces $f$ is independent of $\vec r$, 
+On the other hand, considering a diluted gas in equilibrium
+with no applied external forces, \textit{i.e.}
+$(\partial f/\partial t)=0$ and $f$ is independent of $\vec r$), 
 it can be directly proven that the equilibrium distribution function obtained from
@@
 Subsequently, the Maxwell-Boltzmann distrubution function can be obtained by
-taking the logarithm to Eq.~(\ref{eq:neccondf0}), and exploiting mechanical
+taking the logarithm of Eq.~(\ref{eq:neccondf0}), and exploiting mechanical
 conserved quantities (see \cite[ch. 4.2]{bib:huang}).
 
-Before introducing our proposal, we call the reader's attention to the following. In defining
-$H_B$ Eq.~(\ref{eq:hbfunctional}), it is assumed that the distribution function 
-$f_B$ is spattially homogeneous. Clearly this assumption simplifies the demostration 
-of the $H$ theorem, nevertheless, it introduces as well a limitate-conception
-of an out of equilibrium condition in the gas. Given the relatively simple nature of a dilute 
-gas, one of the main characteristics of an out of equilibrium condition would be the existence of
+Before introducing our proposed $H$-functional, we remark the following. In defining
+$H_B$, Eq.~(\ref{eq:hbfunctional}), it is assumed that the distribution function 
+$f$ is spatially homogeneous. This assumption simplifies the demostration 
+of the $H$ theorem, however, it introduces as well a limited conception
+of the out-of-equilibrium condition of the gas. Given the relatively simple nature of a dilute 
+gas, one of the salient features of an out-of-equilibrium condition is the existence of
 inhomogeneities in the system.
@@
 that the distribution function is homogeneous and that the hypothesis of
-molecular chaos is fullfiled. In order to extend the procedure for systems whose
-distribution functions are not necessarily homogeneous, would possibilities allows 
-for studying systems been in a more general out of equilibrium condition, we introduce a modified $H$-functional 
-as follows. For the sake of clearity and simplicity we will use primed functions
-and quantities to denotate the classical case and differentiate it from the quantum
-case.
+molecular chaos is fullfiled. In order to extend the previous procedure to systems whose
+distribution functions are non-homogeneous, which might allow for studying systems
+being in a more general out-of-equilibrium condition, we introduce a modified $H$-functional 
+in the following. For the sake of clearity and simplicity, we will use primed functions
+and quantities to denote the classical case, so as to differentiate them from the quantum
+analogues.
 
-Our proposed functional, denoted as $\mathcal{H}'$, describes a dilute classical
-gas occupying a volume $V$. For our theoretical treatment we will consider this volume
-divided into $K$ cells, which, without loss of generality,
-have identical volumes $\delta V_M = V/K,\ M=1,\,\dots\,,K$.
-Locally the $\mathcal{H}'$-functional is denoted by $\mathcal{H}'_M$, and 
-the local homogeneous distribution function, $f'_{M}(\vec{v},t)$; local 
-number of particles, $\mathcal{N}'_M$; local temperature, $T'_M$; and a local 
-energy, $\mathcal{E}'_M$. As a whole, the system has a
-energy $E$, and a global number of free classical particles $N$. Let us assume that the system is 
-to be isolated. We also assume that the number of particles
+Our proposed classical $H$-functional, denoted as $\mathcal{H}'$, describes a dilute classical
+gas occupying a volume $V$. In our theoretical treatment, we divide this volume
+into $K$ cells, which, without loss of generality,
+have identical volumes, $\delta V_M = V/K,\ M=1,\,\dots\,,K$.
+Each cell of index $M$ has the following local functions, properties, and variables:
+an $H$-functional, $\mathcal{H}'_M$, a homogeneous distribution function, $f'_{M}(\vec{v},t)$,
+a number of particles, $\mathcal{N}'_M$, a temperature, $T'_M$, and an
+energy, $\mathcal{E}'_M$. Taken as a whole, the system has a
+energy $E$, and a global number of free classical particles $N$. We also assume that the system is 
+perfectly isolated, and that the number of particles
 in each cell is sufficiently large, so as to obtain accurate averages.
-We start our analysis by proposing the following inhomogeneous functional
+We start our analysis by proposing the following inhomogeneous $H$-functional:
 %
@@
 %
-Notice that assuming homogeneous distribution functions per cell means that we are accepting valid the 
+Notice that assuming that every $f'_M$ is homogeneous implies that we are accepting the validity of the
 local equilibrium hypothesis. In addition, the set $\{f'_{M}(\vec{v},t)\}$ must satisfy the following restrictions:
@@
 
-We now use the variational method in order to find the extremal of $\mathcal{H}'$, consistent with the restrictions
-(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) by using the corresponding Lagrange
-multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields
+We now use the variational method to find the extremal of $\mathcal{H}'$,
+consistent with the restrictions
+(\ref{eq:cellrestrictions}) and (\ref{eq:micro}), together with the corresponding Lagrange
+multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields:
 %
@@
 of particles and the local
-energy do not depend on the cell number, on a statistical sense, this is
+energy do not depend on the cell number, in a statistical sense, this is
 %
@@
 In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
-Moreover, the distribution function is homogeneous, hence $f'_M$
+Moreover, the global distribution function is homogeneous, hence $f'_M$
 does not depend on the cell number $M$ (\textit{i.e.} $f'_M(\vec v,t)=f'(\vec v,t)$,
@@
 from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$. Here we have used Eqs. (\ref{eq:Ecell}) and (\ref{eq:Ncell}).
-Eq.~(\ref{eq:cHdef}), in terms of Eq.~(\ref{eq:hbfunctional}), can be cast into
+Eq.~(\ref{eq:cHdef}), in terms of Eq.~(\ref{eq:hbfunctional}), can be rewritten as:
 %
@@
 %
-In order to identify $\mathcal{H}'$ with the entropy, still we need to show that $\mathcal{H}'(t)$ is extensive,
-with respect to $Kf'_(\vec v,t)$.
-This is shown by analyzing the following expression
+
+In order to identify $\mathcal{H}'$ with the entropy, we need to show that $\mathcal{H}'(t)$ is extensive,
+with respect to $Kf'(\vec v,t)$.
+This is shown by analyzing the following expression:
 %
@@
 %
-We observe that if the number of particles in the $\mu$-space, $f'(\vec{v},t)$, is much large than
-the number of cells $K$, then the first term of Eq.~(\ref{eq:sum-h}) is negligible, and consequently:
+We observe that if the number of particles in the $\mu$-space, $f'(\vec{v},t)$, is much larger than
+the number of cells, $K$, then the first term of Eq.~(\ref{eq:sum-h}) is negligible, and consequently:
 %
 \begin{equation}\label{aditive-property}
-    \int \big[Kf'(\vec{v},t)\big] \ln \big[Kf'(\vec{v},t)\big] \mathrm{d}^{3}v \simeq
+    \int \big[Kf'(\vec{v},t)\big] \ln \big[Kf'(\vec{v},t)\big] \mathrm{d}^{3}v \approx
     K\int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v,
@@
 system, which reduces to the Boltzmann $H$-functional.
-\commb{Hence, $\mathcal{H}'$ can also be identified with a density of entropy for the system.}
+Therefore, $\mathcal{H}'$ can be identified with the entropy density of the system.
 
-Furthermore, in equilibrium the Lagrange multipliers are position- and time-independent,
+Furthermore, in equilibrium, the Lagrange multipliers are position- and time-independent,
 thus $f'_M(\vec{v})$
@@
 %
-since $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$.
-$N$ is a constant and we are mainly intersted in the time-derivative of $\mathcal{H}''$, thus
-$C$ can be conveniently omitted. In other words, $\mathcal{H}'$ leads the Maxwell-Boltzmann
-distribution function for systems in equilibrium.
+Since $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$ (a constant)
+and because we are mainly interested in the time-derivative of $\mathcal{H}''$,
+$C$ can be conveniently omitted. In other words, $\mathcal{H}'$ leads to the Maxwell-Boltzmann
+distribution function of systems in equilibrium.
 
@@
 
-Along this section, we consider a classical gas
+Throughout this section, we will consider a classical gas
 whose initial condition is close to the equilibrium,
@@
 the time-evolution of the system.
-Also we will use the following definitions for the deviations of concentration 
-and energy from the equilibrium values:
+Also, we will use the following definitions for the deviations of concentration 
+and energy, relative to the equilibrium values:
 \begin{subequations}\label{eq:restrictionoutsideclassical}
@@
 Here $\bar {\mathcal{N}}' =N/K$ and $\bar{\mathcal{E}}'=E/K$ are
-the cell particle number and the cell energy in equilibrium, respectively, they are given by:
+the cell particle number and the cell energy in equilibrium, respectively, which are given by:
 %
@@
 are considered deviations relative to
-$\bar{\mathcal{N}}'$ and $\bar{\mathcal{E}}'$, respectively, and for systems that are
+$\bar{\mathcal{N}}'$ and $\bar{\mathcal{E}}'$, respectively. For systems that are
 sufficiently close to equilibrium, it is reasonable to expect on the one hand that
@@
 $\delta'_M(t)\ll\bar{\mathcal{E}}'$, and on the other hand that $\Delta'_M$ and
-$\delta'_M$ are sufficiently large compared with fluctuations of $\bar{\mathcal{N}}'$
+$\delta'_M$ are sufficiently large compared with the fluctuations of $\bar{\mathcal{N}}'$
 and $\bar{\mathcal{E}}'$. Under the same perspective, we can assume that every local
@@
 
-With the previous considerations, in the following we provide an alternative demonstration
-of the $H$-theorem, considering the $H$-functional, $\mathcal{H}'$, defined by Eq.~(\ref{eq:cHdef}).
+With the previous considerations, in the following we proof an alternative
+$H$-theorem, considering the $H$-functional, $\mathcal{H}'$, defined by Eq.~(\ref{eq:cHdef}).
 
@@
 \begin{equation}\label{eq:dHpdtleq0}
- \frac{d\mathcal{H}'}{dt}\leq0.
+ \frac{d\mathcal{H}'}{dt}\leq0.\qquad\qquad\textrm{QED.}
 \end{equation}
@@
 
-In summary, if a gas occupying a volume V (which is divided into $K$ small cells), with a
+In summary, considering a gas occupying a volume V (which is divided into $K$ small cells), with a
 total energy $E$ and $N$ classical free particles, whose initial state is not in the equilibrium,
@@
 In Eq.~(\ref{eq:CH3}), $f'_M(\vec{v},t)$ is the Maxwell-Boltzmann distribution
-function, which in general is different for different cells \textit{i.e.} the complete system
-can be non-homogeneous, and each $f'_M(\vec v,t)$ is compatible
+function, which in general is different for different cells ---\textit{i.e.} the complete system
+can be non-homogeneous---, and each $f'_M(\vec v,t)$ is compatible
 with the cell properties, such as number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,

### 2021-01-25 tex/disscussion.tex
Rewording: All sections finished but 5.

@@
 
-\section{Discussion and conclusion}\label{sec:disscussion}
+\section{Comments and remarks}\label{sec:disscussion}
 
-Traditionally, the proof of the classical $H$-theorem involves assuming that the
-distribution function is spatially homogeneous, and the proof starts with
-considering particular systems that are out of equilibrium, but in such a manner
-that their distribution functions are still homogeneous. This out-of-equilibrium systems
-are conceived considering the micro-states of the system, and the $H$-theorem determines
-the microscopic evolution of the system towards a state with the most-probable distribution function,
-given a thermodynamic state. 
-This approach does not cover systems with spatial inhomogeneities.
-To address this issue, we
+The demonstration of the classical $H$-theorem usually starts by assuming that the gas,
+despite being initially out of equilibrium, can be described by a
+spatially homogeneous distribution function. 
+Subsequently, the time-evolution of the system occurs in such a manner that $dH/dt \leq 0$. 
+Therefore, this approach does not cover the evolution to equilibrium of systems with spatial inhomogeneities.
+To address this issue, in this article we
 proposed a framework that may be useful to describe the time-evolution of systems
-initially non-homogeneous. For this, we divided the
-system into small cells, so as to conceive a system wherein we accept the local equilibrium
-hypothesis to be valid for each cell, but in such a manner that the global system is not homogeneous. Systems
-that satisfy the previous conditions will evolve towards global equilibrium, and the evolution is
-determined by the relations $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef}),
+initially non-homogeneous. To this end, we divided the
+system into small cells, so as to conceive a system wherein the local equilibrium
+hypothesis is valid in each cell, but in such a manner that the total system is not homogeneous. Systems
+that satisfy the previous conditions will evolve towards equilibrium, and the evolution
+occurs according to $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef}),
 and $d\mathcal{H}/dt\leq0$, Eq.~(\ref{eq:qHdef}), for classical and quantum gases, respectively.
-In some aspects, this can be considered as an extension or as a potential equivalent of the $H$ theorem
-for systems out-of-equilibrium in the spatial sense, \textit{i.e.} systems whose thermodynamic
-variables are not spatially homogeneous.
+Consequently, this approach can be considered as an extension of the $H$-theorem
+for more realistic out-of-equilibrium systems.
 
@@
 functions both in out-of-equilibrium states (locally) and when the system attains the global
-equilibrium state. This enables the use of all the machinery of thermodynamics, including the
-proper definition of state variables as functions of time for each cell. In addition, our approach
-does not require to know the specific mechanisms by which energy or particles are transferred from one
-cell to its neighbors. Instead, the time-evolution from the out-of-equilibrium to equilibrium states are
-governed by monotonic functions that account for deviations from the global equilibrium.
+equilibrium state. \commb{The relaxation process of the system is described by
+monotonic functions that account for deviations from the global equilibrium.}
 
-The specific forms of the monotonic functions $g'_M$ and $g_{nM}$ for classical and quantum systems will be the
-subject of our future work. However, some of their properties can be foreseen, \textit{e.g.} they must
-be consistent with the system relaxation times, they must be consistent with the mechanisms of 
-energy transfer between open systems, they must be related to the cell thermodynamic variables, etc.
+It is clear that for describing the relaxation process of a concrete system,
+it is necessary to know, at least with some approximation level, the specific forms of the
+monotonic functions $g'_M$ and $g_{nM}$, for classical and quantum systems.
+Whereas the complete analysis of these functions is out of the scope of the present work,
+some of their properties can be foreseen, \textit{e.g.} they must
+be consistent both with the system relaxation times and with the mechanisms of 
+energy transfer between cells.
 
@@
 over a system divided into cells enables their use for defining the entropy
-of other out-of-equilibrium systems. In particular, and 
+of out-of-equilibrium systems, other than diluted gases. In particular, and 
 derived from our previous work (\textit{e.g.} \cite{bib:nicolas2020,bib:nicolas2016}), the
@@
 
-In conclusion, we have proposed a framework for studying classical and quantum out-of-equilibrium
-diluted gases, whose initial
-state is inhomogeneous in nature. This framework consists, briefly, of dividing the system into
-a set of small cells, and assuming that each cell is in equilibrium. Subsequently,
-$H$-functionals, $\mathcal H$, can be defined
-such that the time evolution towards equilibrium resembles the Boltzmann $H$-theorem: $d\mathcal H/dt\leq0$,
-and the equality condition is satisfied when the system attains the global equlibrium.
-The distribution functions,
-stemming from the $\mathcal H$ functionals, at any time, $t$, are the respective equilibrium distribution functions,
-which can be Maxwell-Boltzmann, Fermi-Dirac or Bose-Einstein, depending on the classical or quantum nature
-of the gas under consideration. Furthermore,
-the distribution function of each cell evolves so as to match the global distribution function in equilibrium.
-A proof of the relation $d\mathcal H/dt\leq0$ is performed through the variational method,
-the functional form of the distribution
-functions, in terms of a monotonic functions $g$, are defined, and some of the properties of $g$ are discussed.
-The functionals $\mathcal H$ can be associated with the entropy of the system under consideration, hence we
-refer to this entropy as the variational entropy of the system. Finally,
-a potential application of this framework is outlined briefly.
+In summary, we proposed a variational procedure to demonstrate the classical and
+quantum $H$-theorems, which made possible to describe, at a mesoscopic local view
+(cell-scale), the time-evolution of an out-of-equilibrium and spatially
+non-homogeneous system, towards the equilibrium condition. In principle, this
+approach would render possible to investigate the transport phenomena, inherent to the
+equilibration process, occurring in a system that started from an spatially inhomogeneous
+out-of-equilibrium initial condition.
 

### 2021-01-25 tex/introduction.tex
Rewording: sec. 6 (disc. and concl).
Rewording: All sections finished but 5.

--- 2020-12-21 tex/introduction.tex
@@
 
-In the literature, the theoretical basis that allows us to describe equilibrium
-systems in the classical scheme is well-established.
-The procedures to describe
-equilibrium systems are well-known and work in almost every system in nature
-(Thermodynamics for a phenomenological description and statistical mechanics
-for a microscopic description). In contrast, a unified mathematical model to
-describe the evolution of any system to an equilibrium state has not developed.
-Nevertheless, developments can be found in the literature and describe the
-evolution to equilibrium to almost any system (the internal behavior of glasses
-\cite{bib:zanotto2018} and the entropy measurement
-\cite{bib:schmelzer2018,bib:nemilov2018} are some examples where the recent
-developments do not work.) Examples of those developments can be the Onsager
-formulation (called linear thermodynamics)
-\cite{bib:keizer1987,bib:onsager1931}, and the Kinetic Theory of dilute gases
-developed by Boltzmann.
+The theoretical basis and the procedures that allow to describe classical equilibrium systems
+are well-stablished. These procedures can be applied to treat a wide range of the systems
+present in nature, and include both the macroscopic phenomenological methods (thermodynamics)
+as well as the microscopic description (statistical mechanics). For instance,
+in the Kinetic Theory of gases, the behavior of a dilute classical gas
+is described through the Boltzmann transport equation \cite{bib:huang},
+and the time-evolution of a system towards the equilibrium is proved by
+means of the Boltzmann $H$-theorem, wherein the
+spatial-homogeneous distribution function hypothesis plays a crucial role.
 
-In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas
-was described by Boltzmann through the Boltzmann transport equation.
-Furthermore, the evolution of an equilibrium state for any system was proved by
-means the Boltzmann's $H$-theorem (or classical $H$-theorem) and the
-spatial-homogeneous distribution function hypothesis.
-However, there are some aspects in the classical $H$-theorem and the
-Boltzmann's $H$-functional that deserve some review to accomplish better
-consistency, for example, the modification of the $H$-theorem in stochastic
-trajectories, violations in the second law of thermodynamics, and the relation
-between Shannon's measure of information and the Boltzmann's entropy %[6-10].
-\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.
-
-To obtain the well-known correspondence principle between classical mechanics
-and quantum mechanics, a quantum version of the Boltzmann's $H$-theorem (or
-quantum $H$-theorem) and the quantum analogous to the Boltzmann transport
-equation is necessary such that in the appropriate limit, the  classical $H$
-theorem and the Boltzmann transport equation should be recovered.
-
-One of the first developments in this field was developed by Tolman
-\cite{bib:tolman}. He proposes one of the first quantum versions of the
-Boltzmann's $H$-theorem using the probability transition relation, the random
-phases hypothesis, and a homogeneous distribution function. The equation of
-motion for the occupation number (necessary to prove the quantum $H$-theorem in
-the same way that Boltzmann's procedure) was obtained by applying the time
-perturbation theory. However, to obtain the flux relations in analogy to the
-Boltzmann's method, a quantum version of the transport equation that includes a
-non-homogeneous distribution function is required.
-
-On the other hand, some works in the quantum operator formalism attempt to
-describe quantum transport phenomena using the Hamiltonian of the system and
-the master equation (analogous to the Boltzmann transport equation)
-\cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. %[12-15]
-Besides, some authors construct the $H$-functional and develop the proof of the
+In contrast, for quantum equilibrium systems, the construction of a framework,
+with the same level of
+success and universality as the classical version, is still an open problem. For instance,
+in order to obtain a complete correspondence principle, between classical mechanics
+and quantum mechanics, the quantum analogous of both the Boltzmann $H$-theorem
+and the Boltzmann transport equation have not attained a fully accepted form.
+In this context, Tolman was one of the earliest physicists to propose a quantum
+$H$-theorem \cite{bib:tolman}, using probability transition relations, the random
+phases hypothesis, and a spatially homogeneous distribution function.
+Also, Tolman proposed a potential quantum analogous of the transport equation,
+in terms of the occupation number,
+by applying time perturbation theory. Further attempts, under the quantum operator
+formalism, have addressed the description of quantum transport phenomena by means of the Hamiltonian
+of the system and the master equation (which here is the analogous
+of the Boltzmann transport equation)
+\cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. However, these approaches 
+are not consistent with the classical-quantum correspondence principle.
+In the same line of thought, some authors have proposed $H$-functionals
+and attempted to proof a
 quantum $H$-theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
-However, the homogeneous distribution function hypothesis and the
-correspondence principle are uncleared or not discussed. Moreover, the validity
-of the quantum $H$-theorem and the second law of thermodynamics, and the
+However, whether or not the homogeneous distribution function hypothesis is assumed, or
+if their framework obeys the correspondence principle is unclear or not discussed whatsoever.
+Furthermore, the general validity
+of the quantum $H$-theorem, the second law of thermodynamics, and the
 interpretation of the quantum entropy is still in discussion %[19-26].
@@
 
+On the other hand, the framework to describe spatially non-homogeneous systems is still
+under construction, although several approaches have been developed. For instance,
+the celebrated Onsager formulation (\textit{aka} linear thermodynamics)
+\cite{bib:keizer1987,bib:onsager1931} have been successful in describing irreversible chemical 
+and physical phenomena. Nonetheless, some aspects such as describing the internal behavior of glasses
+\cite{bib:zanotto2018} and the entropy measurement
+\cite{bib:schmelzer2018,bib:nemilov2018} cannot be completely addressed with linear thermodynamics.
+
+In addition, some aspects regarding the classical $H$-theorem and the
+Boltzmann $H$-functional require to be revised, for the purpose of improving their mutual consistency.
+For example, the modification of the $H$-theorem to include phenomena stemming from stochastic
+trajectories, violations of the second law of thermodynamics, and the relation
+between Shannon's measure of information and the Boltzmann's entropy
+\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.
+
 In order to contribute to the construction of a consistent classical and
-quantum $H$-theorem, and a new formalism to describe out of equilibrium
-systems, we develop a new theoretical basis to describe them. In particular, we
-are interested in including a non-homogeneous distribution function in the
-mathematical model, and so including non-homogeneous systems in the $H$
-theorem. Besides, we are interested in developing the quantum analogous of
-Boltzmann's $H$-theorem considering a non-spatial-homogeneous distribution
-function and solve the problem of the correspondence principle mentioned above.
+quantum $H$-theorem, within a formalism that describes out-of-equilibrium
+non-homogeneous systems, in this article we propose a new theoretical framework.
+Specifically, for both classical and quantum systems, we include non-homogeneous distribution functions in
+the $H$-functionals, and consider non-homogeneous systems in the proofs of the resulting
+$H$-theorems. Our proposed $H$-functionals obey
+the correspondence principle, but more importantly, these functionals describe
+the time-evolution of spatially non-homogeneous
+systems towards equilibrium.
 
@@
 we provide an alternative method to obtain the Maxwell-Boltzmann
-distribution using the variational method, and an alternative demonstration of
-the $H$-theorem for classical systems. In Section~\ref{sec:quantum}
+distribution using the variational method, and the demonstration of
+an alternative $H$-theorem for classical systems. In Section~\ref{sec:quantum}
 we review the Tolman proposal for the quantum
@@
 considered the analogous of the classical Boltzmann transport equation. Finally,
-we discuss some key ideas derived from our approach in Section~\ref{sec:disscussion},
-and provide some conclusions in Section~\ref{sec:conclusions}.
+we discuss some key ideas derived from our approach and close with a brief summary
+in Section~\ref{sec:disscussion}.

### 2021-01-25 tex/quantum.tex
Se agregó un breve desarrolo para la sección de relajación en términos de la funcional H evaluada en diferentes puntos. Corregí un texto en rojo en la sección quantum donde estaban mal identificados los signos para bosones y fermiones. También agregué en rojo la subsección OUT OF EQUILIBRIUM, NON-HOMOGENEOUS SYSTEMS debido a que ahí no se hace mención sobre sistemas fuera de equilibrio sino hasta que demostramos el teorema H.
Cambié lo de la subsección en rojo, error de numeración
Rewording: All sections finished but 5.

@@
 
-The classical $H$-theorem is even to these days one of the pillars whereon the classical statistical
-physics is founded. Unfortunately, regardless of several attempts have been made \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:han2015,bib:das2018,bib:vonneumann2010}, 
-the generality of the classical $H$-theorem has no an equally robust quantum match. In this section, we shall
+The classical $H$-theorem is even to these days one of the pillars on which the classical statistical
+physics is founded. Unfortunately, regardless of several attempts that have been made \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:han2015,bib:das2018,bib:vonneumann2010}, 
+the generality of the classical $H$-theorem has no equally robust quantum match. In this section, we shall
 propose and analize an alternative quantum $H$-functional using the variational method.
-We start by briefly outline a typical textbook demonstration of the quantum $H$-theorem
-\cite{bib:tolman}, then, we present the analysis of our proposed $H$-functional.
+We start by briefly outlining a typical textbook demonstration of the quantum $H$-theorem
+\cite{bib:tolman}, and subsequently we present the analysis of our proposed $H$-functional.
 
@@
 where $G$ describes the total number of accessible quantum states of the gas that satisfy the
-above mentioned conditions \cite{bib:tolman}.
+above conditions \cite{bib:tolman}.
 The quantum $H$-theorem can be demostrated as follows.  $G$ can be divided into groups of neighboring states,
-$g_k$, and a certain occupation numbers, $n_k$, can be associated with each of these
+$g_k$, and certain occupation numbers, $n_k$, can be associated with each of these
 groups. Thus, the above functional takes the form:
@@
 
-Assuming that the energy exchange between particles is produced by collisions between particles.
-By using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
+Assuming that the energy exchange between particles is produced by collisions between particles,
+and using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
 %
@@
 In Eq.~(\ref{eq:amnkldef}), $\Delta\epsilon$ is the net energy change occuring during the
-collision and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where
+collision and $|I_1-I_2|^2=|V_{mn,kl}|^2$, where
 $V_{mn,kl}$ is the element of the transition matrix of a binary collision.
 It is important to remark that in deriving Eq.~(\ref{eq:changen}), the equal a priori
-probabilities and the random a priori phases hypotheses were taken valid. The 
-the random a priori phases hypothesis can be
-considered an analogous to the molecular chaos hypothesis
-\cite{bib:das2018} as the mechanism to introduce stochasticity into the system.
+probabilities and the random a priori phases hypotheses were taken as valid. The 
+random a priori phases hypothesis can be
+considered an analogue of the molecular chaos hypothesis
+\cite{bib:das2018}, as it is the mechanism by which stochasticity is introduced into the system.
 
-Substituting Eq.~(\ref{eq:changen}) into Eq.~(\ref{eq:quantumdHdt}) after some algebra, it can be proven that
+Substituting Eq.~(\ref{eq:changen}) into Eq.~(\ref{eq:quantumdHdt}), it is straightforward to prove that
 %
@@
 %
-Here, the upper and lower sign refer to \commr{bosons and fermions respectively}.
+Here, the upper and lower sign refer to fermions and bosons respectively.
 In addition, when needed, each cell will have an associated local chemical potential,
@@
 $\{\alpha_M\}$ and $\{\beta_M\}$, it is straightforward to obtain
-(see also the discussion of Eq.~(\ref{eq:deltaHpdeltafpj})):
+(see also the discussion related to Eq.~(\ref{eq:deltaHpdeltafpj})):
 %
@@
 %
-Thus, in this zero-order approximation the form of equilibrium distribution functions are conserved.
+Thus, in this zero-order approximation, the form of equilibrium distribution functions are conserved.
 
@@
         		\left(\frac{1}{\exp\big(-\bar{\alpha}-\bar{\beta}\epsilon_{n}\big)\pm 1} \right)
-           \commr{\ln \left(\frac{1}{\exp\big(-\bar{\alpha}-\bar{\beta}\epsilon_{n}\big)\pm 1} \right)}
+           \ln \left(\frac{1}{\exp\big(-\bar{\alpha}-\bar{\beta}\epsilon_{n}\big)\pm 1} \right)
          \right]\nonumber \\
@@
 %
-This quantity is what we refer to as ``variational entropy'',
+This quantity is what we will refer to as ``variational entropy'',
 and this name reflects the fact that it was obtained \textit{via}
-the variational method. {\color{blue} We will use it at the end of this article to analyze relaxation processes.}
+the variational method.
 
@@
 For quantum systems, we also accept the validity of the local equilibrium hypothesis
-for every cell in the system. This allows us to define non-homogeneous systems,
+for every cell in the system. This allows to define non-homogeneous systems,
 wherein thermodynamic quantities are well-defined on a cell-per-cell basis,
@@
 In the present context, $|\Delta_M|$ and $|\delta_M|$ are sufficiently large so they are not
-fluctuations of the system, but enough small so that the local equilibrium hypothesis is
+fluctuations of the system, and sufficiently small so that the local equilibrium hypothesis is
 valid for $t>0$ (we set $t_0=0$, and $t_0$ is the initial time at which the system is prepared).
@@
 Approximate the logarithmic terms, corresponding to
-Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$:
+Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$ as:
 %

### 2021-01-25 tex/introduction.tex
cambié las ubicaciones de las citas y agregué unas que faltaban

@@
 of the Boltzmann transport equation)
-\cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. However, these approaches 
+\cite{bib:grabert1974,bib:wang2014,%
+bib:angel2017,
+bib:amato2020,bib:nicacio2015,bib:hussein2014}. However, these approaches 
 are not consistent with the classical-quantum correspondence principle.
@@
 and attempted to proof a
-quantum $H$-theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
+quantum $H$-theorem \cite{bib:gorban2014,bib:bennaim2017,bib:silva2010,%
+bib:deroeck2006,bib:acharya2019,bib:kastner2017,bib:han2015,bib:das2018,bib:vonneumann2010}.
 However, whether or not the homogeneous distribution function hypothesis is assumed, or
@@
 interpretation of the quantum entropy is still in discussion %[19-26].
-\cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:acharya2019,%
-bib:kastner2017,bib:gring2012,bib:han2015,bib:das2018}.
+\cite{bib:silva2010,bib:acharya2019,bib:kastner2017,bib:han2015,bib:brown2008,%
+bib:dragoljub2009,bib:vonneumann2010,%
+bib:syros1999,bib:lesovik2016,
+bib:lesovik2019}.
 
@@
 For example, the modification of the $H$-theorem to include phenomena stemming from stochastic
-trajectories, violations of the second law of thermodynamics, and the relation
-between Shannon's measure of information and the Boltzmann's entropy
-\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.
+trajectories, violations of the second law of thermodynamics, the relation
+between Shannon's measure of information and the Boltzmann's entropy, 
+and calculation of thermodynamical quantities and thermalization for some systems
+\cite{bib:gorban2014,bib:li2019,bib:gring2012,bib:nemilov2018,bib:wang2014}.
 

### 2021-01-25 tex/introduction.tex
Typos: intro

@@
 between Shannon's measure of information and the Boltzmann's entropy, 
-and calculation of thermodynamical quantities and thermalization for some systems
+and the calculation of thermodynamical quantities and thermalization of certain systems
 \cite{bib:gorban2014,bib:li2019,bib:gring2012,bib:nemilov2018,bib:wang2014}.

### 2021-01-25 tex/classical.tex
Additions to intro.

@@
 %
-Here $\vec r_M$ is the center of the cell of index $M$, and $f'_M(\vec v,t)\neq f'_N(\vec v,t)$
+Here, $\vec r_M$ is the center of the cell of index $M$, and $f'_M(\vec v,t)\neq f'_N(\vec v,t)$
 for $M\neq N$. This extended definition
@@
 %
-Notice that assuming that every $f'_M$ is homogeneous implies that we are accepting the validity of the
+Notice that assuming every $f'_M$ to be homogeneous implies that we are accepting the validity of the
 local equilibrium hypothesis. In addition, the set $\{f'_{M}(\vec{v},t)\}$ must satisfy the following restrictions:

### 2021-01-25 tex/introduction.tex
Additions to intro.

@@
 
-The theoretical basis and the procedures that allow to describe classical equilibrium systems
+The theoretical basis and the procedures that allow to describe equilibrium systems
 are well-stablished. These procedures can be applied to treat a wide range of the systems
 present in nature, and include both the macroscopic phenomenological methods (thermodynamics)
-as well as the microscopic description (statistical mechanics). For instance,
+as well as the microscopic description (statistical mechanics).
+(Of course, regarding out-of equilibrium systems, the present situation is still a challenge.) 
+For instance,
 in the Kinetic Theory of gases, the behavior of a dilute classical gas
@@
 
-In contrast, for quantum equilibrium systems, the construction of a framework,
+On the other hand, for quantum equilibrium systems, the construction of a framework,
 with the same level of
@@
 we provide an alternative method to obtain the Maxwell-Boltzmann
-distribution using the variational method, and the demonstration of
-an alternative $H$-theorem for classical systems. In Section~\ref{sec:quantum}
+distribution using the variational method, followed by our proposal of
+an alternative $H$-functional for classical systems, and the demonstration of
+the respective $H$-theorem . In Section~\ref{sec:quantum}
 we review the Tolman proposal for the quantum
 version of the $H$-theorem (quantum $H$-theorem)
-\commr{and how the Bose-Einstein and Fermi-Dirac distributions are treated within
-this framework},
-and our alternative proof of the quantum $H$-theorem. In 
+and how the Bose-Einstein and Fermi-Dirac distributions are treated within
+this framework, subsequently we present our proposal for a quantum $H$-functional
+and the proof of the corresponding quantum $H$-theorem. In 
 Section~\ref{sec:qccorrespondence}
 the classical-quantum correspondence between the
-quantum $H$-functional and the classical $H$-functional is analyzed.
+quantum and classical $H$-functionals is analyzed.
 In Section~\ref{sec:relaxproc} we explore how relaxation processes occur
-in a quantum ideal gas; here we provide a time-evolution equation that could be
-considered the analogous of the classical Boltzmann transport equation. Finally,
+in a quantum ideal gas; \commb{here we provide a time-evolution equation that could be
+considered the analogous of the classical Boltzmann transport equation.} Finally,
 we discuss some key ideas derived from our approach and close with a brief summary

### 2021-01-26 tex/introduction.tex
Verstion submitted to Entropy on 20210126

@@
 is described through the Boltzmann transport equation \cite{bib:huang},
-and the time-evolution of a system towards the equilibrium is proved by
-means of the Boltzmann $H$-theorem, wherein the
-spatial-homogeneous distribution function hypothesis plays a crucial role.
+and the time-evolution of a system towards the equilibrium is finely accounted for
+through the Boltzmann $H$-theorem.
 
-On the other hand, for quantum equilibrium systems, the construction of a framework,
-with the same level of
-success and universality as the classical version, is still an open problem. For instance,
+On the other hand, for quantum out-of-equilibrium systems, the construction of a
+kinetic framework, with the same level of
+success and universality as the classical version, still presents some fundamental challenges. For instance,
 in order to obtain a complete correspondence principle, between classical mechanics
@@
 $H$-theorem \cite{bib:tolman}, using probability transition relations, the random
-phases hypothesis, and a spatially homogeneous distribution function.
+phases hypothesis, and an $H$-functional defined in terms of a spatially homogeneous distribution function.
 Also, Tolman proposed a potential quantum analogous of the transport equation,
-in terms of the occupation number,
+in terms of the occupation numbers,
 by applying time perturbation theory. Further attempts, under the quantum operator
 formalism, have addressed the description of quantum transport phenomena by means of the Hamiltonian
-of the system and the master equation (which here is the analogous
+of the system and the master equation (which is, in these works, the analogous
 of the Boltzmann transport equation)
@@
 However, whether or not the homogeneous distribution function hypothesis is assumed, or
-if their framework obeys the correspondence principle is unclear or not discussed whatsoever.
-Furthermore, the general validity
-of the quantum $H$-theorem, the second law of thermodynamics, and the
-interpretation of the quantum entropy is still in discussion %[19-26].
+if their framework fulfills the correspondence principle is unclear or not discussed whatsoever.
+After the pioneering work of Tolman, at several stages, there has been some discussion regarding
+the general validity of the quantum $H$-theorem,
+some possible violations of the second law of thermodynamics, and the
+interpretation of the quantum entropy  %[19-26].
 \cite{bib:silva2010,bib:acharya2019,bib:kastner2017,bib:han2015,bib:brown2008,%
@@
 under construction, although several approaches have been developed. For instance,
-the celebrated Onsager formulation (\textit{aka} linear thermodynamics)
+the celebrated Onsager formulation (linear thermodynamics)
 \cite{bib:keizer1987,bib:onsager1931} have been successful in describing irreversible chemical 
-and physical phenomena. Nonetheless, some aspects such as describing the internal behavior of glasses
+and physical phenomena. Nonetheless, some topics, such as describing the internal behavior of glasses
 \cite{bib:zanotto2018} and the entropy measurement
-\cite{bib:schmelzer2018,bib:nemilov2018} cannot be completely addressed with linear thermodynamics.
+\cite{bib:schmelzer2018,bib:nemilov2018}, cannot be completely addressed with linear thermodynamics.
 
@@
 the $H$-functionals, and consider non-homogeneous systems in the proofs of the resulting
-$H$-theorems. Our proposed $H$-functionals obey
+$H$-theorems. Our proposed $H$-functionals satisfy
 the correspondence principle, but more importantly, these functionals describe
@@
 and how the Bose-Einstein and Fermi-Dirac distributions are treated within
-this framework, subsequently we present our proposal for a quantum $H$-functional
+this framework, subsequently we present our proposal for a quantum $H$-functional,
 and the proof of the corresponding quantum $H$-theorem. In 
@@
 In Section~\ref{sec:relaxproc} we explore how relaxation processes occur
-in a quantum ideal gas; \commb{here we provide a time-evolution equation that could be
-considered the analogous of the classical Boltzmann transport equation.} Finally,
+in a quantum ideal gas and, on the basis of what we call variational entropy,
+we propose a time-evolution equation for the distribution function. Finally,
 we discuss some key ideas derived from our approach and close with a brief summary

### 2021-01-26 tex/quantum.tex
Verstion submitted to Entropy on 20210126

@@
 
-Consider a diluted gas of $N$ non-interacting quantum particles (either bosons or fermions),
+Consider a dilute gas of $N$ non-interacting quantum particles (either bosons or fermions),
 contained by a vessel of volume
@@
 
-Consider a diluted gas enclosed by a perfectly isolated vessel of volume $V$, with
+Consider a dilute gas enclosed by a perfectly isolated vessel of volume $V$, with
 total energy $E$, and total number of quantum particles $N$, which can be free fermions or
@@
 We propose the following functional as an alternative $H$-functional
-for quantum non-homogeneous diluted gases:
+for quantum non-homogeneous dilute gases:
 %
@@
 %============================================================
-If the system is in equilibrium (we will show that it will also be valid at $t\to\infty$), the temperature becomes homogeneous throughout
+If the system is in equilibrium, the temperature becomes homogeneous throughout
 the complete system. Also, the local number of particles, the local
@@
 %
-By means of the above equation, we can recover the distribution function and the entropy of a diluted quantum gas in equilibrium as follows. Setting $\bar\alpha=\mu/kT$ and $\bar\beta=-1/kT$,
+By means of the above equation, we can recover the distribution function and the entropy of a dilute quantum gas in equilibrium as follows. Setting $\bar\alpha=\mu/kT$ and $\bar\beta=-1/kT$,
 and substituting them in Eq.~(\ref{eq:qfneq}), it renders the Fermi-Dirac and Bose-Einstein
@@
 
-\commr{
-An additional consideration is necessary for treating gases composed of fermions,
-which must obey the relation $f_{Mn}(t)\leq1$, hence
+An additional consideration is necessary for treating Fermi gases.
+Since, for these systems, $f_{Mn}(t)\leq1$, we have
 %
@@
 $\bar f_{n}=1$ is certainly satisfied if the system is at zero temperature. In this state,
-all energy levels below and including the Fermi energy are occupied because of
-the exclusion principle, therefore the system will
+all energy levels below and including the Fermi energy are occupied, thus the system will
 necessarily be homogeneous , and consequently $g_{nM}=0$. In this article, we will
 omit this scenario and will only discuss Fermi gases whose temperature is non zero.
-}
 
@@
 
-Succinctly, considering a diluted quantum gas contained in a vessel of volume $V$ (divided into
+Succinctly, considering a dilute quantum gas contained in a vessel of volume $V$ (divided into
 $K$ small cells), with total energy $E$ and $N$ quantum free particles, which initially

### 2021-01-26 tex/classical.tex

@@
 position $\vec r$ and velocity $\vec v$, and are contained in a $\mu$-space volume element $d^3rd^3v$.
-Assuming a deterministic Newtonian description of the molecules motion, 
+Assuming a deterministic Newtonian description of the molecules' motion, 
 as well as the invariance of the $\mu$-space volume measure, one arrives to the Boltzmann rate equation:
@@
 originally introduced by Boltzmann in 1872,
-describes a diluted gas occupying a volume $V$, at temperature $T$, with total energy $E$, and total
+describes a dilute gas occupying a volume $V$, at temperature $T$, with total energy $E$, and total
 number of free classical particles $N$.
@@
 
-The physically correct espontaneous time-evolution of an out of equilibrium diluted gas 
+The physically correct espontaneous time-evolution of an out of equilibrium dilute gas 
 is corroborated by the $H$-theorem.
@@
 
-On the other hand, considering a diluted gas in equilibrium
+On the other hand, considering a dilute gas in equilibrium
 with no applied external forces, \textit{i.e.}
-$(\partial f/\partial t)=0$ and $f$ is independent of $\vec r$), 
+$(\partial f/\partial t)=0$ and $f$ is independent of $\vec r$, 
 it can be directly proven that the equilibrium distribution function obtained from
@@
 gas, one of the salient features of an out-of-equilibrium condition is the existence of
-inhomogeneities in the system.
+inhomogeneities in the system, which is not considered in the above.
 
@@
 that the distribution function is homogeneous and that the hypothesis of
-molecular chaos is fullfiled. In order to extend the previous procedure to systems whose
+molecular chaos is fulfilled. In order to extend the previous procedure to systems whose
 distribution functions are non-homogeneous, which might allow for studying systems
 being in a more general out-of-equilibrium condition, we introduce a modified $H$-functional 
-in the following. For the sake of clearity and simplicity, we will use primed functions
+in the following. For the sake of clarity and simplicity, we will use primed functions
 and quantities to denote the classical case, so as to differentiate them from the quantum
@@
 %
-where $C$ is a constant. We notice that the use of the variational procedure on
+where $C$ is a constant. We notice that applying the variational procedure on
 $\mathcal{H}'$ predicts that when the equilibrium is reached, the
@@
 %
-\begin{equation}
+\begin{equation}\label{CH3}
    \mathcal{H}''(t)=\sum_{M=1}^{K}\int \left[f'_M(\vec{v},t)
-    \ln f'_M(\vec{v},t)-f'_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
+    \ln f'_M(\vec{v},t)-f'_M(\vec{v},t)\right]\mathrm{d}^3v.
 \end{equation}

### 2021-01-26 tex/disscussion.tex

@@
 functions both in out-of-equilibrium states (locally) and when the system attains the global
-equilibrium state. \commb{The relaxation process of the system is described by
-monotonic functions that account for deviations from the global equilibrium.}
+equilibrium state. The relaxation process of the system is described by
+monotonic functions that account for deviations from the global equilibrium.
 
@@
 out-of-equlibrium. Because the functionals $\mathcal{H}'$ and $\mathcal{H}$ can be related to
-the entropy of diluted gases, either classical or quantum, the fact that these functionals are defined
+the entropy of dilute gases, either classical or quantum, the fact that these functionals are defined
 over a system divided into cells enables their use for defining the entropy
-of out-of-equilibrium systems, other than diluted gases. In particular, and 
+of out-of-equilibrium systems, other than dilute gases. In particular, and 
 derived from our previous work (\textit{e.g.} \cite{bib:nicolas2020,bib:nicolas2016}), the
@@
 growth of complex physical systems, such as fractals. Possibly, studying these systems might also
-shed light on the functional form of $g'_M$ and $g_{nM}$.
+shed light on the explicit functional form of $g'_M$ and $g_{nM}$.
 

### 2021-01-26 tex/qccorrespondence.tex

@@
 method can be applied to $H$-functionals, which correctly describes the behavior of
-classical and quantum diluted gases, as to their respective time-evolution.
+classical and quantum dilute gases, as to their respective time-evolution.
 Both $H$-functionals defined in Eqs.~(\ref{eq:cHdef}) and

### 2021-01-26 tex/relaxproc.tex
Additions to intro.

@@
 \section{Relaxation processes in degenerated quantum gases}\label{sec:relaxproc}
-In order to obtain an equation of motion for the quantum gas, we evaluate 
-$\Delta H=H(t_2)-H(t_1)$ where the $H$ functional is evaluated in diferent times such that $t_2 > t_1$.
-We begin using the following notation $f_{nM}(t_2)\equiv f'_{nM}$ and 
-$f_{nM}(t_1) \equiv f''_{nM}$. 
-Now, using (\ref{eq:qHdef}) to calculate $\Delta H$ with the previous notation, we obtain
-\begin{eqnarray} \label{eq:deltaHdef}
-	\Delta H &=& \sum_{M}^{K}\sum_n [f'_{nM} \ln f'_{nM} - f''_{nM}	\ln f''_{nM}\nonumber \\ 
-			&&\pm (1\mp f'_{nM}) \ln(1\mp f'_{nM}) \mp (1\mp f''_{nM}) \ln (1\mp f''_{nM})]\delta V.  
+
+In order to obtain a time-evolution equation for an out-of-equilibrium quantum gas,
+we propose the following approach. We start by evaluating $\Delta\mathcal{H}=
+\mathcal{H}(t_2)-\mathcal{H}(t_1)$,
+where our quantum $H$-functional ---Eq.~(\ref{eq:qHdef})--- is evaluated at diferent times $t_1$ and $t_2$, with
+$t_2 > t_1$. This yields:
+%
+\begin{eqnarray}\label{eq:deltaHdef}
+	\Delta\mathcal H &=& \sum_{M=1}^{K}\sum_n [f''_{nM} \ln f''_{nM} - f'_{nM}	\ln f'_{nM}\nonumber \\ 
+			&&\pm (1\mp f''_{nM}) \ln(1\mp f''_{nM}) \mp (1\mp f'_{nM}) \ln (1\mp f'_{nM})]\delta V_M. 
 \end{eqnarray}
-In order to obtain a general equation of motion for the distribution function 
-$f_{nM}$, we only take the limit when $(t_2-t_1) \rightarrow 0 $. 
-However, our current work is focused on the deviation $g_{nM}$. So, keeping the same workline
-we apply the first order approximation (\ref{eq:qFirstOrd}) to (\ref{eq:deltaHdef}) to obtain the following expression
+%
+In the above equation, and for the rest of this section, we use the short-hand notation
+$f_{nM}(t_2)\equiv f''_{nM}$ and 
+$f_{nM}(t_1) \equiv f'_{nM}$.
+Subsequently, in Eq.~(\ref{eq:deltaHdef}), we replace the distribution functions by their expresions in terms
+of deviations from equilibrium ---Eq.~(\ref{eq:qFirstOrd})---, which renders:
 \begin{eqnarray}\label{eq:deltaH1}
-	\Delta H &=& \sum_{M}^{K} \sum_n [\bar{f}_n(1+g'_{nM}) \ln \bar{f}_{n}(1+g'_{nM})  -  
-			\bar{f}_n(1+g''_{nM}) \ln \bar{f}_n(1+g''_{nM}) \nonumber \\
-		    &&  \pm (1 \mp \bar{f}_n \{1+g'_{nM}\}) \ln (1\mp \bar{f}_n\{1+g'_{nM} \})\nonumber \\
-		    &&  \mp (1\mp \bar{f}_n \{ 1+g''_{nM} \}) \ln (1\mp \bar{f}_n \{1+g''_{nM} \})]\delta V.\nonumber \\
+	\Delta\mathcal H &=& \sum_{M=1}^{K} \sum_n\Big[\bar{f}_n(1+g''_{nM}) \ln \bar{f}_{n}(1+g''_{nM})  -  
+			\bar{f}_n(1+g'_{nM}) \ln \bar{f}_n(1+g'_{nM})\Big. \nonumber \\
+		    &&  \pm \big(1 \mp \bar{f}_n \{1+g''_{nM}\}\big) \ln \big(1\mp \bar{f}_n\{1+g''_{nM} \}\big)\nonumber \\
+		    &&  \mp \big(1\mp \bar{f}_n \{ 1+g'_{nM} \}\big) \ln\big(1\mp \bar{f}_n \{1+g'_{nM} \}\big)\Big.\Big]
+		    \delta V_M.
 \end{eqnarray}
-The equation (\ref{eq:deltaH1}) can be rewritten, after some approximations on the logarithm functions, as
+Afterwards, we expand the logarithmic terms up to the first-order in $g'_{nM}$ and $g''_{nM}$
+and rearrange the result, which gives:
+%
 \begin{equation}\label{eq:deltaH2}
-\Delta H = \sum_{M}^{K}\sum_n [\bar{f}_n(1+\ln \bar{f}_n)-\bar{f}_n \{\ln(1\mp \bar{f}_n)\mp 1 \}](g'_{nM}-g''_{nM})\delta V.
+\Delta\mathcal H = \sum_{M}^{K}\sum_n [\bar{f}_n(1
+	+\ln \bar{f}_n)-\bar{f}_n \{\ln(1\mp \bar{f}_n)\mp 1 \}](g''_{nM}-g'_{nM})\delta V_M.
 \end{equation}
-The equation (\ref{eq:deltaH2}) must be analyzed taking the limit $(t_2-t_1) \rightarrow 0$ 
-to obtain the equation of motion for the deviation $g_{nM}$.
-This analysis will not be treated in this article.
+%
+Finally we divide Eq.~(\ref{eq:deltaH2}) by $\Delta t\equiv t_2-t_1$, and take the limit
+$\Delta t\to0$ to obtain:
+%
+\begin{equation}\label{eq:deltaH3}
+\frac{d\mathcal H}{dt} = \sum_{M}^{K}\sum_n [\bar{f}_n(1
+	+\ln \bar{f}_n)-\bar{f}_n \{\ln(1\mp \bar{f}_n)\mp 1 \}]\left(\frac{g_{nM}}{dt}\right)\delta V_M.
+\end{equation}
+%
+
+Eq.~(\ref{eq:deltaH3}) is, within our framework, the time-evolution equation for $g_{nM}$.
+Clearly, for describing a concrete situation, it is required to provide a specific approximation for the deviation function $g_{nM}$, however this will be the subject of future work.
 

### 2021-03-10 tex/classical.tex
Removed comments (pdf remains the same)

@@
-%\section{Classical}
-%\subsection{Boltzmann transport equation}
-%\subsection{H theorem and the Maxwell-Boltzmann distribution}
-%Todo esto de arriba en la forma tradicional y muy resumido. Solo remarcando lo fundamental.
-%\subsection{Out of equilibrium, non-homogeneous distributions}.
-%Entrar a la generalizacion, ahi si de manera muy detallada y con amplitud discutida
-%fisicamente, del tratamiento con las celdas. Es decir, la demostracion del teorema H
-%clasico pero con el enfoque de funciones de distribucion no homogeneas y por tanto
-%con el tratamiento de celdas.
-
-%----------------------------------
-
 \section{Classical scheme}\label{sec:classical}
@@
 %
-\begin{comment}
-\begin{eqnarray}
-    \ln f'_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
-       -1\quad\Rightarrow\quad f_J(\vec{v}')\\
-    & = & e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1}\nonumber\\
-    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{eq:relacion1},
-\end{eqnarray}
-\end{comment}
 \begin{equation}\label{eq:relacion1}
@@
 %
-%---------------------------------------
-%The text below is not needed, also it may be wrong. $g(t_0)>0$ indicates $f(t_0)>\bar f$, but this does
-%not necessarily imply excess of particles and/or energy. E.g., a particular combination of a
-%few missing particles and a considerable energy excess might render the corresponding
-% $g(t_0)$ to be greater than zero.
-\begin{comment}
-Cells with ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)$, $I=1,\,\dots\,,L$, 
-have an excess of particles and/or energy (relative to the mean value in equilibrium).
-In contrast, cells with ${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)$, $J=1,\,\dots\,,P$,
-have a scarcity of particles and/or energy
-(relative to the mean value in equilibrium).
-In addition, $\dot{g}^{+}_{J}$ represents the
-change on the deviation on cells that have an excess of particles or energy
-while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that
-have missing particles or energy.
-On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells
-that have an excess of particles or energy. In contrast, $g^{-}_{J}$ represents
-the value of the deviation on cells that have missing particles or energy.
-Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles
-and/or energy and so, $g^{+}_{J}>0$. On the other hand, $\dot{g}^{-}_{J}>0$
-describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.
-
-We sort the previous ideas in the following form
-%
-\begin{equation}\label{separacionclassical}
-\begin{array}{rl}
-  g^{+}_{J}=+|g^{+}_{J}|; & \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|,\\
-  g^{-}_{J}=-|g^{-}_{J}|; & \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}|,
- \end{array}
-\end{equation}
-%
-and consequently, (\ref{classicalH3}) obtains the following form
-
-We can observe that $\bar{f}(\vec{v})$ in (\ref{eq:classicalH4}) is always
-positive. Also the deviation and its derivative are positive, then all the
-expression is positive. However, the global sign makes the derivative of the
-$\mathcal{H}'$ with respect to time is always less to zero. With this we proved that
-$\frac{d\mathcal{H}'}{dt}<0$. If the system is in equilibrium,
-$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{d\mathcal{H}'}{dt}=0$.
-\end{comment}
-%-----------------------------
 %
@@
 $T_M$, and Legendre multipliers $\alpha_M$ and $\beta_M$.
-
-\begin{comment}
-Joining both results, we can say the following statement:
-Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal
-volume elements), total energy $E$ and total number of free particles $N$.
-Consider also the system has inhomogeneities and suffers a relaxation process.
-If we define the following functional
-%
-\begin{equation}\label{CH3}
-   \mathcal{H}'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v,
-\end{equation}
-%
-where $f_M(\vec{v},t)$ is the local distribution function of each cell in the
-system, then in the first-order approximation
-%
-\begin{equation}
-    \frac{d\mathcal{H}'}{dt} \leq 0.
-\end{equation}
-%
-This statement will be the classical $H$-theorem with inhomogeneities.
-
-In the next section, we present a proposal quantum version of the
-$H$-theorem defined by Tolman and our proposal quantum version of the $H$
-theorem applying the method of the volume divided into cells.
-\end{comment}

### 2021-03-10 tex/classical.tex
Cambridge corrections applied.

@@
 
-Boltzmann Kinetic theory of gases represents a fundamental connection
+The Boltzmann kinetic theory of gases represents a fundamental connection
 between the microscopic nature of matter and the phenomenological macroscopic laws of classical thermodynamics.
-The stochasticity introduced by the molecular chaos hypothesis in the, otherwise deterministic, kinetics of the
-particles allows the demostration 
-of the celebrated Boltzmann $H$-theorem. In contrast, in this article we propose an alternative approach, which is 
-developed by means of a variational procedure applied to an $H$-functional.
-Let us start this section by briefly accounting for the important elements of the 
-standard derivation of the Boltzmann transport equation and demostration the $H$-theorem,
+The stochasticity introduced by the molecular chaos hypothesis in the otherwise deterministic kinetics of the
+particles allows for the demonstration 
+of the celebrated Boltzmann $H$-theorem. In contrast, in this article, we propose an alternative approach 
+developed using a variational procedure applied to an $H$-functional.
+We start this section by briefly accounting for the important elements of the 
+standard derivation of the Boltzmann transport equation and demonstrating the $H$-theorem,
 such as they are presented in classical textbooks \cite{bib:huang}.
@@
 
-In Boltzmann kinetic theory of gases, one starts by defining the distribution function, 
+The first step in the Boltzmann kinetic theory of gases is defining the distribution function, 
 $f(\vec{r},\vec{v},t)$, as the average number of molecules that, at time $t$, have 
 position $\vec r$ and velocity $\vec v$, and are contained in a $\mu$-space volume element $d^3rd^3v$.
-Assuming a deterministic Newtonian description of the molecules' motion, 
-as well as the invariance of the $\mu$-space volume measure, one arrives to the Boltzmann rate equation:
+Assuming a deterministic Newtonian description of molecular motion, 
+as well as the invariance of the $\mu$-space volume measure, one arrives at the Boltzmann rate equation:
 %
@@
 external force applied to the system, and $f_1$ and $f_2$ ($\tilde f_1$ and $\tilde f_2$) are the distribution
-functions of particles 1 and 2, respectively, before (after) the collision.
+functions of particles 1 and 2, respectively, before (and after) the collision.
 
 Particle dynamics and the effects of external forces are described
-by the left-hand side of Eq.~(\ref{eq:transport}), and the right-hand side
-is derived considering binary collisions between particles and accepting
+by the left-hand side of Eq.~(\ref{eq:transport}). The right-hand side
+is derived by considering binary collisions between particles and accepting
 the \textit{molecular chaos hypothesis}, \textit{i.e.} it is assumed that
@@
 %-------------------------
-\subsection{A brief summary of the H-theorem and the Maxwell-Boltzmann distribution}
+\subsection{A summary of the H-theorem and the Maxwell-Boltzmann distribution}
 
@@
 number of free classical particles $N$.
-In order to clearly distinguish the Boltzmann functional $H_B$, we denote hereafter
+To clearly distinguish the Boltzmann functional $H_B$, we denote hereafter
 the Maxwell-Boltzmann distribution function as $f_B$.
 
-The physically correct espontaneous time-evolution of an out of equilibrium dilute gas 
+The physically correct spontaneous time-evolution of an out of equilibrium dilute gas 
 is corroborated by the $H$-theorem.
-This theorem establishes that if the two conditions, a) the homogeneous
+This theorem establishes that if a) the homogeneous
 function $f(\vec{v},t)$ satisfies the Boltzmann
@@
 then the system is in the equilibrium state. 
-The $H$-theorem is straightforward to prove through Eq.~(\ref{eq:transport}) \cite{bib:huang}, 
-and it assures the consistency between our microscopic approach to describe the system's spontaneous 
+The $H$-theorem is straightforward to prove using Eq.~(\ref{eq:transport}) \cite{bib:huang}, 
+and it assures the consistency between our microscopic approach to describe the system’s spontaneous 
 time-evolution and the phenomenological observations
-established by the $2^{\textrm{nd}}$ law of classical thermodynamics; in fact,
+established by the second law of classical thermodynamics; in fact,
 $H_B$ can be associated with an entropy density.
@@
 $(\partial f/\partial t)=0$ and $f$ is independent of $\vec r$, 
-it can be directly proven that the equilibrium distribution function obtained from
+we can directly prove that the equilibrium distribution function obtained from
 Eq.~(\ref{eq:transport}) is precisely the Maxwell-Boltzmann
-distribution function. The proof of the above commences by finding the sufficient
+distribution function. The proof of the above first requires identification of the sufficient
 condition for $f$ to render a null r.h.s. of Eq. (\ref{eq:transport}). Such an
-$f$, which we denote here as $f_0$, must satisfy:
+$f$, which we denote here as $f_0$, must satisfy
 %
@@
 %
-Subsequently, the Maxwell-Boltzmann distrubution function can be obtained by
-taking the logarithm of Eq.~(\ref{eq:neccondf0}), and exploiting mechanical
-conserved quantities (see \cite[ch. 4.2]{bib:huang}).
+Subsequently, the Maxwell-Boltzmann distribution function can be obtained by
+taking the logarithm of Eq.~(\ref{eq:neccondf0}) and conserved mechanical
+quantities (see \cite[ch. 4.2]{bib:huang}).
 
-Before introducing our proposed $H$-functional, we remark the following. In defining
+Before introducing our proposed $H$-functional, we must state that in defining
 $H_B$, Eq.~(\ref{eq:hbfunctional}), it is assumed that the distribution function 
-$f$ is spatially homogeneous. This assumption simplifies the demostration 
-of the $H$ theorem, however, it introduces as well a limited conception
+$f$ is spatially homogeneous. This assumption simplifies the demonstration 
+of the $H$ theorem. However, it also introduces a limited conception
 of the out-of-equilibrium condition of the gas. Given the relatively simple nature of a dilute 
@@
 the validity of the $H$-theorem relies significantly on assuming
-that the distribution function is homogeneous and that the hypothesis of
-molecular chaos is fulfilled. In order to extend the previous procedure to systems whose
-distribution functions are non-homogeneous, which might allow for studying systems
-being in a more general out-of-equilibrium condition, we introduce a modified $H$-functional 
-in the following. For the sake of clarity and simplicity, we will use primed functions
-and quantities to denote the classical case, so as to differentiate them from the quantum
+that the distribution function is homogeneous and the
+molecular chaos hypothesis is fulfilled. To extend the previous procedure to systems with non-homogeneous
+distribution functions, which might allow for the study of systems
+in a more general out-of-equilibrium condition, we introduce a modified $H$-functional. 
+For the sake of clarity and simplicity, we will use primed functions
+and quantities to denote the classical case to differentiate them from the quantum
 analogues.
 
-Our proposed classical $H$-functional, denoted as $\mathcal{H}'$, describes a dilute classical
+Our proposed classical $H$-functional, denoted as $\mathcal{H}’ $, describes a dilute classical
 gas occupying a volume $V$. In our theoretical treatment, we divide this volume
@@
 an $H$-functional, $\mathcal{H}'_M$, a homogeneous distribution function, $f'_{M}(\vec{v},t)$,
-a number of particles, $\mathcal{N}'_M$, a temperature, $T'_M$, and an
-energy, $\mathcal{E}'_M$. Taken as a whole, the system has a
+number of particles, $\mathcal{N}'_M$, temperature, $T'_M$, and 
+energy, $\mathcal{E}'_M$. Taken as a whole, the system has an
 energy $E$, and a global number of free classical particles $N$. We also assume that the system is 
@@
 for $M\neq N$. This extended definition
-allows us to omit the symbol $\delta V_M$ on all integrals performed over the cell volume.
-In terms of $f'_M(\vec{v},t)$ and a local variable of energy, $\epsilon(\vec{v})$, we have:
+allows us to omit the symbol $\delta V_M$ in all integrals performed over the cell volume.
+In terms of $f'_M(\vec{v},t)$ and a local variable of energy, $\epsilon(\vec{v})$, we have
 %
@@
 We now use the variational method to find the extremal of $\mathcal{H}'$,
-consistent with the restrictions
-(\ref{eq:cellrestrictions}) and (\ref{eq:micro}), together with the corresponding Lagrange
-multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields:
+consistent with restrictions
+(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) together with the corresponding Lagrange
+multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields
 %
@@
 Solving the last line for $f'_J(\vec{v}')$ renders
-%
 \begin{equation}\label{eq:relacion1}
-	f'_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
+	f'_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right)
 \end{equation}
 %
-where $C$ is a constant. We notice that applying the variational procedure on
-$\mathcal{H}'$ predicts that when the equilibrium is reached, the
+where $C$ is a constant. We notice that by applying the variational procedure on
+$\mathcal{H} ‘$, we predict that when equilibrium is reached, the
 distribution function of each cell has the form of the Maxwell-Boltzmann
@@
 of particles and the local
-energy do not depend on the cell number, in a statistical sense, this is
+energy do not depend on the cell number. In a statistical sense, this is
 %
@@
 does not depend on the cell number $M$ (\textit{i.e.} $f'_M(\vec v,t)=f'(\vec v,t)$,
-$\forall M$). Several properties follow directly from this, \textit{e.g.}
+$\forall M$). Several properties arise directly from this, \textit{e.g.}
 from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$. Here we have used Eqs. (\ref{eq:Ecell}) and (\ref{eq:Ncell}).
@@
 
-In order to identify $\mathcal{H}'$ with the entropy, we need to show that $\mathcal{H}'(t)$ is extensive,
+To identify $\mathcal{H}'$ with the entropy, we need to show that $\mathcal{H}'(t)$ is extensive,
 with respect to $Kf'(\vec v,t)$.
@@
 %
-\textit{i.e.} $\mathcal{H}'_M$ is extensive, and the sum $\sum_M\mathcal{H}'_M$ is the $H$-functional of the complete
+\textit{i.e.}, $\mathcal{H}'_M$ is extensive, and the sum $\sum_M\mathcal{H}'_M$ is the $H$-functional of the complete
 system, which reduces to the Boltzmann $H$-functional.
-Therefore, $\mathcal{H}'$ can be identified with the entropy density of the system.
+Therefore, $\mathcal{H}’ $ can be identified with the entropy density of the system.
 
@@
 %
-Since $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$ (a constant)
+Since $\sum_{M=1}^{K} \int f'_M(\vec{v},t)\mathrm{d}^3v =N$ (a constant),
 and because we are mainly interested in the time-derivative of $\mathcal{H}''$,
@@
 Throughout this section, we will consider a classical gas
-whose initial condition is close to the equilibrium,
+with an initial condition close to the equilibrium,
 which ensures that the local equilibrium hypothesis remains valid during
@@
 Here $\bar {\mathcal{N}}' =N/K$ and $\bar{\mathcal{E}}'=E/K$ are
-the cell particle number and the cell energy in equilibrium, respectively, which are given by:
+the cell particle number and the cell energy in equilibrium, respectively, which are given by
 %
@@
 $\bar{\mathcal{N}}'$ and $\bar{\mathcal{E}}'$, respectively. For systems that are
-sufficiently close to equilibrium, it is reasonable to expect on the one hand that
+sufficiently close to equilibrium, it is reasonable to expect first that
 $\Delta'_M(t)\ll\bar{\mathcal{N}}'$ and
-$\delta'_M(t)\ll\bar{\mathcal{E}}'$, and on the other hand that $\Delta'_M$ and
-$\delta'_M$ are sufficiently large compared with the fluctuations of $\bar{\mathcal{N}}'$
-and $\bar{\mathcal{E}}'$. Under the same perspective, we can assume that every local
+$\delta'_M(t)\ll\bar{\mathcal{E}}'$, and second that $\Delta'_M$ and
+$\delta'_M$ are sufficiently large compared to the fluctuations of $\bar{\mathcal{N}}'$
+and $\bar{\mathcal{E}}'$. Similarly, we can assume that every local
 distribution function can be written as
@@
 
-With the previous considerations, in the following we proof an alternative
+With the previous considerations, in the following, we proof an alternative
 $H$-theorem, considering the $H$-functional, $\mathcal{H}'$, defined by Eq.~(\ref{eq:cHdef}).
 
-We commence by differentiating Eq.~(\ref{eq:cHdef}) with respect to time,
+We commence by differentiating Eq.~(\ref{eq:cHdef}) with respect to time:
 %
@@
 %
-(Here and throughout the article, we use the standard notation $\dot h\equiv(dh/dt)$.)
+(Starting here, we use the standard notation $\dot h\equiv(dh/dt)$).
 Substituting Eq.~(\ref{eq:firstorder}) into Eq.~(\ref{eq:dH1}) yields
@@
 Substituting $\ln \bar f'(\vec{v})=\exp\left(\alpha+\beta \epsilon(\vec{v})\right)$,
-see Eq.~(\ref{eq:fbardef}) and text following it, and omitting $C$ we obtain
+see Eq.~(\ref{eq:fbardef}) and subsequent text, and omitting $C$ we obtain
 %
@@
 From the definitions of $\mathcal{N}'_M$ and $\mathcal{E}'_M$ 
----Eqs.~(\ref{eq:restrictionoutsideclassical})---, and $f'_{M}(\vec{v},t)$
----Eq.~(\ref{eq:firstorder})---, it is straightforward to show that
+---Eqs.~(\ref{eq:restrictionoutsideclassical})---and $f'_{M}(\vec{v},t)$
+---Eq.~(\ref{eq:firstorder})---it is straightforward to show that
 %
@@
 %
-Therefore, because of
+Therefore, due to
 Eqs.~(\ref{eq:classsumdotdeltaseq0}) and (\ref{eq:sumderdeltaseq0}), Eq.~(\ref{eq:dH1-2}) simplifies to
@@
   \frac{d\mathcal{H}'}{dt} = \sum_{M=1}^{K}\int\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.
-\end{equation}
-%
-In order to clearly determine the time-evolution of Eq.~(\ref{eq:dHpdtbefgdecclass}),
+\end{equation}%
+To clearly determine the time-evolution of Eq.~(\ref{eq:dHpdtbefgdecclass}),
 we split the summation over $M$ into two terms:
@@
         +\sum_J^{P}\int
-      \bar f'(\vec{v}){g'_J}^{-}(\vec{v},t){{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)\mathrm{d}^3v,
+      \bar f'(\vec{v}){g'_J}^{-}(\vec{v},t){{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)\mathrm{d}^3v
 \end{equation}
@@
 or ${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)$, respectively.
-%
-%
+
 If the system's initial state is sufficiently close to equilibrium, it is
-physically suitable
+physically appropriate
 to assume that $\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$ in a monotonous manner,
@@
 In summary, considering a gas occupying a volume V (which is divided into $K$ small cells), with a
-total energy $E$ and $N$ classical free particles, whose initial state is not in the equilibrium,
-but sufficiently close to equilibrium, then the funcional
+total energy $E$ and $N$ classical free particles, whose initial state is not in equilibrium,
+but sufficiently close to equilibrium, then the functional
 %
@@
 %
----where $f'_M(\vec{v},t)$ is the cell distribution function---,
+---where $f'_M(\vec{v},t)$ is the cell distribution function---
 satisfies $d\mathcal{H}'/dt\leq0$, and the equality relation is attained at $t\to\infty$.
 In Eq.~(\ref{eq:CH3}), $f'_M(\vec{v},t)$ is the Maxwell-Boltzmann distribution
-function, which in general is different for different cells ---\textit{i.e.} the complete system
-can be non-homogeneous---, and each $f'_M(\vec v,t)$ is compatible
+function, which in general is different for different cells---\textit{i.e.}, the complete system
+can be non-homogeneous---and each $f'_M(\vec v,t)$ is compatible
 with the cell properties, such as number of particles, $\mathcal{N}'_M$, energy, $\mathcal{E}'_M$, temperature,

### 2021-03-10 tex/disscussion.tex
Removed comments (pdf remains the same)
Cambridge corrections applied.

@@
-%--------------------------------------
-% Me parece que sera necesario incluir la discusin sobre los procesos de relajacin.
-%Sin esa parte el contenido original del artculo podria parecer insuficiente para
-%una buena publicacion%
-
 \section{Comments and remarks}\label{sec:disscussion}
 
-The demonstration of the classical $H$-theorem usually starts by assuming that the gas,
+The demonstration of the classical $H$-theorem usually begins by assuming that the gas,
 despite being initially out of equilibrium, can be described by a
@@
 Subsequently, the time-evolution of the system occurs in such a manner that $dH/dt \leq 0$. 
-Therefore, this approach does not cover the evolution to equilibrium of systems with spatial inhomogeneities.
-To address this issue, in this article we
-proposed a framework that may be useful to describe the time-evolution of systems
-initially non-homogeneous. To this end, we divided the
-system into small cells, so as to conceive a system wherein the local equilibrium
-hypothesis is valid in each cell, but in such a manner that the total system is not homogeneous. Systems
+Therefore, this approach does not describe the evolution to equilibrium of systems with spatial inhomogeneities.
+To address this issue, in this article, we
+proposed a framework that may be useful to describe the time-evolution of 
+initially non-homogeneous systems. To this end, we divided the
+system into small cells to conceive a system wherein the local equilibrium
+hypothesis is valid in each cell but in such a manner that the total system is not homogeneous. Systems
 that satisfy the previous conditions will evolve towards equilibrium, and the evolution
-occurs according to $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef}),
+occurs according to $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef})
 and $d\mathcal{H}/dt\leq0$, Eq.~(\ref{eq:qHdef}), for classical and quantum gases, respectively.
-Consequently, this approach can be considered as an extension of the $H$-theorem
+Consequently, this approach can be considered an extension of the $H$-theorem
 for more realistic out-of-equilibrium systems.
@@
 correctly recover the most-probable distribution
-functions both in out-of-equilibrium states (locally) and when the system attains the global
+functions in out-of-equilibrium states (locally) and when the system attains the global
 equilibrium state. The relaxation process of the system is described by
@@
 It is clear that for describing the relaxation process of a concrete system,
-it is necessary to know, at least with some approximation level, the specific forms of the
+it is necessary to know, at least to some approximation, the specific forms of the
 monotonic functions $g'_M$ and $g_{nM}$, for classical and quantum systems.
-Whereas the complete analysis of these functions is out of the scope of the present work,
-some of their properties can be foreseen, \textit{e.g.} they must
-be consistent both with the system relaxation times and with the mechanisms of 
+Whereas the complete analysis of these functions is beyond the scope of the present work,
+some of their properties can be predicted, \textit{e.g.} they must
+be consistent both with the system relaxation times and the mechanisms of 
 energy transfer between cells.
@@
 An important aspect of the framework proposed in this work is related to the entropy of systems
-out-of-equlibrium. Because the functionals $\mathcal{H}'$ and $\mathcal{H}$ can be related to
+out-of-equilibrium. Because the functionals $\mathcal{H}'$ and $\mathcal{H}$ can be related to
 the entropy of dilute gases, either classical or quantum, the fact that these functionals are defined
 over a system divided into cells enables their use for defining the entropy
-of out-of-equilibrium systems, other than dilute gases. In particular, and 
+of out-of-equilibrium systems, other than dilute gases. Specifically, and 
 derived from our previous work (\textit{e.g.} \cite{bib:nicolas2020,bib:nicolas2016}), the
@@
 In summary, we proposed a variational procedure to demonstrate the classical and
-quantum $H$-theorems, which made possible to describe, at a mesoscopic local view
+quantum $H$-theorems, which allowed us to describe, at a mesoscopic local view
 (cell-scale), the time-evolution of an out-of-equilibrium and spatially
-non-homogeneous system, towards the equilibrium condition. In principle, this
-approach would render possible to investigate the transport phenomena, inherent to the
-equilibration process, occurring in a system that started from an spatially inhomogeneous
+non-homogeneous system moving towards the equilibrium condition. In principle, this
+approach would permit the investigation of the transport phenomena inherent to the
+equilibration process, occurring in a system with a spatially inhomogeneous
 out-of-equilibrium initial condition.
-
-%-----------------------------------------------

### 2021-03-10 tex/introduction.tex
hice un ligero cambio en las citas
Cambridge corrections applied.

@@
 
-The theoretical basis and the procedures that allow to describe equilibrium systems
-are well-stablished. These procedures can be applied to treat a wide range of the systems
-present in nature, and include both the macroscopic phenomenological methods (thermodynamics)
-as well as the microscopic description (statistical mechanics).
-(Of course, regarding out-of equilibrium systems, the present situation is still a challenge.)
+The theoretical bases and the procedures that allow us to describe equilibrium systems
+are well-established. These procedures can be applied to a wide range of natural systems,
+including both the macroscopic phenomenological methods (thermodynamics)
+and the microscopic description (statistical mechanics).
+(Out-of-equilibrium systems, of course, are still a challenge.)
 For instance,
-in the Kinetic Theory of gases, the behavior of a dilute classical gas
+in the kinetic theory of gases, the behavior of a dilute classical gas
 is described through the Boltzmann transport equation \cite{bib:huang},
-and the time-evolution of a system towards the equilibrium is finely accounted for
+and the time-evolution of a system towards equilibrium is finely accounted for
 through the Boltzmann $H$-theorem.
 
-On the other hand, for quantum out-of-equilibrium systems, the construction of a
-kinetic framework, with the same level of
-success and universality as the classical version, still presents some fundamental challenges. For instance,
-in order to obtain a complete correspondence principle, between classical mechanics
-and quantum mechanics, the quantum analogous of both the Boltzmann $H$-theorem
-and the Boltzmann transport equation have not attained a fully accepted form.
+However, for quantum out-of-equilibrium systems, the construction of a
+kinetic framework with the same level of
+success and universality as the classical version still presents some fundamental challenges. For instance,
+to obtain a complete correspondence principle between classical mechanics
+and quantum mechanics, the form of the quantum analogues of both the Boltzmann $H$-theorem
+and the Boltzmann transport equation is inadequate.
 In this context, Tolman was one of the earliest physicists to propose a quantum
-$H$-theorem \cite{bib:tolman}, using probability transition relations, the random
+$H$-theorem \cite{bib:tolman}, using a probability transition relationship, the random
 phases hypothesis, and an $H$-functional defined in terms of a spatially homogeneous distribution function.
-Also, Tolman proposed a potential quantum analogous of the transport equation,
+Tolman also proposed a potential quantum analogue of the transport equation,
 in terms of the occupation numbers,
-by applying time perturbation theory. Further attempts, under the quantum operator
-formalism, have addressed the description of quantum transport phenomena by means of the Hamiltonian
-of the system and the master equation (which is, in these works, the analogous
+by applying time perturbation theory. Additional attempts, under quantum operator
+formalism, have addressed the description of quantum transport phenomena through the Hamiltonian
+of the system and the master equation (which is, in these works, the analogue
 of the Boltzmann transport equation)
-\cite{bib:grabert1974,bib:wang2014,%
+\cite{bib:grabert1974,%
 bib:angel2017,
@@
 are not consistent with the classical-quantum correspondence principle.
-In the same line of thought, some authors have proposed $H$-functionals
+Similarly, some authors have proposed $H$-functionals
 and attempted to proof a
@@
 bib:deroeck2006,bib:acharya2019,bib:kastner2017,bib:han2015,bib:das2018,bib:vonneumann2010}.
-However, whether or not the homogeneous distribution function hypothesis is assumed, or
-if their framework fulfills the correspondence principle is unclear or not discussed whatsoever.
-After the pioneering work of Tolman, at several stages, there has been some discussion regarding
+However, whether or not the homogeneous distribution function hypothesis is assumed or
+if its framework fulfills the correspondence principle is unclear or not discussed.
+Since the pioneering work of Tolman, at several stages, there has been some discussion regarding
 the general validity of the quantum $H$-theorem,
@@
 
-On the other hand, the framework to describe spatially non-homogeneous systems is still
+Nonetheless, the framework to describe spatially non-homogeneous systems is still
 under construction, although several approaches have been developed. For instance,
 the celebrated Onsager formulation (linear thermodynamics)
-\cite{bib:keizer1987,bib:onsager1931} have been successful in describing irreversible chemical 
-and physical phenomena. Nonetheless, some topics, such as describing the internal behavior of glasses
+\cite{bib:keizer1987,bib:onsager1931} has been successful in describing irreversible chemical 
+and physical phenomena. However, some descriptions, such as those the internal behavior of gases
 \cite{bib:zanotto2018} and the entropy measurement
@@
 In addition, some aspects regarding the classical $H$-theorem and the
-Boltzmann $H$-functional require to be revised, for the purpose of improving their mutual consistency.
-For example, the modification of the $H$-theorem to include phenomena stemming from stochastic
-trajectories, violations of the second law of thermodynamics, the relation
-between Shannon's measure of information and the Boltzmann's entropy, 
-and the calculation of thermodynamical quantities and thermalization of certain systems
+Boltzmann $H$-functional require revision to improve their mutual consistency.
+One example is the modification of the $H$-theorem to include phenomena stemming from stochastic
+trajectories, violations of the second law of thermodynamics, the relationship
+between Shannon’s measure of information and the Boltzmann’s entropy, 
+and the calculation of thermodynamical quantities and thermalization of specific systems
 \cite{bib:gorban2014,bib:li2019,bib:gring2012,bib:nemilov2018,bib:wang2014}.
 
-In order to contribute to the construction of a consistent classical and
+To contribute to the construction of a consistent classical and
 quantum $H$-theorem, within a formalism that describes out-of-equilibrium
-non-homogeneous systems, in this article we propose a new theoretical framework.
+non-homogeneous systems, we propose a new theoretical framework.
 Specifically, for both classical and quantum systems, we include non-homogeneous distribution functions in
@@
 The organization of this article is as follows.
-In Section~\ref{sec:classical}, we highlight, to our purposes, the most
+In Section~\ref{sec:classical}, we highlight, for our purposes, the most
 fundamental assumptions required to proof the Boltzmann $H$-theorem,
 we provide an alternative method to obtain the Maxwell-Boltzmann
-distribution using the variational method, followed by our proposal of
-an alternative $H$-functional for classical systems, and the demonstration of
-the respective $H$-theorem . In Section~\ref{sec:quantum}
+distribution using the variational method, propose
+an alternative $H$-functional for classical systems, and demonstrate
+the respective $H$-theorem. In Section~\ref{sec:quantum},
 we review the Tolman proposal for the quantum
@@
 and how the Bose-Einstein and Fermi-Dirac distributions are treated within
-this framework, subsequently we present our proposal for a quantum $H$-functional,
+this framework. Subsequently, we present our proposal for a quantum $H$-functional
 and the proof of the corresponding quantum $H$-theorem. In 
 Section~\ref{sec:qccorrespondence}
-the classical-quantum correspondence between the
-quantum and classical $H$-functionals is analyzed.
+we analyze the classical-quantum correspondence between the
+quantum and classical $H$-functionals.
 In Section~\ref{sec:relaxproc} we explore how relaxation processes occur
-in a quantum ideal gas and, on the basis of what we call variational entropy,
-we propose a time-evolution equation for the distribution function. Finally,
-we discuss some key ideas derived from our approach and close with a brief summary
+in a quantum ideal gas and, based on what we call variational entropy,
+propose a time-evolution equation for the distribution function. Finally,
+we discuss some key ideas resulting from our approach and close with a summary
 in Section~\ref{sec:disscussion}.

### 2021-03-10 tex/qccorrespondence.tex
Cambridge corrections applied.

@@
-\section{Quantum Classical correspondence}\label{sec:qccorrespondence}
+\section{Quantum–classical correspondence}\label{sec:qccorrespondence}
 
-In sections \ref{sec:classical} and \ref{sec:quantum}, we have seen that the variational
+In sections \ref{sec:classical} and \ref{sec:quantum}, we saw that the variational
 method can be applied to $H$-functionals, which correctly describes the behavior of
-classical and quantum dilute gases, as to their respective time-evolution.
+classical and quantum dilute gases, with regard to their respective time-evolution.
 Both $H$-functionals defined in Eqs.~(\ref{eq:cHdef}) and
-(\ref{eq:qHdef}) also recovers the well-known distribution functions, either Maxwell-Boltzmann
+(\ref{eq:qHdef}) also recover the well-known distribution functions, either Maxwell-Boltzmann
 for a classical gas, Fermi-Dirac for a Fermi gas, or Bose-Einstein for a Bose gas. Nevertheless,
 the functionals (\ref{eq:cHdef}) and
-(\ref{eq:qHdef}) are seemingly different, and in this section we will show that both are
+(\ref{eq:qHdef}) are seemingly different, and in this section, we will show they are
 related by the correspondence principle.
@@
 degeneration parameter $\xi\equiv\exp\big(-(\epsilon-\mu)/(k_BT)\big) \gg 1$.
-Alternatively, a more general approach to show the quantum classical correspondence
-consists of analyzing the collapse from the quantum to the classical $H$-functionals,
-within the appropiated limit.
-For the case treated here, this limit is $f_{nM} \approx 0$ because of the following reasons.
-Systems at very low temperatures, where the quantum effects cannot be ignored, are obviously excluded
-from the current analysis. For systems at sufficently large temperature, the particles
+Alternatively, a more general approach to show the quantum–classical correspondence
+consists of analyzing the collapse from the quantum to the classical $H$-functionals
+within the appropriated limit.
+For the case treated here, this limit is $f_{nM} \approx 0$ for several reasons.
+Systems at very low temperatures, in which the quantum effects cannot be ignored, are obviously excluded
+from the current analysis. In systems at sufficiently high temperatures, the particles
 occupy almost exclusively high-energy levels. 
-Furthermore, the energy spectrum approaches to be a continuum, as is expected by taking the limit
-$\hbar\to0$, and the number of particles per level are very close to zero.
+Furthermore, the energy spectrum approaches a continuum, as is expected by taking the limit
+$\hbar\to0$, and the number of particles per level is very close to zero.
 
-Subsequently, we substitute $f_{nM}\approx0$ into Eq.~(\ref{eq:qHdef}) to obtain:
+Subsequently, we substitute $f_{nM}\approx0$ into Eq.~(\ref{eq:qHdef}) to obtain
 %
@@
 
-Finally, the sum over the quatum energy levels can be replaced by an integral over the velocities
+Finally, the sum over the quantum energy levels can be replaced by an integral over the velocities
 by invoking both the uncertainty principle and the fact that, for free particles, the continuum
@@
 can be written as a function of the velocity. Hence the quantum $H$-functional
-transforms, in the classical limit, to:
+transforms, in the classical limit, to
 %
@@
 a function of $\vec v$.
-
-%--------------------------------------

### 2021-03-10 tex/quantum.tex
Removed comments (pdf remains the same)
Cambridge corrections applied.

@@
 
-The classical $H$-theorem is even to these days one of the pillars on which the classical statistical
-physics is founded. Unfortunately, regardless of several attempts that have been made \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:han2015,bib:das2018,bib:vonneumann2010}, 
+The classical $H$-theorem is still considered one pillar on which classical statistical
+physics is founded. Unfortunately, despite multiple attempts \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:han2015,bib:das2018,bib:vonneumann2010}, 
 the generality of the classical $H$-theorem has no equally robust quantum match. In this section, we shall
-propose and analize an alternative quantum $H$-functional using the variational method.
+propose and analyze an alternative quantum $H$-functional using the variational method.
 We start by briefly outlining a typical textbook demonstration of the quantum $H$-theorem
-\cite{bib:tolman}, and subsequently we present the analysis of our proposed $H$-functional.
+\cite{bib:tolman}, and subsequently present the analysis of our proposed $H$-functional.
 
@@
 contained by a vessel of volume
-$V$, at some temperature $T$, and total energy E. Starting from the Boltzmann definition
+$V$, temperature $T$, and total energy E. Starting from the Boltzmann definition
 of entropy, the quantum $H$ functional is
@@
 above conditions \cite{bib:tolman}.
-The quantum $H$-theorem can be demostrated as follows.  $G$ can be divided into groups of neighboring states,
+The quantum $H$-theorem can be demonstrated as follows.  $G$ can be divided into groups of neighboring states,
 $g_k$, and certain occupation numbers, $n_k$, can be associated with each of these
-groups. Thus, the above functional takes the form:
+groups. Thus, the above functional takes the form
 %
@@
 
-Assuming that the energy exchange between particles is produced by collisions between particles,
+Assuming that the energy exchange between particles is produced by interparticle collisions,
 and using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
@@
 %
-In Eq.~(\ref{eq:amnkldef}), $\Delta\epsilon$ is the net energy change occuring during the
+In Eq.~(\ref{eq:amnkldef}), $\Delta\epsilon$ is the net energy change occurring during the
 collision and $|I_1-I_2|^2=|V_{mn,kl}|^2$, where
@@
 It is important to remark that in deriving Eq.~(\ref{eq:changen}), the equal a priori
-probabilities and the random a priori phases hypotheses were taken as valid. The 
-random a priori phases hypothesis can be
+probabilities and the random a priori phase hypotheses were assumed valid. The 
+random a priori phase hypothesis can be
 considered an analogue of the molecular chaos hypothesis
@@
 
-In equilibrium (at $t\to\infty$), $dn_{\kappa}/dt=0$, hence from Eq.~(\ref{eq:changen})
+At equilibrium (at $t\to\infty$), $dn_{\kappa}/dt=0$, hence from Eq.~(\ref{eq:changen})
 %
@@
 %
-Considering that the energy is conserved during the collsion, the Bose-Einstein or Fermi-Dirac distribution
+Considering that energy is conserved during the collision, the Bose-Einstein or Fermi-Dirac distribution
 functions can be recovered from Eq.~(\ref{eq:boseiferdirdistfnctsrc}):
@@
 
-\textit{I.e.}, in equilibrium $dH_T/dt=0$ and the distribution function obtained from
+\textit In other words, at equilibrium $dH_T/dt=0$, the distribution function obtained from
 Eq.~(\ref{eq:quantumh}) is the expected distribution function.
 
-%----------------------------------------------
-\begin{comment}
-In addition, Tolman also showed that (\ref{eq:quantumh}) can be reduced, in the
-high-energy limit, to
-%
-\begin{equation}
-    H_T = \sum_{\kappa} (n_{\kappa} \ln n_{\kappa} - n_{\kappa} \ln g_{\kappa}). \label{reduce-h}
-\end{equation}
-%
-This expression also can be obtained also from the Boltzmann $H$-functional through defining
-%
-\begin{equation}
-    f=\frac{n_{\kappa}}{ g_{\kappa}}.
-\end{equation}
-\end{comment}
-%
-%--------------------------
-
-\subsection{Out of equilibrium, non-homogeneous quantum systems}
+\subsection{Out-of-equilibrium, non-homogeneous quantum systems}
 
@@
 bosons. For our purposes, the volume $V$ is divided into $K$ small cells, each of which
-has a constant volume $\delta V_M=V/K$ ($M=1,\,\dots\,,K$), a temperature $T_M$, an energy $\epsilon_M$,
-a number of particles $\mathcal{N}_M$, and a distribution function,
-$\{f_{Mn}(t)\}$. Hereafter we will use a short-hand notation:
+has constant volume $\delta V_M=V/K$ ($M=1,\,\dots\,,K$), temperature $T_M$, energy $\epsilon_M$,
+number of particles $\mathcal{N}_M$, and distribution function,
+$\{f_{Mn}(t)\}$. Hereafter we will use the following short-hand notation:
 %
@@
 where $\vec r_M$ is the radius vector pointing at the center of the $M$-th cell.
-$f_{Mn}(t)$ represent the number
-of particles contained in the $M$-th cell that occupy the energy level $\epsilon_n$ at
+$f_{Mn}(t)$ represents the number
+of particles contained in the $M$-th cell that occupies the energy level $\epsilon_n$ at
 time $t$. Since the particles are considered to be free, the energy levels should not
 depend on the cell properties, \textit{i.e.} the energy spectrum, $\{\epsilon_n\}$, is the same for all
-cells; thus there is no need to label $\epsilon_n$ with an index $M$.
+cells; thus, there is no need to label $\epsilon_n$ with an index $M$.
 
@@
 %
-Here, the upper and lower sign refer to fermions and bosons respectively.
-In addition, when needed, each cell will have an associated local chemical potential,
-$\alpha_M$, and a local $H$-functional, which is defined by:
+Here, the upper and lower signs refer to fermions and bosons, respectively.
+In addition, when needed, each cell has an associated local chemical potential,
+$\alpha_M$, and a local $H$-functional, which is defined by
 %
@@
 %
-Therefore, $\mathcal{N}_M$ and $\mathcal{E}_M$ as functions of time are given by:
+Therefore, $\mathcal{N}_M$ and $\mathcal{E}_M$ as functions of time are given by
 %
@@
 Lagrange multipliers
-$\{\alpha_M\}$ and $\{\beta_M\}$, it is straightforward to obtain
+$\{\alpha_M\}$ and $\{\beta_M\}$, we readily obtain
 (see also the discussion related to Eq.~(\ref{eq:deltaHpdeltafpj})):
@@
 %
-Thus, in this zero-order approximation, the form of equilibrium distribution functions are conserved.
+Thus, in this zero-order approximation, the form of equilibrium distribution functions is conserved.
 
@@
 the complete system. Also, the local number of particles, the local
-energy, and the Lagrange multipliers do not depend on the cell number and they should be homogeneous,
-this is 
+energy, and the Lagrange multipliers do not depend on the cell number, and they should be homogeneous.
+This is represented by
 %
@@
 Substituting Eqs.~(\ref{eq:qEqAlphaBeta}) into Eq.~(\ref{eq:qfMn}) yields the distribution
-function of each cell in equilibrium to be
+function of each cell in equilibrium:
 %
@@
 %
-By means of the above equation, we can recover the distribution function and the entropy of a dilute quantum gas in equilibrium as follows. Setting $\bar\alpha=\mu/kT$ and $\bar\beta=-1/kT$,
-and substituting them in Eq.~(\ref{eq:qfneq}), it renders the Fermi-Dirac and Bose-Einstein
+Using the above equation, we can recover the distribution function and the entropy of a dilute quantum gas in equilibrium as follows. Setting $\bar\alpha=\mu/kT$ and $\bar\beta=-1/kT$,
+and substituting them into Eq.~(\ref{eq:qfneq}), it renders the Fermi-Dirac and Bose-Einstein
 distribution functions:
@@
 and substituting Eq.~(\ref{eq:qfneq}) into the negative of Eq.(\ref{eq:qHdef}), 
-the entropy of a quantum ideal gas is obtained
+the entropy of a quantum ideal gas is 
 %
@@
 %
-This quantity is what we will refer to as ``variational entropy'',
-and this name reflects the fact that it was obtained \textit{via}
+This quantity is what we will refer to as ``variational entropy,''
+and this name reflects the fact that it was obtained via
 the variational method.
@@
 For quantum systems, we also accept the validity of the local equilibrium hypothesis
-for every cell in the system. This allows to define non-homogeneous systems,
-wherein thermodynamic quantities are well-defined on a cell-per-cell basis,
-and in terms of the equilibrium properties, we have
+for every cell in the system. This allows us to define non-homogeneous systems,
+wherein thermodynamic quantities are well-defined on a per-cell basis.
+In terms of the equilibrium properties, we have
 
@@
 
-In the present context, $|\Delta_M|$ and $|\delta_M|$ are sufficiently large so they are not
+In the present context, $|\Delta_M|$ and $|\delta_M|$ are sufficiently large to not be
 fluctuations of the system, and sufficiently small so that the local equilibrium hypothesis is
 valid for $t>0$ (we set $t_0=0$, and $t_0$ is the initial time at which the system is prepared).
-Therefore, it is reasonable to re-write the distribution functions as:
+Therefore, the distribution functions can be rewritten as
 %
@@
 from which it follows, by substituting Eq.~(\ref{eq:qFirstOrd}) into
-Eq.~(\ref{eq:restrictionoutside}), that $\Delta_M$ and $\delta_M$ satisfy:
+Eq.~(\ref{eq:restrictionoutside}), that $\Delta_M$ and $\delta_M$ satisfy
 %
@@
 %
-$\bar f_{n}=1$ is certainly satisfied if the system is at zero temperature. In this state,
+$\bar f_{n}=1$ is certainly satisfied if the system temperature is zero. In this state,
 all energy levels below and including the Fermi energy are occupied, thus the system will
-necessarily be homogeneous , and consequently $g_{nM}=0$. In this article, we will
-omit this scenario and will only discuss Fermi gases whose temperature is non zero.
+necessarily be homogeneous, and consequently, $g_{nM}=0$. In this article, we will
+omit this scenario and will only discuss Fermi gases with non-zero temperatures.
 
 %\begin{proof}
-To proof the quantum $H$-theorem, start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
+To proof the quantum $H$-theorem, we start by taking the time-derivative of Eq. (\ref{eq:qHdef}):
 %
@@
 %
-Subsequently, substitute Eq.~(\ref{eq:qFirstOrd}) in the above equation to obtain:
+Subsequently, we substitute Eq.~(\ref{eq:qFirstOrd}) into the above equation to obtain
 %
@@
 %
-Approximate the logarithmic terms, corresponding to
-Fermi and Bose gases, through a Taylor series around $\bar f_n g_{nM}=0$ as:
+The logarithmic terms, corresponding to
+Fermi and Bose gases, are approximated through a Taylor series around $\bar f_n g_{nM}=0$ as
 %
@@
 respectively. Eq.~(\ref{eq:qlnApproxFermions}) is valid because, for non-extremely degenerated Fermi gases, 
-$1-\bar{f}_n \gg \bar{f}_n|g_{nM}|$, and Eq.~(\ref{eq:qlnApproxBosons}) is fulfilled because,
+$1-\bar{f}_n \gg \bar{f}_n|g_{nM}|$ and Eq.~(\ref{eq:qlnApproxBosons}) is fulfilled because,
 for Boson gases,
@@
 
-Combine Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}):
+Combining Eqs. (\ref{eq:cambioH1}) and (\ref{eq:qlnApprox}),
 %
@@
     \frac{d\mathcal{H}}{dt}&=&\sum_n \sum_{M=1}^{K} \bar{f}_n\left \{ (\ln \bar{f}_n+ g_{nM})\dot{g}_{nM}\right\} \de V_M \delta \epsilon_n \nonumber \\
-    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M,
+    &&-\sum_{n}\sum_{M=1}^{K}\bar f_n\left\{ \left( \ln[1\mp\bar{f}_n]\mp \left[\frac{\bar{f}_n}{1\mp\bar{f}_n} \right] g_{nM}\right)\dot{g}_{nM} \right \}\de V_M
 \end{eqnarray}
 %
-and substitute Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
+and substituting Eq.~(\ref{eq:qfneq}) into Eq.~(\ref{eq:cambioH2}):
 %
@@
 Since both the total number of particles and the total energy 
-of the system are constant, it follows,
-from Eqs.~(\ref{eq:qGlobRest}) and (\ref{eq:restrictionoutside}) that:
+of the system are constant, it follows
+from Eqs.~(\ref{eq:qGlobRest}) and (\ref{eq:restrictionoutside}) that
 %
@@
 
-Substitute the previous expression in Eq.~(\ref{eq:cambioH3}) to obtain:
+Substitute the previous expression in Eq.~(\ref{eq:cambioH3}) to obtain
 %
@@
 
-For obtaining the right-most hand side of Eq.~(\ref{eq:cambioH4}), 
-we have used the relation $g_{nM}\dot{g}_{nM}\leq0$ for $t>0$. This can be proven 
+To obtain the far right side of Eq.~(\ref{eq:cambioH4}), 
+we have used the relationship $g_{nM}\dot{g}_{nM}\leq0$ for $t>0$. This can be proven 
 by simply arguing that, in the initial
 state, if a cell is described by $g_{nM}(t_0)>0$ then $g_{nM}(t)\geq0$ and $\dot g_{nM}(t)\leq0$, and if
-$g_{nM}(t_0)<0$ then $g_{nM}(t)\leq0$ and $\dot g_{nM}(t)\geq0$. Here we have used the facts that
+$g_{nM}(t_0)<0$ then $g_{nM}(t)\leq0$ and $\dot g_{nM}(t)\geq0$. Here we have exploited the fact that
 the system in equilibrium is homogeneous, and that, by accepting the local equilibrium hypothesis, $g_{Mn}(t)$
-is a monotonic function and $g_{nM}\to0$ as $t\to\infty$ as the system aproaches to the equilibrium state.
+is a monotonic function and $g_{nM}\to0$ as $t\to\infty$ as the system approaches the equilibrium state.
 Another approach to prove Eq.~(\ref{eq:cambioH4}) consists of splitting the cells into two subsets,
@@
 
-Succinctly, considering a dilute quantum gas contained in a vessel of volume $V$ (divided into
+Briefly, considering a dilute quantum gas contained in a vessel of volume $V$ (divided into
 $K$ small cells), with total energy $E$ and $N$ quantum free particles, which initially
-is out of equilibrium ---but in such a manner that the local equilibrium hypothesis is valid---,
+is out of equilibrium---but in such a manner that the local equilibrium hypothesis is valid---
 the functional
@@
 the equilibrium state. In Eq.~(\ref{eq:qHfunctTheo}), and for a Fermi (Bose)
-gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distrubution function for each cell.
+gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distribution function for each cell.
 Locally, each cell is in equilibrium, although the complete system may be non-homogeneous,
-and is caracterized by the respective $f_{Mn}$, number of particles $\mathcal{N}_M$, energy $\mathcal{E}_M$,
-temperature $T_M$, and Legendre Multipliers $\alpha_M$ and $\beta_M$.
+and is characterized by the respective $f_{Mn}$, number of particles $\mathcal{N}_M$, energy $\mathcal{E}_M$,
+temperature $T_M$, and Legendre multipliers $\alpha_M$ and $\beta_M$.

### 2021-03-10 tex/relaxproc.tex
Cambridge corrections applied.

@@
 
-In order to obtain a time-evolution equation for an out-of-equilibrium quantum gas,
+To obtain a time-evolution equation for an out-of-equilibrium quantum gas,
 we propose the following approach. We start by evaluating $\Delta\mathcal{H}=
 \mathcal{H}(t_2)-\mathcal{H}(t_1)$,
-where our quantum $H$-functional ---Eq.~(\ref{eq:qHdef})--- is evaluated at diferent times $t_1$ and $t_2$, with
-$t_2 > t_1$. This yields:
+where our quantum $H$-functional---Eq.~(\ref{eq:qHdef})---is evaluated at different times $t_1$ and $t_2$, with
+$t_2 > t_1$. This yields
 %
@@
 $f_{nM}(t_1) \equiv f'_{nM}$.
-Subsequently, in Eq.~(\ref{eq:deltaHdef}), we replace the distribution functions by their expresions in terms
-of deviations from equilibrium ---Eq.~(\ref{eq:qFirstOrd})---, which renders:
+Subsequently, in Eq.~(\ref{eq:deltaHdef}), we replace the distribution functions with their expressions in terms
+of deviations from equilibrium---Eq.~(\ref{eq:qFirstOrd})---which renders:
 \begin{eqnarray}\label{eq:deltaH1}
@@
 \end{eqnarray}
-Afterwards, we expand the logarithmic terms up to the first-order in $g'_{nM}$ and $g''_{nM}$
-and rearrange the result, which gives:
+Subsequently, we expand the logarithmic terms up to the first-order in $g'_{nM}$ and $g''_{nM}$
+and rearrange the result, which gives
 %
@@
 Finally we divide Eq.~(\ref{eq:deltaH2}) by $\Delta t\equiv t_2-t_1$, and take the limit
-$\Delta t\to0$ to obtain:
+$\Delta t\to0$ to obtain
 %
@@
 Eq.~(\ref{eq:deltaH3}) is, within our framework, the time-evolution equation for $g_{nM}$.
-Clearly, for describing a concrete situation, it is required to provide a specific approximation for the deviation function $g_{nM}$, however this will be the subject of future work.
-
-  %----------------------------------------
-
-%-------------------------------------------------
+Clearly, to describe a realistic situation, providing a specific approximation for the deviation function $g_{nM}$ is required. This subject will be explored in future work.
